### 2021-01-06 inference.tex
add inference

\documentclass{article}

\usepackage{amsmath, amsthm, array, stmaryrd}
\usepackage{booktabs}
\usepackage{todonotes}

\title{Notes on Control Plane Inference}
\author{Eric Campbell}

\newcommand{\NumRows}{\mathsf{numrows}}
\newcommand{\MatchRow}{\mathsf{keys}}
\newcommand{\ActionRow}{\mathsf{action}}
\newcommand{\ActionDataRow}{\mathsf{data}}

\begin{document}

\maketitle

\section{Preliminaries}

Let a data plane program $p \in \mathsf{DProg}$ be defined by the following grammar:

\[\begin{array}{lcl}
    p & ::= & \mathsf{skip}\\
      & \mid & f := e \\
      & \mid & \mathsf{assert}~b \\
      & \mid & \mathsf{assume}~b \\
      & \mid & p;p \\
      & \mid & p [] p \\
      & \mid & t.apply() \\
  \end{array}
\]

Where $b$ is a boolean logic over bitvectors expressions $e$, and
$t$ is a table as defined in Avenir.

Let $\delta$ be Control plane edits defined by the grammar
\[\begin{array}{lcl}
    \delta & ::= & \mathsf{Add}_t(\vec m,a(\vec k)) \\
           & \mid & \mathsf{Del}_t(\vec m) \\             
  \end{array}\]

Write $\delta.\mathit{keys}$ to be the list of keys $\vec m$. When
$\delta$ is an \textsf{Add} operation, write $\delta.\mathit{action}$
to indicate the action $a$, and write $\delta.\mathit{data}$ to
indicate the argument list $\vec k$. For any list $\vec x$ write
$\vec x[i]$ for element $x_i$.

Define a control plane instance $\sigma$ to be a list of edits
$\delta$, and is minimal if there are no deletions and no shadowed
insertions. An insertion $\delta = \mathsf{Add}_t(\vec m, a(\vec k))$
is \emph{shadowed} in an instance $\sigma$ if $\sigma[i] = \delta$ and
every packet that matches $\delta$ matches a preceeding rule
$\sigma[j]$, $0 \leq k < i$. An instance $\sigma$ is minimal for a
program $p$ if for all $\sigma'$ such that
$\llbracket p[\sigma] \rrbracket = \llbracket p[\sigma'] \rrbracket$,
then $|\sigma| \leq |\sigma'|$.

The operation $\sigma(t)$ selects all the operations on table
$t$ in $\sigma$. To apply the edits $\sigma$ to a program $p$ as is
done in Avenir, write $p[\sigma] \in \mathsf{DProg}$.

Let $\varphi$ be a control plane constraint modelled by minimal lists
of edits. For a table $t$ let $t[i]$ be a record
$\{\MatchRow = \vec m; \ActionRow = \vec a; \ActionDataRow = \vec k\}$
representing the $i$th row in that table.  We can describe $\varphi$
by the boolean combinations of (in)equalities over the following
variables.
\begin{center}
\begin{tabular}{lll}
  \toprule
  Variable & Semantics on minimal $\sigma$\\
  \midrule
  $t.\NumRows$ & Denotes the number of edits in table $t$\\
  $t[i].\MatchRow(j)$ & In table $t$ the $j$th key in row $i$\\
  $t[i].\ActionRow$ & In table $t$ the action selected in row $i$\\
  $t[i].\ActionDataRow(j)$ & in table $t$ the $i$th action argument in row $i$\\
  \bottomrule
\end{tabular}
\end{center}

Then the grammar we get is:
\[
  \begin{array}{lcl}
    \varphi &::=& \bot \\
            & \mid & \varphi \Rightarrow \varphi \\
            & \mid & e < e  \\
    e       & ::= & n \\
            & \mid & [n]_w \\
            & \mid & t \\
            & \mid & e[i] \\
            & \mid & e.\NumRows \\
            & \mid & e.\MatchRow \\
            & \mid & e.\ActionRow \\
            & \mid & e.\ActionDataRow \\
  \end{array} 
\]

We use $\psi$ to denote a formula that intermingles the control plane
variables above with dataplane variables like \texttt{ipv4.src\_addr}.

We write the modelling relation $\sigma \models \varphi$ to indicate
that $\sigma$ satisfies $\varphi$ in the standard
way with respect to the above semantics.

\[
  \begin{array}{rcl}
    \sigma \models \bot &\iff & \mathit{never} \\
    \sigma \models \varphi_1 \Rightarrow \varphi_2 & \iff & \sigma \not\models \varphi_1,~\text{or}~\sigma \models \varphi_2 \\
    \sigma \models e_1 < e_2 &\iff& \llbracket e_1 \rrbracket \sigma < \llbracket e_2 \rrbracket \sigma \\\\
    % \llbracket n \rrbracket \sigma & = & n \\
    % \llbracket [n]_w \rrbracket \sigma & = & [n]_w \\
    % \llbracket t \rrbracket \sigma & = & \sigma(t) \\
    % \llbracket e[i] \rrbracket \sigma & = & \sigma(t)[i], \mathit{where},~\llbracket e \rrbracket \sigma = \sigma(t),  \\
    % \llbracket e.\NumRows \rrbracket \sigma & = &  |\sigma(t)|, \mathit{where}~\llbracket e \rrbracket \sigma = \sigma(t)[i], \\    
    % \vdots
  \end{array}
\]

\section{The Problem}

Given a program $p$, compute $\varphi$ such that for all
instantiations $\sigma$, $\sigma \models \varphi$ if (and
only if) one of the following (equivalent expressions) holds:
\begin{enumerate}
\item $\forall \mathit{pkt}. \llbracket p[\sigma]\rrbracket~\mathit{pkt} \neq \bot$
\item $\mathit{wp}(p[\sigma], \top)$ is valid
\end{enumerate}

Our proposed solution is to instrument $p$ with ghost variables to
(completely) describe the unknowns in the program. Write the
instrumented program as $\mathcal{I}(p)$. An immediate strategy is to
replace any table application $t.\mathit{apply}()$, where
$t = (\mathsf{keys}= \vec k, \mathsf{actions} = \overrightarrow{a(\vec
  d)})$ which statically has a maxmimum number of rows $n$ with a
table of the following form:
\[
  \begin{array}{ll}
    \multicolumn{2}{l}{\mathsf{assume} (t.\mathsf{numrows} = 0) []}\\
    {[]}_{1 \leq r \leq n}
    & \mathsf{assume}(t.\mathsf{numrows} \geq r)\\
    & \mathsf{assume}(\bigwedge_l k_l = t[r].\MatchRow(l)); \\
    & {[]}_j (\mathsf{assume}(j = t[r].\ActionRow); \\
    & \quad\; \mathsf{assume}(\bigwedge_i d_i = t[r].\ActionDataRow(i)); \\
    & \quad\; a_j(\vec d)) \\
    ) 
  \end{array}
\]
This is the approach taken by Avenir's strawman simple synthesis
algorithm. P4v doesn't do such an enumeratively crazy instrumentation
procedure---it uses only one table row $t[0]$ rather than $n$ table
rows. More care is needed to figure out whether these are different or
the same.

Once we have this instrumentation, we can compute a verification
condition (VC), to constrain the ``safe'' behavior of the instrumented
program $\mathcal{I}(p)$. For the time being, we'll use the weakest
precondition predicate transformer
$\mathit{wp}(\mathcal{I}(p), \top)$, and notate the result as
$\psi_p$. Notice that this formula $\psi_p$ may contain both
dataplane variables, like $\texttt{ipv4.srcAddr}$ and control plane
ghost variables, like $t[i].\MatchRow(l)$. Write
$\psi_p[\vec{\texttt{d}},\vec{\mathsf{c}}]$ to denote that the free
variables of $\psi_p$ are the union of the dataplane variables
$\{\texttt{d}_1, \ldots,\texttt{d}_n\}$ and the control plane variables
$\{\mathsf{c}_1, \ldots, \mathsf{c}_m\}$.

Now we can reformulate the problem entirely logically. Given a program
$p$ and its corresponding instrumented verification condition
$\psi_p$, compute a controller constraint $\varphi$ such that the following equation is valid
\[\forall \vec{\texttt{d}}.~\psi_p[\vec{\texttt{d}},\vec{\mathsf{c}}] \iff \varphi[\vec{\mathsf{c'}}] \]
Now we only need to investigate a quantifier elimination algorithm for
the variables $\vec{\texttt{d}}$. The standard algorithm (replace
every $\texttt{d}_i$ of sort $\texttt{(\_~BitVec~n)}$ with every
element in $\mathbf{2}^n$) for fixed-width bitvector theories is
intractable: a standard ethernet frame is $1500$ bits, and the number
of corresponding disjunctions is $2^{1500} = 10^{451.5\ldots}$, which
is far greater than the $10^{82}$ atoms in the universe.

However, I believe we can support a complete collection of rewrite
rules to eliminate these variables based leveraging the known
invariants about the program \& formula structures.

\section{An Incremental Problem Statement}

Writing an instrumentation function for these indexed variables seems
difficult. Instead lets just concentrate on a single new insertion at
a time into a table $t$. This way, we can simply just insert the
instrumented edit
\[?\delta_t = \mathsf{Add}(\langle t.\MatchRow(0),\ldots,
  t.\MatchRow(n-1)\rangle, t.\ActionRow(t.\ActionDataRow(0), \ldots,
  t.\ActionDataRow(m-1)))\] where $n$ is the number of keys in table
$t$, and $m$ is the maximum number of action data parameters that any
action in $t$ has. For this incremental solution we can omit the
subscripts, because there's only one rule being added.

\begin{quote}
  \it Given a program $p$, a table $t$, and a minimal instantiation
  $\sigma$ compute the weakest formula $\varphi$ such that
  $\mathit{wp}(p[\sigma;?\delta_t],\top) \Rightarrow \varphi$.
\end{quote}

\subsection{One Possible Solution}

We can write a formula $\psi = \mathit{wp}(p[\sigma;\varphi],\top)$, as
\[\forall d_1,\ldots, d_n. \psi[c_1,\ldots,c_n,d_1,\ldots,d_m]\]
where the variables $d_i$ are dataplane variables like
\texttt{ipv4.src\_addr} and the variables $d_1, \ldots, d_m$ are
controllable variables like $t.\MatchRow(j)$, and
$\{d_1,\ldots, d_m, c_1,\ldots, c_n\}$ are all the free variables of $\varphi_{\mathit{vc}}$

Then, compute a solution $\varphi$ by performing quantifier
elimination on the $d_1,\ldots,d_n$.

This is probably untenable to do in the brute-force way. Is there another way?

\subsection{Failed Horn Encoding}

Given the above formula $\psi$, solve following CHC for $H$

\[H \leftarrow \forall \vec d.\psi[\vec d, \vec c]\]

This will trivially give True.

\end{document}

### 2021-02-03 inference.tex
write notes about bug

@@
 
-\usepackage{amsmath, amsthm, array, stmaryrd}
+\usepackage{amsmath, amsthm, array, stmaryrd, latexsym}
 \usepackage{booktabs}
 \usepackage{todonotes}
+\usepackage{float}
 
@@
 \begin{enumerate}
-\item $\forall \mathit{pkt}. \llbracket p[\sigma]\rrbracket~\mathit{pkt} \neq \bot$
+\item $\forall \mathit{pkt}. \llbracket p[\sigma]\rrbracket~\mathit{pkt} \neq \mathsf{error}$
 \item $\mathit{wp}(p[\sigma], \top)$ is valid
@@
 $\psi_p$, compute a controller constraint $\varphi$ such that the following equation is valid
-\[\forall \vec{\texttt{d}}.~\psi_p[\vec{\texttt{d}},\vec{\mathsf{c}}] \iff \varphi[\vec{\mathsf{c'}}] \]
+%% \[(\forall \vec{\texttt{d}}.~\psi_p[\vec{\texttt{d}},\vec{\mathsf{c}}]) \iff \varphi[\vec{\mathsf{c'}}] \]
+\[\varphi[\vec{\mathsf{c}}] \Rightarrow \psi_p[\vec{\texttt{d}},\vec{\mathsf{c}}] \]
+
 Now we only need to investigate a quantifier elimination algorithm for
@@
 
-\section{An Incremental Problem Statement}
+\section{Examples}
 
-Writing an instrumentation function for these indexed variables seems
-difficult. Instead lets just concentrate on a single new insertion at
-a time into a table $t$. This way, we can simply just insert the
-instrumented edit
+There are a few questions we want to investigate with examples.
+\begin{enumerate}
+\item \label{question:rowiter}
+  Do we need to consider multiple rows at the same time? In other words, do
+  we need to instrument each table $t$ with a different row for \(t[0]\ldots
+  t[t.\mathsf{maxrows}]\) or is it sufficient to, like P4v, only consider a
+  single row at a time?
+\item \label{question:elimination}
+  Can we eliminate all references to dataplane state in the conditions we
+  generate?
+\end{enumerate}
+
+\subsection{A Multi-Table Example}
+We will construct a pipeline with 3 tables, \(t_i\) for \(i=0,1,2\),
+where each \(t_i\) matches on \(x_i\) and assigns a value to
+\(y_i\).
+\begin{verbatim}
+table ti {
+  keys : xi
+  actions : ai
+  default : ai(i)
+  max_rows : 2
+}
+action ai(y){
+  yi := y
+}
+\end{verbatim}
+
+The control block applies \(t_0\), and then conditionally applies
+\(t_1\) or \(t_2\). The assertion is that \(t_i\) is even for all \(i=0,1,2\).
+
+\begin{verbatim}
+y0 := 0;
+y1 := 0;
+y2 := 0;
+t0.apply();
+if (y0%3 = 0) {
+  t1.apply()
+} else {
+  t2.apply()
+}
+assert y0%2 = 0;
+assert y1%2 = 0;
+assert y2%2 = 0;
+\end{verbatim}
+
+Now, we can instrument this program using the ghost state defined above.
+\begin{verbatim}
+y0 := 0; y1 := 0; y2 := 0;
+assume t0[0].action <= 0;
+assume t0[1].action <= 0;
+assume t0.numrows <= 2;
+{ assume t0.numrows > 0 && t0[0].keys(x0) = x0;
+  assume t0[0].action = 0;
+  y0 := t0[0].data(y)
+} [] {
+ assume not (t0.numrows > 0 && t0[0].keys(x0) = x0);
+ assume t0.numrows > 1  && t0[1].keys(x0) = x0;
+ assume t0[1].action = 0;
+ y0 := t0[1].data(y)
+} [] {
+ assume not (t0.numrows > 0 && t0[0].keys(x0) = x0);
+ assume not (t0.numrows > 1 && t0[1].keys(x0) = x0);
+ y0 := 0
+};
+{assume y0%3 = 0;
+  assume t1[0].action <= 0;
+  assume t1[1].action <= 0;
+  assume t1.numrows <= 2;
+  { assume t1.numrows > 0 && t1[0].keys(x1) = x1;
+    assume t1[0].action = 0;
+    y1 := t1[0].data(y)
+  } [] {
+    assume not (t1.numrows > 0 && t1[0].keys(x1) = x1);
+    assume t1.numrows > 1  && t1[1].keys(x1) = x1;
+    assume t1[1].action = 0;
+    y1 := t1[1].data(y)
+  } [] {
+    assume not (t1.numrows > 0 && t1[0].keys(x1) = x1);
+    assume not (t1.numrows > 1 && t1[1].keys(x1) = x1);
+    y1 := 1
+  }
+} [] {
+  assume y0%3 != 0;
+  assume t2[0].action <= 0;
+  assume t2[1].action <= 0;
+  assume t2.numrows <= 2;
+  { assume t2.numrows > 0 && t2[0].keys(x2) = x2;
+    assume t2[0].action = 0;
+    y2 := t2[0].data(y)
+  } [] {
+    assume not (t2.numrows > 0 && t2[0].keys(x2) = x2);
+    assume t2.numrows > 1  && t2[1].keys(x2) = x2;
+    assume t2[1].action = 0;
+    y2 := t2[1].data(y)
+  } [] {
+    assume not (t2.numrows > 0 && t2[0].keys(x2) = x2);
+    assume not (t2.numrows > 1 && t2[1].keys(x2) = x2);
+    y2 := 2
+  }
+}
+assert y0 % 2 = y1 % 2 = y2 % 2 = 0
+\end{verbatim}
+
+And finally, we can compute the weakest precondition to produce a (universally
+closed) condition on variables.
+
+\[
+\begin{array}{l}
+  t_0.\NumRows > 0 \wedge t_0[0].\MatchRow(x_0) = x_0 \\
+  \Rightarrow
+  \begin{array}[t]{l}
+    t_0[0].\ActionDataRow(y) \% 3 = 0 \Rightarrow
+    \begin{array}[t]{l}
+      t_1.\NumRows > 0 \wedge t_1[0].\MatchRow(x_1) = x_1 \\
+      \qquad \Rightarrow t_1[0].\ActionDataRow(y) \% 2 = 0 \wedge t_0[0].\ActionDataRow(y)\% 2 = 0 \\
+      \wedge (\neg (t_1.\NumRows > 0 \wedge t_1[0].\MatchRow(x_1) = x_1) \\
+      \phantom{\wedge(}\wedge (t_1.\NumRows > 1 \wedge t_1[1].\MatchRow(x_1) = x_1) \\
+      \phantom{\wedge(}\Rightarrow t_1[1].\ActionDataRow(y) \% 2 = 0 \wedge t_0[1].\ActionDataRow(y)\% 2 = 0 )\\
+      \wedge (\neg (t_1.\NumRows > 0 \wedge t_1[0].\MatchRow(x_1) = x_1) \\
+      \phantom{\wedge (}\wedge \neg (t_1.\NumRows > 1 \wedge t_1[1].\MatchRow(x_1) = x_1)) \\
+    \end{array} \\
+    \wedge\,t_0[0].\ActionDataRow(y) \% 3 \neq 0 \Rightarrow
+    \begin{array}[t]{l}
+      t_2.\NumRows > 0 \wedge t_2[0].\MatchRow(x_2) = x_2 \\
+      \qquad \Rightarrow t_2[0].\ActionDataRow(y) \% 2 = 0 \wedge t_0[0].\ActionDataRow(y)\% 2 = 0 \\
+      \wedge (\neg (t_2.\NumRows > 0 \wedge t_2[0].\MatchRow(x_2) = x_2) \\
+      \phantom{\wedge(}\wedge (t_2.\NumRows > 1 \wedge t_2[1].\MatchRow(x_2) = x_2) \\
+      \phantom{\wedge(}\Rightarrow t_2[1].\ActionDataRow(y) \% 2 = 0 \wedge t_0[1].\ActionDataRow(y)\% 2 = 0 )\\
+      \wedge (\neg (t_2.\NumRows > 0 \wedge t_2[0].\MatchRow(x_2) = x_2) \\
+      \phantom{\wedge (}\wedge \neg (t_2.\NumRows ? 1 \wedge t_2[1].\MatchRow(x_2) = x_2) \\
+      \phantom{\wedge (} t_0[1].\ActionDataRow(y) \% 2 = 0)
+    \end{array}
+  \end{array} \\
+  t_0.\NumRows > 1 \wedge t_0[1].\MatchRow(x_0) = x_0 \\
+  \wedge \neg (t_0.\NumRows > 0 \wedge t_0[0].\MatchRow(x_0) = x_0 ) \\
+  \Rightarrow
+  \begin{array}[t]{l}
+    t_0[1].\ActionDataRow(y) \% 3 = 0 \Rightarrow
+    \begin{array}[t]{l}
+      t_1.\NumRows > 0 \wedge t_1[0].\MatchRow(x_1) = x_1 \\
+      \qquad \Rightarrow t_1[0].\ActionDataRow(y) \% 2 = 0 \wedge t_0[1].\ActionDataRow(y)\% 2 = 0 \\
+      \wedge (\neg (t_1.\NumRows > 0 \wedge t_1[0].\MatchRow(x_1) = x_1) \\
+      \phantom{\wedge(}\wedge (t_1.\NumRows > 1 \wedge t_1[1].\MatchRow(x_1) = x_1) \\
+      \phantom{\wedge(}\Rightarrow t_1[1].\ActionDataRow(y) \% 2 = 0 \wedge t_0[1].\ActionDataRow(y)\% 2 = 0 )\\
+      \wedge ((t_1.\NumRows > 0 \wedge t_1[0].\MatchRow(x_1) = x_1) \\
+      \phantom{\wedge (}\vee (t_1.\NumRows ? 1 \wedge t_1[1].\MatchRow(x_1) = x_1)) \\
+    \end{array} \\
+    \wedge\,t_0[1].\ActionDataRow(y) \% 3 \neq 0 \Rightarrow
+    \begin{array}[t]{l}
+      t_2.\NumRows > 0 \wedge t_2[0].\MatchRow(x_2) = x_2 \\
+      \qquad \Rightarrow t_2[0].\ActionDataRow(y) \% 2 = 0 \wedge t_0[1].\ActionDataRow(y)\% 2 = 0 \\
+      \wedge (\neg (t_2.\NumRows > 0 \wedge t_2[0].\MatchRow(x_2) = x_2) \\
+      \phantom{\wedge(}\wedge (t_2.\NumRows > 1 \wedge t_2[1].\MatchRow(x_2) = x_2) \\
+      \phantom{\wedge(}\Rightarrow t_2[1].\ActionDataRow(y) \% 2 = 0 \wedge t_0[1].\ActionDataRow(y)\% 2 = 0 )\\
+      \wedge (\neg (t_2.\NumRows > 0 \wedge t_2[0].\MatchRow(x_2) = x_2) \\
+      \phantom{\wedge (}\wedge \neg (t_2.\NumRows ? 1 \wedge t_2[1].\MatchRow(x_2) = x_2) \\
+      \phantom{\wedge (} t_0[1].\ActionDataRow(y) \% 2 = 0)
+    \end{array}
+  \end{array} \\
+  \neg (t_0.\NumRows > 0 \wedge t_0[0].\MatchRow(x_0) = x_0) \\
+  \wedge \neg (t_0.\NumRows > 1 \wedge t_0[1].\MatchRow(x_0) = x_0 ) \\
+  \Rightarrow
+  \begin{array}[t]{l}
+    t_1.\NumRows > 0 \wedge t_1[0].\MatchRow(x_1) = x_1 \\
+    \qquad \Rightarrow t_1[0].\ActionDataRow(y) \% 2 = 0 \\
+    \wedge (\neg (t_1.\NumRows > 0 \wedge t_1[0].\MatchRow(x_1) = x_1) \\
+    \phantom{\wedge(}\wedge (t_1.\NumRows > 1 \wedge t_1[1].\MatchRow(x_1) = x_1) \\
+    \phantom{\wedge(}\Rightarrow t_1[1].\ActionDataRow(y) \% 2 = 0 )\\
+    \wedge ((t_1.\NumRows > 0 \wedge t_1[0].\MatchRow(x_1) = x_1) \\
+    \phantom{\wedge (}\vee (t_1.\NumRows ? 1 \wedge t_1[1].\MatchRow(x_1) = x_1)) \\
+    \end{array}
+\end{array}
+\]
+
+\subsubsection{Question~\ref{question:rowiter}}
+
+For this example, its clear that we dont need the cases for both \(t_i[0]\) and
+\(t_i[1]\) for $i = 1,2,3$, so we can equivalently compute the following
+(universally closed) verification condition.
+
+\[
+\begin{array}{l}
+  t_0.\MatchRow(x_0) = x_0 \\
+  \Rightarrow
+  \begin{array}[t]{l}
+    t_0.\ActionDataRow(y) \% 3 = 0 \Rightarrow
+    \begin{array}[t]{l}
+      t_1.\MatchRow(x_1) = x_1 \\
+      \qquad \Rightarrow t_1.\ActionDataRow(y) \% 2 = 0 \wedge t_0.\ActionDataRow(y)\% 2 = 0 \\
+      \wedge \neg (t_1.\MatchRow(x_1) = x_1 \Rightarrow \bot) \\
+    \end{array} \\
+    \wedge\,t_0.\ActionDataRow(y) \% 3 \neq 0 \Rightarrow
+    \begin{array}[t]{l}
+      t_2.\MatchRow(x_2) = x_2 \\
+      \qquad \Rightarrow t_2.\ActionDataRow(y) \% 2 = 0 \wedge t_0.\ActionDataRow(y)\% 2 = 0 \\
+      \wedge (\neg (t_2.\MatchRow(x_2) = x_2) \\
+      \qquad \Rightarrow t_0.\ActionDataRow(y) \% 2 = 0)
+    \end{array}
+  \end{array} \\\\
+  \wedge \neg (t_0.\MatchRow(x_0) = x_0)
+  \Rightarrow
+  \begin{array}[t]{l}
+    t_1.\MatchRow(x_1) = x_1 \Rightarrow t_1.\ActionDataRow(y) \% 2 = 0 \\
+    \wedge \, \neg(t_1.\MatchRow(x_1) = x_1 \Rightarrow \bot)\\
+    \end{array}
+\end{array}
+\]
+which is similar to p4v's approach.
+
+\subsubsection{Question~\ref{question:elimination}}
+
+Looking at these examples, its not clear that we can eliminate the dataplane
+state, which occurs only in the form of match conditions. Specifically,
+dataplane state occurs only in the formula as expressions the form
+\(t_i.\MatchRow(x_i) = x_i \Rightarrow \varphi\), where \(\varphi\) is some
+further formula that may intermingle control and dataplane state.
+
+Inspecting the formula above, my intuition tells me that we should be able to
+come up with a rewrite rule that replaces these occurences with \(\top\),
+resulting in a formula with shape.
+
+\[
+\begin{array}{l}
+  \begin{array}[t]{l}
+    t_0.\ActionDataRow(y) \% 3 = 0 \Rightarrow
+    \begin{array}[t]{l}
+      t_1.\ActionDataRow(y) \% 2 = 0 \wedge t_0.\ActionDataRow(y)\% 2 = 0 \\
+    \end{array} \\
+    \wedge\,t_0.\ActionDataRow(y) \% 3 \neq 0 \Rightarrow
+    \begin{array}[t]{l}
+      t_2.\ActionDataRow(y) \% 2 = 0 \wedge t_0.\ActionDataRow(y)\% 2 = 0 \\
+    \end{array}
+  \end{array} \\
+  \,\,\,\wedge\, t_1.\ActionDataRow(y) \% 2 = 0 \\
+\end{array}
+\]
+
+which constraints only controllable variables.
+
+However, we have completely lost the fact that no packet may miss in table
+\(t_1\), which is encoded in the raw formula. More care is needed here.
+
+\subsection{Existential Example}
+
+For a while I've been concerned with the issue of control plane conditions of
+the form \[\forall \rho_1 \in t_1, \exists \rho_2 \in t_2,\] because I was
+afraid that this row existence wouldn't be captured in the predicate transformer
+semantics. This is related to Question~\ref{question:rowiter}---the multiplicity
+of rows is intended to enable this kind of fine-grained approach.
+
+It seems that we \emph{can} recover the existential information from the weakest
+preconditions constructed from instrumenting with only a single rule.
+
+Consider the following program that focuses on a single header instance \texttt{h}.
+\begin{figure}[H]
+  \begin{minipage}{0.32\textwidth}
+\begin{verbatim}
+control {
+  out := 0;
+  h.setInvalid();
+  t1.apply();
+  if(h.isValid()){
+    t2.apply();
+    assert out!=0;
+  }
+}
+\end{verbatim}
+  \end{minipage} \begin{minipage}{0.39\textwidth}
+\begin{verbatim}
+table t1 {
+  keys = x : exact;
+  acts : set_h
+  default_action : nop
+}
+table t2 {
+  keys = h.f:exact;
+  acts = {nop; set_out}
+  default_action = nop;:
+}
+\end{verbatim}
+  \end{minipage} \begin{minipage}{0.27\textwidth}
+\begin{verbatim}
+action set_h(f)  {
+  h.setValid();
+  h.f := x;
+}
+action set_out (o) {
+  out := o
+}
+action nop () {}
+\end{verbatim}
+\end{minipage}
+\end{figure}
+
+Notice that the control plane must satisfy the following condtion for the assertion to never fail:
+\[ \forall \rho_1 \in t_1.~\exists \rho_2 \in t_2.~\rho_2.\MatchRow(h.f) = \rho_1.\ActionDataRow(f) \wedge \rho_2.\ActionDataRow(o) \neq 0 \]
+
+If we follow the single-row-per-table instrumentation strategy, we compute the
+following condition:
+
+\[\begin{array}{l}
+t_1.\MatchRow(x) = x \Rightarrow \\
+\quad t_2.\MatchRow(h.f) = t_1.\ActionDataRow(f) \\
+\quad \wedge t_2.\ActionDataRow(o) \neq 0
+\end{array}\]
+
+which bears a striking resemblence to the condition above. We can typset them side-by-side to highlight the similarities.
+
+\[\begin{array}{lcl}
+t_1.\MatchRow(x) = x \Rightarrow && \forall \rho_1 \in t_1. \\
+\quad t_2.\MatchRow(h.f) = t_1.\ActionDataRow(f) & \leadsto & \quad \exists \rho_2 \in t_2.~\rho_2.\MatchRow(h.f) = \rho_1.\ActionDataRow(f) \\
+\qquad \wedge t_2.\ActionDataRow(o) \neq 0 && \qquad \wedge\rho_2.\ActionDataRow(o) \neq 0
+\end{array}
+\]
+
+This provides evidence for the p4v approach, in repsonse to Question~\ref{question:rowiter}.
+
+\subsection{Hypothesizing Transformations}
+
+In response to Question~\ref{question:eliminations}, we can hypothesize the
+following transformation rules in response to the above example, where
+\(t.\mathsf{o}\) represents an arbitrary access to table \(t\), \(\texttt{d}\)
+represents dataplane state,
+
+\[\begin{array}{l>{\leadsto}cr}
+t.\MatchRow(x) = x \rightarrow \phi && \forall \rho \in t. \phi \\
+t.\mathsf{o} = t'.\mathsf{o'} \wedge \phi && \exists \rho \in t'. \phi \\
+t.\mathsf{o} = n && \rho.\mathsf{o} = n \\
+\end{array}\]
+
+An example of each of these can be seen in the example above.
+
+However, these rules as stated are incorrect, they must aqcuire and maintain
+some context. For example, in the final rule, mapping tests liket $t.\mathsf(o)
+= 0$, we need to know which \(\rho\) row-variable corresponds to the \(t\).
+
+Not only that, but we also need to know how to bind the variables. For example,
+how would we translate a rule like
+\[t_1.\MatchRow = 6 \Rightarrow t_2.\MatchRow = 7\]
+
+The set of hypothesized rules suggests that this should become
+\[\forall \rho_1 \in t_1. \rho_1.\MatchRow = 6 \Rightarrow \exists \rho_2 \in t_2. \rho_2.\MatchRow = 7\]
+
+Some random thoughts
+\begin{enumerate}
+\item Context seems important
+\item Can we generalize this into only introducing one operator? It seems like
+  both of the quantifier introduction rules could be replaced by a local
+  \(\exists\)-quantifiaction,
+\item We need more examples and a more formal approach. Can we derive these rules?
+\end{enumerate}
+
+\subsection{Additional Examples}
+More examples are needed to investigate
+Questions~\ref{question:rowiter}\&\ref{question:elimination}.
+
+\clearpage
+\section{(OUTDATED) An Incremental Problem Statement}
+
+Writing an instrumentation function for these indexed variables seems difficult.
+Instead lets just concentrate on a single new insertion at a time into a table
+$t$. This way, we can simply just insert the instrumented edit
 \[?\delta_t = \mathsf{Add}(\langle t.\MatchRow(0),\ldots,

### 2021-03-26 inference.tex
formalism

--- 2021-01-06 inference.tex
@@
 
-\usepackage{amsmath, amsthm, array, stmaryrd}
+\usepackage{amsmath, amssymb, amsthm, array, stmaryrd}
+\usepackage{wasysym}
 \usepackage{booktabs}
 \usepackage{todonotes}
+\usepackage{float}
+%% \usepackage[notext,not1,nomath]{stix}
+\usepackage{enumitem}
 
@@
 
+% semantics
+\newcommand{\pkt}{\mathit{pkt}}
+\newcommand{\denote}[1]{\llbracket#1\rrbracket}
+\newcommand{\edenote}[1]{\mathcal{E}\denote{#1}}
+
+%booleans
+\newcommand{\TRUE}{\mathsf{tt}}
+\newcommand{\FALSE}{\mathsf{ff}}
+
+%% bitvectors
+\newcommand{\binop}{\mathbin{\oplus}}
+\newcommand{\unop}{\mathop{\text{\pointer}}}
+
+% sets
+\newcommand{\Value}{\mathsf{Value}}
+\newcommand{\BVExpr}{\mathsf{BVExpr}}
+\newcommand{\BExpr}{\mathsf{BExpr}}
+\newcommand{\Cmd}{\mathsf{Cmd}}
+\newcommand{\Pkt}{\mathsf{Pkt}}
+\newcommand{\Model}{\mathsf{Model}}
+\newcommand{\Hdr}{\mathsf{Hdr}}
+\newcommand{\Field}{\mathsf{Field}}
+\newcommand{\Table}{\mathsf{Table}}
+\newcommand{\Row}{\mathsf{Row}}
+\newcommand{\List}{\mathsf{List}}
+\newcommand{\Var}{\mathsf{Var}}
+\newcommand{\SymbRow}{\mathsf{SymbRow}}
+
+% record fields
+\newcommand{\matches}{\mathsf{matches}}
+\newcommand{\action}{\mathsf{action}}
+\newcommand{\actions}{\mathsf{actions}}
+\newcommand{\keys}{\mathsf{keys}}
+\newcommand{\data}{\mathsf{data}}
+\newcommand{\id}{\mathsf{id}}
+
+% commands
+\newcommand{\assert}{\mathop{\mathsf{assert}}}
+\newcommand{\assume}{\mathop{\mathsf{assume}}}
+\newcommand{\apply}{\mathsf{apply}}
+\newcommand{\choice}{\mathbin{\rotatebox[origin=c]{90}{$\sqsubset\!\!\!\sqsupset$}}}
+\newcommand{\SKIP}{\mathsf{skip}}
+
+%math
+\newcommand{\argmin}{\mathop{\textrm{argmin}}}
+
+% Old
 \newcommand{\NumRows}{\mathsf{numrows}}
@@
 
-Let a data plane program $p \in \mathsf{DProg}$ be defined by the following grammar:
+\subsection{Syntax}
 
-\[\begin{array}{lcl}
-    p & ::= & \mathsf{skip}\\
-      & \mid & f := e \\
-      & \mid & \mathsf{assert}~b \\
-      & \mid & \mathsf{assume}~b \\
-      & \mid & p;p \\
-      & \mid & p [] p \\
-      & \mid & t.apply() \\
+\begin{figure}[htp]
+  \[\begin{array}{lcll}
+    \multicolumn{3}{l}{\Value} \\
+    v & ::= & [n]_w  & \textit{Bitvector Literal}\\
+    \multicolumn{3}{l}{\SymbRow} \\
+    \rho & ::= & \left\{
+    \begin{array}{l}
+      \matches = \List[x];\\
+      \action = x;\\
+      \data = \List[x]
+    \end{array}\right\} & \textit{Symbolic Row}\\
+    \multicolumn{3}{l}{\BVExpr} \\
+    e & ::= & v & \textit{Value}\\
+      & \mid & h.f & \textit{Header Access} \\
+      & \mid & \rho.\matches[i] & \textit{Symbolic Match} \\
+      & \mid & \rho.\action & \textit{Symbolic Action}\\
+      & \mid & \rho.\data[i] & \textit{Symbolic Action Data}\\
+      & \mid & x & \textit{Variable}\\
+      & \mid & e \binop e & \textit{Arbitrary Binary Operation}\\
+    & \mid & \unop e & \textit{Arbitrary Unary Operation}\\
+    \multicolumn{3}{l}{\Table} \\
+    t & ::= & \left\{\begin{array}{l}
+    t.\id = v; \\
+    t.\keys = \List[h.f]; \\
+    t.\actions = \List[\lambda \vec x. c]\\
+    \end{array}\right\} & \textit{Table}\\
+    \multicolumn{3}{l}{\BExpr} \\
+    b & ::= & \FALSE  & \textit{Falsity}\\
+      & \mid & b \Rightarrow b & \textit{Implication} \\
+      & \mid & e = e & \textit{Bitvector Equality}\\
+    & \mid & \forall \rho \in t.~b & \textit{Row Universality}\\
+    \multicolumn{3}{l}{\Cmd} \\
+    c & ::=  & h.f := e & \textit{Assignment}\\
+      & \mid & \assert b & \textit{Assertion}\\
+      & \mid & \assume b & \textit{Assumption}\\
+      & \mid & t.\apply() & \textit{Table Application}\\
+      & \mid & c;c & \textit{Sequential Composition} \\
+    & \mid & c \choice c & \textit{Nondeterministic Choice}\\
+    \multicolumn{3}{l}{
+      \begin{array}{ll}
+        x \in \Var & h \in \Hdr\\
+        n,w,i \in \mathbb{N} & f \in \Field
+    \end{array}} & \textit{Sets }\\
   \end{array}
 \]
+\caption{Grammar of Dataplane Programs}
+\label{fig:grammar}
+\end{figure}
 
-Where $b$ is a boolean logic over bitvectors expressions $e$, and
-$t$ is a table as defined in Avenir.
+The following sections describe aspects of the grammar shown in Figure~\ref{fig:grammar}.
 
-Let $\delta$ be Control plane edits defined by the grammar
-\[\begin{array}{lcl}
-    \delta & ::= & \mathsf{Add}_t(\vec m,a(\vec k)) \\
-           & \mid & \mathsf{Del}_t(\vec m) \\             
+\paragraph{Packets}
+Our dataplane programs will reason about structured representations of packets,
+that is, they will be transformations on sets of packet headers. In programs,
+packet headers $h \in \Hdr$ are essentially structs with fields $f \in
+\Field$. Write $h.f$ for a header $h$ and a field $f$ is the value $v$ of $f$
+in header $h$.
+
+\paragraph{Tables}
+A table $t \in \Table$ is a record with a few fields described below:
+\begin{enumerate}[align=left]
+  \item[($t.\id$)] A table's unique identifier. A well-formed program can only have
+    one occurence of each table identifier.
+  \item[($t.\keys$)] A list of header field accesses indicating the match keys.
+  \item[($t.\actions$)] A list of actions that can be executed. An action is a lambda
+    expression $\lambda \vec x. c$ where the variables $x$ are in scope for the
+    commands $c$, which $c$ are straight-line code only (see below).
+\end{enumerate}
+For now we assume that the default action is the no-operation command (see
+$\SKIP$ below). We use $|l|$ to indicate the number of elements in a list or vector $l$
+such as $t.\keys$ or $t.\actions$.
+
+\paragraph{Symbolic Table Rows}
+A symbolic table row $\rho \in \SymbRow$ represents a symbolic row. Like tables,
+rows are also records with the following fields \todo{Well formedness constraint?}
+\begin{enumerate}[align=left]
+  \item[($\rho.\matches$)] A list variable of length representing the possible match keys.
+  \item[($\rho.\action$)] A variable representing the action choice
+  \item[($\rho.\data$)] A list variable representing the possible action data
+\end{enumerate}
+
+To represent the $i$th match variable in $\rho$, use standard list
+access notation: $\rho.\matches[i]$.
+
+Note that symbolic table rows are only meant to be used in the instrumentation
+stage and not by a programmer.
+
+\paragraph{Values}
+Values $v \in \BVExpr$ are bitvector literals, written $[n]_w$, where $n$
+is a natural value and $w$ is the width. We say that $[n]_w = [m]_l$ is
+undefined if $w \neq l$, and otherwise is true iff $n \equiv m \mod 2^w$.
+
+\paragraph{Bitvector Expressions}
+Bitvector expressions $e \in \BVExpr$ can be values $v$, header accesses $h.f$,
+symbolic row variables, variables $x$, or any
+(currently unspecified) binary or unary bitvector operation.
+
+\paragraph{Boolean Expressions}
+Boolean expressions $b \in \BExpr$ can be $\FALSE$, implication ($b_1
+\Rightarrow b_2$), equality of bitvector expressions ($e_1 = e_2$ for two
+bitvector expressions $e_1$ and $e_2$), or universal quantification $\forall
+\rho \in t. b$, where $\rho$ is a symbolic row in table $t$. Note that universal
+quantification is meant to only be used by the instrumentation mechanism and not
+a programmer.
+
+We can represent other standard boolean operators by using standard encodings:
+\[\begin{array}{l}
+  \TRUE \triangleq \TRUE \Rightarrow \TRUE \\
+  \neg b \triangleq b \Rightarrow \FALSE \\
+  b_1 \vee b_2 \triangleq \neg b_1 \Rightarrow b_2 \\
+  b_1 \wedge b_2 \triangleq \neg(\neg b_1 \vee \neg b_2) \\
+  \vec e = \vec e' \triangleq \bigwedge_i e_i = e'_i
 \end{array}\]
 
-Write $\delta.\mathit{keys}$ to be the list of keys $\vec m$. When
-$\delta$ is an \textsf{Add} operation, write $\delta.\mathit{action}$
-to indicate the action $a$, and write $\delta.\mathit{data}$ to
-indicate the argument list $\vec k$. For any list $\vec x$ write
-$\vec x[i]$ for element $x_i$.
+\paragraph{Commands}
+Data plane programs $c \in \mathsf{Cmd}$ can be assignments ($h.f := e$),
+assumptions ($\assume b$), assertions ($\assert b$), table application
+($t.\apply()$), sequential composition ($c_1;c_2$ for two commands $c_1$ and
+$c_2$), and nondeterministic choice ($c_1 \choice c_2$ for two commands $c_1$
+and $c_2$).
 
-Define a control plane instance $\sigma$ to be a list of edits
-$\delta$, and is minimal if there are no deletions and no shadowed
-insertions. An insertion $\delta = \mathsf{Add}_t(\vec m, a(\vec k))$
-is \emph{shadowed} in an instance $\sigma$ if $\sigma[i] = \delta$ and
-every packet that matches $\delta$ matches a preceeding rule
-$\sigma[j]$, $0 \leq k < i$. An instance $\sigma$ is minimal for a
-program $p$ if for all $\sigma'$ such that
-$\llbracket p[\sigma] \rrbracket = \llbracket p[\sigma'] \rrbracket$,
-then $|\sigma| \leq |\sigma'|$.
+For actions, we restrict the language to be generated by only assignment, assumption, assertion and sequence, excluding table application and nondeterministic choice.
 
-The operation $\sigma(t)$ selects all the operations on table
-$t$ in $\sigma$. To apply the edits $\sigma$ to a program $p$ as is
-done in Avenir, write $p[\sigma] \in \mathsf{DProg}$.
+We define $\SKIP$ as the following syntactic sugar $\assume \TRUE$.
 
-Let $\varphi$ be a control plane constraint modelled by minimal lists
-of edits. For a table $t$ let $t[i]$ be a record
-$\{\MatchRow = \vec m; \ActionRow = \vec a; \ActionDataRow = \vec k\}$
-representing the $i$th row in that table.  We can describe $\varphi$
-by the boolean combinations of (in)equalities over the following
-variables.
-\begin{center}
-\begin{tabular}{lll}
-  \toprule
-  Variable & Semantics on minimal $\sigma$\\
-  \midrule
-  $t.\NumRows$ & Denotes the number of edits in table $t$\\
-  $t[i].\MatchRow(j)$ & In table $t$ the $j$th key in row $i$\\
-  $t[i].\ActionRow$ & In table $t$ the action selected in row $i$\\
-  $t[i].\ActionDataRow(j)$ & in table $t$ the $i$th action argument in row $i$\\
-  \bottomrule
-\end{tabular}
-\end{center}
+\subsection{Semantics}
 
-Then the grammar we get is:
-\[
-  \begin{array}{lcl}
-    \varphi &::=& \bot \\
-            & \mid & \varphi \Rightarrow \varphi \\
-            & \mid & e < e  \\
-    e       & ::= & n \\
-            & \mid & [n]_w \\
-            & \mid & t \\
-            & \mid & e[i] \\
-            & \mid & e.\NumRows \\
-            & \mid & e.\MatchRow \\
-            & \mid & e.\ActionRow \\
-            & \mid & e.\ActionDataRow \\
+First we describe the semantic objects packet and state before giving semantics
+to bitvector expressions, boolean expressions, and commands.
+
+\paragraph{Packets}
+A packet $\pkt$ is a map from header fields to values. Use the syntax $\pkt[h.f
+  \mapsto v]$ to represent the packet that is equivalent to $\pkt$ on all header
+fields except $h.f$, which maps to $v$.
+
+\paragraph{State}
+On-switch state is represented by a mapping from tables identifiers $t.\id$ to
+lists of rows $[r_1,\ldots,r_n] \in \Row(t)^*$.
+
+A similar to a symbolic row $\rho$, a concrete row $r$ comprises a list of match
+values $r.\matches$, an action identifier $r.\action$, and a list of action data
+$r.\data$. A concrete row $r$ is in the set $\Row(t)$ if it is well formed for a
+table $t$ iff the following conditions hold:
+\begin{enumerate}
+\item $|r.\matches| = |t.\keys|$
+\item $r.\action < |t.\actions|$
+\item $|r.\data| = |\vec x|$ where $\lambda \vec x. c = t.\actions[r.\action]$
+\end{enumerate}\
+
+\paragraph{Row Models}
+We also need models $\mu \in \Model$ to indicate which symbolic rows corresponds
+to which concrete rows, i.e. $\Model$ is a map from symbolic rows to concrete
+rows; write $\mu(\rho) = r$ for a model $\mu$, a symbolic row $\rho$, and a
+concrete row $r$.
+
+\paragraph{Expressions}
+The semantics of expressions is shown in Figure~\label{fig:sem-expr}. Note that
+here we implicitly use the strict semantics for definedness, that is, operators
+only succeed when all of their arguments are defined, otherwise it returns
+undefined.
+
+\begin{figure}[htp]
+  \[\begin{array}{l>{\triangleq}cl}
+  \multicolumn{3}{l}{\edenote{e} : \Pkt \to \Model \rightharpoonup \Value }\\
+    \edenote{v}~\pkt~\mu &&
+    v \\
+    \edenote{h.f}~\pkt~\mu &&
+    pkt(h.f)\\
+    \edenote{\rho.\matches[i]}~\pkt~\mu &&
+    \mu(\rho).\matches[i] \\
+    \edenote{\rho.\action}~\pkt~\mu &&
+    \mu(\rho).\action \\
+    \edenote{\rho.\data[i]} &&
+    \mu(\rho).\data[i]\\
+    \edenote{e_1 \binop e_2}~\pkt~\mu&&
+    \edenote{e_1}~\pkt~\mu \binop
+    \edenote{e_2}~\pkt~\mu \\
+    \edenote{\unop e}~\pkt~\mu &&
+    \unop \left(\edenote{e}~\pkt~\mu\right)
   \end{array}
   \]
+  \caption{Semantics of Expressions}
+  \label{fig:sem-expr}
+\end{figure}
 
-We use $\psi$ to denote a formula that intermingles the control plane
-variables above with dataplane variables like \texttt{ipv4.src\_addr}.
+\paragraph{Boolean Satisfaction}
+The satisfaction relation $\pkt,\mu \models_\sigma b$ is defined in
+Figure~\ref{fig:bool-satis} in terms of the three kinds of variable contexts,
+packets, models, and states. We also use the strict semantics here, most
+importantly in the equality rule. Sometimes two expressions are considered equal
+if they are both undefined; we do not take this approach, considering the
+equality undefined if either argument is undefined.
 
-We write the modelling relation $\sigma \models \varphi$ to indicate
-that $\sigma$ satisfies $\varphi$ in the standard
-way with respect to the above semantics.
-
-\[
-  \begin{array}{rcl}
-    \sigma \models \bot &\iff & \mathit{never} \\
-    \sigma \models \varphi_1 \Rightarrow \varphi_2 & \iff & \sigma \not\models \varphi_1,~\text{or}~\sigma \models \varphi_2 \\
-    \sigma \models e_1 < e_2 &\iff& \llbracket e_1 \rrbracket \sigma < \llbracket e_2 \rrbracket \sigma \\\\
-    % \llbracket n \rrbracket \sigma & = & n \\
-    % \llbracket [n]_w \rrbracket \sigma & = & [n]_w \\
-    % \llbracket t \rrbracket \sigma & = & \sigma(t) \\
-    % \llbracket e[i] \rrbracket \sigma & = & \sigma(t)[i], \mathit{where},~\llbracket e \rrbracket \sigma = \sigma(t),  \\
-    % \llbracket e.\NumRows \rrbracket \sigma & = &  |\sigma(t)|, \mathit{where}~\llbracket e \rrbracket \sigma = \sigma(t)[i], \\    
-    % \vdots
+\begin{figure}
+  \[\begin{array}{l>{\iff}cl}
+  \pkt,\mu \models_\sigma \FALSE && \textit{pigs fly} \\
+  \pkt,\mu \models_\sigma b_1 \Rightarrow b_2 &&
+  \pkt,\mu \not\models_\sigma b_1~\textit{or}~
+  \pkt,\mu \models_\sigma b_2\\
+  \pkt,\mu \models_\sigma e_1 = e_2 &&
+  \edenote{e_1}~\pkt = \edenote{e_2}~\pkt, \qquad
+  \text{(note strict semantics)} \\
+  \pkt,\sigma \models \edenote{\forall \rho \in t. b} &&
+  \forall r \in \sigma(t.\id).
+  \pkt,\sigma,\mu[\rho \mapsto r] \models b
   \end{array}
   \]
+  \caption{Satisfaction for $\BExpr$. Write $\pkt \models_\sigma b$ as syntactic
+    sugar when the model $\mu$ is empty.}
+  \label{fig:bool-satis}
+\end{figure}
 
-\section{The Problem}
+\paragraph{Command Evaluation}
+Commands are evaluated similar to Avenir in Figure~\ref{fig:com-sem}. The
+notable differences are the inclusion of assertions, the use of a binary
+nondeterminism operator, and a cleaner representation of table application.
 
-Given a program $p$, compute $\varphi$ such that for all
-instantiations $\sigma$, $\sigma \models \varphi$ if (and
-only if) one of the following (equivalent expressions) holds:
-\begin{enumerate}
-\item $\forall \mathit{pkt}. \llbracket p[\sigma]\rrbracket~\mathit{pkt} \neq \bot$
-\item $\mathit{wp}(p[\sigma], \top)$ is valid
-\end{enumerate}
+\begin{figure}[htp]
+  \[\begin{array}{l >{\triangleq}cl}
+  \multicolumn{3}{l}{\denote{c} : \Pkt \to 2^\Pkt_\bot }\\
+    \denote{h.f := e}^\sigma~\pkt && \{\pkt\{h \mapsto \edenote{e}\pkt\}\} \\
+    \denote{\assume b}^\sigma~\pkt &&
+    \begin{cases}
+      \{\pkt\} & \pkt \models_\sigma b \\
+      \{\} & \textit{otherwise} \\
+    \end{cases}\\
+    \denote{\assert b}^\sigma \pkt &&
+    \begin{cases}
+      \{pkt\} & \pkt \models_\sigma b \\
+      \bot & \textit{otherwise}
+    \end{cases} \\
+    \denote{c_1;c_2}^\sigma \pkt&&
+    \bigcup_{\pkt' \in \denote{c_1}\pkt}\denote{c_2}~\pkt' \\
+    \denote{c_1 \choice c_2}^\sigma \pkt &&
+    \denote{c_1}^\sigma~\pkt \cup \denote{c_2}^\sigma~\pkt \\
+    \denote{t.\apply()}^\sigma \pkt &&
+    \begin{cases}
+      \denote{c[\vec d/\vec x]}^\sigma~\pkt,
+      & \text{where}~\lambda \vec x.~c = r.\action, \vec d = r.\data \\
+      & \text{where}~r = \argmin_i\{\pkt \models_\sigma \sigma(t.\id)[i].\matches = t.\keys \} \\
 
-Our proposed solution is to instrument $p$ with ghost variables to
-(completely) describe the unknowns in the program. Write the
-instrumented program as $\mathcal{I}(p)$. An immediate strategy is to
-replace any table application $t.\mathit{apply}()$, where
-$t = (\mathsf{keys}= \vec k, \mathsf{actions} = \overrightarrow{a(\vec
-  d)})$ which statically has a maxmimum number of rows $n$ with a
-table of the following form:
-\[
-  \begin{array}{ll}
-    \multicolumn{2}{l}{\mathsf{assume} (t.\mathsf{numrows} = 0) []}\\
-    {[]}_{1 \leq r \leq n}
-    & \mathsf{assume}(t.\mathsf{numrows} \geq r)\\
-    & \mathsf{assume}(\bigwedge_l k_l = t[r].\MatchRow(l)); \\
-    & {[]}_j (\mathsf{assume}(j = t[r].\ActionRow); \\
-    & \quad\; \mathsf{assume}(\bigwedge_i d_i = t[r].\ActionDataRow(i)); \\
-    & \quad\; a_j(\vec d)) \\
-    ) 
-  \end{array}
-\]
-This is the approach taken by Avenir's strawman simple synthesis
-algorithm. P4v doesn't do such an enumeratively crazy instrumentation
-procedure---it uses only one table row $t[0]$ rather than $n$ table
-rows. More care is needed to figure out whether these are different or
-the same.
+      \{\pkt\}, & \mathit{otherwise}
+    \end{cases}
 
-Once we have this instrumentation, we can compute a verification
-condition (VC), to constrain the ``safe'' behavior of the instrumented
-program $\mathcal{I}(p)$. For the time being, we'll use the weakest
-precondition predicate transformer
-$\mathit{wp}(\mathcal{I}(p), \top)$, and notate the result as
-$\psi_p$. Notice that this formula $\psi_p$ may contain both
-dataplane variables, like $\texttt{ipv4.srcAddr}$ and control plane
-ghost variables, like $t[i].\MatchRow(l)$. Write
-$\psi_p[\vec{\texttt{d}},\vec{\mathsf{c}}]$ to denote that the free
-variables of $\psi_p$ are the union of the dataplane variables
-$\{\texttt{d}_1, \ldots,\texttt{d}_n\}$ and the control plane variables
-$\{\mathsf{c}_1, \ldots, \mathsf{c}_m\}$.
+  \end{array}\]
+  \caption{Semantics of Commands}
+  \label{fig:com-sem}
+\end{figure}
 
-Now we can reformulate the problem entirely logically. Given a program
-$p$ and its corresponding instrumented verification condition
-$\psi_p$, compute a controller constraint $\varphi$ such that the following equation is valid
-\[\forall \vec{\texttt{d}}.~\psi_p[\vec{\texttt{d}},\vec{\mathsf{c}}] \iff \varphi[\vec{\mathsf{c'}}] \]
-Now we only need to investigate a quantifier elimination algorithm for
-the variables $\vec{\texttt{d}}$. The standard algorithm (replace
-every $\texttt{d}_i$ of sort $\texttt{(\_~BitVec~n)}$ with every
-element in $\mathbf{2}^n$) for fixed-width bitvector theories is
-intractable: a standard ethernet frame is $1500$ bits, and the number
-of corresponding disjunctions is $2^{1500} = 10^{451.5\ldots}$, which
-is far greater than the $10^{82}$ atoms in the universe.
+\subsection{Encoding}
 
-However, I believe we can support a complete collection of rewrite
-rules to eliminate these variables based leveraging the known
-invariants about the program \& formula structures.
+%% \section{OLD}
 
-\section{An Incremental Problem Statement}
+%% Let $\delta$ be Control plane edits defined by the grammar
+%% \[\begin{array}{lcl}
+%%     \delta & ::= & \mathsf{Add}_t(\vec m,a(\vec k)) \\
+%%            & \mid & \mathsf{Del}_t(\vec m) \\
+%%   \end{array}\]
 
-Writing an instrumentation function for these indexed variables seems
-difficult. Instead lets just concentrate on a single new insertion at
-a time into a table $t$. This way, we can simply just insert the
-instrumented edit
-\[?\delta_t = \mathsf{Add}(\langle t.\MatchRow(0),\ldots,
-  t.\MatchRow(n-1)\rangle, t.\ActionRow(t.\ActionDataRow(0), \ldots,
-  t.\ActionDataRow(m-1)))\] where $n$ is the number of keys in table
-$t$, and $m$ is the maximum number of action data parameters that any
-action in $t$ has. For this incremental solution we can omit the
-subscripts, because there's only one rule being added.
+%% Write $\delta.\mathit{keys}$ to be the list of keys $\vec m$. When
+%% $\delta$ is an \textsf{Add} operation, write $\delta.\mathit{action}$
+%% to indicate the action $a$, and write $\delta.\mathit{data}$ to
+%% indicate the argument list $\vec k$. For any list $\vec x$ write
+%% $\vec x[i]$ for element $x_i$.
 
-\begin{quote}
-  \it Given a program $p$, a table $t$, and a minimal instantiation
-  $\sigma$ compute the weakest formula $\varphi$ such that
-  $\mathit{wp}(p[\sigma;?\delta_t],\top) \Rightarrow \varphi$.
-\end{quote}
+%% Define a control plane instance $\sigma$ to be a list of edits
+%% $\delta$, and is minimal if there are no deletions and no shadowed
+%% insertions. An insertion $\delta = \mathsf{Add}_t(\vec m, a(\vec k))$
+%% is \emph{shadowed} in an instance $\sigma$ if $\sigma[i] = \delta$ and
+%% every packet that matches $\delta$ matches a preceeding rule
+%% $\sigma[j]$, $0 \leq k < i$. An instance $\sigma$ is minimal for a
+%% program $p$ if for all $\sigma'$ such that
+%% $\llbracket p[\sigma] \rrbracket = \llbracket p[\sigma'] \rrbracket$,
+%% then $|\sigma| \leq |\sigma'|$.
 
-\subsection{One Possible Solution}
+%% The operation $\sigma(t)$ selects all the operations on table
+%% $t$ in $\sigma$. To apply the edits $\sigma$ to a program $p$ as is
+%% done in Avenir, write $p[\sigma] \in \mathsf{DProg}$.
 
-We can write a formula $\psi = \mathit{wp}(p[\sigma;\varphi],\top)$, as
-\[\forall d_1,\ldots, d_n. \psi[c_1,\ldots,c_n,d_1,\ldots,d_m]\]
-where the variables $d_i$ are dataplane variables like
-\texttt{ipv4.src\_addr} and the variables $d_1, \ldots, d_m$ are
-controllable variables like $t.\MatchRow(j)$, and
-$\{d_1,\ldots, d_m, c_1,\ldots, c_n\}$ are all the free variables of $\varphi_{\mathit{vc}}$
+%% Let $\varphi$ be a control plane constraint modelled by minimal lists
+%% of edits. For a table $t$ let $t[i]$ be a record
+%% $\{\MatchRow = \vec m; \ActionRow = \vec a; \ActionDataRow = \vec k\}$
+%% representing the $i$th row in that table.  We can describe $\varphi$
+%% by the boolean combinations of (in)equalities over the following
+%% variables.
+%% \begin{center}
+%% \begin{tabular}{lll}
+%%   \toprule
+%%   Variable & Semantics on minimal $\sigma$\\
+%%   \midrule
+%%   $t.\NumRows$ & Denotes the number of edits in table $t$\\
+%%   $t[i].\MatchRow(j)$ & In table $t$ the $j$th key in row $i$\\
+%%   $t[i].\ActionRow$ & In table $t$ the action selected in row $i$\\
+%%   $t[i].\ActionDataRow(j)$ & in table $t$ the $i$th action argument in row $i$\\
+%%   \bottomrule
+%% \end{tabular}
+%% \end{center}
 
-Then, compute a solution $\varphi$ by performing quantifier
-elimination on the $d_1,\ldots,d_n$.
+%% Then the grammar we get is:
+%% \[
+%%   \begin{array}{lcl}
+%%     \varphi &::=& \bot \\
+%%             & \mid & \varphi \Rightarrow \varphi \\
+%%             & \mid & e < e  \\
+%%     e       & ::= & n \\
+%%             & \mid & [n]_w \\
+%%             & \mid & t \\
+%%             & \mid & e[i] \\
+%%             & \mid & e.\NumRows \\
+%%             & \mid & e.\MatchRow \\
+%%             & \mid & e.\ActionRow \\
+%%             & \mid & e.\ActionDataRow \\
+%%   \end{array}
+%% \]
 
-This is probably untenable to do in the brute-force way. Is there another way?
+%% We use $\psi$ to denote a formula that intermingles the control plane
+%% variables above with dataplane variables like \texttt{ipv4.src\_addr}.
 
-\subsection{Failed Horn Encoding}
+%% We write the modelling relation $\sigma \models \varphi$ to indicate
+%% that $\sigma$ satisfies $\varphi$ in the standard
+%% way with respect to the above semantics.
 
-Given the above formula $\psi$, solve following CHC for $H$
+%% \[
+%%   \begin{array}{rcl}
+%%     \sigma \models \bot &\iff & \mathit{never} \\
+%%     \sigma \models \varphi_1 \Rightarrow \varphi_2 & \iff & \sigma \not\models \varphi_1,~\text{or}~\sigma \models \varphi_2 \\
+%%     \sigma \models e_1 < e_2 &\iff& \llbracket e_1 \rrbracket \sigma < \llbracket e_2 \rrbracket \sigma \\\\
+%%     % \llbracket n \rrbracket \sigma & = & n \\
+%%     % \llbracket [n]_w \rrbracket \sigma & = & [n]_w \\
+%%     % \llbracket t \rrbracket \sigma & = & \sigma(t) \\
+%%     % \llbracket e[i] \rrbracket \sigma & = & \sigma(t)[i], \mathit{where},~\llbracket e \rrbracket \sigma = \sigma(t),  \\
+%%     % \llbracket e.\NumRows \rrbracket \sigma & = &  |\sigma(t)|, \mathit{where}~\llbracket e \rrbracket \sigma = \sigma(t)[i], \\
+%%     % \vdots
+%%   \end{array}
+%% \]
 
-\[H \leftarrow \forall \vec d.\psi[\vec d, \vec c]\]
+%% \section{The Problem}
 
-This will trivially give True.
+%% Given a program $p$, compute $\varphi$ such that for all
+%% instantiations $\sigma$, $\sigma \models \varphi$ if (and
+%% only if) one of the following (equivalent expressions) holds:
+%% \begin{enumerate}
+%% \item $\forall \mathit{pkt}. \llbracket p[\sigma]\rrbracket~\mathit{pkt} \neq \mathsf{error}$
+%% \item $\mathit{wp}(p[\sigma], \top)$ is valid
+%% \end{enumerate}
+
+%% Our proposed solution is to instrument $p$ with ghost variables to
+%% (completely) describe the unknowns in the program. Write the
+%% instrumented program as $\mathcal{I}(p)$. An immediate strategy is to
+%% replace any table application $t.\mathit{apply}()$, where
+%% $t = (\mathsf{keys}= \vec k, \mathsf{actions} = \overrightarrow{a(\vec
+%%   d)})$ which statically has a maxmimum number of rows $n$ with a
+%% table of the following form:
+%% \[
+%%   \begin{array}{ll}
+%%     \multicolumn{2}{l}{\mathsf{assume} (t.\mathsf{numrows} = 0) []}\\
+%%     {[]}_{1 \leq r \leq n}
+%%     & \mathsf{assume}(t.\mathsf{numrows} \geq r)\\
+%%     & \mathsf{assume}(\bigwedge_l k_l = t[r].\MatchRow(l)); \\
+%%     & {[]}_j (\mathsf{assume}(j = t[r].\ActionRow); \\
+%%     & \quad\; \mathsf{assume}(\bigwedge_i d_i = t[r].\ActionDataRow(i)); \\
+%%     & \quad\; a_j(\vec d)) \\
+%%     )
+%%   \end{array}
+%% \]
+%% This is the approach taken by Avenir's strawman simple synthesis
+%% algorithm. P4v doesn't do such an enumeratively crazy instrumentation
+%% procedure---it uses only one table row $t[0]$ rather than $n$ table
+%% rows. More care is needed to figure out whether these are different or
+%% the same.
+
+%% Once we have this instrumentation, we can compute a verification
+%% condition (VC), to constrain the ``safe'' behavior of the instrumented
+%% program $\mathcal{I}(p)$. For the time being, we'll use the weakest
+%% precondition predicate transformer
+%% $\mathit{wp}(\mathcal{I}(p), \top)$, and notate the result as
+%% $\psi_p$. Notice that this formula $\psi_p$ may contain both
+%% dataplane variables, like $\texttt{ipv4.srcAddr}$ and control plane
+%% ghost variables, like $t[i].\MatchRow(l)$. Write
+%% $\psi_p[\vec{\texttt{d}},\vec{\mathsf{c}}]$ to denote that the free
+%% variables of $\psi_p$ are the union of the dataplane variables
+%% $\{\texttt{d}_1, \ldots,\texttt{d}_n\}$ and the control plane variables
+%% $\{\mathsf{c}_1, \ldots, \mathsf{c}_m\}$.
+
+%% Now we can reformulate the problem entirely logically. Given a program
+%% $p$ and its corresponding instrumented verification condition
+%% $\psi_p$, compute a controller constraint $\varphi$ such that the following equation is valid
+%% %% \[(\forall \vec{\texttt{d}}.~\psi_p[\vec{\texttt{d}},\vec{\mathsf{c}}]) \iff \varphi[\vec{\mathsf{c'}}] \]
+%% \[\varphi[\vec{\mathsf{c}}] \Rightarrow \psi_p[\vec{\texttt{d}},\vec{\mathsf{c}}] \]
+
+%% Now we only need to investigate a quantifier elimination algorithm for
+%% the variables $\vec{\texttt{d}}$. The standard algorithm (replace
+%% every $\texttt{d}_i$ of sort $\texttt{(\_~BitVec~n)}$ with every
+%% element in $\mathbf{2}^n$) for fixed-width bitvector theories is
+%% intractable: a standard ethernet frame is $1500$ bits, and the number
+%% of corresponding disjunctions is $2^{1500} = 10^{451.5\ldots}$, which
+%% is far greater than the $10^{82}$ atoms in the universe.
+
+%% However, I believe we can support a complete collection of rewrite
+%% rules to eliminate these variables based leveraging the known
+%% invariants about the program \& formula structures.
+
+%% \section{Examples}
+
+%% There are a few questions we want to investigate with examples.
+%% \begin{enumerate}
+%% \item \label{question:rowiter}
+%%   Do we need to consider multiple rows at the same time? In other words, do
+%%   we need to instrument each table $t$ with a different row for \(t[0]\ldots
+%%   t[t.\mathsf{maxrows}]\) or is it sufficient to, like P4v, only consider a
+%%   single row at a time?
+%% \item \label{question:elimination}
+%%   Can we eliminate all references to dataplane state in the conditions we
+%%   generate?
+%% \end{enumerate}
+
+%% \subsection{A Multi-Table Example}
+%% We will construct a pipeline with 3 tables, \(t_i\) for \(i=0,1,2\),
+%% where each \(t_i\) matches on \(x_i\) and assigns a value to
+%% \(y_i\).
+%% \begin{verbatim}
+%% table ti {
+%%   keys : xi
+%%   actions : ai
+%%   default : ai(i)
+%%   max_rows : 2
+%% }
+%% action ai(y){
+%%   yi := y
+%% }
+%% \end{verbatim}
+
+%% The control block applies \(t_0\), and then conditionally applies
+%% \(t_1\) or \(t_2\). The assertion is that \(t_i\) is even for all \(i=0,1,2\).
+
+%% \begin{verbatim}
+%% y0 := 0;
+%% y1 := 0;
+%% y2 := 0;
+%% t0.apply();
+%% if (y0%3 = 0) {
+%%   t1.apply()
+%% } else {
+%%   t2.apply()
+%% }
+%% assert y0%2 = 0;
+%% assert y1%2 = 0;
+%% assert y2%2 = 0;
+%% \end{verbatim}
+
+%% Now, we can instrument this program using the ghost state defined above.
+%% \begin{verbatim}
+%% y0 := 0; y1 := 0; y2 := 0;
+%% assume t0[0].action <= 0;
+%% assume t0[1].action <= 0;
+%% assume t0.numrows <= 2;
+%% { assume t0.numrows > 0 && t0[0].keys(x0) = x0;
+%%   assume t0[0].action = 0;
+%%   y0 := t0[0].data(y)
+%% } [] {
+%%  assume not (t0.numrows > 0 && t0[0].keys(x0) = x0);
+%%  assume t0.numrows > 1  && t0[1].keys(x0) = x0;
+%%  assume t0[1].action = 0;
+%%  y0 := t0[1].data(y)
+%% } [] {
+%%  assume not (t0.numrows > 0 && t0[0].keys(x0) = x0);
+%%  assume not (t0.numrows > 1 && t0[1].keys(x0) = x0);
+%%  y0 := 0
+%% };
+%% {assume y0%3 = 0;
+%%   assume t1[0].action <= 0;
+%%   assume t1[1].action <= 0;
+%%   assume t1.numrows <= 2;
+%%   { assume t1.numrows > 0 && t1[0].keys(x1) = x1;
+%%     assume t1[0].action = 0;
+%%     y1 := t1[0].data(y)
+%%   } [] {
+%%     assume not (t1.numrows > 0 && t1[0].keys(x1) = x1);
+%%     assume t1.numrows > 1  && t1[1].keys(x1) = x1;
+%%     assume t1[1].action = 0;
+%%     y1 := t1[1].data(y)
+%%   } [] {
+%%     assume not (t1.numrows > 0 && t1[0].keys(x1) = x1);
+%%     assume not (t1.numrows > 1 && t1[1].keys(x1) = x1);
+%%     y1 := 1
+%%   }
+%% } [] {
+%%   assume y0%3 != 0;
+%%   assume t2[0].action <= 0;
+%%   assume t2[1].action <= 0;
+%%   assume t2.numrows <= 2;
+%%   { assume t2.numrows > 0 && t2[0].keys(x2) = x2;
+%%     assume t2[0].action = 0;
+%%     y2 := t2[0].data(y)
+%%   } [] {
+%%     assume not (t2.numrows > 0 && t2[0].keys(x2) = x2);
+%%     assume t2.numrows > 1  && t2[1].keys(x2) = x2;
+%%     assume t2[1].action = 0;
+%%     y2 := t2[1].data(y)
+%%   } [] {
+%%     assume not (t2.numrows > 0 && t2[0].keys(x2) = x2);
+%%     assume not (t2.numrows > 1 && t2[1].keys(x2) = x2);
+%%     y2 := 2
+%%   }
+%% }
+%% assert y0 % 2 = y1 % 2 = y2 % 2 = 0
+%% \end{verbatim}
+
+%% And finally, we can compute the weakest precondition to produce a (universally
+%% closed) condition on variables.
+
+%% \[
+%% \begin{array}{l}
+%%   t_0.\NumRows > 0 \wedge t_0[0].\MatchRow(x_0) = x_0 \\
+%%   \Rightarrow
+%%   \begin{array}[t]{l}
+%%     t_0[0].\ActionDataRow(y) \% 3 = 0 \Rightarrow
+%%     \begin{array}[t]{l}
+%%       t_1.\NumRows > 0 \wedge t_1[0].\MatchRow(x_1) = x_1 \\
+%%       \qquad \Rightarrow t_1[0].\ActionDataRow(y) \% 2 = 0 \wedge t_0[0].\ActionDataRow(y)\% 2 = 0 \\
+%%       \wedge (\neg (t_1.\NumRows > 0 \wedge t_1[0].\MatchRow(x_1) = x_1) \\
+%%       \phantom{\wedge(}\wedge (t_1.\NumRows > 1 \wedge t_1[1].\MatchRow(x_1) = x_1) \\
+%%       \phantom{\wedge(}\Rightarrow t_1[1].\ActionDataRow(y) \% 2 = 0 \wedge t_0[1].\ActionDataRow(y)\% 2 = 0 )\\
+%%       \wedge (\neg (t_1.\NumRows > 0 \wedge t_1[0].\MatchRow(x_1) = x_1) \\
+%%       \phantom{\wedge (}\wedge \neg (t_1.\NumRows > 1 \wedge t_1[1].\MatchRow(x_1) = x_1)) \\
+%%     \end{array} \\
+%%     \wedge\,t_0[0].\ActionDataRow(y) \% 3 \neq 0 \Rightarrow
+%%     \begin{array}[t]{l}
+%%       t_2.\NumRows > 0 \wedge t_2[0].\MatchRow(x_2) = x_2 \\
+%%       \qquad \Rightarrow t_2[0].\ActionDataRow(y) \% 2 = 0 \wedge t_0[0].\ActionDataRow(y)\% 2 = 0 \\
+%%       \wedge (\neg (t_2.\NumRows > 0 \wedge t_2[0].\MatchRow(x_2) = x_2) \\
+%%       \phantom{\wedge(}\wedge (t_2.\NumRows > 1 \wedge t_2[1].\MatchRow(x_2) = x_2) \\
+%%       \phantom{\wedge(}\Rightarrow t_2[1].\ActionDataRow(y) \% 2 = 0 \wedge t_0[1].\ActionDataRow(y)\% 2 = 0 )\\
+%%       \wedge (\neg (t_2.\NumRows > 0 \wedge t_2[0].\MatchRow(x_2) = x_2) \\
+%%       \phantom{\wedge (}\wedge \neg (t_2.\NumRows ? 1 \wedge t_2[1].\MatchRow(x_2) = x_2) \\
+%%       \phantom{\wedge (} t_0[1].\ActionDataRow(y) \% 2 = 0)
+%%     \end{array}
+%%   \end{array} \\
+%%   t_0.\NumRows > 1 \wedge t_0[1].\MatchRow(x_0) = x_0 \\
+%%   \wedge \neg (t_0.\NumRows > 0 \wedge t_0[0].\MatchRow(x_0) = x_0 ) \\
+%%   \Rightarrow
+%%   \begin{array}[t]{l}
+%%     t_0[1].\ActionDataRow(y) \% 3 = 0 \Rightarrow
+%%     \begin{array}[t]{l}
+%%       t_1.\NumRows > 0 \wedge t_1[0].\MatchRow(x_1) = x_1 \\
+%%       \qquad \Rightarrow t_1[0].\ActionDataRow(y) \% 2 = 0 \wedge t_0[1].\ActionDataRow(y)\% 2 = 0 \\
+%%       \wedge (\neg (t_1.\NumRows > 0 \wedge t_1[0].\MatchRow(x_1) = x_1) \\
+%%       \phantom{\wedge(}\wedge (t_1.\NumRows > 1 \wedge t_1[1].\MatchRow(x_1) = x_1) \\
+%%       \phantom{\wedge(}\Rightarrow t_1[1].\ActionDataRow(y) \% 2 = 0 \wedge t_0[1].\ActionDataRow(y)\% 2 = 0 )\\
+%%       \wedge ((t_1.\NumRows > 0 \wedge t_1[0].\MatchRow(x_1) = x_1) \\
+%%       \phantom{\wedge (}\vee (t_1.\NumRows ? 1 \wedge t_1[1].\MatchRow(x_1) = x_1)) \\
+%%     \end{array} \\
+%%     \wedge\,t_0[1].\ActionDataRow(y) \% 3 \neq 0 \Rightarrow
+%%     \begin{array}[t]{l}
+%%       t_2.\NumRows > 0 \wedge t_2[0].\MatchRow(x_2) = x_2 \\
+%%       \qquad \Rightarrow t_2[0].\ActionDataRow(y) \% 2 = 0 \wedge t_0[1].\ActionDataRow(y)\% 2 = 0 \\
+%%       \wedge (\neg (t_2.\NumRows > 0 \wedge t_2[0].\MatchRow(x_2) = x_2) \\
+%%       \phantom{\wedge(}\wedge (t_2.\NumRows > 1 \wedge t_2[1].\MatchRow(x_2) = x_2) \\
+%%       \phantom{\wedge(}\Rightarrow t_2[1].\ActionDataRow(y) \% 2 = 0 \wedge t_0[1].\ActionDataRow(y)\% 2 = 0 )\\
+%%       \wedge (\neg (t_2.\NumRows > 0 \wedge t_2[0].\MatchRow(x_2) = x_2) \\
+%%       \phantom{\wedge (}\wedge \neg (t_2.\NumRows ? 1 \wedge t_2[1].\MatchRow(x_2) = x_2) \\
+%%       \phantom{\wedge (} t_0[1].\ActionDataRow(y) \% 2 = 0)
+%%     \end{array}
+%%   \end{array} \\
+%%   \neg (t_0.\NumRows > 0 \wedge t_0[0].\MatchRow(x_0) = x_0) \\
+%%   \wedge \neg (t_0.\NumRows > 1 \wedge t_0[1].\MatchRow(x_0) = x_0 ) \\
+%%   \Rightarrow
+%%   \begin{array}[t]{l}
+%%     t_1.\NumRows > 0 \wedge t_1[0].\MatchRow(x_1) = x_1 \\
+%%     \qquad \Rightarrow t_1[0].\ActionDataRow(y) \% 2 = 0 \\
+%%     \wedge (\neg (t_1.\NumRows > 0 \wedge t_1[0].\MatchRow(x_1) = x_1) \\
+%%     \phantom{\wedge(}\wedge (t_1.\NumRows > 1 \wedge t_1[1].\MatchRow(x_1) = x_1) \\
+%%     \phantom{\wedge(}\Rightarrow t_1[1].\ActionDataRow(y) \% 2 = 0 )\\
+%%     \wedge ((t_1.\NumRows > 0 \wedge t_1[0].\MatchRow(x_1) = x_1) \\
+%%     \phantom{\wedge (}\vee (t_1.\NumRows ? 1 \wedge t_1[1].\MatchRow(x_1) = x_1)) \\
+%%     \end{array}
+%% \end{array}
+%% \]
+
+%% \subsubsection{Question~\ref{question:rowiter}}
+
+%% For this example, its clear that we dont need the cases for both \(t_i[0]\) and
+%% \(t_i[1]\) for $i = 1,2,3$, so we can equivalently compute the following
+%% (universally closed) verification condition.
+
+%% \[
+%% \begin{array}{l}
+%%   t_0.\MatchRow(x_0) = x_0 \\
+%%   \Rightarrow
+%%   \begin{array}[t]{l}
+%%     t_0.\ActionDataRow(y) \% 3 = 0 \Rightarrow
+%%     \begin{array}[t]{l}
+%%       t_1.\MatchRow(x_1) = x_1 \\
+%%       \qquad \Rightarrow t_1.\ActionDataRow(y) \% 2 = 0 \wedge t_0.\ActionDataRow(y)\% 2 = 0 \\
+%%       \wedge \neg (t_1.\MatchRow(x_1) = x_1 \Rightarrow \bot) \\
+%%     \end{array} \\
+%%     \wedge\,t_0.\ActionDataRow(y) \% 3 \neq 0 \Rightarrow
+%%     \begin{array}[t]{l}
+%%       t_2.\MatchRow(x_2) = x_2 \\
+%%       \qquad \Rightarrow t_2.\ActionDataRow(y) \% 2 = 0 \wedge t_0.\ActionDataRow(y)\% 2 = 0 \\
+%%       \wedge (\neg (t_2.\MatchRow(x_2) = x_2) \\
+%%       \qquad \Rightarrow t_0.\ActionDataRow(y) \% 2 = 0)
+%%     \end{array}
+%%   \end{array} \\\\
+%%   \wedge \neg (t_0.\MatchRow(x_0) = x_0)
+%%   \Rightarrow
+%%   \begin{array}[t]{l}
+%%     t_1.\MatchRow(x_1) = x_1 \Rightarrow t_1.\ActionDataRow(y) \% 2 = 0 \\
+%%     \wedge \, \neg(t_1.\MatchRow(x_1) = x_1 \Rightarrow \bot)\\
+%%     \end{array}
+%% \end{array}
+%% \]
+%% which is similar to p4v's approach.
+
+%% \subsubsection{Question~\ref{question:elimination}}
+
+%% Looking at these examples, its not clear that we can eliminate the dataplane
+%% state, which occurs only in the form of match conditions. Specifically,
+%% dataplane state occurs only in the formula as expressions the form
+%% \(t_i.\MatchRow(x_i) = x_i \Rightarrow \varphi\), where \(\varphi\) is some
+%% further formula that may intermingle control and dataplane state.
+
+%% Inspecting the formula above, my intuition tells me that we should be able to
+%% come up with a rewrite rule that replaces these occurences with \(\top\),
+%% resulting in a formula with shape.
+
+%% \[
+%% \begin{array}{l}
+%%   \begin{array}[t]{l}
+%%     t_0.\ActionDataRow(y) \% 3 = 0 \Rightarrow
+%%     \begin{array}[t]{l}
+%%       t_1.\ActionDataRow(y) \% 2 = 0 \wedge t_0.\ActionDataRow(y)\% 2 = 0 \\
+%%     \end{array} \\
+%%     \wedge\,t_0.\ActionDataRow(y) \% 3 \neq 0 \Rightarrow
+%%     \begin{array}[t]{l}
+%%       t_2.\ActionDataRow(y) \% 2 = 0 \wedge t_0.\ActionDataRow(y)\% 2 = 0 \\
+%%     \end{array}
+%%   \end{array} \\
+%%   \,\,\,\wedge\, t_1.\ActionDataRow(y) \% 2 = 0 \\
+%% \end{array}
+%% \]
+
+%% which constraints only controllable variables.
+
+%% However, we have completely lost the fact that no packet may miss in table
+%% \(t_1\), which is encoded in the raw formula. More care is needed here.
+
+%% \subsection{Existential Example}
+
+%% For a while I've been concerned with the issue of control plane conditions of
+%% the form \[\forall \rho_1 \in t_1, \exists \rho_2 \in t_2,\] because I was
+%% afraid that this row existence wouldn't be captured in the predicate transformer
+%% semantics. This is related to Question~\ref{question:rowiter}---the multiplicity
+%% of rows is intended to enable this kind of fine-grained approach.
+
+%% It seems that we \emph{can} recover the existential information from the weakest
+%% preconditions constructed from instrumenting with only a single rule.
+
+%% Consider the following program that focuses on a single header instance \texttt{h}.
+%% \begin{figure}[H]
+%%   \begin{minipage}{0.32\textwidth}
+%% \begin{verbatim}
+%% control {
+%%   out := 0;
+%%   h.setInvalid();
+%%   t1.apply();
+%%   if(h.isValid()){
+%%     t2.apply();
+%%     assert out!=0;
+%%   }
+%% }
+%% \end{verbatim}
+%%   \end{minipage} \begin{minipage}{0.39\textwidth}
+%% \begin{verbatim}
+%% table t1 {
+%%   keys = x : exact;
+%%   acts : set_h
+%%   default_action : nop
+%% }
+%% table t2 {
+%%   keys = h.f:exact;
+%%   acts = {nop; set_out}
+%%   default_action = nop;:
+%% }
+%% \end{verbatim}
+%%   \end{minipage} \begin{minipage}{0.27\textwidth}
+%% \begin{verbatim}
+%% action set_h(f)  {
+%%   h.setValid();
+%%   h.f := x;
+%% }
+%% action set_out (o) {
+%%   out := o
+%% }
+%% action nop () {}
+%% \end{verbatim}
+%% \end{minipage}
+%% \end{figure}
+
+%% Notice that the control plane must satisfy the following condtion for the assertion to never fail:
+%% \[ \forall \rho_1 \in t_1.~\exists \rho_2 \in t_2.~\rho_2.\MatchRow(h.f) = \rho_1.\ActionDataRow(f) \wedge \rho_2.\ActionDataRow(o) \neq 0 \]
+
+%% If we follow the single-row-per-table instrumentation strategy, we compute the
+%% following condition:
+
+%% \[\begin{array}{l}
+%% t_1.\MatchRow(x) = x \Rightarrow \\
+%% \quad t_2.\MatchRow(h.f) = t_1.\ActionDataRow(f) \\
+%% \quad \wedge t_2.\ActionDataRow(o) \neq 0
+%% \end{array}\]
+
+%% which bears a striking resemblence to the condition above. We can typset them side-by-side to highlight the similarities.
+
+%% \[\begin{array}{lcl}
+%% t_1.\MatchRow(x) = x \Rightarrow && \forall \rho_1 \in t_1. \\
+%% \quad t_2.\MatchRow(h.f) = t_1.\ActionDataRow(f) & \leadsto & \quad \exists \rho_2 \in t_2.~\rho_2.\MatchRow(h.f) = \rho_1.\ActionDataRow(f) \\
+%% \qquad \wedge t_2.\ActionDataRow(o) \neq 0 && \qquad \wedge\rho_2.\ActionDataRow(o) \neq 0
+%% \end{array}
+%% \]
+
+%% This provides evidence for the p4v approach, in repsonse to Question~\ref{question:rowiter}.
+
+%% \subsection{Hypothesizing Transformations}
+
+%% In response to Question~\ref{question:eliminations}, we can hypothesize the
+%% following transformation rules in response to the above example, where
+%% \(t.\mathsf{o}\) represents an arbitrary access to table \(t\), \(\texttt{d}\)
+%% represents dataplane state,
+
+%% \[\begin{array}{l>{\leadsto}cr}
+%% t.\MatchRow(x) = x \rightarrow \phi && \forall \rho \in t. \phi \\
+%% t.\mathsf{o} = t'.\mathsf{o'} \wedge \phi && \exists \rho \in t'. \phi \\
+%% t.\mathsf{o} = n && \rho.\mathsf{o} = n \\
+%% \end{array}\]
+
+%% An example of each of these can be seen in the example above.
+
+%% However, these rules as stated are incorrect, they must aqcuire and maintain
+%% some context. For example, in the final rule, mapping tests liket $t.\mathsf(o)
+%% = 0$, we need to know which \(\rho\) row-variable corresponds to the \(t\).
+
+%% Not only that, but we also need to know how to bind the variables. For example,
+%% how would we translate a rule like
+%% \[t_1.\MatchRow = 6 \Rightarrow t_2.\MatchRow = 7\]
+
+%% The set of hypothesized rules suggests that this should become
+%% \[\forall \rho_1 \in t_1. \rho_1.\MatchRow = 6 \Rightarrow \exists \rho_2 \in t_2. \rho_2.\MatchRow = 7\]
+
+%% Some random thoughts
+%% \begin{enumerate}
+%% \item Context seems important
+%% \item Can we generalize this into only introducing one operator? It seems like
+%%   both of the quantifier introduction rules could be replaced by a local
+%%   \(\exists\)-quantifiaction,
+%% \item We need more examples and a more formal approach. Can we derive these rules?
+%% \end{enumerate}
+
+%% \subsection{Additional Examples}
+%% More examples are needed to investigate
+%% Questions~\ref{question:rowiter}\&\ref{question:elimination}.
+
+%% \clearpage
+%% \section{(OUTDATED) An Incremental Problem Statement}
+
+%% Writing an instrumentation function for these indexed variables seems difficult.
+%% Instead lets just concentrate on a single new insertion at a time into a table
+%% $t$. This way, we can simply just insert the instrumented edit
+%% \[?\delta_t = \mathsf{Add}(\langle t.\MatchRow(0),\ldots,
+%%   t.\MatchRow(n-1)\rangle, t.\ActionRow(t.\ActionDataRow(0), \ldots,
+%%   t.\ActionDataRow(m-1)))\] where $n$ is the number of keys in table
+%% $t$, and $m$ is the maximum number of action data parameters that any
+%% action in $t$ has. For this incremental solution we can omit the
+%% subscripts, because there's only one rule being added.
+
+%% \begin{quote}
+%%   \it Given a program $p$, a table $t$, and a minimal instantiation
+%%   $\sigma$ compute the weakest formula $\varphi$ such that
+%%   $\mathit{wp}(p[\sigma;?\delta_t],\top) \Rightarrow \varphi$.
+%% \end{quote}
+
+%% \subsection{One Possible Solution}
+
+%% We can write a formula $\psi = \mathit{wp}(p[\sigma;\varphi],\top)$, as
+%% \[\forall d_1,\ldots, d_n. \psi[c_1,\ldots,c_n,d_1,\ldots,d_m]\]
+%% where the variables $d_i$ are dataplane variables like
+%% \texttt{ipv4.src\_addr} and the variables $d_1, \ldots, d_m$ are
+%% controllable variables like $t.\MatchRow(j)$, and
+%% $\{d_1,\ldots, d_m, c_1,\ldots, c_n\}$ are all the free variables of $\varphi_{\mathit{vc}}$
+
+%% Then, compute a solution $\varphi$ by performing quantifier
+%% elimination on the $d_1,\ldots,d_n$.
+
+%% This is probably untenable to do in the brute-force way. Is there another way?
+
+%% \subsection{Failed Horn Encoding}
+
+%% Given the above formula $\psi$, solve following CHC for $H$
+
+%% \[H \leftarrow \forall \vec d.\psi[\vec d, \vec c]\]
+
+%% This will trivially give True.
 

### 2021-03-27 inference.tex
attempt outline of instrumentation c[?]. Stuck on scoping; looks like DRS

--- 2021-01-06 inference.tex
@@
 
-\newcommand{\NumRows}{\mathsf{numrows}}
-\newcommand{\MatchRow}{\mathsf{keys}}
-\newcommand{\ActionRow}{\mathsf{action}}
-\newcommand{\ActionDataRow}{\mathsf{data}}
+% semantics
+\newcommand{\pkt}{\mathit{pkt}}
+\newcommand{\denote}[1]{\llbracket#1\rrbracket}
+\newcommand{\edenote}[1]{\mathcal{E}\denote{#1}}
+
+%booleans
+\newcommand{\TRUE}{\mathsf{tt}}
+\newcommand{\FALSE}{\mathsf{ff}}
+
+%% bitvectors
+\newcommand{\binop}{\mathbin{\oplus}}
+\newcommand{\unop}{\mathop{\text{\pointer}}}
+
+% sets
+\newcommand{\Value}{\mathsf{Value}}
+\newcommand{\BVExpr}{\mathsf{BVExpr}}
+\newcommand{\BExpr}{\mathsf{BExpr}}
+\newcommand{\Cmd}{\mathsf{Cmd}}
+\newcommand{\Pkt}{\mathsf{Pkt}}
+\newcommand{\Model}{\mathsf{Model}}
+\newcommand{\Hdr}{\mathsf{Hdr}}
+\newcommand{\Field}{\mathsf{Field}}
+\newcommand{\Table}{\mathsf{Table}}
+\newcommand{\Row}{\mathsf{Row}}
+\newcommand{\List}{\mathsf{List}}
+\newcommand{\Var}{\mathsf{Var}}
+\newcommand{\SymbRow}{\mathsf{SymbRow}}
+
+% record fields
+\newcommand{\matches}{\mathsf{matches}}
+\newcommand{\action}{\mathsf{action}}
+\newcommand{\actions}{\mathsf{actions}}
+\newcommand{\keys}{\mathsf{keys}}
+\newcommand{\data}{\mathsf{data}}
+\newcommand{\id}{\mathsf{id}}
+
+% commands
+\newcommand{\assert}{\mathop{\mathsf{assert}}}
+\newcommand{\assume}{\mathop{\mathsf{assume}}}
+\newcommand{\apply}{\mathsf{apply}}
+\newcommand{\choice}{\mathbin{\rotatebox[origin=c]{90}{$\sqsubset\!\!\!\sqsupset$}}}
+\newcommand{\SKIP}{\mathsf{skip}}
+
+%math
+\newcommand{\argmin}{\mathop{\textrm{argmin}}}
+
+% Theorems
+\newtheorem{theorem}{Theorem}
 
@@
 
-Let a data plane program $p \in \mathsf{DProg}$ be defined by the following grammar:
+\subsection{Syntax}
 
-\[\begin{array}{lcl}
-    p & ::= & \mathsf{skip}\\
-      & \mid & f := e \\
-      & \mid & \mathsf{assert}~b \\
-      & \mid & \mathsf{assume}~b \\
-      & \mid & p;p \\
-      & \mid & p [] p \\
-      & \mid & t.apply() \\
+\begin{figure}[htp]
+  \[\begin{array}{lcll}
+    \multicolumn{3}{l}{\Value} \\
+    v & ::= & [n]_w  & \textit{Bitvector Literal}\\
+    \multicolumn{3}{l}{\SymbRow} \\
+    \rho & ::= & \left\{
+    \begin{array}{l}
+      \matches = \List[x];\\
+      \action = x;\\
+      \data = \List[x]
+    \end{array}\right\} & \textit{Symbolic Row}\\
+    \multicolumn{3}{l}{\BVExpr} \\
+    e & ::= & v & \textit{Value}\\
+      & \mid & h.f & \textit{Header Access} \\
+      & \mid & \rho.\matches[i] & \textit{Symbolic Match} \\
+      & \mid & \rho.\action & \textit{Symbolic Action}\\
+      & \mid & \rho.\data[i] & \textit{Symbolic Action Data}\\
+      & \mid & x & \textit{Variable}\\
+      & \mid & e \binop e & \textit{Arbitrary Binary Operation}\\
+    & \mid & \unop e & \textit{Arbitrary Unary Operation}\\
+    \multicolumn{3}{l}{\Table} \\
+    t & ::= & \left\{\begin{array}{l}
+    t.\id = v; \\
+    t.\keys = \List[h.f]; \\
+    t.\actions = \List[\lambda \vec x. c]\\
+    \end{array}\right\} & \textit{Table}\\
+    \multicolumn{3}{l}{\BExpr} \\
+    b & ::= & \FALSE  & \textit{Falsity}\\
+      & \mid & b \Rightarrow b & \textit{Implication} \\
+      & \mid & e = e & \textit{Bitvector Equality}\\
+    & \mid & \forall \rho \in t.~b & \textit{Row Universality}\\
+    \multicolumn{3}{l}{\Cmd} \\
+    c & ::=  & h.f := e & \textit{Assignment}\\
+      & \mid & \assert b & \textit{Assertion}\\
+      & \mid & \assume b & \textit{Assumption}\\
+      & \mid & t.\apply() & \textit{Table Application}\\
+      & \mid & c;c & \textit{Sequential Composition} \\
+    & \mid & c \choice c & \textit{Nondeterministic Choice}\\
+    \multicolumn{3}{l}{
+      \begin{array}{ll}
+        x \in \Var & h \in \Hdr\\
+        n,w,i \in \mathbb{N} & f \in \Field
+    \end{array}} & \textit{Sets }\\
   \end{array}
 \]
+\caption{Grammar of Dataplane Programs}
+\label{fig:grammar}
+\end{figure}
 
-Where $b$ is a boolean logic over bitvectors expressions $e$, and
-$t$ is a table as defined in Avenir.
+The following sections describe aspects of the grammar shown in Figure~\ref{fig:grammar}.
 
-Let $\delta$ be Control plane edits defined by the grammar
-\[\begin{array}{lcl}
-    \delta & ::= & \mathsf{Add}_t(\vec m,a(\vec k)) \\
-           & \mid & \mathsf{Del}_t(\vec m) \\             
+\paragraph{Packets}
+Our dataplane programs will reason about structured representations of packets,
+that is, they will be transformations on sets of packet headers. In programs,
+packet headers $h \in \Hdr$ are essentially structs with fields $f \in
+\Field$. Write $h.f$ for a header $h$ and a field $f$ is the value $v$ of $f$
+in header $h$.
+
+\paragraph{Tables}
+A table $t \in \Table$ is a record with a few fields described below:
+\begin{enumerate}[align=left]
+  \item[($t.\id$)] A table's unique identifier. A well-formed program can only have
+    one occurence of each table identifier.
+  \item[($t.\keys$)] A list of header field accesses indicating the match keys.
+  \item[($t.\actions$)] A list of actions that can be executed. An action is a lambda
+    expression $\lambda \vec x. c$ where the variables $x$ are in scope for the
+    commands $c$, which $c$ are straight-line code only (see below).
+\end{enumerate}
+For now we assume that the default action is the no-operation command (see
+$\SKIP$ below). We use $|l|$ to indicate the number of elements in a list or vector $l$
+such as $t.\keys$ or $t.\actions$.
+
+\paragraph{Symbolic Table Rows}
+A symbolic table row $\rho \in \SymbRow$ represents a symbolic row. Like tables,
+rows are also records with the following fields \todo{Well formedness constraint?}
+\begin{enumerate}[align=left]
+  \item[($\rho.\matches$)] A list variable of length representing the possible match keys.
+  \item[($\rho.\action$)] A variable representing the action choice
+  \item[($\rho.\data$)] A list variable representing the possible action data
+\end{enumerate}
+
+To represent the $i$th match variable in $\rho$, use standard list
+access notation: $\rho.\matches[i]$.
+
+Note that symbolic table rows are only meant to be used in the instrumentation
+stage and not by a programmer.
+
+\paragraph{Values}
+Values $v \in \BVExpr$ are bitvector literals, written $[n]_w$, where $n$
+is a natural value and $w$ is the width. We say that $[n]_w = [m]_l$ is
+undefined if $w \neq l$, and otherwise is true iff $n \equiv m \mod 2^w$.
+
+\paragraph{Bitvector Expressions}
+Bitvector expressions $e \in \BVExpr$ can be values $v$, header accesses $h.f$,
+symbolic row variables, variables $x$, or any
+(currently unspecified) binary or unary bitvector operation.
+
+\paragraph{Boolean Expressions}
+Boolean expressions $b \in \BExpr$ can be $\FALSE$, implication ($b_1
+\Rightarrow b_2$), equality of bitvector expressions ($e_1 = e_2$ for two
+bitvector expressions $e_1$ and $e_2$), or universal quantification $\forall
+\rho \in t. b$, where $\rho$ is a symbolic row in table $t$. Note that universal
+quantification is meant to only be used by the instrumentation mechanism and not
+a programmer.
+
+We can represent other standard boolean operators by using standard encodings:
+\[\begin{array}{l}
+  \TRUE \triangleq \TRUE \Rightarrow \TRUE \\
+  \neg b \triangleq b \Rightarrow \FALSE \\
+  b_1 \vee b_2 \triangleq \neg b_1 \Rightarrow b_2 \\
+  b_1 \wedge b_2 \triangleq \neg(\neg b_1 \vee \neg b_2) \\
+  \vec e = \vec e' \triangleq \bigwedge_i e_i = e'_i
 \end{array}\]
 
-Write $\delta.\mathit{keys}$ to be the list of keys $\vec m$. When
-$\delta$ is an \textsf{Add} operation, write $\delta.\mathit{action}$
-to indicate the action $a$, and write $\delta.\mathit{data}$ to
-indicate the argument list $\vec k$. For any list $\vec x$ write
-$\vec x[i]$ for element $x_i$.
+\paragraph{Commands}
+Data plane programs $c \in \mathsf{Cmd}$ can be assignments ($h.f := e$),
+assumptions ($\assume b$), assertions ($\assert b$), table application
+($t.\apply()$), sequential composition ($c_1;c_2$ for two commands $c_1$ and
+$c_2$), and nondeterministic choice ($c_1 \choice c_2$ for two commands $c_1$
+and $c_2$).
 
-Define a control plane instance $\sigma$ to be a list of edits
-$\delta$, and is minimal if there are no deletions and no shadowed
-insertions. An insertion $\delta = \mathsf{Add}_t(\vec m, a(\vec k))$
-is \emph{shadowed} in an instance $\sigma$ if $\sigma[i] = \delta$ and
-every packet that matches $\delta$ matches a preceeding rule
-$\sigma[j]$, $0 \leq k < i$. An instance $\sigma$ is minimal for a
-program $p$ if for all $\sigma'$ such that
-$\llbracket p[\sigma] \rrbracket = \llbracket p[\sigma'] \rrbracket$,
-then $|\sigma| \leq |\sigma'|$.
+For actions, we restrict the language to be generated by only assignment, assumption, assertion and sequence, excluding table application and nondeterministic choice.
 
-The operation $\sigma(t)$ selects all the operations on table
-$t$ in $\sigma$. To apply the edits $\sigma$ to a program $p$ as is
-done in Avenir, write $p[\sigma] \in \mathsf{DProg}$.
+We define $\SKIP$ as the following syntactic sugar $\assume \TRUE$.
 
-Let $\varphi$ be a control plane constraint modelled by minimal lists
-of edits. For a table $t$ let $t[i]$ be a record
-$\{\MatchRow = \vec m; \ActionRow = \vec a; \ActionDataRow = \vec k\}$
-representing the $i$th row in that table.  We can describe $\varphi$
-by the boolean combinations of (in)equalities over the following
-variables.
-\begin{center}
-\begin{tabular}{lll}
-  \toprule
-  Variable & Semantics on minimal $\sigma$\\
-  \midrule
-  $t.\NumRows$ & Denotes the number of edits in table $t$\\
-  $t[i].\MatchRow(j)$ & In table $t$ the $j$th key in row $i$\\
-  $t[i].\ActionRow$ & In table $t$ the action selected in row $i$\\
-  $t[i].\ActionDataRow(j)$ & in table $t$ the $i$th action argument in row $i$\\
-  \bottomrule
-\end{tabular}
-\end{center}
+\subsection{Semantics}
 
-Then the grammar we get is:
-\[
-  \begin{array}{lcl}
-    \varphi &::=& \bot \\
-            & \mid & \varphi \Rightarrow \varphi \\
-            & \mid & e < e  \\
-    e       & ::= & n \\
-            & \mid & [n]_w \\
-            & \mid & t \\
-            & \mid & e[i] \\
-            & \mid & e.\NumRows \\
-            & \mid & e.\MatchRow \\
-            & \mid & e.\ActionRow \\
-            & \mid & e.\ActionDataRow \\
+First we describe the semantic objects packet and state before giving semantics
+to bitvector expressions, boolean expressions, and commands.
+
+\paragraph{Packets}
+A packet $\pkt$ is a map from header fields to values. Use the syntax $\pkt[h.f
+  \mapsto v]$ to represent the packet that is equivalent to $\pkt$ on all header
+fields except $h.f$, which maps to $v$.
+
+\paragraph{State}
+On-switch state is represented by a mapping from tables identifiers $t.\id$ to
+lists of rows $[r_1,\ldots,r_n] \in \Row(t)^*$.
+
+A similar to a symbolic row $\rho$, a concrete row $r$ comprises a list of match
+values $r.\matches$, an action identifier $r.\action$, and a list of action data
+$r.\data$. A concrete row $r$ is in the set $\Row(t)$ if it is well formed for a
+table $t$ iff the following conditions hold:
+\begin{enumerate}
+\item $|r.\matches| = |t.\keys|$
+\item $r.\action < |t.\actions|$
+\item $|r.\data| = |\vec x|$ where $\lambda \vec x. c = t.\actions[r.\action]$
+\end{enumerate}\
+
+\paragraph{Row Models}
+We also need models $\mu \in \Model$ to indicate which symbolic rows corresponds
+to which concrete rows, i.e. $\Model$ is a map from symbolic rows to concrete
+rows; write $\mu(\rho) = r$ for a model $\mu$, a symbolic row $\rho$, and a
+concrete row $r$.
+
+\paragraph{Expressions}
+The semantics of expressions is shown in Figure~\label{fig:sem-expr}. Note that
+here we implicitly use the strict semantics for definedness, that is, operators
+only succeed when all of their arguments are defined, otherwise it returns
+undefined.
+
+\begin{figure}[htp]
+  \[\begin{array}{l>{\triangleq}cl}
+  \multicolumn{3}{l}{\edenote{e} : \Pkt \to \Model \rightharpoonup \Value }\\
+    \edenote{v}~\pkt~\mu &&
+    v \\
+    \edenote{h.f}~\pkt~\mu &&
+    pkt(h.f)\\
+    \edenote{\rho.\matches[i]}~\pkt~\mu &&
+    \mu(\rho).\matches[i] \\
+    \edenote{\rho.\action}~\pkt~\mu &&
+    \mu(\rho).\action \\
+    \edenote{\rho.\data[i]} &&
+    \mu(\rho).\data[i]\\
+    \edenote{e_1 \binop e_2}~\pkt~\mu&&
+    \edenote{e_1}~\pkt~\mu \binop
+    \edenote{e_2}~\pkt~\mu \\
+    \edenote{\unop e}~\pkt~\mu &&
+    \unop \left(\edenote{e}~\pkt~\mu\right)
   \end{array}
   \]
+  \caption{Semantics of Expressions}
+  \label{fig:sem-expr}
+\end{figure}
 
-We use $\psi$ to denote a formula that intermingles the control plane
-variables above with dataplane variables like \texttt{ipv4.src\_addr}.
+\paragraph{Boolean Satisfaction}
+The satisfaction relation $\pkt,\mu \models_\sigma b$ is defined in
+Figure~\ref{fig:bool-satis} in terms of the three kinds of variable contexts,
+packets, models, and states. We also use the strict semantics here, most
+importantly in the equality rule. Sometimes two expressions are considered equal
+if they are both undefined; we do not take this approach, considering the
+equality undefined if either argument is undefined.
 
-We write the modelling relation $\sigma \models \varphi$ to indicate
-that $\sigma$ satisfies $\varphi$ in the standard
-way with respect to the above semantics.
-
-\[
-  \begin{array}{rcl}
-    \sigma \models \bot &\iff & \mathit{never} \\
-    \sigma \models \varphi_1 \Rightarrow \varphi_2 & \iff & \sigma \not\models \varphi_1,~\text{or}~\sigma \models \varphi_2 \\
-    \sigma \models e_1 < e_2 &\iff& \llbracket e_1 \rrbracket \sigma < \llbracket e_2 \rrbracket \sigma \\\\
-    % \llbracket n \rrbracket \sigma & = & n \\
-    % \llbracket [n]_w \rrbracket \sigma & = & [n]_w \\
-    % \llbracket t \rrbracket \sigma & = & \sigma(t) \\
-    % \llbracket e[i] \rrbracket \sigma & = & \sigma(t)[i], \mathit{where},~\llbracket e \rrbracket \sigma = \sigma(t),  \\
-    % \llbracket e.\NumRows \rrbracket \sigma & = &  |\sigma(t)|, \mathit{where}~\llbracket e \rrbracket \sigma = \sigma(t)[i], \\    
-    % \vdots
+\begin{figure}
+  \[\begin{array}{l>{\iff}cl}
+  \pkt,\mu \models_\sigma \FALSE && \textit{pigs fly} \\
+  \pkt,\mu \models_\sigma b_1 \Rightarrow b_2 &&
+  \pkt,\mu \not\models_\sigma b_1~\textit{or}~
+  \pkt,\mu \models_\sigma b_2\\
+  \pkt,\mu \models_\sigma e_1 = e_2 &&
+  \edenote{e_1}~\pkt = \edenote{e_2}~\pkt, \qquad
+  \text{(note strict semantics)} \\
+  \pkt,\sigma \models \edenote{\forall \rho \in t. b} &&
+  \forall r \in \sigma(t.\id).
+  \pkt,\sigma,\mu[\rho \mapsto r] \models b
   \end{array}
   \]
+  \caption{Satisfaction for $\BExpr$. Write $\pkt \models_\sigma b$ as syntactic
+    sugar when the model $\mu$ is empty.}
+  \label{fig:bool-satis}
+\end{figure}
 
-\section{The Problem}
+\paragraph{Command Evaluation}
+Commands are evaluated similar to Avenir in Figure~\ref{fig:com-sem}. The
+notable differences are the inclusion of assertions, the use of a binary
+nondeterminism operator, and a cleaner representation of table application.
 
-Given a program $p$, compute $\varphi$ such that for all
-instantiations $\sigma$, $\sigma \models \varphi$ if (and
-only if) one of the following (equivalent expressions) holds:
-\begin{enumerate}
-\item $\forall \mathit{pkt}. \llbracket p[\sigma]\rrbracket~\mathit{pkt} \neq \bot$
-\item $\mathit{wp}(p[\sigma], \top)$ is valid
-\end{enumerate}
+\begin{figure}[htp]
+  \[\begin{array}{l >{\triangleq}cl}
+  \multicolumn{3}{l}{\denote{c} : \Pkt \to 2^\Pkt_\bot }\\
+    \denote{h.f := e}^\sigma~\pkt && \{\pkt\{h \mapsto \edenote{e}\pkt\}\} \\
+    \denote{\assume b}^\sigma~\pkt &&
+    \begin{cases}
+      \{\pkt\} & \pkt \models_\sigma b \\
+      \{\} & \textit{otherwise} \\
+    \end{cases}\\
+    \denote{\assert b}^\sigma \pkt &&
+    \begin{cases}
+      \{pkt\} & \pkt \models_\sigma b \\
+      \bot & \textit{otherwise}
+    \end{cases} \\
+    \denote{c_1;c_2}^\sigma \pkt&&
+    \bigcup_{\pkt' \in \denote{c_1}\pkt}\denote{c_2}~\pkt' \\
+    \denote{c_1 \choice c_2}^\sigma \pkt &&
+    \denote{c_1}^\sigma~\pkt \cup \denote{c_2}^\sigma~\pkt \\
+    \denote{t.\apply()}^\sigma \pkt &&
+    \begin{cases}
+      \denote{c[\vec d/\vec x]}^\sigma~\pkt,
+      & \text{where}~\lambda \vec x.~c = r.\action, \vec d = r.\data \\
+      & \text{where}~r = \argmin_i\{\pkt \models_\sigma \sigma(t.\id)[i].\matches = t.\keys \} \\
 
-Our proposed solution is to instrument $p$ with ghost variables to
-(completely) describe the unknowns in the program. Write the
-instrumented program as $\mathcal{I}(p)$. An immediate strategy is to
-replace any table application $t.\mathit{apply}()$, where
-$t = (\mathsf{keys}= \vec k, \mathsf{actions} = \overrightarrow{a(\vec
-  d)})$ which statically has a maxmimum number of rows $n$ with a
-table of the following form:
-\[
-  \begin{array}{ll}
-    \multicolumn{2}{l}{\mathsf{assume} (t.\mathsf{numrows} = 0) []}\\
-    {[]}_{1 \leq r \leq n}
-    & \mathsf{assume}(t.\mathsf{numrows} \geq r)\\
-    & \mathsf{assume}(\bigwedge_l k_l = t[r].\MatchRow(l)); \\
-    & {[]}_j (\mathsf{assume}(j = t[r].\ActionRow); \\
-    & \quad\; \mathsf{assume}(\bigwedge_i d_i = t[r].\ActionDataRow(i)); \\
-    & \quad\; a_j(\vec d)) \\
-    ) 
+      \{\pkt\}, & \mathit{otherwise}
+    \end{cases}
+
+  \end{array}\]
+  \caption{Semantics of Commands}
+  \label{fig:com-sem}
+\end{figure}
+
+\section{Encoding}
+
+Now we will define a collection of source-to-source translations that deal with
+the controller interface.
+
+\section{Eliminating Tables}
+
+The first thing to do is to define a source-to-source translation that
+eliminates the tables. The observation here is that tables essentially represent
+if-else statements, so once we have a state $\sigma$ in hand we can rewrite the
+program to a command $c[\sigma] \in \Cmd$ that uses no tables. This translation
+is defined in Figure~\ref{fig:table-elim}
+
+\begin{figure}
+  \[\begin{array}{r>{\triangleq}cl}
+  (h.f := e)[\sigma] && h.f := e   \\
+  (\assume b)[\sigma] && \assume b \\
+  (\assert b)[\sigma] && \assert b \\
+  (c_1;c_2)[\sigma] && c_1[\sigma];c_2[\sigma] \\
+  (c_1 \choice c_2)[\sigma] && c_1[\sigma] \choice c_2[\sigma] \\
+  (t.\apply())[\sigma] &&
+  \begin{array}[t]{l}
+    \{\\
+    \quad \assume {r_1.\matches = t.\keys};\\
+    \quad t.\action[r_1.\action](r_1.\data)\\
+    \}~\choice\\
+    \multicolumn{1}{c}{\vdots}\\
+    \choice~\{\\
+    \quad \assume {r_n.\matches = t.\keys};\\
+    \quad t.\action[r_n.\action](r_n.\data)\\
+    \}\\
+    \textrm{where}~[r_1;\ldots;r_n] = \sigma(t.\id)
+  \end{array}
   \end{array}
   \]
-This is the approach taken by Avenir's strawman simple synthesis
-algorithm. P4v doesn't do such an enumeratively crazy instrumentation
-procedure---it uses only one table row $t[0]$ rather than $n$ table
-rows. More care is needed to figure out whether these are different or
-the same.
+  \caption{Table Elimination}
+  \label{fig:table-elim}
+\end{figure}
 
-Once we have this instrumentation, we can compute a verification
-condition (VC), to constrain the ``safe'' behavior of the instrumented
-program $\mathcal{I}(p)$. For the time being, we'll use the weakest
-precondition predicate transformer
-$\mathit{wp}(\mathcal{I}(p), \top)$, and notate the result as
-$\psi_p$. Notice that this formula $\psi_p$ may contain both
-dataplane variables, like $\texttt{ipv4.srcAddr}$ and control plane
-ghost variables, like $t[i].\MatchRow(l)$. Write
-$\psi_p[\vec{\texttt{d}},\vec{\mathsf{c}}]$ to denote that the free
-variables of $\psi_p$ are the union of the dataplane variables
-$\{\texttt{d}_1, \ldots,\texttt{d}_n\}$ and the control plane variables
-$\{\mathsf{c}_1, \ldots, \mathsf{c}_m\}$.
+\begin{theorem}[Table Elimination]
+  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$,
+  and packet $\pkt \in \Pkt$, then for any state $\tau : \Value \to \Row^*$,
+  \[\denote{c}^\sigma~\pkt = \denote{c[\sigma]}^\tau~\pkt\]
+\end{theorem}
+\begin{proof}
+\todo[inline]{By induction on the structure of $c$.}
+\end{proof}
 
-Now we can reformulate the problem entirely logically. Given a program
-$p$ and its corresponding instrumented verification condition
-$\psi_p$, compute a controller constraint $\varphi$ such that the following equation is valid
-\[\forall \vec{\texttt{d}}.~\psi_p[\vec{\texttt{d}},\vec{\mathsf{c}}] \iff \varphi[\vec{\mathsf{c'}}] \]
-Now we only need to investigate a quantifier elimination algorithm for
-the variables $\vec{\texttt{d}}$. The standard algorithm (replace
-every $\texttt{d}_i$ of sort $\texttt{(\_~BitVec~n)}$ with every
-element in $\mathbf{2}^n$) for fixed-width bitvector theories is
-intractable: a standard ethernet frame is $1500$ bits, and the number
-of corresponding disjunctions is $2^{1500} = 10^{451.5\ldots}$, which
-is far greater than the $10^{82}$ atoms in the universe.
+\section{Instrumenting Tables}
 
-However, I believe we can support a complete collection of rewrite
-rules to eliminate these variables based leveraging the known
-invariants about the program \& formula structures.
+The next step is to define a source to source translation that eliminates tables
+in the absence of concrete table state, which we will write $c[?] \in \Cmd$.
+This is the first step towards the goal of constraining control plane behavior.
+Want the instrumented program $c[?]$ to fail on exactly the same situations that
+$c[\sigma]$ does, however $c[?]$ is constrained to have no table applications.
+The only remaining mechanism is to use $\assume$ and $\assert$ statements to
+inspect $\sigma$. We define $c[?]$ in Figure~\ref{fig:table-instrument}.
 
-\section{An Incremental Problem Statement}
+\begin{figure}[htp]
+  \[\begin{array}{r >{\triangleq}cl}
+  (h.f := e)[?] && h.f := e \\
+  (\assume b)[?] && \assume b \\
+  (\assert b)[?] && \assert b\\
+  (c_1;c_2)[?] && c_1[?];c_2[?]\\
+  (c_1 \choice c_2)[?] && c_1[?] \choice c_2[?]\\
+  (t.\apply())[?] &&
+  \begin{array}[t]{l}
+  \{ \\
+  \quad \assume \rho.\matches = t.\keys; \\
+  \quad \{\assume \rho.\action = 0; t.\action[0](\rho.\data)\}\\
+  \quad \choice \cdots \\
+  \quad \choice \{\assume \rho.action = |t.\action| - 1; t.\action[t.\action - 1](\rho.\data)\}\\
+  \quad \}\\
+  \} \choice \{ \\
+  \quad \assume \forall \rho \in t. \rho.\matches \neq t.\keys; \\
+  \quad \SKIP \\
+  \quad \}\\
+  \}
+  \end{array}
+  \end{array}
+  \]
+\end{figure}
 
-Writing an instrumentation function for these indexed variables seems
-difficult. Instead lets just concentrate on a single new insertion at
-a time into a table $t$. This way, we can simply just insert the
-instrumented edit
-\[?\delta_t = \mathsf{Add}(\langle t.\MatchRow(0),\ldots,
-  t.\MatchRow(n-1)\rangle, t.\ActionRow(t.\ActionDataRow(0), \ldots,
-  t.\ActionDataRow(m-1)))\] where $n$ is the number of keys in table
-$t$, and $m$ is the maximum number of action data parameters that any
-action in $t$ has. For this incremental solution we can omit the
-subscripts, because there's only one rule being added.
+\todo[inline]{This is wrong because $\rho$ is unbound in the first branch}
 
-\begin{quote}
-  \it Given a program $p$, a table $t$, and a minimal instantiation
-  $\sigma$ compute the weakest formula $\varphi$ such that
-  $\mathit{wp}(p[\sigma;?\delta_t],\top) \Rightarrow \varphi$.
-\end{quote}
+\begin{theorem}[Instrumentation Sufficiency]
+  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, and
+  packet $\pkt \in \Pkt$, $\denote{c}^\sigma~\pkt = \pkt'$ if and only if
+  $\pkt' \in \denote{c[?]}^\sigma~\pkt$
+\end{theorem}
 
-\subsection{One Possible Solution}
-
-We can write a formula $\psi = \mathit{wp}(p[\sigma;\varphi],\top)$, as
-\[\forall d_1,\ldots, d_n. \psi[c_1,\ldots,c_n,d_1,\ldots,d_m]\]
-where the variables $d_i$ are dataplane variables like
-\texttt{ipv4.src\_addr} and the variables $d_1, \ldots, d_m$ are
-controllable variables like $t.\MatchRow(j)$, and
-$\{d_1,\ldots, d_m, c_1,\ldots, c_n\}$ are all the free variables of $\varphi_{\mathit{vc}}$
-
-Then, compute a solution $\varphi$ by performing quantifier
-elimination on the $d_1,\ldots,d_n$.
-
-This is probably untenable to do in the brute-force way. Is there another way?
-
-\subsection{Failed Horn Encoding}
-
-Given the above formula $\psi$, solve following CHC for $H$
-
-\[H \leftarrow \forall \vec d.\psi[\vec d, \vec c]\]
-
-This will trivially give True.
+\begin{proof}
+\todo[inline]{By induction on the structure of $c$}
+\end{proof}
 

### 2021-03-27 inference.tex
attempt with choose; need lexical scope

@@
 \usepackage{float}
+\usepackage{scalerel}
 %% \usepackage[notext,not1,nomath]{stix}
@@
 \newcommand{\pkt}{\mathit{pkt}}
-\newcommand{\denote}[1]{\llbracket#1\rrbracket}
+\newcommand{\error}{\mathsf{error}}
+\newcommand{\denote}[1]{\left\llbracket#1\right\rrbracket}
 \newcommand{\edenote}[1]{\mathcal{E}\denote{#1}}
@@
 \newcommand{\apply}{\mathsf{apply}}
-\newcommand{\choice}{\mathbin{\rotatebox[origin=c]{90}{$\sqsubset\!\!\!\sqsupset$}}}
+\newcommand{\choiceop}{\rotatebox[origin=c]{90}{$\sqsubset\!\!\!\sqsupset$}}
+\newcommand{\choice}{\mathbin{\choiceop}}
+\DeclareMathOperator*{\bigchoice}{\scalerel*{\choiceop}{\sum}}
+\renewcommand{\choose}[2]{\mathsf{choose}~#1~\mathsf{from}~#2}
 \newcommand{\SKIP}{\mathsf{skip}}
 
-%math
+% math
 \newcommand{\argmin}{\mathop{\textrm{argmin}}}
+
+% defined functions
+\newcommand{\tables}{\mathop{\mathsf{tables}}}
 
@@
       & \mid & \assume b & \textit{Assumption}\\
+      & \mid & \choose \rho t & \textit{Selection} \\
       & \mid & t.\apply() & \textit{Table Application}\\
@@
 assumptions ($\assume b$), assertions ($\assert b$), table application
-($t.\apply()$), sequential composition ($c_1;c_2$ for two commands $c_1$ and
+($t.\apply()$), nondeterministic selection ($\choose \rho t$), sequential composition ($c_1;c_2$ for two commands $c_1$ and
 $c_2$), and nondeterministic choice ($c_1 \choice c_2$ for two commands $c_1$
@@
 We define $\SKIP$ as the following syntactic sugar $\assume \TRUE$.
+
+Define $\tables(c) \subset \Table$ to be the set of all tables hat occur in $c$.
 
@@
 
+A state map $\sigma : \Value \to \Row^*$ is well-formed for a command $c$ iff the
+following conditions hold:
+\begin{enumerate}
+\item[\textsc{WellFormedRows}] For every $t\in \tables(c)$, every row in $\sigma(t.\id)$ is well-formed.
+\item[\textsc{NoShadow}] For every $t \in \tables(c)$, with $r_1,\ldots,r_n =
+  \sigma(t.\id)$, for every $1 \leq i \leq n$, there does not exist $1 \leq j <
+  i$ such that $r_i.\matches = r_j.\matches$.
+\end{enumerate}
+\todo[inline]{Tables(c) is undefined}
+
 \paragraph{Row Models}
@@
 
-\begin{figure}[htp]
+\begin{figure}[H]
   \[\begin{array}{l>{\triangleq}cl}
@@
   \pkt,\mu \models_\sigma e_1 = e_2 &&
-  \edenote{e_1}~\pkt = \edenote{e_2}~\pkt, \qquad
+  \edenote{e_1}~\pkt~\mu = \edenote{e_2}~\pkt~\mu, \qquad
   \text{(note strict semantics)} \\
-  \pkt,\sigma \models \edenote{\forall \rho \in t. b} &&
+  \pkt,\mu \models_\sigma \forall \rho \in t. ~b &&
   \forall r \in \sigma(t.\id).
@@
   \[\begin{array}{l >{\triangleq}cl}
-  \multicolumn{3}{l}{\denote{c} : \Pkt \to 2^\Pkt_\bot }\\
-    \denote{h.f := e}^\sigma~\pkt && \{\pkt\{h \mapsto \edenote{e}\pkt\}\} \\
-    \denote{\assume b}^\sigma~\pkt &&
+  \multicolumn{3}{l}{\denote{c}^\sigma : \Model \to \Pkt \to \mathcal{P}(\Model \times \Pkt)_\error }\\
+    \denote{h.f := e}^\sigma~\mu~\pkt && \{(\mu, \pkt\{h \mapsto \edenote{e}~\pkt~\mu\})\} \\
+    \denote{\assume b}^\sigma~\mu~\pkt &&
     \begin{cases}
-      \{\pkt\} & \pkt \models_\sigma b \\
+      \{(\mu,\pkt)\} & \pkt,\mu \models_\sigma b \\
       \{\} & \textit{otherwise} \\
     \end{cases}\\
-    \denote{\assert b}^\sigma \pkt &&
+    \denote{\assert b}^\sigma~\mu~\pkt &&
     \begin{cases}
-      \{pkt\} & \pkt \models_\sigma b \\
-      \bot & \textit{otherwise}
+      \{(\mu,\pkt)\} & \pkt,\mu \models_\sigma b \\
+      \error & \textit{otherwise}
     \end{cases} \\
-    \denote{c_1;c_2}^\sigma \pkt&&
-    \bigcup_{\pkt' \in \denote{c_1}\pkt}\denote{c_2}~\pkt' \\
-    \denote{c_1 \choice c_2}^\sigma \pkt &&
-    \denote{c_1}^\sigma~\pkt \cup \denote{c_2}^\sigma~\pkt \\
-    \denote{t.\apply()}^\sigma \pkt &&
+    \denote{\choose \rho t}^\sigma~\mu~\pkt &&
+    \bigcup_{r \in \sigma(t)}\{(\pkt,\mu\{\rho \mapsto r\})\}
+    \\
+    \denote{c_1;c_2}^\sigma~\mu~\pkt&&
+    \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2}^\sigma~\mu'~\pkt' \\
+    \denote{c_1 \choice c_2}^\sigma~\mu~\pkt &&
+    \denote{c_1}^\sigma~\mu~\pkt \cup \denote{c_2}^\sigma~\mu~\pkt \\
+    \denote{t.\apply()}^\sigma~\mu~\pkt &&
     \begin{cases}
-      \denote{c[\vec d/\vec x]}^\sigma~\pkt,
+      \denote{c[\vec d/\vec x]}^\sigma~\mu~\pkt,
       & \text{where}~\lambda \vec x.~c = r.\action, \vec d = r.\data \\
-      & \text{where}~r = \argmin_i\{\pkt \models_\sigma \sigma(t.\id)[i].\matches = t.\keys \} \\
-
-      \{\pkt\}, & \mathit{otherwise}
+      & \text{where}~r = \argmin_i\{\pkt,\mu \models_\sigma \sigma(t.\id)[i].\matches = t.\keys \} \\
+      \{(\mu,\pkt)\}, & \mathit{otherwise}
     \end{cases}
-
   \end{array}\]
@@
   \label{fig:com-sem}
+  \todo[inline]{$\Model$ then $\Pkt$ everywhere.}
 \end{figure}
@@
   (c_1 \choice c_2)[\sigma] && c_1[\sigma] \choice c_2[\sigma] \\
+  (\choose \rho t)[\sigma] && \textrm{\textcolor{red}{unnecessary, but similar to t.apply}}\\
   (t.\apply())[\sigma] &&
@@
 \begin{theorem}[Table Elimination]
-  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$,
-  and packet $\pkt \in \Pkt$, then for any state $\tau : \Value \to \Row^*$,
-  \[\denote{c}^\sigma~\pkt = \denote{c[\sigma]}^\tau~\pkt\]
+  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, model
+  $\mu \in \Model$, and packet $\pkt \in \Pkt$, then for any state $\tau :
+  \Value \to \Row^*$,
+  \[\denote{c}^\sigma~\mu~\pkt = \denote{c[\sigma]}^\tau~\mu~\pkt\]
 \end{theorem}
@@
   (c_1 \choice c_2)[?] && c_1[?] \choice c_2[?]\\
+  (\choose \rho t)[?] &&
+  \textrm{\textcolor{red}{unnecessary}} \\
   (t.\apply())[?] &&
-  \begin{array}[t]{l}
+  \begin{array}[t]{l} \displaystyle
     \{ \\
+    \quad \choose \rho t; \\
     \quad \assume \rho.\matches = t.\keys; \\
-  \quad \{\assume \rho.\action = 0; t.\action[0](\rho.\data)\}\\
-  \quad \choice \cdots \\
-  \quad \choice \{\assume \rho.action = |t.\action| - 1; t.\action[t.\action - 1](\rho.\data)\}\\
-  \quad \}\\
+    \quad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
+    \left\{\begin{array}{l}
+    \assume \rho.\action = i;\\
+    t.\action[i](\rho.\data)
+    \end{array}\right\}\\
     \} \choice \{ \\
@@
   \]
+  \caption{Symbolic Table instrumentation}
+  \label{fig:table-instrument}
 \end{figure}
 
-\todo[inline]{This is wrong because $\rho$ is unbound in the first branch}
-
-\begin{theorem}[Instrumentation Sufficiency]
-  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, and
-  packet $\pkt \in \Pkt$, $\denote{c}^\sigma~\pkt = \pkt'$ if and only if
-  $\pkt' \in \denote{c[?]}^\sigma~\pkt$
+\begin{theorem}[Instrumentation Expressivity]
+  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, model
+  $\mu \in \Model$, and packet $\pkt \in \Pkt$, $\denote{c}^\sigma~\mu~\pkt =
+  \denote{c[?]}^\sigma~\mu~\pkt$.
 \end{theorem}
@@
 \begin{proof}
-\todo[inline]{By induction on the structure of $c$}
+  Proceed by induction on the structure of $c$, leaving $\sigma$, $\mu$, $\pkt$ general.
+  \begin{enumerate}[align=left]
+  \item[$(c = h.f := e)$]
+    Let $\sigma$,$\mu$,$\pkt$ be given.
+    By definition $h.f:=e[?] = h.f:=e$. The result follows.
+  \item[$(c = \assume b)$] Similar. $[?]$ is no-op.
+  \item[$(c = \assert b)$] Similar. $[?]$ is no-op.
+  \item[$(c = \choose \rho t)$] \textcolor{red}{Admitted}.
+  \item[$(c = c_1;c_2)$]
+    Let $\sigma,\mu,\pkt$ be given.
+    By definition $(c_1;c_2)[?] = c_1[?]; c_2[?]$.
+    Compute as follows:
+    \[\begin{array}{>{\displaystyle}ll}
+    \denote{(c_1;c_2)[?]}^\sigma~\mu~\pkt \\
+    = \denote{c_1[?];c_2[?]}^\sigma~\mu\pkt
+    & \text{by definition of } [?]  \\
+    = \bigcup_{(\mu',\pkt') \in \denote{c_1[?]}^\sigma\,\mu\,\pkt}\denote{c_2[?]}^\sigma~\mu'~\pkt'
+    & \text{by definition of }\denote{-} \\
+    = \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2[?]}^\sigma~\mu'~\pkt'
+    & \text{by } \mathit{IH}(c_1, \mu, \pkt) \\
+    = \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2}^\sigma~\mu'~\pkt'
+    & \text{by } \mathit{IH}(c_2, \mu', \pkt') \\
+    = \denote{c_1;c_2}^\sigma~\mu~\pkt
+    & \text{by definition of } \denote{-}
+    \end{array}
+    \]
+  \item[$(c = c_1 \choice c_2)$] Similar, by IHs.
+
+  \item[$(c = t.\apply())$]
+    Let $\sigma$,$\mu$,$\pkt$ be given.
+    Let $r_1,\ldots,r_n = \sigma(t)$.
+    There are two cases, either the packet hits a row, or hits no row. Explore each case separately.
+    \begin{enumerate}
+    \item[\textsc{Hit}] In this case, there is some $r_i$ such that $\pkt,\mu
+      \models_\sigma r_i.\matches = t.\keys$. By the well-formedness assumption,
+      we know that this $r_i$ is unique, and that we can write
+      $t.\actions[r_i.\action]$ as $\lambda \vec x. c$, and $|r_i.\data| = |\vec x|$.
+
+      Compute as follows:
+      \[\begin{array}{>{\displaystyle}ll}
+      \denote{t.\apply()}^\sigma~\mu~\pkt \\
+      = \denote{c[r_i.data/\vec x]}^\sigma~\mu~\pkt & \text{by definition of} \models \text{, and the uniqueness of }r_i  \\
+      \end{array}
+      \]
+
+      Now, we compute the denotation of $t.\apply()[?]$ to see that it is equivalent to the above.
+      The definition of $t.\apply()[?]$ gives us two alternatives. Consider the second, first:
+      \[\begin{array}{>{\displaystyle}ll}
+        \denote{\assume \forall \rho \in t. \rho.\matches \neq t.\keys;\SKIP}^\sigma~\mu~\pkt \\
+        = \denote{\assume \forall \rho \in t. \rho.\matches \neq t.\keys}^\sigma~\mu~\pkt
+        & \textrm{\textcolor{red}{Redo with defaults}} \\
+        = \begin{cases}
+          \{(\mu,\pkt)\}, & \pkt,\mu \models_\sigma \rho.\matches \neq t.\keys \\
+          \{\}, &\textit{otherwise}
+        \end{cases}
+        & \textrm{By definition of }\denote{-}\\
+        =  \{\}
+        & \textrm{by \textsc{Hit} assumption}
+      \end{array}
+      \]
+
+      Now let's consider the first branch, considering each command sequentially.
+      \[\begin{array}{>{\displaystyle}ll}
+      \denote{\choose \rho t; \assume \rho.\matches = t.\keys}^\sigma~\mu~\pkt \\
+      = \bigcup_{r \in \sigma(t)}\denote{\assume\rho.\matches = t.\keys}^\sigma~\mu\{\rho \mapsto r\}~\pkt
+      & \textrm{by definition of }\denote{-}\\
+      = \{(\mu\{\rho\mapsto r\},\pkt)
+      \mid r \in \sigma(t), \pkt,\mu\{\rho\mapsto r\} \models_\sigma \rho.\matches = t.\keys
+      \}
+      & \textrm{by definitions and set theory} \\
+      = \{(\mu\{\rho \mapsto r\}, \pkt)
+      \mid r \in \sigma(t), \pkt,\mu \models_\sigma r.\matches = t.\keys\}
+      & \textrm{by definitions \textcolor{red}{TODO Lemma}} \\
+      = \{(\mu\{\rho \mapsto r_i\}, \pkt)\}
+      & \textrm{by \textsc{Hit} and \textsc{NoShadow}} \\
+      \end{array}\]
+
+      Now by the sequential composition rule, we can determine the result of the
+      first branch by computing as follows:
+      \[\begin{array}{>{\displaystyle}ll}
+      \denote{\bigchoice_{i=0}^{|t.\action| -1}
+        \left\{\begin{array}{l}
+        \assume \rho.\action = i;\\
+        t.\actions[i](\rho.\data)
+        \end{array}\right\}
+      }^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt \\
+      = \denote{\denote{c[r_i.\data/\vec x]}^\sigma~(\mu\{\rho \mapsto r\})~\pkt}\\
+      & \textrm{skipping lots of definition steps} \\
+      \end{array}
+      \]
+      \todo[inline]{Need Scoping Rules for $\rho$}
+
+    \item[\textsc{Miss}]
+      \todo[inline]{Should be easy}
+
+    \end{enumerate}
+
+  \end{enumerate}
 \end{proof}

### 2021-03-27 inference.tex
Outline proofs; Scope isn't managed correctly

@@
 \DeclareMathOperator*{\bigchoice}{\scalerel*{\choiceop}{\sum}}
-\renewcommand{\choose}[2]{\mathsf{choose}~#1~\mathsf{from}~#2}
+\renewcommand{\choose}[2]{\mathop{\mathsf{choose}~#1~\mathsf{from}~#2~\mathsf{in}}}
 \newcommand{\SKIP}{\mathsf{skip}}
@@
       & \mid & \assume b & \textit{Assumption}\\
-      & \mid & \choose \rho t & \textit{Selection} \\
       & \mid & t.\apply() & \textit{Table Application}\\
@@
     & \mid & c \choice c & \textit{Nondeterministic Choice}\\
+    & \mid & \choose \rho t c & \textit{Selection} \\
     \multicolumn{3}{l}{
@@
 assumptions ($\assume b$), assertions ($\assert b$), table application
-($t.\apply()$), nondeterministic selection ($\choose \rho t$), sequential composition ($c_1;c_2$ for two commands $c_1$ and
+($t.\apply()$), lexically-scoped nondeterministic selection ($\choose \rho t c$), sequential composition ($c_1;c_2$ for two commands $c_1$ and
 $c_2$), and nondeterministic choice ($c_1 \choice c_2$ for two commands $c_1$
@@
 We also need models $\mu \in \Model$ to indicate which symbolic rows corresponds
-to which concrete rows, i.e. $\Model$ is a map from symbolic rows to concrete
-rows; write $\mu(\rho) = r$ for a model $\mu$, a symbolic row $\rho$, and a
-concrete row $r$.
+to which concrete rows, i.e. $\Model = \SymbRow \rightharpoonup \Row$. We write
+$\mu\{\rho \mapsto r\}$ to extend $\mu$ by mapping $\rho$ to $r$, and write $\mu
+\setminus \rho$, to remove $\rho$ from the map.
 
@@
 Commands are evaluated similar to Avenir in Figure~\ref{fig:com-sem}. The
-notable differences are the inclusion of assertions, the use of a binary
-nondeterminism operator, and a cleaner representation of table application.
+notable differences are the inclusion of assertions, the nondeterminism
+operators, and a cleaner representation of table application.
 
@@
     \end{cases} \\
-    \denote{\choose \rho t}^\sigma~\mu~\pkt &&
-    \bigcup_{r \in \sigma(t)}\{(\pkt,\mu\{\rho \mapsto r\})\}
+    \denote{\choose \rho t c}^\sigma~\mu~\pkt &&
+    \{(\mu'\setminus \rho,\pkt')
+    \mid (\mu',\pkt') \in \denote{c}^\sigma\,(\mu\{\rho\mapsto r\})\,\pkt, r \in \sigma(t)\}
     \\
@@
       & \text{where}~\lambda \vec x.~c = r.\action, \vec d = r.\data \\
-      & \text{where}~r = \argmin_i\{\pkt,\mu \models_\sigma \sigma(t.\id)[i].\matches = t.\keys \} \\
+      & \text{where}~r = \argmin_i\{\pkt,\mu \models_\sigma b_i\} \\
+      & \text{where}~b_i = \sigma(t.\id)[i].\matches = t.\keys \\
       \{(\mu,\pkt)\}, & \mathit{otherwise}
@@
   (c_1 \choice c_2)[\sigma] && c_1[\sigma] \choice c_2[\sigma] \\
-  (\choose \rho t)[\sigma] && \textrm{\textcolor{red}{unnecessary, but similar to t.apply}}\\
+  (\choose \rho t c)[\sigma] && \textrm{\textcolor{red}{unnecessary, but similar to t.apply}}\\
   (t.\apply())[\sigma] &&
@@
   (c_1 \choice c_2)[?] && c_1[?] \choice c_2[?]\\
-  (\choose \rho t)[?] &&
+  (\choose \rho t c)[?] &&
   \textrm{\textcolor{red}{unnecessary}} \\
@@
     \{ \\
-    \quad \choose \rho t; \\
-    \quad \assume \rho.\matches = t.\keys; \\
-    \quad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
+    \quad \choose \rho t \;\{ \\
+    \qquad \assume \rho.\matches = t.\keys; \\
+    \qquad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
     \left\{\begin{array}{l}
@@
     \end{array}\right\}\\
+    \quad \} \\
     \} \choice \{ \\
@@
 
-      Now let's consider the first branch, considering each command sequentially.
-      \[\begin{array}{>{\displaystyle}ll}
-      \denote{\choose \rho t; \assume \rho.\matches = t.\keys}^\sigma~\mu~\pkt \\
-      = \bigcup_{r \in \sigma(t)}\denote{\assume\rho.\matches = t.\keys}^\sigma~\mu\{\rho \mapsto r\}~\pkt
-      & \textrm{by definition of }\denote{-}\\
-      = \{(\mu\{\rho\mapsto r\},\pkt)
-      \mid r \in \sigma(t), \pkt,\mu\{\rho\mapsto r\} \models_\sigma \rho.\matches = t.\keys
-      \}
-      & \textrm{by definitions and set theory} \\
-      = \{(\mu\{\rho \mapsto r\}, \pkt)
-      \mid r \in \sigma(t), \pkt,\mu \models_\sigma r.\matches = t.\keys\}
-      & \textrm{by definitions \textcolor{red}{TODO Lemma}} \\
-      = \{(\mu\{\rho \mapsto r_i\}, \pkt)\}
-      & \textrm{by \textsc{Hit} and \textsc{NoShadow}} \\
-      \end{array}\]
-
-      Now by the sequential composition rule, we can determine the result of the
-      first branch by computing as follows:
-      \[\begin{array}{>{\displaystyle}ll}
-      \denote{\bigchoice_{i=0}^{|t.\action| -1}
+      Now let's consider the first branch, i.e.
+      \[
+      \denote{\choose \rho t \left\{
+        \begin{array}{>{\displaystyle}l}
+          \assume \rho.\matches = t.\keys;\\
+          \bigchoice_{i=0}^{|t.\action| -1}
           \left\{\begin{array}{l}
@@
           \end{array}\right\}
-      }^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt \\
-      = \denote{\denote{c[r_i.\data/\vec x]}^\sigma~(\mu\{\rho \mapsto r\})~\pkt}\\
-      & \textrm{skipping lots of definition steps} \\
-      \end{array}
+        \end{array}\right\}
+      }^\sigma~\mu~\pkt\]
+
+      Compute the following inner result, by definitions and set theory
+      \[
+      \begin{array}{>{\displaystyle}ll}
+      \bigcup_{r \in \sigma(t)}\denote{\begin{array}{>{\displaystyle}l}
+          \assume \rho.\matches = t.\keys;\\
+          \bigchoice_{i=0}^{|t.\action| -1}
+          \left\{\begin{array}{l}
+          \assume \rho.\action = i;\\
+          t.\actions[i](\rho.\data)
+          \end{array}\right\}
+      \end{array}}^\sigma~\mu\{\rho \mapsto r\}~\pkt\\
+      = \bigcup_{
+        \substack{r \in \sigma(t) \\
+          (\mu',\pkt') \in \denote{\assume \rho.\matches = t.\keys}^\sigma\,(\mu\{\rho \mapsto r\})\,\pkt}}
+      \bigcup_{i=0}^{|t.\action| -1}\denote{
+       \begin{array}{l}
+          \assume \rho.\action = i;\\
+          t.\actions[i](\rho.\data)
+      \end{array}}^\sigma~\mu'~\pkt'
+
+      \\\\
+      (\textrm{by \textsc{Hit}, \textsc{NoShadow}, and the definition of }\denote{-}) \\
+      \\
+      = \bigcup_{i=0}^{|t.\action| -1}\denote{
+       \begin{array}{l}
+          \assume \rho.\action = i;\\
+          t.\actions[i](\rho.\data)
+      \end{array}}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt \\
+      = \bigcup_{j=0}^{|t.\action| -1}\denote{
+       \begin{array}{l}
+          \assume r_i.\action = j;\\
+          t.\actions[j](r_i.\data)
+      \end{array}}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt \\
+      \textrm{\textcolor{red}{TODO LEMMA}} \\
+      = \denote{t.\actions[r_i.\action](r_i.\data)}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt \\
+      = \denote{c[r_i.\data/\vec x]}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt
+      \end{array}\]
+
+      Since $\rho$ is fresh in $c$, and consequently in $c[r_i.\data/\vec x]$,
+      then by \textcolor{red}{Lemma},
+      \[\{(\mu'\setminus \rho,\pkt')
+      \mid (\mu', \pkt') \in \denote{c[r_i.\data/\vec x]}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt\}
+       = \denote{c[r_i.\data/\vec x]}^\sigma~\mu~\pkt
        \]
-      \todo[inline]{Need Scoping Rules for $\rho$}
+       \todo[inline]{Relies on the fact that $\rho \not \in \mu$, or that $\rho$ is globally fresh}
+
+       The result follows.
 

### 2021-03-27 inference.tex
partially fix scoping issue, still need to generalize pop-up rule

@@
 \newtheorem{theorem}{Theorem}
+\newtheorem{lemma}{Lemma}
 
@@
 
-For actions, we restrict the language to be generated by only assignment, assumption, assertion and sequence, excluding table application and nondeterministic choice.
+We define $\SKIP$ as the following syntactic sugar $\assume \TRUE$.
 
-We define $\SKIP$ as the following syntactic sugar $\assume \TRUE$.
+For actions, we restrict the language to be generated by only non-symbolic
+assignment, sequence, and $\SKIP$.
 
@@
 \end{enumerate}
-\todo[inline]{Tables(c) is undefined}
-
 \paragraph{Row Models}
@@
 
+\begin{lemma}[Models Invariant over Actions]
+  \label{lem:models-invariant-over-actions}
+  For an action command $c$, $\mu,\mu'
+  \in \Model$, $\sigma \in \Value \to
+  \Row^*$, $\pkt \in \Pkt$, there exists some $\pkt'$
+  \[\denote{c}^\sigma\,\mu\,\pkt = \{\mu,\pkt'\}\]
+  \end{lemma}
+
+\begin{proof}
+  Proceed by induction over $c$.
+  \begin{enumerate}[align=left]
+  \item[($c = \SKIP$)] Immediate.
+  \item[$(c = h.f = e)$] Immediate.
+  \item[$(c = c_1;c_2)$]
+    By IH($c_1$) there exists some $\pkt''$ such that
+    $\denote{c_1}^\sigma\,\mu\,\pkt = \{(\mu,\pkt'')\}$.
+    By IH($c_2$) there exists some $\pkt'$ such that
+    $\denote{c_2}^\sigma,\mu\,pkt'' = \{(\mu, \pkt')\}$.
+    The result follows by definition of $\denote{-}$.
+    \end{enumerate}
+\end{proof}
+
 \begin{theorem}[Instrumentation Expressivity]
@@
       = \denote{t.\actions[r_i.\action](r_i.\data)}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt \\
-      = \denote{c[r_i.\data/\vec x]}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt
+      = \denote{c[r_i.\data/\vec x]}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt \\
       \end{array}\]
 
-      Since $\rho$ is fresh in $c$, and consequently in $c[r_i.\data/\vec x]$,
-      then by \textcolor{red}{Lemma},
-      \[\{(\mu'\setminus \rho,\pkt')
-      \mid (\mu', \pkt') \in \denote{c[r_i.\data/\vec x]}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt\}
-       = \denote{c[r_i.\data/\vec x]}^\sigma~\mu~\pkt
+      By Lemma~\ref{lem:models-invariant-over-actions}, there's some $\pkt'$ such that
+      \[\{\mu\{\rho \mapsto r_i\}, \pkt'\} =
+      \denote{c[r_i.\data/\vec x]}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt
       \]
-       \todo[inline]{Relies on the fact that $\rho \not \in \mu$, or that $\rho$ is globally fresh}
+
+      which means that the result of the first branch is
+      \[\{\mu, \pkt'\} =
+      \denote{c[r_i.\data/\vec x]}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt
+      \]
+
+      Since $\rho$ doesn't occur in $c$, then also
+      \[\{\mu, \pkt'\} = \denote{c[r_i.\data/\vec x]}^\sigma~\mu~\pkt \]
 

### 2021-03-29 inference.tex
Add subgrammars

@@
 \newcommand{\Cmd}{\mathsf{Cmd}}
+\newcommand{\Action}{\mathsf{Action}}
+\newcommand{\Instr}{\mathsf{Instr}}
+\newcommand{\Prog}{\mathsf{Prog}}
+\newcommand{\GCL}{\mathsf{GCL}}
 \newcommand{\Pkt}{\mathsf{Pkt}}
@@
 \newcommand{\argmin}{\mathop{\textrm{argmin}}}
+\newcommand{\WP}{\textit{wp}}
 
@@
 \newtheorem{lemma}{Lemma}
+\newtheorem{conjecture}{Conjecture}
 
@@
 \begin{figure}[htp]
-  \[\begin{array}{lcll}
+  \[\begin{array}{lclll}
     \multicolumn{3}{l}{\Value} \\
-    v & ::= & [n]_w  & \textit{Bitvector Literal}\\
+    v & ::= & [n]_w  & \textit{Bitvector Literal} & (a,g,p,\iota) \\
     \multicolumn{3}{l}{\SymbRow} \\
@@
       \data = \List[x]
-    \end{array}\right\} & \textit{Symbolic Row}\\
+    \end{array}\right\} & \textit{Symbolic Row} & (\iota)\\
     \multicolumn{3}{l}{\BVExpr} \\
-    e & ::= & v & \textit{Value}\\
-      & \mid & h.f & \textit{Header Access} \\
-      & \mid & \rho.\matches[i] & \textit{Symbolic Match} \\
-      & \mid & \rho.\action & \textit{Symbolic Action}\\
-      & \mid & \rho.\data[i] & \textit{Symbolic Action Data}\\
-      & \mid & x & \textit{Variable}\\
-      & \mid & e \binop e & \textit{Arbitrary Binary Operation}\\
-    & \mid & \unop e & \textit{Arbitrary Unary Operation}\\
+    e & ::= & v & \textit{Value} & (a,g,p,\iota)\\
+      & \mid & h.f & \textit{Header Access} & (a,g,p,\iota)\\
+      & \mid & \rho.\matches[i] & \textit{Symbolic Match} & (\iota)\\
+      & \mid & \rho.\action & \textit{Symbolic Action} & (\iota) \\
+      & \mid & \rho.\data[i] & \textit{Symbolic Action Data} & (\iota) \\
+      & \mid & x & \textit{Action Data Variable} & (a^*) \\
+      & \mid & e \binop e & \textit{Arbitrary Binary Operation} & (a,g,p,\iota) \\
+    & \mid & \unop e & \textit{Arbitrary Unary Operation} & (a,g,p,\iota) \\
     \multicolumn{3}{l}{\Table} \\
@@
     t.\keys = \List[h.f]; \\
-    t.\actions = \List[\lambda \vec x. c]\\
-    \end{array}\right\} & \textit{Table}\\
+    t.\actions = \List[\lambda \vec x.\,a]\\
+    \end{array}\right\} & \textit{Table} & (p)\\
     \multicolumn{3}{l}{\BExpr} \\
-    b & ::= & \FALSE  & \textit{Falsity}\\
-      & \mid & b \Rightarrow b & \textit{Implication} \\
-      & \mid & e = e & \textit{Bitvector Equality}\\
-    & \mid & \forall \rho \in t.~b & \textit{Row Universality}\\
-    \multicolumn{3}{l}{\Cmd} \\
-    c & ::=  & h.f := e & \textit{Assignment}\\
-      & \mid & \assert b & \textit{Assertion}\\
-      & \mid & \assume b & \textit{Assumption}\\
-      & \mid & t.\apply() & \textit{Table Application}\\
-      & \mid & c;c & \textit{Sequential Composition} \\
-    & \mid & c \choice c & \textit{Nondeterministic Choice}\\
-    & \mid & \choose \rho t c & \textit{Selection} \\
+    b & ::= & \FALSE  & \textit{Absurdity}& (a,g,p,\iota)\\
+      & \mid & b \Rightarrow b & \textit{Implication} & (a,g,p,\iota)\\
+      & \mid & e = e & \textit{Bitvector Equality} & (a,g,p,\iota)\\
+    & \mid & \forall \rho \in t.~b & \textit{Row Universality} & (\iota)\\
+    \multicolumn{4}{l}{c \in \Cmd, a \in \Action, g \in \GCL, p \in \Prog, \iota \in \Instr } \\
+    c & ::=  & h.f := e & \textit{Assignment} & (g,p,\iota) \\
+      & \mid & \assert b & \textit{Assertion} & (g,p,\iota)\\
+      & \mid & \assume b & \textit{Assumption} & (g,p,\iota) \\
+      & \mid & t.\apply() & \textit{Table Application} & (p)\\
+      & \mid & c;c & \textit{Sequential Composition} & (a,g,p,\iota) \\
+    & \mid & c \choice c & \textit{Nondeterministic Choice} & (g,p,\iota)\\
+    & \mid & \choose \rho t c & \textit{Selection} & (\iota) \\
     \multicolumn{3}{l}{
@@
 \]
-\caption{Grammar of Dataplane Programs}
+\caption{Grammar of Commands $c \in \Cmd$. Legal action indicated by $a \in
+  \Action$, legal instrumented command variants indicated by $\iota \in \Instr$,
+  dataplane programs indicated by $p \in \Prog$, and table-free commands
+  indicated by $g \in \GCL$.}
 \label{fig:grammar}
@@
 assumptions ($\assume b$), assertions ($\assert b$), table application
-($t.\apply()$), lexically-scoped nondeterministic selection ($\choose \rho t c$), sequential composition ($c_1;c_2$ for two commands $c_1$ and
-$c_2$), and nondeterministic choice ($c_1 \choice c_2$ for two commands $c_1$
-and $c_2$).
+($t.\apply()$), lexically-scoped nondeterministic selection ($\choose \rho t
+c$), sequential composition ($c_1;c_2$ for two commands $c_1$ and $c_2$), and
+nondeterministic choice ($c_1 \choice c_2$ for two commands $c_1$ and $c_2$).
 
@@
 
-For actions, we restrict the language to be generated by only non-symbolic
-assignment, sequence, and $\SKIP$.
+We express a collection of invariants on commands as subsets of the set $\Cmd$.
+It is convenient to describe these as subgrammars. First, action-commands $a \in
+\Action$ contain only non-symbolic assignment and sequence. If an action
+command has variables, indicated by the starred variant $a*$, it is called
+\emph{open}, if it has none it is called \emph{closed}.
 
-Define $\tables(c) \subset \Table$ to be the set of all tables hat occur in $c$.
+Second, $p \in \Prog \subset \Cmd$ represents the expected set of programs extracted from a
+P4 interface, most notably, it contains tables $t$ and no symbolic selection
+statements or symbolic expersions.
+
+Conversely $\iota \in \Instr \subset \Cmd$ represents a program instrumented
+with symbolic variables. A program $\iota$ cannot apply tables using
+$t.\apply()$, but instead uses a havoc statement $\choose \rho t c$, to
+introduce symbolic variables $\rho$ in $c$.
+
+Finally $g \in \GCL \subset \Prog \subset \Cmd$ represents the core algebraic
+flavor of GCL without havoc, symbolic tables, or table application.
+
+Define $\tables(c) \subset \Table$ to be the set of all tables that occur in $c$.
 
@@
   \[\begin{array}{r>{\triangleq}cl}
+  \multicolumn{3}{l}{(\cdot)[\sigma] : \Cmd \to \Cmd} \\
+  \multicolumn{3}{l}{(\cdot)[\sigma] : \Prog \to \GCL} \\
   (h.f := e)[\sigma] && h.f := e   \\
@@
   \[\begin{array}{r >{\triangleq}cl}
+  \multicolumn{3}{l}{(\cdot)[?] : \Cmd \to \Cmd} \\
+  \multicolumn{3}{l}{(\cdot)[?] : \Prog \to \Instr} \\
   (h.f := e)[?] && h.f := e \\
@@
   \label{lem:models-invariant-over-actions}
-  For an action command $c$, $\mu,\mu'
+  For a closed action command $c$, $\mu,\mu'
   \in \Model$, $\sigma \in \Value \to
@@
   \begin{enumerate}[align=left]
-  \item[($c = \SKIP$)] Immediate.
   \item[$(c = h.f = e)$] Immediate.
@@
 
-\begin{theorem}[Instrumentation Expressivity]
+\begin{theorem}[Instrumentation Equivalence]
+  \label{thm:instr-equiv}
   For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, model
@@
       By Lemma~\ref{lem:models-invariant-over-actions}, there's some $\pkt'$ such that
-      \[\{\mu\{\rho \mapsto r_i\}, \pkt'\} =
+      \[\{(\mu\{\rho \mapsto r_i\}, \pkt')\} =
       \denote{c[r_i.\data/\vec x]}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt
@@
       which means that the result of the first branch is
-      \[\{\mu, \pkt'\} =
-      \denote{c[r_i.\data/\vec x]}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt
-      \]
+      \[\{(\mu, \pkt')\} \]
 
-      Since $\rho$ doesn't occur in $c$, then also
-      \[\{\mu, \pkt'\} = \denote{c[r_i.\data/\vec x]}^\sigma~\mu~\pkt \]
+      Since $\rho$ doesn't occur in $c$, then there's some $\mu'$ such that also
+      \[\{\mu', \pkt'\} = \denote{c[r_i.\data/\vec x]}^\sigma~\mu~\pkt \]
+
+      and $\mu = \mu'$ by Lemma~\ref{lem:models-invariant-over-actions}.
 
@@
 
-    \item[\textsc{Miss}]
-      \todo[inline]{Should be easy}
+    \item[\textsc{Miss}] In this case, there is no $r_i$ such that $\pkt,\mu
+      \models_\sigma r_i.\matches = t.\keys$, which means that the LHS of the
+      toplevel choice operator should evaluate to $\{\}$ \todo{compute this},
+      and the RHS evaluates to $\{\mu,\pkt\}$\todo{compute this}, which is
+      precisely the result we get when executing
+      $\denote{t.apply()}^\sigma~\mu~\pkt$.
 
@@
 
+\section{The Research Problem}
+
+Now we can formally define the research problems that we want to solve. We have
+some ideal goals, that are probably not solvable in general.
+
+\begin{quote}
+  \textbf{Conservative Problem.} Given a command $c \in \Cmd$, compute the weakest test
+  $b$ such that if a state $\sigma : \Value \to \Row^*$ the following
+  judgement is defined and satisfied \[\emptyset, \emptyset \models_\sigma b,\] then \[\forall \pkt \in \Pkt.~
+  \denote{c}^\sigma \pkt \neq \error \]
+\end{quote}
+
+However, this $b$ may be $\FALSE$ for many realistic programs $c$, so we may
+prefer to solve the following problem:
+
+\begin{quote}
+  \textbf{Flexible Problem.} Given a command $c \in \Cmd$, compute a test
+  $b_{\mathit{cp}}$, and a test $b_{\mathit{pkt}}$ such that
+  for every packet $\pkt \in \Pkt$ such that
+  \[\emptyset, \emptyset \models_\emptyset b_{\mathit{pkt}},\]
+  and for every state $\sigma : \Value \to \Row^*$ such that
+  \[\pkt, \emptyset \models_\sigma b_{\mathit{cp}},\]
+  then \[\denote{c}^\sigma ~\emptyset~\pkt \neq \error \]
+\end{quote}
+
+\section{Reduction to QE}
+
+We can observe that the \textbf{Conservative Problem} is simply solvable. We
+sketch the argument here.
+
+\begin{conjecture}[Quantifer Elimination]
+  The \textbf{\textrm{Conservative Problem}} is reducable to Quantifer Elimination.
+\end{conjecture}
+
+\begin{proof}[Proof Sketch.]
+  Given a command $c \in \Cmd$. Compute the instrumented command $c[?]$. We know
+  it is equivalent to $c$ by Theorem~\ref{thm:instr-equiv}. Then $\forall
+  \overrightarrow{h.f}.\WP(c[?],\TRUE)[\overrightarrow{h.f}]$,\todo{how to write this?} is such a
+  formula $b$ by Dijkstra '75. To make $b$ a controller formula, then we simply
+  need to remove the dataplane variables via a QE procedure.
+\end{proof}
+
+This doesn't work in general. Standard QE procedures are exponential at best, and
+enumerating all eth/ipv4 src/dst addresses already takes an intractable amount
+of time.
+
+\section{Can Templates save us?}
+
+One option is to use a kind of constrained template-based synthesis, where we
+generate a catalogue of temaplates based on syntactic features of the target table.
+
+For example, given a single table like the following:
+
+\[t = \{\keys: [h.f]; \actions: [\lambda x.~~h.g:= x]\}\]
+
+and the simple program \[t.\apply()\]
+
+We may be able to generate a template that looks like
+
+\[\begin{array}{>{\displaystyle}c}
+  {[\exists \rho \in t.~ \rho.\keys[0]~\square_{\{=,\neq\}}~?_{v_1}\wedge \rho.\data[0]~\square_{\{=,\neq\}}~?_{v_2} ]?} \\
+  \wedge\\
+  {[\forall \rho \in t.~\rho.\keys[0]~\square_{\{=,\neq\}}~?_{v_3} \Rightarrow \rho.\data[0]~\square_{\{=,\neq\}}~?_{v_4}]?}
+\end{array}\]
+
+where $[b]?$ indicates a hole to either use template $b$ or replace it with
+$\TRUE$, $\square_S$ indicates a template for an operation, where $S$ is the set
+of allowed operations, and $?_{v_i}$ indicates a hole that should evaluate to an element
+of $\Value$.
+
+Now we may wonder: \emph{can we procedurally generate templates in a (nearly) complete way}?
+
+To do so, we need to consider the following dimensions:
+\begin{enumerate}[align=left]
+\item[\textbf{Keys}.] Should generalize by changing the LHS comparisons to $
+  \bigwedge_i[\rho.\keys[i]~\square_{\{=,\neq\}}~?_{v_i}]?$
+\item[\textbf{Action Data}.] Should generalize by changing the RHS comparisons to $\bigwedge_i[\rho.\data[i]~\square_{\{=,\neq\}}~?_{v_i}]?$
+\item[\textbf{Actions}.] We need to generalize the template across multiple actions. This should be as simple as adding $\rho.\action = i$ on the RHS of each implication above.
+\item[\textbf{Sequence}.] I'm not sure how to generalize sequentially composed tables.
+  Perhaps a larger catalogue of templates conditioned on structural information:
+  if we notice that $t_1$ has a dataflow path to the keys of $t_2$, we should
+  generate a join condition template, otherwise we may be able to simply ``and'' the templates.
+
+\item[\textbf{Nondeterminism} ($c_1 \choice c_2$).]. To solve the \textbf{Conservative
+  Problem} we simply and the templates together, but to solve the
+  \textbf{Flexible Problem} we may be able to constrain the problem (by
+  constraining the packets) to either $c_1$ or $c_2$. This would require
+  computing a $b_{\textit{pkt}}$.
+\end{enumerate}
+
 \end{document}

### 2021-03-29 inference.tex
clean up satisfaction

@@
 \newcommand{\WP}{\textit{wp}}
+\newcommand{\satisfy}[3]{({#1,#3}) \models_{#2}}
+\newcommand{\notsatisfy}[3]{({#1,#3})\not\models_{#2}}
 
@@
 \paragraph{Boolean Satisfaction}
-The satisfaction relation $\pkt,\mu \models_\sigma b$ is defined in
+The satisfaction relation $\satisfy \sigma \pkt \mu b$ is defined in
 Figure~\ref{fig:bool-satis} in terms of the three kinds of variable contexts,
@@
   \[\begin{array}{l>{\iff}cl}
-  \pkt,\mu \models_\sigma \FALSE && \textit{pigs fly} \\
-  \pkt,\mu \models_\sigma b_1 \Rightarrow b_2 &&
-  \pkt,\mu \not\models_\sigma b_1~\textit{or}~
-  \pkt,\mu \models_\sigma b_2\\
-  \pkt,\mu \models_\sigma e_1 = e_2 &&
+  \satisfy \sigma \pkt \mu \FALSE && \textit{pigs fly} \\
+  \satisfy \sigma \pkt \mu b_1 \Rightarrow b_2 &&
+  \notsatisfy \sigma \pkt \mu  b_1~\textit{or}~
+  \satisfy \sigma \pkt \mu b_2\\
+  \satisfy \sigma \pkt \mu e_1 = e_2 &&
   \edenote{e_1}~\pkt~\mu = \edenote{e_2}~\pkt~\mu, \qquad
   \text{(note strict semantics)} \\
-  \pkt,\mu \models_\sigma \forall \rho \in t. ~b &&
+  \satisfy \sigma \pkt \mu \forall \rho \in t. ~b &&
   \forall r \in \sigma(t.\id).
-  \pkt,\sigma,\mu[\rho \mapsto r] \models b
+  \satisfy \sigma \pkt {\mu\{\rho \mapsto r\}} b
   \end{array}
   \]
-  \caption{Satisfaction for $\BExpr$. Write $\pkt \models_\sigma b$ as syntactic
-    sugar when the model $\mu$ is empty.}
+  \todo[inline]{Clarify $\error$ states}
+  \todo[inline]{@Nate I know you wanted a standard satisfaction relation, but this was easier to think about in the context of the domain. The difference boils down to a naming convention}
+  \caption{Satisfaction for $\BExpr$.}
   \label{fig:bool-satis}
@@
     \begin{cases}
-      \{(\mu,\pkt)\} & \pkt,\mu \models_\sigma b \\
+      \{(\mu,\pkt)\} & \satisfy \sigma \pkt \mu b \\
       \{\} & \textit{otherwise} \\
@@
     \begin{cases}
-      \{(\mu,\pkt)\} & \pkt,\mu \models_\sigma b \\
+      \{(\mu,\pkt)\} & \satisfy \sigma \pkt \mu b \\
       \error & \textit{otherwise}
@@
       & \text{where}~\lambda \vec x.~c = r.\action, \vec d = r.\data \\
-      & \text{where}~r = \argmin_i\{\pkt,\mu \models_\sigma b_i\} \\
+      & \text{where}~r = \argmin_i\{\satisfy \sigma \pkt \mu b_i\} \\
       & \text{where}~b_i = \sigma(t.\id)[i].\matches = t.\keys \\
@@
     \begin{enumerate}
-    \item[\textsc{Hit}] In this case, there is some $r_i$ such that $\pkt,\mu
-      \models_\sigma r_i.\matches = t.\keys$. By the well-formedness assumption,
+    \item[\textsc{Hit}] In this case, there is some $r_i$ such that $\satisfy \sigma \pkt \mu r_i.\matches = t.\keys$. By the well-formedness assumption,
       we know that this $r_i$ is unique, and that we can write
@@
         = \begin{cases}
-          \{(\mu,\pkt)\}, & \pkt,\mu \models_\sigma \rho.\matches \neq t.\keys \\
+          \{(\mu,\pkt)\}, & \satisfy\sigma\pkt\mu \rho.\matches \neq t.\keys \\
           \{\}, &\textit{otherwise}
@@
 
-    \item[\textsc{Miss}] In this case, there is no $r_i$ such that $\pkt,\mu
-      \models_\sigma r_i.\matches = t.\keys$, which means that the LHS of the
-      toplevel choice operator should evaluate to $\{\}$ \todo{compute this},
-      and the RHS evaluates to $\{\mu,\pkt\}$\todo{compute this}, which is
-      precisely the result we get when executing
+    \item[\textsc{Miss}] In this case, there is no $r_i$ such that
+      $\satisfy\sigma\pkt\mu r_i.\matches = t.\keys$, which means that the LHS
+      of the toplevel choice operator should evaluate to $\{\}$ \todo{compute
+        this}, and the RHS evaluates to $\{\mu,\pkt\}$\todo{compute this}, which
+      is precisely the result we get when executing
       $\denote{t.apply()}^\sigma~\mu~\pkt$.
@@
   $b$ such that if a state $\sigma : \Value \to \Row^*$ the following
-  judgement is defined and satisfied \[\emptyset, \emptyset \models_\sigma b,\] then \[\forall \pkt \in \Pkt.~
+  judgement is defined and satisfied \[\satisfy\sigma\emptyset\emptyset b,\] then \[\forall \pkt \in \Pkt.~
   \denote{c}^\sigma \pkt \neq \error \]
@@
   for every packet $\pkt \in \Pkt$ such that
-  \[\emptyset, \emptyset \models_\emptyset b_{\mathit{pkt}},\]
+  \[\satisfy \emptyset \pkt \emptyset b_{\mathit{pkt}},\]
   and for every state $\sigma : \Value \to \Row^*$ such that
-  \[\pkt, \emptyset \models_\sigma b_{\mathit{cp}},\]
+  \[\satisfy \sigma \pkt \emptyset b_{\mathit{cp}},\]
   then \[\denote{c}^\sigma ~\emptyset~\pkt \neq \error \]
@@
 
-To do so, we need to consider the following dimensions:
+To do so, we need to extend our templates along the following dimensions:
 \begin{enumerate}[align=left]
@@
 
-\item[\textbf{Nondeterminism} ($c_1 \choice c_2$).]. To solve the \textbf{Conservative
+\item[\textbf{Nondeterminism} ($c_1 \choice c_2$).] To solve the \textbf{Conservative
   Problem} we simply and the templates together, but to solve the

### 2021-03-29 inference.tex
add p4v discussion (sans proof)

@@
 \newcommand{\notsatisfy}[3]{({#1,#3})\not\models_{#2}}
+\newcommand{\pv}{\mathit{p4v}}
 
@@
   (c_1 \choice c_2)[\sigma] && c_1[\sigma] \choice c_2[\sigma] \\
-  (\choose \rho t c)[\sigma] && \textrm{\textcolor{red}{unnecessary, but similar to t.apply}}\\
+  (\choose \rho t c)[\sigma] && \textrm{\textcolor{red}{undefined}}\\
   (t.\apply())[\sigma] &&
@@
   (\choose \rho t c)[?] &&
-  \textrm{\textcolor{red}{unnecessary}} \\
+  \textrm{\textcolor{red}{undefined}} \\
   (t.\apply())[?] &&
@@
 
+\section{P4v comparison}
+
+Ive said before that the P4v-style instrumentation is an overapproximation of
+the control plane behavior. It's a slightly inaccurate claim in that it
+correctly captures all possible control plane behaviors, but it doesn't have a
+handle on constrain control plane behavior.
+
+A simplified version of P4v's encoding is shown in Figure~\ref{fig:p4vlike}. It
+omits the key-reads commands and hit/miss predicates.
+
+\begin{figure}
+  \[\begin{array}{l >{\triangleq}cl}
+  \multicolumn{3}{l}{\pv : \Cmd \to \Cmd} \\
+  \multicolumn{3}{l}{\pv : \Prog \to \Instr} \\
+  \pv(h.f := e) && h.f := e \\
+  \pv(\assume b) && \assume b \\
+  \pv(\assert b) && \assert b \\
+  \pv(c_1;c_2) && \pv(c_1;c_2) \\
+  \pv(c_1 \choice c_2) && \pv(c_1) \choice \pv(c_2) \\
+  \pv(\choose \rho t c) && \textcolor{red}{\text{undefined}} \\
+  \pv(t.\apply()) && \\
+  \multicolumn{3}{>{\qquad}r}{
+    \displaystyle
+    \bigchoice_{i = 0}^{|t.\action|-1}t.\action[i]()
+  }
+  \end{array}\]
+  \caption{A \texttt{p4v}-like instrumentation. For simplicity of reasoning
+    about variables, assume that actions take no action data. }
+  \label{fig:p4vlike}
+\end{figure}
+
+The theorem below demonstrates the over-approximation. Compare with the Instrumentation Equivalence Theorem~\ref{thm:instr-equiv}.
+
+\begin{theorem}[\texttt{p4v} over-approximates]
+  \label{thm:p4v-over}
+  For every $c \in \Prog$, well-formed state $\sigma : \Value \to \Row^* $,
+  model $\mu \in \Model$, and packet $\pkt \in \Pkt$,
+  $\denote{c}^\sigma\,\mu\,\pkt \subseteq \denote{\pv(c)}^\sigma\,\mu\,\pkt$
+\end{theorem}
+
+\begin{proof}
+\todo[inline]{By induction on c}.
+\end{proof}
+
+We can witness non-equality by the following counter-example. Consider a program
+$t.\apply()$ where $t = \{\keys=[h.f];\actions=[\lambda().\SKIP, \lambda(). h.f :=
+  99]\}$. Consider an input packet with $\pkt(h.f) = 47$, and an empty input
+model and state that selects a single action.
+
+Then $\denote{t.\apply()}^\emptyset\,\emptyset\,\pkt = \{(\mu,\pkt)\}$.
+
+Whereas
+\[\begin{array}{l}
+\denote{\pv(t.\apply())}^\emptyset\,\emptyset\,\pkt \\
+= \denote{
+    \displaystyle
+    \bigchoice_{i = 0}^{|t.\action|-1} t.\action[i]()
+}^\emptyset\,\emptyset\,\pkt \\
+= \denote{\SKIP}^\emptyset\,\emptyset\,\pkt \cup \denote{h.f := 99}^\emptyset\,\emptyset\,\pkt\\
+= \{(\emptyset,\pkt),(\emptyset,\pkt\{h.f\mapsto 99\})\}
+\end{array}\]
+
+Finally, $\{(\emptyset,\pkt)\} \subset
+\{(\emptyset,\pkt),(\emptyset,\pkt\{h.f\mapsto 99\})\}$, exemplifies a strict
+subset behavior described by Theorem~\ref{thm:p4v-over}.
+
 \end{document}

### 2021-03-29 inference.tex
choose-> havoc
fix note

@@
 \DeclareMathOperator*{\bigchoice}{\scalerel*{\choiceop}{\sum}}
-\renewcommand{\choose}[2]{\mathop{\mathsf{choose}~#1~\mathsf{from}~#2~\mathsf{in}}}
+\renewcommand{\choose}[2]{\mathop{\mathsf{havoc}~#1~\mathsf{from}~#2~\mathsf{in}}}
 \newcommand{\SKIP}{\mathsf{skip}}
@@
   \]
-  \todo[inline]{Clarify $\error$ states}
-  \todo[inline]{@Nate I know you wanted a standard satisfaction relation, but this was easier to think about in the context of the domain. The difference boils down to a naming convention}
+  \todo[inline]{Clarify $\error$ states} \todo[inline]{A goal here is to have a
+    standard (i.e. single-model) satisfaction relation, but this was easier to
+    think about in the context of the domain (esp. with scoped havoc). The
+    difference boils down to a naming convention}
   \caption{Satisfaction for $\BExpr$.}

### 2021-03-31 inference.tex
outline p4v. think it doesnt work, because we need to know which row each packet hits in each table

@@
 \newcommand{\notsatisfy}[3]{({#1,#3})\not\models_{#2}}
-\newcommand{\pv}{\mathit{p4v}}
+\newcommand{\nondet}{\mathsf{nondet}}
+\newcommand{\ghostly}{\mathsf{ghostly}}
 
@@
 
+\section{Ghost State}
+
+Following p4v, we want construct a ghostly header $\gamma$ from a program $p$
+and use it to instrument the program to capture all dataplane behaviors.
+
+\[
+\begin{array}{l>{\triangleq}cl}
+  \multicolumn{3}{l}{\ghostly :: \Cmd \to \Cmd} \\
+  \multicolumn{3}{l}{\ghostly :: \Prog \to \GCL} \\
+  \ghostly(h.f := e)
+  && h.f := e \\
+  \ghostly(\assert b)
+  && \assert b\\
+  \ghostly(\assume b)
+  && \assume b \\
+  \ghostly(c_1;c_2)
+  && \ghostly(c_1);\ghostly(c_2) \\
+  \ghostly(c_1\choice c_2)
+  && \ghostly(c_1) \choice \ghostly(c_1)\\
+  \ghostly(t.\apply())
+  &&\\
+  \multicolumn{3}{c}{
+    \begin{array}[t]{>{\displaystyle\qquad}l}
+       \gamma.t_{\mathit{reach}} := [1]_1; \\
+       \gamma.t_{\mathit{read}(0)} := t.\keys[0]; \\
+       \vdots \\
+       \gamma.t_{\mathit{read}(|t.\keys|-1)} := t.\keys[|t.\keys|-1]; \\[1em]
+       \left(\gamma.t_{\mathit{hit}} := [0]_1 \right)
+       \choice \left(\bigchoice_{j = 1}^{|t.\action|-1}{
+         \left(\begin{array}{l}
+           \gamma.t_{\mathit{hit}} := [1]_1 ;\\
+           \gamma.t_{\mathit{action}} := [1]_1;\\
+           t.\action[j](\overrightarrow{\gamma.t_{\mathit{data}}})
+         \end{array}\right)
+       }\right)
+     \end{array}}
+\end{array}
+\]
+
+And now what we want to prove is
+
+\begin{theorem}{Instrumentation sufficiency}
+  For every well-formed program $c \in \Cmd$, every state $\sigma : \Value \to
+  \Row^*$, there is some $b_\sigma \in \BExpr$ such that for every model $\mu
+  \in \Model$ and every $\pkt \in \Pkt$,
+  \[\denote{c}^\sigma\,\mu\,\pkt =_{\not\gamma} \denote{\ghostly(c);\assume b}^\sigma\,\mu\,\pkt\]
+\end{theorem}
+
+\begin{proof}
+  Proceed by induction on $c$:
+  \begin{enumerate}[align=left]
+  \item[$(h.f:=e)$] Let $b_\sigma = \TRUE$; immediate.
+  \item[$(\assume b)$] Let $b_\sigma = \TRUE$; immediate.
+  \item[$(\assert b)$] Let $b_\sigma = \TRUE$; immediate.
+  \item[$(c_1;c_2)$] By IHs. $b_\sigma = b_1 \wedge b_2$ because $b_1$ and $b_2$ refer to disjoint variables.
+  \item[$(c_1 \choice c_2)$] \textit{sim}.
+  \item[$(t.\apply())$]
+
+    Let $b_\sigma$ be defined as follows:
+    \[
+    b_\sigma = \bigvee_{r \in \sigma(t.\id)}
+    \begin{array}{>{\displaystyle}l}
+      \gamma
+    \end{array}
+    \]
+
+    We can compute as follows:
+    \[\begin{array}{>{\displaystyle}ll}
+    \denote{\ghostly(c)}^\sigma\,\mu\,\pkt\\
+    =\denote{
+        \begin{array}{>{\displaystyle}l}
+          \gamma.t_{\mathit{reach}} := [1]_1; \\
+          \gamma.t_{\mathit{read}(0)} := t.\keys[0]; \\
+          \vdots \\
+          \gamma.t_{\mathit{read}(|t.\keys|-1)} := t.\keys[|t.\keys|-1]; \\[1em]
+          \left(\gamma.t_{\mathit{hit}} := [0]_1 \right)
+          \choice \left(\bigchoice_{j = 1}^{|t.\action|-1}
+            \left(\begin{array}{l}
+              \gamma.t_{\mathit{hit}} := [1]_1 ;\\
+              \gamma.t_{\mathit{action}} := j;\\
+              t.\action[j](\overrightarrow{\gamma.t_{\mathit{data}}})
+            \end{array}\right)
+          \right)
+    \end{array}}^\sigma\,\mu\,\pkt \\[2em]
+    =\denote{
+        \begin{array}{>{\displaystyle}l}
+          \left(\gamma.t_{\mathit{hit}} := [0]_1 \right)
+          \choice \left(\bigchoice_{j = 1}^{|t.\action|-1}
+            \left(\begin{array}{l}
+              \gamma.t_{\mathit{hit}} := [1]_1 ;\\
+              \gamma.t_{\mathit{action}} := j;\\
+              t.\action[j](\overrightarrow{\gamma.t_{\mathit{data}}})
+            \end{array}\right)
+          \right)
+    \end{array}}^\sigma\!\mu\,\pkt' \\
+    \qquad\qquad \text{where}~\pkt' = \pkt\left\{\begin{array}{l}
+    \gamma.t_{\mathit{reach}} \mapsto [1]_1,\\
+    \overrightarrow{\gamma.t_{\textit{key}} \mapsto \edenote{t.\keys}\,\pkt\,\mu}
+    \end{array}\right\} \\
+    =\denote{
+      \left(\gamma.t_{\mathit{hit}} := [0]_1 \right)}^\sigma\!\mu\,\pkt'
+    \cup
+    \denote{\left(\bigchoice_{j = 1}^{|t.\action|-1}
+            \left(\begin{array}{l}
+              \gamma.t_{\mathit{hit}} := [1]_1 ;\\
+              \gamma.t_{\mathit{action}} := j;\\
+              t.\action[j](\overrightarrow{\gamma.t_{\mathit{data}}})
+            \end{array}\right)
+            \right)
+    }^\sigma\!\mu\,\pkt' \\
+    = \left\{(\mu,\pkt'\left\{\gamma.t_{\mathit{hit}} \mapsto [0]_1\right\})\right\}
+    \cup
+        \denote{\left(\bigchoice_{j = 1}^{|t.\action|-1}
+          \left(\begin{array}{l}
+            \gamma.t_{\mathit{hit}} := [1]_1 ;\\
+            \gamma.t_{\mathit{action}} := j;\\
+            t.\action[j](\overrightarrow{\gamma.t_{\mathit{data}}})
+          \end{array}\right)
+          \right)
+        }^\sigma\!\mu\,\pkt' \\
+  = \left\{(\mu,\pkt'\left\{\gamma.t_{\mathit{hit}} \mapsto [0]_1\right\})\right\}
+  \cup
+  \left(\bigcup_{j = 1}^{|t.\action|-1}
+  \denote{t.\action[j](\overrightarrow{\gamma.t_{\mathit{data}}})}^\sigma\,\mu\,\pkt'\left\{
+  \begin{array}{l}
+    \gamma.t_{\textit{hit}} \mapsto [1]_1,\\
+    \gamma.t_{\textit{action}} \mapsto j
+  \end{array}
+  \right\}
+  \right) \\
+
+    \end{array}
+\]
+
+  \end{enumerate}
+\end{proof}
+
+\section{Tables as nondeterministic choice}
+
+Ive said before that the nondeterministic choice is an overapproximation of
+the control plane behavior. It's a slightly inaccurate claim in that it
+correctly captures all possible control plane behaviors, but it doesn't have a
+handle on constrain control plane behavior.
+
+A simplified version of P4v's encoding is shown in Figure~\ref{fig:p4vlike}. It
+omits the key-reads commands and hit/miss predicates.
+
+\begin{figure}
+  \[\begin{array}{l >{\triangleq}cl}
+  \multicolumn{3}{l}{\nondet : \Cmd \to \Cmd} \\
+  \multicolumn{3}{l}{\nondet : \Prog \to \Instr} \\
+  \nondet(h.f := e) && h.f := e \\
+  \nondet(\assume b) && \assume b \\
+  \nondet(\assert b) && \assert b \\
+  \nondet(c_1;c_2) && \nondet(c_1;c_2) \\
+  \nondet(c_1 \choice c_2) && \nondet(c_1) \choice \nondet(c_2) \\
+  \nondet(\choose \rho t c) && \textcolor{red}{\text{undefined}} \\
+  \nondet(t.\apply()) && \\
+  \multicolumn{3}{>{\qquad}r}{
+    \displaystyle
+    \bigchoice_{i = 0}^{|t.\action|-1}t.\action[i]()
+  }
+  \end{array}\]
+  \caption{A \texttt{p4v}-like instrumentation. For simplicity of reasoning
+    about variables, assume that actions take no action data. }
+  \label{fig:p4vlike}
+\end{figure}
+
+The theorem below demonstrates the over-approximation. Compare with the Instrumentation Equivalence Theorem~\ref{thm:instr-equiv}.
+
+\begin{theorem}[\texttt{p4v} over-approximates]
+  \label{thm:p4v-over}
+  For every $c \in \Prog$, well-formed state $\sigma : \Value \to \Row^* $,
+  model $\mu \in \Model$, and packet $\pkt \in \Pkt$,
+  $\denote{c}^\sigma\,\mu\,\pkt \subseteq \denote{\nondet(c)}^\sigma\,\mu\,\pkt$
+\end{theorem}
+
+\begin{proof}
+\todo[inline]{By induction on c}.
+\end{proof}
+
+We can witness non-equality by the following counter-example. Consider a program
+$t.\apply()$ where $t = \{\keys=[h.f];\actions=[\lambda().\SKIP, \lambda(). h.f :=
+  99]\}$. Consider an input packet with $\pkt(h.f) = 47$, and an empty input
+model and state that selects a single action.
+
+Then $\denote{t.\apply()}^\emptyset\,\emptyset\,\pkt = \{(\mu,\pkt)\}$.
+
+Whereas
+\[\begin{array}{l}
+\denote{\nondet(t.\apply())}^\emptyset\,\emptyset\,\pkt \\
+= \denote{
+    \displaystyle
+    \bigchoice_{i = 0}^{|t.\action|-1} t.\action[i]()
+}^\emptyset\,\emptyset\,\pkt \\
+= \denote{\SKIP}^\emptyset\,\emptyset\,\pkt \cup \denote{h.f := 99}^\emptyset\,\emptyset\,\pkt\\
+= \{(\emptyset,\pkt),(\emptyset,\pkt\{h.f\mapsto 99\})\}
+\end{array}\]
+
+Finally, $\{(\emptyset,\pkt)\} \subset
+\{(\emptyset,\pkt),(\emptyset,\pkt\{h.f\mapsto 99\})\}$, exemplifies a strict
+subset behavior described by Theorem~\ref{thm:p4v-over}.
+
 \section{The Research Problem}
@@
 
-\section{P4v comparison}
-
-Ive said before that the P4v-style instrumentation is an overapproximation of
-the control plane behavior. It's a slightly inaccurate claim in that it
-correctly captures all possible control plane behaviors, but it doesn't have a
-handle on constrain control plane behavior.
-
-A simplified version of P4v's encoding is shown in Figure~\ref{fig:p4vlike}. It
-omits the key-reads commands and hit/miss predicates.
-
-\begin{figure}
-  \[\begin{array}{l >{\triangleq}cl}
-  \multicolumn{3}{l}{\pv : \Cmd \to \Cmd} \\
-  \multicolumn{3}{l}{\pv : \Prog \to \Instr} \\
-  \pv(h.f := e) && h.f := e \\
-  \pv(\assume b) && \assume b \\
-  \pv(\assert b) && \assert b \\
-  \pv(c_1;c_2) && \pv(c_1;c_2) \\
-  \pv(c_1 \choice c_2) && \pv(c_1) \choice \pv(c_2) \\
-  \pv(\choose \rho t c) && \textcolor{red}{\text{undefined}} \\
-  \pv(t.\apply()) && \\
-  \multicolumn{3}{>{\qquad}r}{
-    \displaystyle
-    \bigchoice_{i = 0}^{|t.\action|-1}t.\action[i]()
-  }
-  \end{array}\]
-  \caption{A \texttt{p4v}-like instrumentation. For simplicity of reasoning
-    about variables, assume that actions take no action data. }
-  \label{fig:p4vlike}
-\end{figure}
-
-The theorem below demonstrates the over-approximation. Compare with the Instrumentation Equivalence Theorem~\ref{thm:instr-equiv}.
-
-\begin{theorem}[\texttt{p4v} over-approximates]
-  \label{thm:p4v-over}
-  For every $c \in \Prog$, well-formed state $\sigma : \Value \to \Row^* $,
-  model $\mu \in \Model$, and packet $\pkt \in \Pkt$,
-  $\denote{c}^\sigma\,\mu\,\pkt \subseteq \denote{\pv(c)}^\sigma\,\mu\,\pkt$
-\end{theorem}
-
-\begin{proof}
-\todo[inline]{By induction on c}.
-\end{proof}
-
-We can witness non-equality by the following counter-example. Consider a program
-$t.\apply()$ where $t = \{\keys=[h.f];\actions=[\lambda().\SKIP, \lambda(). h.f :=
-  99]\}$. Consider an input packet with $\pkt(h.f) = 47$, and an empty input
-model and state that selects a single action.
-
-Then $\denote{t.\apply()}^\emptyset\,\emptyset\,\pkt = \{(\mu,\pkt)\}$.
-
-Whereas
-\[\begin{array}{l}
-\denote{\pv(t.\apply())}^\emptyset\,\emptyset\,\pkt \\
-= \denote{
-    \displaystyle
-    \bigchoice_{i = 0}^{|t.\action|-1} t.\action[i]()
-}^\emptyset\,\emptyset\,\pkt \\
-= \denote{\SKIP}^\emptyset\,\emptyset\,\pkt \cup \denote{h.f := 99}^\emptyset\,\emptyset\,\pkt\\
-= \{(\emptyset,\pkt),(\emptyset,\pkt\{h.f\mapsto 99\})\}
-\end{array}\]
-
-Finally, $\{(\emptyset,\pkt)\} \subset
-\{(\emptyset,\pkt),(\emptyset,\pkt\{h.f\mapsto 99\})\}$, exemplifies a strict
-subset behavior described by Theorem~\ref{thm:p4v-over}.
-
 \end{document}

### 2021-03-31 inference.tex
add symbolic table encoding with havoc

@@
 \newcommand{\data}{\mathsf{data}}
+\newcommand{\params}{\mathsf{params}}
 \newcommand{\id}{\mathsf{id}}
@@
 \DeclareMathOperator*{\bigchoice}{\scalerel*{\choiceop}{\sum}}
-\renewcommand{\choose}[2]{\mathop{\mathsf{havoc}~#1~\mathsf{from}~#2~\mathsf{in}}}
+\newcommand{\havoc}{\texttt{<?>}}
 \newcommand{\SKIP}{\mathsf{skip}}
@@
 \newcommand{\ghostly}{\mathsf{ghostly}}
+\newcommand{\state}[1]{\llparenthesis#1\rrparenthesis}
 
@@
       & \mid & h.f & \textit{Header Access} & (a,g,p,\iota)\\
+      & \mid & \havoc & \textit{Havoc} & (a,g,p,\iota) \\
       & \mid & \rho.\matches[i] & \textit{Symbolic Match} & (\iota)\\
@@
     & \mid & c \choice c & \textit{Nondeterministic Choice} & (g,p,\iota)\\
-    & \mid & \choose \rho t c & \textit{Selection} & (\iota) \\
     \multicolumn{3}{l}{
@@
   (c_1 \choice c_2)[?] && c_1[?] \choice c_2[?]\\
-  (\choose \rho t c)[?] &&
-  \textrm{\textcolor{red}{undefined}} \\
   (t.\apply())[?] &&
@@
     \{ \\
-    \quad \choose \rho t \;\{ \\
-    \qquad \assume \rho.\matches = t.\keys; \\
-    \qquad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
+    \quad \rho.\matches := \havoc; \\
+    \quad \rho.\action := \havoc;\\
+    \quad \rho.\data := \havoc; \\
+    \quad \assume |\rho.\matches| = |t.\keys|; \\
+    \quad \assume \rho.\action < |t.\actions|; \\
+    \quad \assume |\rho.\data| < |t.\action[\rho.\action].\params| \\
+    \quad \assume \rho.\matches = t.\keys; \\
+    \quad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
     \left\{\begin{array}{l}
@@
     \end{array}\right\}\\
-    \quad \} \\
     \} \choice \{ \\
@@
     \quad \SKIP \\
-    \quad \}\\
     \}
@@
   \caption{Symbolic Table instrumentation}
+  \label{fig:table-instrument}
+\end{figure}
+
+\begin{figure}[htp]
+  \[\begin{array}{r >{\triangleq}cl}
+  \multicolumn{3}{l}{\state\sigma_c : (\Value \to \Row^*) \to \BExpr} \\
+  \state\sigma_c = \bigwedge_{ t \in \tables(c) } \bigvee_{r_i \in \sigma(t.\id)}
+  \rho = r_i
+  \end{array}
+  \]
+  \caption{Symbolic State Encoding}
   \label{fig:table-instrument}

### 2021-03-31 inference.tex
polish

@@
       & \mid & c \choice c & \textit{Nondeterministic Choice} & (g,p,\iota)\\
+      & \mid & \{c\} & \textit{Scope} & (g,p,\iota)\\
     \multicolumn{3}{l}{
@@
 assumptions ($\assume b$), assertions ($\assert b$), table application
-($t.\apply()$), lexically-scoped nondeterministic selection ($\choose \rho t
-c$), sequential composition ($c_1;c_2$ for two commands $c_1$ and $c_2$), and
+($t.\apply()$), sequential composition ($c_1;c_2$ for two commands $c_1$ and $c_2$), and
 nondeterministic choice ($c_1 \choice c_2$ for two commands $c_1$ and $c_2$).
@@
 with symbolic variables. A program $\iota$ cannot apply tables using
-$t.\apply()$, but instead uses a havoc statement $\choose \rho t c$, to
+$t.\apply()$, but instead uses havoc statements, to
 introduce symbolic variables $\rho$ in $c$.
@@
   \[\begin{array}{l>{\triangleq}cl}
-  \multicolumn{3}{l}{\edenote{e} : \Pkt \to \Model \rightharpoonup \Value }\\
-    \edenote{v}~\pkt~\mu &&
+  \multicolumn{3}{l}{\edenote{e} : \Model \to \Pkt \rightharpoonup \Value }\\
+    \edenote{v}~\mu~\pkt &&
     v \\
-    \edenote{h.f}~\pkt~\mu &&
+    \edenote{\havoc_w}~\mu~\pkt && \{[n]_w \mid n \in \mathbb N\}\\
+    \edenote{h.f}~\mu~\pkt &&
     pkt(h.f)\\
-    \edenote{\rho.\matches[i]}~\pkt~\mu &&
+    \edenote{\rho.\matches[i]}~\mu~\pkt &&
     \mu(\rho).\matches[i] \\
-    \edenote{\rho.\action}~\pkt~\mu &&
+    \edenote{\rho.\action}~\mu~\pkt &&
     \mu(\rho).\action \\
@@
     \mu(\rho).\data[i]\\
-    \edenote{e_1 \binop e_2}~\pkt~\mu&&
-    \edenote{e_1}~\pkt~\mu \binop
-    \edenote{e_2}~\pkt~\mu \\
-    \edenote{\unop e}~\pkt~\mu &&
-    \unop \left(\edenote{e}~\pkt~\mu\right)
+    \edenote{e_1 \binop e_2}~\mu~\pkt&&
+    \edenote{e_1}~\mu~\pkt \binop
+    \edenote{e_2}~\mu~\pkt \\
+    \edenote{\unop e}~\mu~\pkt &&
+    \unop \left(\edenote{e}~\mu~\pkt\right)
   \end{array}
@@
   \multicolumn{3}{l}{\denote{c}^\sigma : \Model \to \Pkt \to \mathcal{P}(\Model \times \Pkt)_\error }\\
-    \denote{h.f := e}^\sigma~\mu~\pkt && \{(\mu, \pkt\{h \mapsto \edenote{e}~\pkt~\mu\})\} \\
-    \denote{\assume b}^\sigma~\mu~\pkt &&
+    \denote{h.f := e}^\sigma\,\mu\,\pkt && \{(\mu, \pkt\{h \mapsto v\} \mid v \in \edenote{e}\,\mu\,\pkt\})\} \\
+    \denote{\assume b}^\sigma\,\mu\,\pkt &&
     \begin{cases}
@@
     \end{cases}\\
-    \denote{\assert b}^\sigma~\mu~\pkt &&
+    \denote{\assert b}^\sigma\,\mu\,\pkt &&
     \begin{cases}
@@
     \end{cases} \\
-    \denote{\choose \rho t c}^\sigma~\mu~\pkt &&
-    \{(\mu'\setminus \rho,\pkt')
-    \mid (\mu',\pkt') \in \denote{c}^\sigma\,(\mu\{\rho\mapsto r\})\,\pkt, r \in \sigma(t)\}
-    \\
-    \denote{c_1;c_2}^\sigma~\mu~\pkt&&
-    \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2}^\sigma~\mu'~\pkt' \\
-    \denote{c_1 \choice c_2}^\sigma~\mu~\pkt &&
-    \denote{c_1}^\sigma~\mu~\pkt \cup \denote{c_2}^\sigma~\mu~\pkt \\
-    \denote{t.\apply()}^\sigma~\mu~\pkt &&
+    \denote{c_1;c_2}^\sigma\,\mu\,\pkt&&
+    \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2}^\sigma\,\mu'\,\pkt' \\
+    \denote{c_1 \choice c_2}^\sigma\,\mu\,\pkt &&
+    \denote{c_1}^\sigma\,\mu\,\pkt \cup \denote{c_2}^\sigma\,\mu\,\pkt \\
+    \denote{\{c\}}^\sigma\,\mu\,\pkt &&
+    \{(\mu,\pkt') \mid (\mu',\pkt') \in \denote{c}^\sigma\,\mu\,\pkt \}\\
+    \denote{t.\apply()}^\sigma\,\mu\,\pkt &&
     \begin{cases}
-      \denote{c[\vec d/\vec x]}^\sigma~\mu~\pkt,
+      \denote{c[\vec d/\vec x]}^\sigma\,\mu\,\pkt,
       & \text{where}~\lambda \vec x.~c = r.\action, \vec d = r.\data \\
@@
   \label{fig:com-sem}
-  \todo[inline]{$\Model$ then $\Pkt$ everywhere.}
 \end{figure}
@@
 
-\section{Eliminating Tables}
+\subsection{Eliminating Tables}
 
@@
   (c_1 \choice c_2)[\sigma] && c_1[\sigma] \choice c_2[\sigma] \\
-  (\choose \rho t c)[\sigma] && \textrm{\textcolor{red}{undefined}}\\
+  (\{c\})[\sigma] && \{c[\sigma]\} \\
   (t.\apply())[\sigma] &&
   \begin{array}[t]{l}
-    \{\\
+    (\\
     \quad \assume {r_1.\matches = t.\keys};\\
     \quad t.\action[r_1.\action](r_1.\data)\\
-    \}~\choice\\
+    )~\choice\\
     \multicolumn{1}{c}{\vdots}\\
-    \choice~\{\\
+    \choice~(\\
     \quad \assume {r_n.\matches = t.\keys};\\
     \quad t.\action[r_n.\action](r_n.\data)\\
-    \}\\
+    )\\
     \textrm{where}~[r_1;\ldots;r_n] = \sigma(t.\id)
@@
 
-\section{Instrumenting Tables}
+\subsection{Instrumenting Tables}
 
@@
 
+For simplicity, We're going to assume that all actions in a given table have the
+same number of action data fields, and each field is the same number of bits.
+This can be realized manually.
+
 \begin{figure}[htp]
@@
   \multicolumn{3}{l}{(\cdot)[?] : \Cmd \to \Cmd} \\
-  \multicolumn{3}{l}{(\cdot)[?] : \Prog \to \Instr} \\
   (h.f := e)[?] && h.f := e \\
@@
   (c_1 \choice c_2)[?] && c_1[?] \choice c_2[?]\\
+  (\{c\})[?] && \{c[?]\}\\
   (t.\apply())[?] &&
@@
     \{ \\
-    \quad \rho.\matches := \havoc; \\
+    \quad \rho.\matches := \overrightarrow\havoc; \\
     \quad \rho.\action := \havoc;\\
-    \quad \rho.\data := \havoc; \\
-    \quad \assume |\rho.\matches| = |t.\keys|; \\
+    \quad \rho.\data := \overrightarrow\havoc; \\
     \quad \assume \rho.\action < |t.\actions|; \\
-    \quad \assume |\rho.\data| < |t.\action[\rho.\action].\params| \\
     \quad \assume \rho.\matches = t.\keys; \\
     \quad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
-    \left\{\begin{array}{l}
+    \left(\begin{array}{l}
     \assume \rho.\action = i;\\
     t.\action[i](\rho.\data)
-    \end{array}\right\}\\
-    \} \choice \{ \\
+    \end{array}\right)\\
+    \} \choice ( \\
     \quad \assume \forall \rho \in t. \rho.\matches \neq t.\keys; \\
     \quad \SKIP \\
-    \}
+    )
   \end{array}
@@
   \]
-  \caption{Symbolic Table instrumentation}
+  \caption{Symbolic Table instrumentation. Typing information is omitted for
+    brevity; each $\havoc$ must have a appropriate bit-width annotation. For
+    simplicity, we assume each action in a table has the same number (and types) of
+    parameters.}
   \label{fig:table-instrument}
@@
 
-\begin{lemma}[Models Invariant over Actions]
-  \label{lem:models-invariant-over-actions}
-  For a closed action command $c$, $\mu,\mu'
-  \in \Model$, $\sigma \in \Value \to
-  \Row^*$, $\pkt \in \Pkt$, there exists some $\pkt'$
-  \[\denote{c}^\sigma\,\mu\,\pkt = \{\mu,\pkt'\}\]
-  \end{lemma}
+%% \begin{lemma}[Models Invariant over Actions]
+%%   \label{lem:models-invariant-over-actions}
+%%   For a closed action command $c$, $\mu,\mu'
+%%   \in \Model$, $\sigma \in \Value \to
+%%   \Row^*$, $\pkt \in \Pkt$, there exists some $\pkt'$
+%%   \[\denote{c}^\sigma\,\mu\,\pkt = \{\mu,\pkt'\}\]
+%%   \end{lemma}
 
-\begin{proof}
-  Proceed by induction over $c$.
-  \begin{enumerate}[align=left]
-  \item[$(c = h.f = e)$] Immediate.
-  \item[$(c = c_1;c_2)$]
-    By IH($c_1$) there exists some $\pkt''$ such that
-    $\denote{c_1}^\sigma\,\mu\,\pkt = \{(\mu,\pkt'')\}$.
-    By IH($c_2$) there exists some $\pkt'$ such that
-    $\denote{c_2}^\sigma,\mu\,pkt'' = \{(\mu, \pkt')\}$.
-    The result follows by definition of $\denote{-}$.
-    \end{enumerate}
-\end{proof}
+%% \begin{proof}
+%%   Proceed by induction over $c$.
+%%   \begin{enumerate}[align=left]
+%%   \item[$(c = h.f = e)$] Immediate.
+%%   \item[$(c = c_1;c_2)$]
+%%     By IH($c_1$) there exists some $\pkt''$ such that
+%%     $\denote{c_1}^\sigma\,\mu\,\pkt = \{(\mu,\pkt'')\}$.
+%%     By IH($c_2$) there exists some $\pkt'$ such that
+%%     $\denote{c_2}^\sigma,\mu\,pkt'' = \{(\mu, \pkt')\}$.
+%%     The result follows by definition of $\denote{-}$.
+%%     \end{enumerate}
+%% \end{proof}
 
+\begin{samepage}
 \begin{theorem}[Instrumentation Equivalence]
@@
   For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, model
-  $\mu \in \Model$, and packet $\pkt \in \Pkt$, $\denote{c}^\sigma~\mu~\pkt =
-  \denote{c[?]}^\sigma~\mu~\pkt$.
+  $\mu \in \Model$, and packet $\pkt \in \Pkt$, \[\denote{c}^\sigma~\mu~\pkt = \denote{c[?];\assume \state\sigma_c}^\sigma~\mu~\pkt.\]
 \end{theorem}
+\end{samepage}
 
 \begin{proof}
-  Proceed by induction on the structure of $c$, leaving $\sigma$, $\mu$, $\pkt$ general.
-  \begin{enumerate}[align=left]
-  \item[$(c = h.f := e)$]
-    Let $\sigma$,$\mu$,$\pkt$ be given.
-    By definition $h.f:=e[?] = h.f:=e$. The result follows.
-  \item[$(c = \assume b)$] Similar. $[?]$ is no-op.
-  \item[$(c = \assert b)$] Similar. $[?]$ is no-op.
-  \item[$(c = \choose \rho t)$] \textcolor{red}{Admitted}.
-  \item[$(c = c_1;c_2)$]
-    Let $\sigma,\mu,\pkt$ be given.
-    By definition $(c_1;c_2)[?] = c_1[?]; c_2[?]$.
-    Compute as follows:
-    \[\begin{array}{>{\displaystyle}ll}
-    \denote{(c_1;c_2)[?]}^\sigma~\mu~\pkt \\
-    = \denote{c_1[?];c_2[?]}^\sigma~\mu\pkt
-    & \text{by definition of } [?]  \\
-    = \bigcup_{(\mu',\pkt') \in \denote{c_1[?]}^\sigma\,\mu\,\pkt}\denote{c_2[?]}^\sigma~\mu'~\pkt'
-    & \text{by definition of }\denote{-} \\
-    = \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2[?]}^\sigma~\mu'~\pkt'
-    & \text{by } \mathit{IH}(c_1, \mu, \pkt) \\
-    = \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2}^\sigma~\mu'~\pkt'
-    & \text{by } \mathit{IH}(c_2, \mu', \pkt') \\
-    = \denote{c_1;c_2}^\sigma~\mu~\pkt
-    & \text{by definition of } \denote{-}
-    \end{array}
-    \]
-  \item[$(c = c_1 \choice c_2)$] Similar, by IHs.
+  \todo[inline]{Proceed by induction on the structure of $c$, leaving $\sigma$, $\mu$, $\pkt$ general}
 
-  \item[$(c = t.\apply())$]
-    Let $\sigma$,$\mu$,$\pkt$ be given.
-    Let $r_1,\ldots,r_n = \sigma(t)$.
-    There are two cases, either the packet hits a row, or hits no row. Explore each case separately.
-    \begin{enumerate}
-    \item[\textsc{Hit}] In this case, there is some $r_i$ such that $\satisfy \sigma \pkt \mu r_i.\matches = t.\keys$. By the well-formedness assumption,
-      we know that this $r_i$ is unique, and that we can write
-      $t.\actions[r_i.\action]$ as $\lambda \vec x. c$, and $|r_i.\data| = |\vec x|$.
+  %% \begin{enumerate}[align=left]
+  %% \item[$(c = h.f := e)$]
+  %%   Let $\sigma$,$\mu$,$\pkt$ be given.
+  %%   By definition $h.f:=e[?] = h.f:=e$. The result follows.
+  %% \item[$(c = \assume b)$] Similar. $[?]$ is no-op.
+  %% \item[$(c = \assert b)$] Similar. $[?]$ is no-op.
+  %% \item[$(c = c_1;c_2)$]
+  %%   Let $\sigma,\mu,\pkt$ be given.
+  %%   By definition $(c_1;c_2)[?] = c_1[?]; c_2[?]$.
+  %%   Compute as follows:
+  %%   \[\begin{array}{>{\displaystyle}ll}
+  %%   \denote{(c_1;c_2)[?]}^\sigma~\mu~\pkt \\
+  %%   = \denote{c_1[?];c_2[?]}^\sigma~\mu\pkt
+  %%   & \text{by definition of } [?]  \\
+  %%   = \bigcup_{(\mu',\pkt') \in \denote{c_1[?]}^\sigma\,\mu\,\pkt}\denote{c_2[?]}^\sigma~\mu'~\pkt'
+  %%   & \text{by definition of }\denote{-} \\
+  %%   = \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2[?]}^\sigma~\mu'~\pkt'
+  %%   & \text{by } \mathit{IH}(c_1, \mu, \pkt) \\
+  %%   = \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2}^\sigma~\mu'~\pkt'
+  %%   & \text{by } \mathit{IH}(c_2, \mu', \pkt') \\
+  %%   = \denote{c_1;c_2}^\sigma~\mu~\pkt
+  %%   & \text{by definition of } \denote{-}
+  %%   \end{array}
+  %%   \]
+  %% \item[$(c = c_1 \choice c_2)$] Similar, by IHs.
 
-      Compute as follows:
-      \[\begin{array}{>{\displaystyle}ll}
-      \denote{t.\apply()}^\sigma~\mu~\pkt \\
-      = \denote{c[r_i.data/\vec x]}^\sigma~\mu~\pkt & \text{by definition of} \models \text{, and the uniqueness of }r_i  \\
-      \end{array}
-      \]
+  %% \item[$(c = t.\apply())$]
+  %%   Let $\sigma$,$\mu$,$\pkt$ be given.
+  %%   Let $r_1,\ldots,r_n = \sigma(t)$.
+  %%   There are two cases, either the packet hits a row, or hits no row. Explore each case separately.
+  %%   \begin{enumerate}
+  %%   \item[\textsc{Hit}]
+  %%     \textitP{}
 
-      Now, we compute the denotation of $t.\apply()[?]$ to see that it is equivalent to the above.
-      The definition of $t.\apply()[?]$ gives us two alternatives. Consider the second, first:
-      \[\begin{array}{>{\displaystyle}ll}
-        \denote{\assume \forall \rho \in t. \rho.\matches \neq t.\keys;\SKIP}^\sigma~\mu~\pkt \\
-        = \denote{\assume \forall \rho \in t. \rho.\matches \neq t.\keys}^\sigma~\mu~\pkt
-        & \textrm{\textcolor{red}{Redo with defaults}} \\
-        = \begin{cases}
-          \{(\mu,\pkt)\}, & \satisfy\sigma\pkt\mu \rho.\matches \neq t.\keys \\
-          \{\}, &\textit{otherwise}
-        \end{cases}
-        & \textrm{By definition of }\denote{-}\\
-        =  \{\}
-        & \textrm{by \textsc{Hit} assumption}
-      \end{array}
-      \]
+  %%   \item[\textsc{Miss}] In this case, there is no $r_i$ such that
+  %%     $\satisfy\sigma\pkt\mu r_i.\matches = t.\keys$, which means that the LHS
+  %%     of the toplevel choice operator should evaluate to $\{\}$ \todo{compute
+  %%       this}, and the RHS evaluates to $\{\mu,\pkt\}$\todo{compute this}, which
+  %%     is precisely the result we get when executing
+  %%     $\denote{t.apply()}^\sigma~\mu~\pkt$.
 
-      Now let's consider the first branch, i.e.
-      \[
-      \denote{\choose \rho t \left\{
-        \begin{array}{>{\displaystyle}l}
-          \assume \rho.\matches = t.\keys;\\
-          \bigchoice_{i=0}^{|t.\action| -1}
-          \left\{\begin{array}{l}
-          \assume \rho.\action = i;\\
-          t.\actions[i](\rho.\data)
-          \end{array}\right\}
-        \end{array}\right\}
-      }^\sigma~\mu~\pkt\]
+  %%   \end{enumerate}
 
-      Compute the following inner result, by definitions and set theory
-      \[
-      \begin{array}{>{\displaystyle}ll}
-      \bigcup_{r \in \sigma(t)}\denote{\begin{array}{>{\displaystyle}l}
-          \assume \rho.\matches = t.\keys;\\
-          \bigchoice_{i=0}^{|t.\action| -1}
-          \left\{\begin{array}{l}
-          \assume \rho.\action = i;\\
-          t.\actions[i](\rho.\data)
-          \end{array}\right\}
-      \end{array}}^\sigma~\mu\{\rho \mapsto r\}~\pkt\\
-      = \bigcup_{
-        \substack{r \in \sigma(t) \\
-          (\mu',\pkt') \in \denote{\assume \rho.\matches = t.\keys}^\sigma\,(\mu\{\rho \mapsto r\})\,\pkt}}
-      \bigcup_{i=0}^{|t.\action| -1}\denote{
-       \begin{array}{l}
-          \assume \rho.\action = i;\\
-          t.\actions[i](\rho.\data)
-      \end{array}}^\sigma~\mu'~\pkt'
-
-      \\\\
-      (\textrm{by \textsc{Hit}, \textsc{NoShadow}, and the definition of }\denote{-}) \\
-      \\
-      = \bigcup_{i=0}^{|t.\action| -1}\denote{
-       \begin{array}{l}
-          \assume \rho.\action = i;\\
-          t.\actions[i](\rho.\data)
-      \end{array}}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt \\
-      = \bigcup_{j=0}^{|t.\action| -1}\denote{
-       \begin{array}{l}
-          \assume r_i.\action = j;\\
-          t.\actions[j](r_i.\data)
-      \end{array}}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt \\
-      \textrm{\textcolor{red}{TODO LEMMA}} \\
-      = \denote{t.\actions[r_i.\action](r_i.\data)}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt \\
-      = \denote{c[r_i.\data/\vec x]}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt \\
-      \end{array}\]
-
-      By Lemma~\ref{lem:models-invariant-over-actions}, there's some $\pkt'$ such that
-      \[\{(\mu\{\rho \mapsto r_i\}, \pkt')\} =
-      \denote{c[r_i.\data/\vec x]}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt
-      \]
-
-      which means that the result of the first branch is
-      \[\{(\mu, \pkt')\} \]
-
-      Since $\rho$ doesn't occur in $c$, then there's some $\mu'$ such that also
-      \[\{\mu', \pkt'\} = \denote{c[r_i.\data/\vec x]}^\sigma~\mu~\pkt \]
-
-      and $\mu = \mu'$ by Lemma~\ref{lem:models-invariant-over-actions}.
-
-      The result follows.
-
-    \item[\textsc{Miss}] In this case, there is no $r_i$ such that
-      $\satisfy\sigma\pkt\mu r_i.\matches = t.\keys$, which means that the LHS
-      of the toplevel choice operator should evaluate to $\{\}$ \todo{compute
-        this}, and the RHS evaluates to $\{\mu,\pkt\}$\todo{compute this}, which
-      is precisely the result we get when executing
-      $\denote{t.apply()}^\sigma~\mu~\pkt$.
-
-    \end{enumerate}
-
-  \end{enumerate}
+  %% \end{enumerate}
 \end{proof}
 
-\section{Ghost State}
+\subsection{Ghost State}
 
@@
 
-\section{Tables as nondeterministic choice}
+\subsection{Tables as nondeterministic choice}
 
@@
   \nondet(c_1 \choice c_2) && \nondet(c_1) \choice \nondet(c_2) \\
-  \nondet(\choose \rho t c) && \textcolor{red}{\text{undefined}} \\
+  %% \nondet(\choose \rho t c) && \textcolor{red}{\text{undefined}} \\
   \nondet(t.\apply()) && \\

### 2021-04-02 inference.tex
add wp
polish example

@@
 
+\usepackage[margin=1in]{geometry}
+\usepackage{halloweenmath}
 \usepackage{amsmath, amssymb, amsthm, array, stmaryrd}
@@
 \newcommand{\id}{\mathsf{id}}
+\newcommand{\reach}{\mathit{\mathit{reach}}}
+\newcommand{\hit}{\mathit{hit}}
+\newcommand{\reads}{\mathit{reads}}
 
@@
 \DeclareMathOperator*{\bigchoice}{\scalerel*{\choiceop}{\sum}}
-\newcommand{\havoc}{\texttt{<?>}}
+\newcommand{\havoc}[1]{\mathop{\textsf{havoc}}#1}
 \newcommand{\SKIP}{\mathsf{skip}}
@@
 \newcommand{\state}[1]{\llparenthesis#1\rrparenthesis}
+\newcommand{\instr}{[\forall]}
+\newcommand{\ginstr}{[\mathghost]}
 
@@
       & \mid & h.f & \textit{Header Access} & (a,g,p,\iota)\\
-      & \mid & \havoc & \textit{Havoc} & (a,g,p,\iota) \\
       & \mid & \rho.\matches[i] & \textit{Symbolic Match} & (\iota)\\
@@
     & \mid & e = e & \textit{Bitvector Equality} & (a,g,p,\iota)\\
-    & \mid & \forall \rho \in t.~b & \textit{Row Universality} & (\iota)\\
+    & \mid & \rho \in t & \textit{Table membership} & (\iota)\\
+    & \mid & \forall \rho.~b & \textit{Row Universality} & (\iota)\\
     \multicolumn{4}{l}{c \in \Cmd, a \in \Action, g \in \GCL, p \in \Prog, \iota \in \Instr } \\
     c & ::=  & h.f := e & \textit{Assignment} & (g,p,\iota) \\
+      & \mid & \havoc \rho & \textit{Row Havoc} & (\iota) \\
       & \mid & \assert b & \textit{Assertion} & (g,p,\iota)\\
@@
     v \\
-    \edenote{\havoc_w}~\mu~\pkt && \{[n]_w \mid n \in \mathbb N\}\\
     \edenote{h.f}~\mu~\pkt &&
@@
   \text{(note strict semantics)} \\
-  \satisfy \sigma \pkt \mu \forall \rho \in t. ~b &&
-  \forall r \in \sigma(t.\id).
+  \satisfy\sigma\pkt\mu \rho \in t. && \mu(\rho) \in \sigma(t.\id) \\
+  \satisfy \sigma \pkt \mu \forall \rho. ~b &&
+  \forall r.
   \satisfy \sigma \pkt {\mu\{\rho \mapsto r\}} b
@@
   \]
-  \todo[inline]{Clarify $\error$ states} \todo[inline]{A goal here is to have a
-    standard (i.e. single-model) satisfaction relation, but this was easier to
-    think about in the context of the domain (esp. with scoped havoc). The
-    difference boils down to a naming convention}
+  \todo[inline]{Clarify $\error$ states}
   \caption{Satisfaction for $\BExpr$.}
@@
     \end{cases} \\
+    \denote{\havoc \rho}^\sigma\,\mu\,\pkt &&
+    \{(\mu\{\rho \mapsto r\}, \pkt) \mid r \in \Row\} \\
     \denote{c_1;c_2}^\sigma\,\mu\,\pkt&&
@@
     )~\choice\\
-    \multicolumn{1}{c}{\vdots}\\
-    \choice~(\\
+    \phantom)~\,\vdots\\
+    \phantom)\,\choice~(\\
     \quad \assume {r_n.\matches = t.\keys};\\
     \quad t.\action[r_n.\action](r_n.\data)\\
+    )\choice (\\
+    \quad \displaystyle \assume {\bigwedge_{i=1}^n r_n.\matches \neq t.\keys};\\
+    \quad \SKIP \\
     )\\
@@
   \]
-  \caption{Table Elimination}
+  \caption{Table Substitution}
   \label{fig:table-elim}
@@
 
-\begin{theorem}[Table Elimination]
+\begin{theorem}[Table Substitution]
   For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, model
@@
 
+We augment the headers for a program $c$ with ghost headers $\gamma_t$ for each
+table $t \in \tables(c)$. The header $\gamma_t$ has defined fields
+$\reach$, which is set to true if the packet reaches table $t$,
+$\hit$ which is set to true the packet hits in the table, and
+$\reads$ which is really a list of $|t.\keys|$ variables that stores the
+values of the packet when evaluating $t.\keys$.
+
 \begin{figure}[htp]
+  \begin{minipage}[t]{0.5\textwidth}
   \[\begin{array}{r >{\triangleq}cl}
-  \multicolumn{3}{l}{(\cdot)[?] : \Cmd \to \Cmd} \\
-  (h.f := e)[?] && h.f := e \\
-  (\assume b)[?] && \assume b \\
-  (\assert b)[?] && \assert b\\
-  (c_1;c_2)[?] && c_1[?];c_2[?]\\
-  (c_1 \choice c_2)[?] && c_1[?] \choice c_2[?]\\
-  (\{c\})[?] && \{c[?]\}\\
-  (t.\apply())[?] &&
+  \multicolumn{3}{l}{(\cdot)\instr : \Cmd \to \Cmd} \\
+  (h.f := e)\instr && h.f := e \\
+  (\assume b)\instr && b \\
+  (\assert b)\instr && \assert b\\
+  (c_1;c_2)\instr && c_1\instr;c_2\instr\\
+  (c_1 \choice c_2)\instr && c_1\instr \choice c_2\instr\\
+  (\{c\})\instr && \{c\instr\}\\
+  (t.\apply())\instr && \\
+  \multicolumn{3}{r}{
+    \qquad
     \begin{array}[t]{l} \displaystyle
       \{ \\
-    \quad \rho.\matches := \overrightarrow\havoc; \\
-    \quad \rho.\action := \havoc;\\
-    \quad \rho.\data := \overrightarrow\havoc; \\
-    \quad \assume \rho.\action < |t.\actions|; \\
+      \quad \havoc \rho;\\
+      \quad \assume \rho \in t; \\
       \quad \assume \rho.\matches = t.\keys; \\
@@
       \} \choice ( \\
-    \quad \assume \forall \rho \in t. \rho.\matches \neq t.\keys; \\
+      \quad \assume \forall \rho. \rho \in t \Rightarrow \rho.\matches \neq t.\keys;\\
       \quad \SKIP \\
       )
-  \end{array}
-  \end{array}
-  \]
-  \caption{Symbolic Table instrumentation. Typing information is omitted for
-    brevity; each $\havoc$ must have a appropriate bit-width annotation. For
-    simplicity, we assume each action in a table has the same number (and types) of
-    parameters.}
+    \end{array}}
+  \end{array}
+  \]
+  \end{minipage}\begin{minipage}[t]{0.45\textwidth}
+\[\begin{array}{r >{\triangleq}cl}
+  \multicolumn{3}{l}{(\cdot)\ginstr : \Cmd \to \Cmd} \\
+  (h.f := e)\ginstr && h.f := e \\
+  (\assume b)\ginstr && \assume b \\
+  (\assert b)\ginstr && \assert b\\
+  (c_1;c_2)\ginstr && c_1\ginstr;c_2\ginstr\\
+  (c_1 \choice c_2)\ginstr && c_1\ginstr \choice c_2\ginstr\\
+  (\{c\})\ginstr && \{c\ginstr\}\\
+  (t.\apply())\ginstr &&\\
+  \multicolumn{3}{l}{
+    \qquad
+    \begin{array}[t]{l} \displaystyle
+    \gamma_t.\reach := [1]_1; \\
+    \gamma_t.\reads := t.\keys; \\
+    (\\
+    \quad \assume \rho.\matches = t.\keys; \\
+    \quad \gamma_t.\hit := [1]_1; \\
+    \quad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
+    \left(\begin{array}{l}
+    \assume \rho.\action = i;\\
+    t.\action[i](\rho.\data)
+    \end{array}\right)\\
+    ) \choice ( \\
+    \quad \assume \rho.\matches \neq t.\keys; \\
+    \quad \gamma_t.\hit := [0]_1;\\
+    \quad \SKIP \\
+    )
+  \end{array}}
+  \end{array}
+  \]
+  \end{minipage}
+  \caption{Symbolic Table instrumentation alternatives. For simplicity, we
+    assume each action in a table has the same number (and types) of parameters.
+    We want to eliminate the universal quantifier in the miss case shown in
+    pn the left. The current attempt, on the right, uses a ghost header
+    $\gamma_t$ for every table $t$. }
   \label{fig:table-instrument}
@@
 \begin{figure}[htp]
-  \[\begin{array}{r >{\triangleq}cl}
+  \[\begin{array}{r cl}
   \multicolumn{3}{l}{\state\sigma_c : (\Value \to \Row^*) \to \BExpr} \\
-  \state\sigma_c = \bigwedge_{ t \in \tables(c) } \bigvee_{r_i \in \sigma(t.\id)}
-  \rho = r_i
+  \state\sigma_c &\triangleq
+  & \displaystyle\bigwedge_{ t \in \tables(c) }
+  \begin{array}{l}
+  \left(\bigvee_{r_i \in \sigma(t.\id)}
+  \rho = r_i\right) \wedge \\
+  \left(\bigwedge_{r_i \in \sigma(t.\id)}
+  \gamma_t.\mathit{reads} \neq r_i.\keys\right)
+  \Rightarrow \gamma_t.\mathit{hit} = [0]_1
+  \end{array}
   \end{array}
@@
 
-%% \begin{lemma}[Models Invariant over Actions]
-%%   \label{lem:models-invariant-over-actions}
-%%   For a closed action command $c$, $\mu,\mu'
-%%   \in \Model$, $\sigma \in \Value \to
-%%   \Row^*$, $\pkt \in \Pkt$, there exists some $\pkt'$
-%%   \[\denote{c}^\sigma\,\mu\,\pkt = \{\mu,\pkt'\}\]
-%%   \end{lemma}
-
-%% \begin{proof}
-%%   Proceed by induction over $c$.
-%%   \begin{enumerate}[align=left]
-%%   \item[$(c = h.f = e)$] Immediate.
-%%   \item[$(c = c_1;c_2)$]
-%%     By IH($c_1$) there exists some $\pkt''$ such that
-%%     $\denote{c_1}^\sigma\,\mu\,\pkt = \{(\mu,\pkt'')\}$.
-%%     By IH($c_2$) there exists some $\pkt'$ such that
-%%     $\denote{c_2}^\sigma,\mu\,pkt'' = \{(\mu, \pkt')\}$.
-%%     The result follows by definition of $\denote{-}$.
-%%     \end{enumerate}
-%% \end{proof}
-
 \begin{samepage}
-\begin{theorem}[Instrumentation Equivalence]
-  \label{thm:instr-equiv}
+\begin{theorem}[Ghostly Instrumentation Equivalence]
+  \label{thm:ghost-instr-equiv}
   For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, model
-  $\mu \in \Model$, and packet $\pkt \in \Pkt$, \[\denote{c}^\sigma~\mu~\pkt = \denote{c[?];\assume \state\sigma_c}^\sigma~\mu~\pkt.\]
+  $\mu \in \Model$, and packet $\pkt \in \Pkt$, \[\denote{c}^\sigma~\mu~\pkt = \denote{c\ginstr;\assume \state\sigma_c}^\sigma~\mu~\pkt.\]
 \end{theorem}
@@
 
+\begin{samepage}
+\begin{theorem}[Quantified Instrumentation Equivalence]
+  \label{thm:forall-instr-equiv}
+  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, model
+  $\mu \in \Model$, and packet $\pkt \in \Pkt$, \[\denote{c}^\sigma~\mu~\pkt = \denote{c\instr}^\sigma~\mu~\pkt.\]
+\end{theorem}
+\end{samepage}
+
 \begin{proof}
@@
 
-  %% \begin{enumerate}[align=left]
-  %% \item[$(c = h.f := e)$]
-  %%   Let $\sigma$,$\mu$,$\pkt$ be given.
-  %%   By definition $h.f:=e[?] = h.f:=e$. The result follows.
-  %% \item[$(c = \assume b)$] Similar. $[?]$ is no-op.
-  %% \item[$(c = \assert b)$] Similar. $[?]$ is no-op.
-  %% \item[$(c = c_1;c_2)$]
-  %%   Let $\sigma,\mu,\pkt$ be given.
-  %%   By definition $(c_1;c_2)[?] = c_1[?]; c_2[?]$.
-  %%   Compute as follows:
-  %%   \[\begin{array}{>{\displaystyle}ll}
-  %%   \denote{(c_1;c_2)[?]}^\sigma~\mu~\pkt \\
-  %%   = \denote{c_1[?];c_2[?]}^\sigma~\mu\pkt
-  %%   & \text{by definition of } [?]  \\
-  %%   = \bigcup_{(\mu',\pkt') \in \denote{c_1[?]}^\sigma\,\mu\,\pkt}\denote{c_2[?]}^\sigma~\mu'~\pkt'
-  %%   & \text{by definition of }\denote{-} \\
-  %%   = \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2[?]}^\sigma~\mu'~\pkt'
-  %%   & \text{by } \mathit{IH}(c_1, \mu, \pkt) \\
-  %%   = \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2}^\sigma~\mu'~\pkt'
-  %%   & \text{by } \mathit{IH}(c_2, \mu', \pkt') \\
-  %%   = \denote{c_1;c_2}^\sigma~\mu~\pkt
-  %%   & \text{by definition of } \denote{-}
-  %%   \end{array}
-  %%   \]
-  %% \item[$(c = c_1 \choice c_2)$] Similar, by IHs.
-
-  %% \item[$(c = t.\apply())$]
-  %%   Let $\sigma$,$\mu$,$\pkt$ be given.
-  %%   Let $r_1,\ldots,r_n = \sigma(t)$.
-  %%   There are two cases, either the packet hits a row, or hits no row. Explore each case separately.
-  %%   \begin{enumerate}
-  %%   \item[\textsc{Hit}]
-  %%     \textitP{}
-
-  %%   \item[\textsc{Miss}] In this case, there is no $r_i$ such that
-  %%     $\satisfy\sigma\pkt\mu r_i.\matches = t.\keys$, which means that the LHS
-  %%     of the toplevel choice operator should evaluate to $\{\}$ \todo{compute
-  %%       this}, and the RHS evaluates to $\{\mu,\pkt\}$\todo{compute this}, which
-  %%     is precisely the result we get when executing
-  %%     $\denote{t.apply()}^\sigma~\mu~\pkt$.
-
-  %%   \end{enumerate}
-
-  %% \end{enumerate}
 \end{proof}
+
+\subsection{Computing Logical Forms}
+
+Now, we would like to be able to translate the result from
+Theorems~\ref{thm:ghost-instr-equiv} and~\ref{thm:forall-instr-equiv} into a
+symbolic representation. To do this, we'll use the weakest precondition, defined
+in Figure~\ref{fig:wp}
+
+\begin{figure}[H]
+  \[
+  \begin{array}{r >{\triangleq}c l}
+    \WP(h.f := e, b)
+    && b[e/h.f] \\
+    \WP(\havoc \rho, b)
+    && \forall \rho.\,b \\
+    \WP(\assert b_a, b)
+    && b_a \wedge b \\
+    \WP(\assume b_a, b)
+    && b_a \Rightarrow b \\
+    \WP(t.\apply,b) && \textrm{\textcolor{red}{undefined}} \\
+    \WP(c_1;c_2, b)
+    && \WP(c_1,\WP(c_2,b)) \\
+    \WP(c_1 \choice c_2, b)
+    && \WP(c_1,b) \wedge \WP(c_2,b)
+  \end{array}
+  \]
+  \caption{Dijkstra's weakest precondition function}
+  \label{fig:wp}
+\end{figure}
+
+\subsection{A Strange Example}
+
+Now we'll present a counterexample showing that even though the
+$\forall$-free instrumentation $c\ginstr$ described in
+Figure~\ref{fig:table-instrument} is sufficient to prove
+Theorem~\ref{thm:ghost-instr-equiv}, it doesn't produce correct preconditions.
+
+\begin{samepage}
+Consider the following example table:
+\[t = \{\id = 47; \keys = [h.z]; \action = [\lambda x.\,h.x := x] \}\]
+
+and the following command
+\[c = h.x := 4; t.\apply(); \assert h.x \neq h.z\]
+\end{samepage}
+
+\subsubsection{Instrumentation}
+
+Now we must instrument this program. The resulting programs for both $c\ginstr$ and $c\instr$ are shown in Figure~\ref{fig:ex-instr}
+
+\begin{figure}[H]
+  \begin{minipage}[t]{0.5\textwidth}
+    \[
+    \begin{array}{l}
+      \multicolumn 1 c {\boxed{c\instr}} \\
+      h.x := 4; \\
+      \{\\
+      \quad \havoc \rho; \\
+      \quad \assume \rho \in t; \\
+      \quad \assume \rho.\matches[0] = h.z; \\
+      \quad \assume \action = 0; \\
+      \quad h.x := \rho.\data[0]; \\
+      \}\choice ( \\
+      \quad \assume \forall \rho.~\rho \in t \Rightarrow \rho.\matches[0] \neq h.z;\\
+      \quad \SKIP
+      )\\
+      \assert h.x \neq h.z
+    \end{array}
+    \]
+  \end{minipage} \begin{minipage}[t]{0.48\textwidth}
+    \[\begin{array}{l}
+    \multicolumn 1 c {\boxed{c\ginstr}} \\
+    h.x := 4;\\
+    \gamma_t.\reach := [1]_1;\\
+    \gamma_t.\reads[0] := [h.z]\\
+    (\\
+    \quad \assume \rho.\matches[0] = h.z; \\
+    \quad \gamma_t.\hit := [1]_1; \\
+    \quad \assume \rho.\action = 0;\\
+    \quad h.x := \rho.\data[0];\\
+    ) \choice (\\
+    \quad \assume \rho.\matches \neq t.\keys; \\
+    \quad \gamma.\hit := [0]_1; \\
+    \quad \SKIP \\
+    );\\
+    \assert h.x \neq h.z \\
+    \end{array}
+    \]
+  \end{minipage}
+  \caption{The left program is the forall instrumentation of program $c$, the
+    right one uses a ghost header. }
+  \label{fig:ex-instr}
+\end{figure}
+
+\subsubsection{Predicate Transformer}
+Now, we compute the predicate semantics for $c\ginstr$, which we will conclude is:
+
+\[\WP(c[?],\TRUE) =\rho.\action = 0 \Rightarrow
+\begin{array}{l}
+  \rho.\matches[0] = h.z \Rightarrow \rho.\data[0] \neq h.z\;\wedge \\
+  \rho.\matches[0] \neq h.z \Rightarrow 4 \neq h.z
+  \end{array}
+\]
+
+The derivation is in Section~\ref{sec:example-ghost-wp-derivation}. The problem is
+that the following state $\sigma$ triggers no assertion violations in the execution, but does violate the specification:
+
+\[\sigma = \left\{ 47 \mapsto \begin{array}{l}
+  [\{\keys = [4]; \action = 0;\data = [3]\}\\
+  \phantom[  \{\keys = [3]; \action = 0; \data = [4]\}
+  ]
+\end{array}
+\right\} \]
+
+First, let's see that $\sigma$ does not trigger an assertion violation. Consider
+the program $c[\sigma]$ below, and let $\pkt \in \Pkt$ be arbitary:
+
+\[
+\begin{array}{l}
+  h.x := 4; (\\
+  \quad (\assume h.z = 4; h.x := 3)\;\choice \\
+  \quad (\assume h.z = 3; h.x := 4)\;\choice \\
+  \quad (\assume h.z \neq 4; \assume h.z \neq 3; \SKIP) \\
+  ); \\
+  \assert h.x \neq h.z
+\end{array}
+\]
+
+There are three cases.
+\begin{enumerate}
+  \item If $\pkt(h.z) = 4$, then when it reaches the assert statement, $\pkt(h.x) =
+    3$, so $h.x \neq h.z$ is true.
+  \item If $\pkt(h.z) = 3$, then when it reaches the assert statement $\pkt(h.x) = 4$, so $h.x \neq h.z$ is true.
+  \item If $\pkt(h.z) \neq 3$ and $\pkt(h.z) \neq 4$, then by the time the
+    packet reaches the assert statement $\pkt(h.x) = 4$, and $h.x \neq h.z$ is
+    true because $\pkt(h.x) = 4 \neq \pkt(h.z)$.
+\end{enumerate}
+
+So we know that $\sigma$ should be allowed by the condition $\WP(c[?],\TRUE)$ for all packets. However the following $\mu$ and $pkt$ demonstrate the falsity of the condition:
+\[\mu = \{\rho \mapsto \sigma(t.\id)[1]\}
+\quad
+\pkt = \{h \mapsto \{z \mapsto 4, x\mapsto 0\}\} \]
+
+\[\begin{array}{l}
+\satisfy\sigma\pkt\mu \left(\rho.\action = 0 \Rightarrow
+\begin{array}{l}
+  \rho.\matches[0] = h.z \Rightarrow \rho.\data[0] \neq h.z\;\wedge \\
+  \rho.\matches[0] \neq h.z \Rightarrow 4 \neq h.z
+\end{array}\right)\\
+\iff\\
+\satisfy\sigma\pkt\mu  0 = 0 \Rightarrow
+\begin{array}{l}
+  (3 = 4 \Rightarrow 3 \neq 4) \wedge
+  (3 \neq 4 \Rightarrow 4 \neq 4)
+\end{array} \\
+\iff \\
+\satisfy\sigma\pkt\mu \TRUE \Rightarrow
+\begin{array}{l}
+  (\FALSE \Rightarrow \TRUE)\wedge (\FALSE \Rightarrow \FALSE)
+\end{array} \\
+\iff\\
+\satisfy\sigma\pkt\mu \FALSE \\
+\iff \\
+\textit{pigs fly}
+\end{array}
+\]
+
+\noindent which demonstrates that $\sigma$ is disallowed by $\WP(c\ginstr,\TRUE)$.
+
+This problem does not arise using the instrumentation $c\instr$. In this case,
+the  weakest precondition is (see Section~\ref{sec:example-forall-wp-derivation} for a derivation)
+\[
+\WP(c\instr,\TRUE) =
+\begin{array}{rl}
+&(\forall \rho.\; \rho \in t \wedge \rho.\matches[0] = h.z \wedge \rho.\action = 0
+\Rightarrow h.z \neq \rho.\data[0])\\
+\wedge
+& (\forall \rho.\;\rho \in t \Rightarrow\rho.\matches[0] \neq h.z) \Rightarrow 4 \neq h.z
+\end{array}
+\]
+\noindent which is satisfied by this example:
+
+\[
+\begin{array}{l}
+  \satisfy\sigma\pkt\mu
+  \begin{array}{rl}
+    &(\forall \rho.\; \rho \in t \wedge \rho.\matches[0] = h.z \wedge \rho.\action = 0
+    \Rightarrow h.z \neq \rho.\data[0])\\
+
+    &\wedge (\forall \rho.\;\rho \in t \Rightarrow\rho.\matches[0] \neq h.z) \Rightarrow 4  \neq h.z
+  \end{array} \\
+  \iff \\
+  \begin{array}{rl}
+    &\satisfy\sigma\pkt\mu \forall \rho.\; \rho \in t \wedge \rho.\matches[0] = h.z \wedge \rho.\action = 0
+    \Rightarrow h.z \neq \rho.\data[0]\\
+    \mathit{and}
+    &\satisfy\sigma\pkt\mu (\forall \rho.\;\rho \in t \Rightarrow\rho.\matches[0] \neq h.z) \Rightarrow 4  \neq h.z
+  \end{array} \\
+  \iff\\
+  \begin{array}{rl}
+    &\satisfy\sigma\pkt{\mu\{\rho \mapsto \sigma(47)[0]\}} \rho.\matches[0] = h.z \wedge \rho.\action = 0
+    \Rightarrow h.z \neq \rho.\data[0]\\
+    \mathit{and}
+  &\satisfy\sigma\pkt{\mu\{\rho \mapsto \sigma(47)[1]\}} \rho.\matches[0] = h.z \wedge   \rho.\action = 0
+    \Rightarrow h.z \neq \rho.\data[0]\\
+    \mathit{and}
+    &\satisfy\sigma\pkt\mu (\forall \rho.\;\rho \in t \Rightarrow\rho.\matches[0] \neq h.z) \Rightarrow 4  \neq h.z
+  \end{array} \\
+  \iff \\
+  \begin{array}{rl}
+    &\satisfy\sigma\pkt{\mu\{\rho \mapsto \sigma(47)[0]\}} 4 = 4 \wedge 0 = 0
+    \Rightarrow 4 \neq 3\\
+    \mathit{and}
+  &\satisfy\sigma\pkt{\mu\{\rho \mapsto \sigma(47)[1]\}} 3 = 4 \wedge \rho.\action = 0
+    \Rightarrow 4 \neq 4\\
+    \mathit{and}
+    &\satisfy\sigma\pkt\mu (\forall \rho.\;\rho \in t \Rightarrow\rho.\matches[0] \neq h.z) \Rightarrow 4  \neq h.z
+  \end{array} \\
+  \iff \\
+  \begin{array}{rl}
+    &\satisfy\sigma\pkt{\mu\{\rho \mapsto \sigma(47)[0]\}}\TRUE \\
+    \mathit{and}
+  &\satisfy\sigma\pkt{\mu\{\rho \mapsto \sigma(47)[1]\}} \TRUE
+    \Rightarrow 4 \neq 4\\
+    \mathit{and}
+    &\satisfy\sigma\pkt\mu (\forall \rho.\;\rho \in t \Rightarrow\rho.\matches[0] \neq h.z) \Rightarrow 4  \neq h.z
+  \end{array} \\
+  \iff \\
+  \begin{array}{rl}
+    &\satisfy\sigma\pkt{\mu\{\rho \mapsto \sigma(47)[0]\}}\TRUE \\
+    \mathit{and}
+  &\satisfy\sigma\pkt{\mu\{\rho \mapsto \sigma(47)[1]\}} \TRUE \\
+    \mathit{and}
+    &\satisfy\sigma\pkt\mu (\forall \rho.\;\rho \in t \Rightarrow\rho.\matches[0] \neq h.z) \Rightarrow 4  \neq h.z
+  \end{array} \\
+  \iff \\
+  \begin{array}{rl}
+    &\notsatisfy\sigma\pkt\mu \forall \rho.\;\rho \in t \Rightarrow\rho.\matches[0] \neq h.z \\
+    \mathit{or}
+    & \satisfy\sigma\pkt\mu ( 4  \neq h.z)
+  \end{array}\\
+  \iff\\
+  \begin{array}{rl}
+    &\notsatisfy\sigma\pkt{\mu\{\rho\mapsto\sigma(47)[0]\}} \rho.\matches[0] \neq h.z \\
+    \mathit{or}
+    &\notsatisfy\sigma\pkt{\mu\{\rho\mapsto\sigma(47)[1]\}} \rho.\matches[0] \neq h.z \\
+    \mathit{or}
+    & \satisfy\sigma\pkt\mu ( 4  \neq h.z)
+  \end{array}\\
+  \iff \\
+  \begin{array}{rll}
+    &\satisfy\sigma\pkt{\mu\{\rho\mapsto\sigma(47)[0]\}} 3 = 4  & \times \\
+    \mathit{or}
+    &\satisfy\sigma\pkt{\mu\{\rho\mapsto\sigma(47)[1]\}} 4 = 4 & \checkmark\\
+    \mathit{or}
+    & \satisfy\sigma\pkt\mu (4 \neq 4) & \times
+  \end{array}\\
+  \end{array}
+\]
+
+\subsubsection{Analysis}
+
+The problem with the ghostly instrumentation, $c\ginstr$, is that it allows for
+packets to execute the default action of a table $t$ whenever \emph{any} row in
+$\sigma(t.\id)$ is missed.
+
+So why we can prove Theorem~\ref{thm:ghost-instr-equiv}? We
+leverage the condition $\state\sigma_c$, which has two parts. The first, shown
+below, constrains the symbolic rows $\rho$ for a specific table $t$ to only
+refer to the rows that are in $\sigma(t.\id)$. In the universally-quantified
+version $c\instr$, this condition is realized symbolically using the constraints
+$\rho \in t$.
+\[ \bigwedge_{t \in \tables(c)} \bigvee_{r_i \in \sigma(t.\id)} \rho_t = r_i\]
+Concretely, for our example this is
+\[ \rho = \{\keys = [4]; \action = 0;\data = [3]\}\vee \rho = \{\keys = [3]; \action = 0; \data = [4]\} \]
+
+The second part of $\state\sigma_c$ ensures that the default action is only
+executed when the packet misses in the table. Shown below, this subformula plays
+the role of the $\forall \rho$ assumption at the start of the default action in
+$c\instr$.
+\[\bigwedge_{t \in \tables(c)} \left(\bigwedge_{r_i \in \sigma(t.\id)} \gamma_t.\reads = r_i.\matches\right) \Rightarrow \gamma_t.\hit = [0]_1 \]
+
+In our example this is
+\[(\gamma_t.\reads[0] \neq 3 \wedge \gamma_t.\reads[0] \neq 4) \Rightarrow \gamma_t.\hit = [0]_1 \]
+
+So we can prove Theorem~\ref{thm:ghost-instr-equiv}, because the state encoding
+$\state\sigma_c$ finitely enumerates the rows of $\sigma$. However, this
+information is not maintained in the weakest precondition calculation
+$\WP(c\ginstr,\TRUE)$.
+
+Of course it's not, because we've omitted $\state\sigma_c$ from the weakest
+precondition calculation! Of course $\WP(c\ginstr,\TRUE)$ isn't going to compute
+the precise constraints we want because the precision comes from $\state\sigma_c$.
+
+One solution is tod to generalize $\state\sigma_c$ to a symbolic representation,
+but this just gets us back to where we started. The naive generalization of the
+$\state\sigma_c$, which we'll write $\state?_c$ is the following:
+
+\[\state?_c \triangleq \bigwedge_{t \in \tables(c)} \rho \in t \wedge
+((\forall \rho. \rho \in t \Rightarrow \gamma_t.\reads = \rho.\keys ) \Rightarrow \gamma_t.\hit = [0]_1),\]
+\noindent which is the exact same structure provided by $c\instr$, except implemented using ghost state rather than at the table application point.
+
+\textbf{What's the move here?} It seems like instrumenting using $c\instr$ gives
+us the VCs we want, and $c\ginstr$ doesn't (yet). This fact is obscured by
+Theorem~\ref{thm:ghost-instr-equiv}. I think we want a theorem that describes
+the semantic completeness of whatever program we are going to pass into $\WP$,
+as Theorem~\ref{thm:forall-instr-equiv} does.
+The end goal is to compute a sufficiently general formula that can describe all legal control-plane behaviors, perhaps something like Theorem~\ref{thm:instr-expressiveness}:
+
+\begin{theorem}[Instrumentation Expressiveness]
+  \label{thm:instr-expressiveness}
+  Given an instrumentation strategy $(\cdot)[?] : \Cmd \to \Cmd$.
+  For every $c \in \Cmd$, and $\sigma : \Value \to \Row^*$
+  \[(\forall \pkt \in \Pkt. \forall \mu \in \Model.~\satisfy\sigma\pkt\mu \WP(c[?], \TRUE)) \iff (\forall \pkt \in \Pkt.\forall \mu \in \Model.~\denote{c}^\sigma\,\mu\,\pkt \neq \error)\]
+\end{theorem}
+
+Now, Theorem~\ref{thm:instr-expressiveness} is a corollary of
+Theorem~\ref{thm:forall-instr-equiv} and Dijsktra '75, with $[?] = \instr$.
+However, its not an immediate corollary of Theorem~\ref{thm:ghost-instr-equiv},
+because (I contend) it shows the equivalence of the program
+$c\ginstr;\assume\state\sigma_c$ and $c$ not the equivalence of $c\ginstr$ and
+$c$.
+
+I see three ways forward. The first is to proceed with $c\instr$. The second is
+to come up with an improvement on $c\ginstr$ that properly encodes misses (e.g.
+by returning to a previous approach that indexed each row $\rho_0, \rho_1,
+\ldots$). The third is to come up with a proper symbolic generalization of state
+$\state\sigma_c$.
+
+\subsubsection{WP derivation for Forall Instrumentation}
+\label{sec:example-forall-wp-derivation}
+
+\[
+\begin{array}{l}
+  \WP\left(
+  \begin{array}{l}
+    h.x := 4; \\
+    \{\\
+    \quad \havoc \rho; \\
+    \quad \assume \rho \in t; \\
+    \quad \assume \rho.\matches[0] = h.z; \\
+    \quad \assume \action = 0; \\
+    \quad h.x := \rho.\data[0]; \\
+    \}\choice ( \\
+    \quad \assume \forall \rho.~\rho \in t \Rightarrow \rho.\matches[0] \neq h.z;\\
+    \quad \SKIP
+    )\\
+    \assert h.x \neq h.z
+  \end{array},
+  \TRUE\right) \\
+  =\WP\left(
+  \begin{array}{l}
+    h.x := 4; \\
+    \{\\
+    \quad \havoc \rho; \\
+    \quad \assume \rho \in t; \\
+    \quad \assume \rho.\matches[0] = h.z; \\
+    \quad \assume \action = 0; \\
+    \quad h.x := \rho.\data[0]; \\
+    \}\choice ( \\
+    \quad \assume \forall \rho.~\rho \in t \Rightarrow \rho.\matches[0] \neq h.z;\\
+    \quad \SKIP
+    )
+  \end{array},
+  h.x \neq h.z\right) \\
+  =\WP\left(h.x := 4,
+  \WP\left(\begin{array}{l}
+    \{\\
+    \quad \havoc \rho; \\
+    \quad \assume \rho \in t; \\
+    \quad \assume \rho.\matches[0] = h.z; \\
+    \quad \assume \action = 0; \\
+    \quad h.x := \rho.\data[0]; \\
+    \}\choice ( \\
+    \quad \assume \forall \rho.~\rho \in t \Rightarrow \rho.\matches[0] \neq h.z;\\
+    \quad \SKIP
+    )
+  \end{array},
+  h.x \neq h.z\right)\right) \\
+  =\WP\left(h.x := 4,
+  \begin{array}{rl}
+    &\WP\left(\begin{array}{l}
+    \havoc \rho; \\
+    \assume \rho \in t; \\
+    \assume \rho.\matches[0] = h.z; \\
+    \assume \action = 0; \\
+    h.x := \rho.\data[0]
+    \end{array}, h.x \neq h.z \right) \\
+    \wedge &
+    \WP\left(\begin{array}{l}
+    \assume \forall \rho.~\rho \in t \Rightarrow \rho.\matches[0] \neq h.z;\\
+    \SKIP
+    \end{array},
+    h.x \neq h.z\right)
+  \end{array} \right) \\
+  =\WP\left(h.x := 4,
+  \begin{array}{rl}
+    &(\forall \rho. \rho \in t \wedge \rho.\matches[0] = h.z \wedge \rho.\action = 0 \Rightarrow \rho.\data[0] \neq h.z) \\
+    \wedge &
+     \left((\forall \rho.~\rho \in t \Rightarrow \rho.\matches[0] \neq h.z)
+     \Rightarrow h.x \neq h.z\right)
+  \end{array} \right) \\
+  =\begin{array}[t]{rl}
+    &(\forall \rho. \rho \in t \wedge \rho.\matches[0] = h.z \wedge \rho.\action = 0 \Rightarrow \rho.\data[0] \neq h.z) \\
+    \wedge &
+     \left((\forall \rho.~\rho \in t \Rightarrow \rho.\matches[0] \neq h.z)
+     \Rightarrow 4 \neq h.z\right)
+  \end{array} \\
+  \end{array}
+\]
+
+\subsubsection{WP derivation for Ghostly Instrumentation}
+\label{sec:example-ghost-wp-derivation}
+
+\[\begin{array}{l}
+  \WP\left(\begin{array}{l}
+    h.x := 4;\\
+    \gamma_t.\reach := [1]_1;\\
+    \gamma_t.\reads[0] := [h.z]\\
+    (\\
+    \quad \assume \rho.\matches[0] = h.z; \\
+    \quad \gamma_t.\hit := [1]_1; \\
+    \quad \assume \rho.\action = 0;\\
+    \quad h.x := \rho.\data[0];\\
+    ) \choice (\\
+    \quad \assume \rho.\matches \neq t.\keys; \\
+    \quad \gamma.\hit := [0]_1; \\
+    \quad \SKIP \\
+    );\\
+    \assert h.x \neq h.z \\
+  \end{array}, \TRUE\right) \\
+  =\WP\left(\begin{array}{l}
+    h.x := 4;\\
+    \gamma_t.\reach := [1]_1;\\
+    \gamma_t.\reads[0] := [h.z]\\
+    (\\
+    \quad \assume \rho.\matches[0] = h.z; \\
+    \quad \gamma_t.\hit := [1]_1; \\
+    \quad \assume \rho.\action = 0;\\
+    \quad h.x := \rho.\data[0];\\
+    ) \choice (\\
+    \quad \assume \rho.\matches \neq t.\keys; \\
+    \quad \gamma.\hit := [0]_1; \\
+    \quad \SKIP \\
+    );\\
+  \end{array},  h.x \neq h.z\right) \\
+  =\WP\left(\begin{array}{l}
+    h.x := 4;\\
+    \gamma_t.\reach := [1]_1;\\
+    \gamma_t.\reads[0] := [h.z]\\
+  \end{array},
+  \WP\left(
+  \begin{array}{rl}
+    &\WP\left(
+    \begin{array}{l}
+    \assume \rho.\matches[0] = h.z; \\
+    \gamma_t.\hit := [1]_1; \\
+    \assume \rho.\action = 0;\\
+    h.x := \rho.\data[0]\\
+    \end{array}, h.x \neq h.z\right)\\
+    \wedge &
+    \WP\left(\begin{array}{rl}
+    \assume \rho.\matches \neq t.\keys; \\
+    \gamma.\hit := [0]_1; \\
+    \SKIP \\
+    \end{array}, h.x\neq h.z \right)
+  \end{array}\right)
+  \right) \\
+  = \WP\left(\begin{array}{l}
+    h.x := 4;\\
+    \gamma_t.\reach := [1]_1;\\
+    \gamma_t.\reads[0] := [h.z]\\
+  \end{array},
+  \left(
+  \begin{array}{rl}
+    &\left(\rho.\matches[0] = h.z \wedge \rho.\action = 0;
+    \rho.\data[0] \neq h.z\right)\\
+    \wedge &
+    \left(\rho.\matches \neq t.\keys \Rightarrow h.x\neq h.z \right)
+  \end{array}\right)
+  \right) \\
+  =
+  \begin{array}{rl}
+    &\left(\rho.\matches[0] = h.z \wedge \rho.\action = 0;
+    \rho.\data[0] \neq h.z\right)\\
+    \wedge &
+    \left(\rho.\matches \neq t.\keys \Rightarrow 4\neq h.z \right)
+  \end{array}
+\end{array}\]
+
+\clearpage
+\section{The Research Problem}
+
+Now we can formally define the research problems that we want to solve. We have
+some ideal goals, that are probably not solvable in general.
+
+\begin{quote}
+  \textbf{Conservative Problem.} Given a command $c \in \Cmd$, compute the weakest test
+  $b$ such that if a state $\sigma : \Value \to \Row^*$ the following
+  judgement is defined and satisfied \[\satisfy\sigma\emptyset\emptyset b,\] then \[\forall \pkt \in \Pkt.~
+  \denote{c}^\sigma \pkt \neq \error \]
+\end{quote}
+
+However, this $b$ may be $\FALSE$ for many realistic programs $c$, so we may
+prefer to solve the following problem:
+
+\begin{quote}
+  \textbf{Flexible Problem.} Given a command $c \in \Cmd$, compute a test
+  $b_{\mathit{cp}}$, and a test $b_{\mathit{pkt}}$ such that
+  for every packet $\pkt \in \Pkt$ such that
+  \[\satisfy \emptyset \pkt \emptyset b_{\mathit{pkt}},\]
+  and for every state $\sigma : \Value \to \Row^*$ such that
+  \[\satisfy \sigma \pkt \emptyset b_{\mathit{cp}},\]
+  then \[\denote{c}^\sigma ~\emptyset~\pkt \neq \error \]
+\end{quote}
+
+\section{Reduction to QE}
+
+We can observe that the \textbf{Conservative Problem} is simply solvable. We
+sketch the argument here.
+
+\begin{conjecture}[Quantifer Elimination]
+  The \textbf{\textrm{Conservative Problem}} is reducable to Quantifer Elimination.
+\end{conjecture}
+
+\begin{proof}[Proof Sketch.]
+  Given a command $c \in \Cmd$. Compute the instrumented command $c[?]$. We know
+  it is equivalent to $c$ by Theorem~\ref{thm:instr-equiv}. Then $\forall
+  \overrightarrow{h.f}.\WP(c[?],\TRUE)[\overrightarrow{h.f}]$,\todo{how to write this?} is such a
+  formula $b$ by Dijkstra '75. To make $b$ a controller formula, then we simply
+  need to remove the dataplane variables via a QE procedure.
+\end{proof}
+
+This doesn't work in general. Standard QE procedures are exponential at best, and
+enumerating all eth/ipv4 src/dst addresses already takes an intractable amount
+of time.
+
+\section{Can Templates save us?}
+
+One option is to use a kind of constrained template-based synthesis, where we
+generate a catalogue of temaplates based on syntactic features of the target table.
+
+For example, given a single table like the following:
+
+\[t = \{\keys: [h.f]; \actions: [\lambda x.~~h.g:= x]\}\]
+
+and the simple program \[t.\apply()\]
+
+We may be able to generate a template that looks like
+
+\[\begin{array}{>{\displaystyle}c}
+  {[\exists \rho \in t.~ \rho.\keys[0]~\square_{\{=,\neq\}}~?_{v_1}\wedge \rho.\data[0]~\square_{\{=,\neq\}}~?_{v_2} ]?} \\
+  \wedge\\
+  {[\forall \rho \in t.~\rho.\keys[0]~\square_{\{=,\neq\}}~?_{v_3} \Rightarrow \rho.\data[0]~\square_{\{=,\neq\}}~?_{v_4}]?}
+\end{array}\]
+
+where $[b]?$ indicates a hole to either use template $b$ or replace it with
+$\TRUE$, $\square_S$ indicates a template for an operation, where $S$ is the set
+of allowed operations, and $?_{v_i}$ indicates a hole that should evaluate to an element
+of $\Value$.
+
+Now we may wonder: \emph{can we procedurally generate templates in a (nearly) complete way}?
+
+To do so, we need to extend our templates along the following dimensions:
+\begin{enumerate}[align=left]
+\item[\textbf{Keys}.] Should generalize by changing the LHS comparisons to $
+  \bigwedge_i[\rho.\keys[i]~\square_{\{=,\neq\}}~?_{v_i}]?$
+\item[\textbf{Action Data}.] Should generalize by changing the RHS comparisons to $\bigwedge_i[\rho.\data[i]~\square_{\{=,\neq\}}~?_{v_i}]?$
+\item[\textbf{Actions}.] We need to generalize the template across multiple actions. This should be as simple as adding $\rho.\action = i$ on the RHS of each implication above.
+\item[\textbf{Sequence}.] I'm not sure how to generalize sequentially composed tables.
+  Perhaps a larger catalogue of templates conditioned on structural information:
+  if we notice that $t_1$ has a dataflow path to the keys of $t_2$, we should
+  generate a join condition template, otherwise we may be able to simply ``and'' the templates.
+
+\item[\textbf{Nondeterminism} ($c_1 \choice c_2$).] To solve the \textbf{Conservative
+  Problem} we simply and the templates together, but to solve the
+  \textbf{Flexible Problem} we may be able to constrain the problem (by
+  constraining the packets) to either $c_1$ or $c_2$. This would require
+  computing a $b_{\textit{pkt}}$.
+\end{enumerate}
 
@@
   \right) \\
-
     \end{array}
@@
 
-\section{The Research Problem}
-
-Now we can formally define the research problems that we want to solve. We have
-some ideal goals, that are probably not solvable in general.
-
-\begin{quote}
-  \textbf{Conservative Problem.} Given a command $c \in \Cmd$, compute the weakest test
-  $b$ such that if a state $\sigma : \Value \to \Row^*$ the following
-  judgement is defined and satisfied \[\satisfy\sigma\emptyset\emptyset b,\] then \[\forall \pkt \in \Pkt.~
-  \denote{c}^\sigma \pkt \neq \error \]
-\end{quote}
-
-However, this $b$ may be $\FALSE$ for many realistic programs $c$, so we may
-prefer to solve the following problem:
-
-\begin{quote}
-  \textbf{Flexible Problem.} Given a command $c \in \Cmd$, compute a test
-  $b_{\mathit{cp}}$, and a test $b_{\mathit{pkt}}$ such that
-  for every packet $\pkt \in \Pkt$ such that
-  \[\satisfy \emptyset \pkt \emptyset b_{\mathit{pkt}},\]
-  and for every state $\sigma : \Value \to \Row^*$ such that
-  \[\satisfy \sigma \pkt \emptyset b_{\mathit{cp}},\]
-  then \[\denote{c}^\sigma ~\emptyset~\pkt \neq \error \]
-\end{quote}
-
-\section{Reduction to QE}
-
-We can observe that the \textbf{Conservative Problem} is simply solvable. We
-sketch the argument here.
-
-\begin{conjecture}[Quantifer Elimination]
-  The \textbf{\textrm{Conservative Problem}} is reducable to Quantifer Elimination.
-\end{conjecture}
-
-\begin{proof}[Proof Sketch.]
-  Given a command $c \in \Cmd$. Compute the instrumented command $c[?]$. We know
-  it is equivalent to $c$ by Theorem~\ref{thm:instr-equiv}. Then $\forall
-  \overrightarrow{h.f}.\WP(c[?],\TRUE)[\overrightarrow{h.f}]$,\todo{how to write this?} is such a
-  formula $b$ by Dijkstra '75. To make $b$ a controller formula, then we simply
-  need to remove the dataplane variables via a QE procedure.
-\end{proof}
-
-This doesn't work in general. Standard QE procedures are exponential at best, and
-enumerating all eth/ipv4 src/dst addresses already takes an intractable amount
-of time.
-
-\section{Can Templates save us?}
-
-One option is to use a kind of constrained template-based synthesis, where we
-generate a catalogue of temaplates based on syntactic features of the target table.
-
-For example, given a single table like the following:
-
-\[t = \{\keys: [h.f]; \actions: [\lambda x.~~h.g:= x]\}\]
-
-and the simple program \[t.\apply()\]
-
-We may be able to generate a template that looks like
-
-\[\begin{array}{>{\displaystyle}c}
-  {[\exists \rho \in t.~ \rho.\keys[0]~\square_{\{=,\neq\}}~?_{v_1}\wedge \rho.\data[0]~\square_{\{=,\neq\}}~?_{v_2} ]?} \\
-  \wedge\\
-  {[\forall \rho \in t.~\rho.\keys[0]~\square_{\{=,\neq\}}~?_{v_3} \Rightarrow \rho.\data[0]~\square_{\{=,\neq\}}~?_{v_4}]?}
-\end{array}\]
-
-where $[b]?$ indicates a hole to either use template $b$ or replace it with
-$\TRUE$, $\square_S$ indicates a template for an operation, where $S$ is the set
-of allowed operations, and $?_{v_i}$ indicates a hole that should evaluate to an element
-of $\Value$.
-
-Now we may wonder: \emph{can we procedurally generate templates in a (nearly) complete way}?
-
-To do so, we need to extend our templates along the following dimensions:
-\begin{enumerate}[align=left]
-\item[\textbf{Keys}.] Should generalize by changing the LHS comparisons to $
-  \bigwedge_i[\rho.\keys[i]~\square_{\{=,\neq\}}~?_{v_i}]?$
-\item[\textbf{Action Data}.] Should generalize by changing the RHS comparisons to $\bigwedge_i[\rho.\data[i]~\square_{\{=,\neq\}}~?_{v_i}]?$
-\item[\textbf{Actions}.] We need to generalize the template across multiple actions. This should be as simple as adding $\rho.\action = i$ on the RHS of each implication above.
-\item[\textbf{Sequence}.] I'm not sure how to generalize sequentially composed tables.
-  Perhaps a larger catalogue of templates conditioned on structural information:
-  if we notice that $t_1$ has a dataflow path to the keys of $t_2$, we should
-  generate a join condition template, otherwise we may be able to simply ``and'' the templates.
-
-\item[\textbf{Nondeterminism} ($c_1 \choice c_2$).] To solve the \textbf{Conservative
-  Problem} we simply and the templates together, but to solve the
-  \textbf{Flexible Problem} we may be able to constrain the problem (by
-  constraining the packets) to either $c_1$ or $c_2$. This would require
-  computing a $b_{\textit{pkt}}$.
-\end{enumerate}
-
 \end{document}

### 2021-04-15 inference.tex
simplify

--- 2021-03-31 inference.tex
@@
 \newcommand{\state}[1]{\llparenthesis#1\rrparenthesis}
+\newcommand{\instr}{[?]}
+\newcommand{\ginstr}{[\mathghost]}
+\newcommand{\qe}{\textsc{Qe}}
 
@@
     & \mid & e = e & \textit{Bitvector Equality} & (a,g,p,\iota)\\
-    & \mid & \forall \rho \in t.~b & \textit{Row Universality} & (\iota)\\
+    & \mid & \rho \in t & \textit{Table membership} & (\iota)\\
+    & \mid & \forall \rho.~b & \textit{Row Universality} & (\iota)\\
     \multicolumn{4}{l}{c \in \Cmd, a \in \Action, g \in \GCL, p \in \Prog, \iota \in \Instr } \\
     c & ::=  & h.f := e & \textit{Assignment} & (g,p,\iota) \\
+      %% & \mid & \havoc \rho & \textit{Row Havoc} & (\iota) \\
       & \mid & \assert b & \textit{Assertion} & (g,p,\iota)\\
@@
       & \mid & c \choice c & \textit{Nondeterministic Choice} & (g,p,\iota)\\
-      & \mid & \{c\} & \textit{Scope} & (g,p,\iota)\\
+      %% & \mid & \{c\} & \textit{Scope} & (g,p,\iota)\\
     \multicolumn{3}{l}{
@@
 
-Second, $p \in \Prog \subset \Cmd$ represents the expected set of programs extracted from a
-P4 interface, most notably, it contains tables $t$ and no symbolic selection
-statements or symbolic expersions.
-
-Conversely $\iota \in \Instr \subset \Cmd$ represents a program instrumented
-with symbolic variables. A program $\iota$ cannot apply tables using
-$t.\apply()$, but instead uses havoc statements, to
-introduce symbolic variables $\rho$ in $c$.
-
-Finally $g \in \GCL \subset \Prog \subset \Cmd$ represents the core algebraic
-flavor of GCL without havoc, symbolic tables, or table application.
-
 Define $\tables(c) \subset \Table$ to be the set of all tables that occur in $c$.
@@
 \begin{enumerate}
+\item[\textsc{Total}] For every $t \in \tables(c)$ $\pkt \in \pkt$ there exists a row $r \in \sigma(t.\id)$ such that $r.matches = \pkt(t.keys)$.
 \item[\textsc{WellFormedRows}] For every $t\in \tables(c)$, every row in $\sigma(t.\id)$ is well-formed.
@@
     \end{cases} \\
+    %% \denote{\havoc \rho}^\sigma\,\mu\,\pkt &&
+    %% \{(\mu\{\rho \mapsto r\}, \pkt) \mid r \in \Row\} \\
     \denote{c_1;c_2}^\sigma\,\mu\,\pkt&&
@@
     \denote{c_1}^\sigma\,\mu\,\pkt \cup \denote{c_2}^\sigma\,\mu\,\pkt \\
-    \denote{\{c\}}^\sigma\,\mu\,\pkt &&
-    \{(\mu,\pkt') \mid (\mu',\pkt') \in \denote{c}^\sigma\,\mu\,\pkt \}\\
+    %% \denote{\{c\}}^\sigma\,\mu\,\pkt &&
+    %% \{(\mu,\pkt') \mid (\mu',\pkt') \in \denote{c}^\sigma\,\mu\,\pkt \}\\
     \denote{t.\apply()}^\sigma\,\mu\,\pkt &&
@@
   \multicolumn{3}{l}{(\cdot)[\sigma] : \Cmd \to \Cmd} \\
-  \multicolumn{3}{l}{(\cdot)[\sigma] : \Prog \to \GCL} \\
   (h.f := e)[\sigma] && h.f := e   \\
@@
   (c_1 \choice c_2)[\sigma] && c_1[\sigma] \choice c_2[\sigma] \\
-  (\{c\})[\sigma] && \{c[\sigma]\} \\
+  %% (\{c\})[\sigma] && \{c[\sigma]\} \\
   (t.\apply())[\sigma] &&
@@
 The next step is to define a source to source translation that eliminates tables
-in the absence of concrete table state, which we will write $c[?] \in \Cmd$.
+in the absence of concrete table state, which we will write $c\instr \in \Cmd$.
 This is the first step towards the goal of constraining control plane behavior.
-Want the instrumented program $c[?]$ to fail on exactly the same situations that
-$c[\sigma]$ does, however $c[?]$ is constrained to have no table applications.
+Want the instrumented program $c\instr$ to fail in exactly the same situations that
+$c\instr$ does, however $c\instr$ is constrained to have no table applications.
 The only remaining mechanism is to use $\assume$ and $\assert$ statements to
@@
 
+We augment the headers for a program $c$ with ghost headers $\gamma_t$ for each
+table $t \in \tables(c)$. The header $\gamma_t$ has defined fields
+$\reach$, which is set to true if the packet reaches table $t$,
+$\hit$ which is set to true the packet hits in the table, and
+$\reads$ which is really a list of $|t.\keys|$ variables that stores the
+values of the packet when evaluating $t.\keys$.
+
 \begin{figure}[htp]
+  %% \begin{minipage}[t]{0.5\textwidth}
+  %% \[\begin{array}{r >{\triangleq}cl}
+  %% \multicolumn{3}{l}{(\cdot)\instr : \Cmd \to \Cmd} \\
+  %% (h.f := e)\instr && h.f := e \\
+  %% (\assume b)\instr && b \\
+  %% (\assert b)\instr && \assert b\\
+  %% (c_1;c_2)\instr && c_1\instr;c_2\instr\\
+  %% (c_1 \choice c_2)\instr && c_1\instr \choice c_2\instr\\
+  %% (\{c\})\instr && \{c\instr\}\\
+  %% (t.\apply())\instr && \\
+  %% \multicolumn{3}{r}{
+  %%   \qquad
+  %%   \begin{array}[t]{l} \displaystyle
+  %%     \{ \\
+  %%     \quad \havoc \rho;\\
+  %%     \quad \assume \rho \in t; \\
+  %%     \quad \assume \rho.\matches = t.\keys; \\
+  %%     \quad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
+  %%     \left(\begin{array}{l}
+  %%       \assume \rho.\action = i;\\
+  %%       t.\action[i](\rho.\data)
+  %%     \end{array}\right)\\
+  %%     \} \choice ( \\
+  %%     \quad \assume \forall \rho. \rho \in t \Rightarrow \rho.\matches \neq t.\keys;\\
+  %%     \quad \SKIP \\
+  %%     )
+  %%   \end{array}}
+  %% \end{array}
+  %% \]
+  %% \end{minipage}\begin{minipage}[t]{0.45\textwidth}
 \[\begin{array}{r >{\triangleq}cl}
-  \multicolumn{3}{l}{(\cdot)[?] : \Cmd \to \Cmd} \\
-  (h.f := e)[?] && h.f := e \\
-  (\assume b)[?] && \assume b \\
-  (\assert b)[?] && \assert b\\
-  (c_1;c_2)[?] && c_1[?];c_2[?]\\
-  (c_1 \choice c_2)[?] && c_1[?] \choice c_2[?]\\
-  (\{c\})[?] && \{c[?]\}\\
-  (t.\apply())[?] &&
+  \multicolumn{3}{l}{(\cdot)\instr : \Cmd \to \Cmd} \\
+  (h.f := e)\instr && h.f := e \\
+  (\assume b)\instr && \assume b \\
+  (\assert b)\instr && \assert b\\
+  (c_1;c_2)\instr && c_1\instr;c_2\instr\\
+  (c_1 \choice c_2)\instr && c_1\instr \choice c_2\instr\\
+  (t.\apply())\instr &&\\
+  \multicolumn{3}{l}{
+    \qquad
     \begin{array}[t]{l} \displaystyle
-    \{ \\
-    \quad \rho.\matches := \overrightarrow\havoc; \\
-    \quad \rho.\action := \havoc;\\
-    \quad \rho.\data := \overrightarrow\havoc; \\
-    \quad \assume \rho.\action < |t.\actions|; \\
-    \quad \assume \rho.\matches = t.\keys; \\
-    \quad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
+    \assume \rho_t.\matches = t.\keys; \\
+    \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
     \left(\begin{array}{l}
-    \assume \rho.\action = i;\\
-    t.\action[i](\rho.\data)
+    \assume \rho_t.\action = i;\\
+    t.\action[i](\rho_t.\data)
     \end{array}\right)\\
-    \} \choice ( \\
-    \quad \assume \forall \rho \in t. \rho.\matches \neq t.\keys; \\
-    \quad \SKIP \\
-    )
-  \end{array}
+  \end{array}}
   \end{array}
   \]
-  \caption{Symbolic Table instrumentation. Typing information is omitted for
-    brevity; each $\havoc$ must have a appropriate bit-width annotation. For
-    simplicity, we assume each action in a table has the same number (and types) of
-    parameters.}
+  \caption{Symbolic Table instrumentation. For simplicity, we
+    assume each action in a table has the same number (and types) of parameters}
   \label{fig:table-instrument}
@@
 \begin{figure}[htp]
-  \[\begin{array}{r >{\triangleq}cl}
+  \[\begin{array}{r cl}
   \multicolumn{3}{l}{\state\sigma_c : (\Value \to \Row^*) \to \BExpr} \\
-  \state\sigma_c = \bigwedge_{ t \in \tables(c) } \bigvee_{r_i \in \sigma(t.\id)}
-  \rho = r_i
+  \state\sigma_c &\triangleq
+  & \displaystyle\bigwedge_{ t \in \tables(c) }
+  \bigvee_{r_i \in \sigma(t.\id)} \rho = r_i
   \end{array}
@@
 
-%% \begin{lemma}[Models Invariant over Actions]
-%%   \label{lem:models-invariant-over-actions}
-%%   For a closed action command $c$, $\mu,\mu'
-%%   \in \Model$, $\sigma \in \Value \to
-%%   \Row^*$, $\pkt \in \Pkt$, there exists some $\pkt'$
-%%   \[\denote{c}^\sigma\,\mu\,\pkt = \{\mu,\pkt'\}\]
-%%   \end{lemma}
-
-%% \begin{proof}
-%%   Proceed by induction over $c$.
-%%   \begin{enumerate}[align=left]
-%%   \item[$(c = h.f = e)$] Immediate.
-%%   \item[$(c = c_1;c_2)$]
-%%     By IH($c_1$) there exists some $\pkt''$ such that
-%%     $\denote{c_1}^\sigma\,\mu\,\pkt = \{(\mu,\pkt'')\}$.
-%%     By IH($c_2$) there exists some $\pkt'$ such that
-%%     $\denote{c_2}^\sigma,\mu\,pkt'' = \{(\mu, \pkt')\}$.
-%%     The result follows by definition of $\denote{-}$.
-%%     \end{enumerate}
-%% \end{proof}
-
 \begin{samepage}
@@
   For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, model
-  $\mu \in \Model$, and packet $\pkt \in \Pkt$, \[\denote{c}^\sigma~\mu~\pkt = \denote{c[?];\assume \state\sigma_c}^\sigma~\mu~\pkt.\]
+  $\mu \in \Model$, and packet $\pkt \in \Pkt$, \[\denote{c}^\sigma~\mu~\pkt = \denote{\assume \state\sigma_c; c\instr}^\sigma~\mu~\pkt.\]
 \end{theorem}
+\begin{proof}
+\todo[inline]{do the proof}
+\end{proof}
 \end{samepage}
 
-\begin{proof}
-  \todo[inline]{Proceed by induction on the structure of $c$, leaving $\sigma$, $\mu$, $\pkt$ general}
+\subsection{Predicate Semantics}
 
-  %% \begin{enumerate}[align=left]
-  %% \item[$(c = h.f := e)$]
-  %%   Let $\sigma$,$\mu$,$\pkt$ be given.
-  %%   By definition $h.f:=e[?] = h.f:=e$. The result follows.
-  %% \item[$(c = \assume b)$] Similar. $[?]$ is no-op.
-  %% \item[$(c = \assert b)$] Similar. $[?]$ is no-op.
-  %% \item[$(c = c_1;c_2)$]
-  %%   Let $\sigma,\mu,\pkt$ be given.
-  %%   By definition $(c_1;c_2)[?] = c_1[?]; c_2[?]$.
-  %%   Compute as follows:
-  %%   \[\begin{array}{>{\displaystyle}ll}
-  %%   \denote{(c_1;c_2)[?]}^\sigma~\mu~\pkt \\
-  %%   = \denote{c_1[?];c_2[?]}^\sigma~\mu\pkt
-  %%   & \text{by definition of } [?]  \\
-  %%   = \bigcup_{(\mu',\pkt') \in \denote{c_1[?]}^\sigma\,\mu\,\pkt}\denote{c_2[?]}^\sigma~\mu'~\pkt'
-  %%   & \text{by definition of }\denote{-} \\
-  %%   = \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2[?]}^\sigma~\mu'~\pkt'
-  %%   & \text{by } \mathit{IH}(c_1, \mu, \pkt) \\
-  %%   = \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2}^\sigma~\mu'~\pkt'
-  %%   & \text{by } \mathit{IH}(c_2, \mu', \pkt') \\
-  %%   = \denote{c_1;c_2}^\sigma~\mu~\pkt
-  %%   & \text{by definition of } \denote{-}
-  %%   \end{array}
-  %%   \]
-  %% \item[$(c = c_1 \choice c_2)$] Similar, by IHs.
+Now, we would like to be able to translate the result from
+Theorem~\ref{thm:instr-equiv} into a
+symbolic representation. To do this, we'll use the weakest precondition, defined
+in Figure~\ref{fig:wp}
 
-  %% \item[$(c = t.\apply())$]
-  %%   Let $\sigma$,$\mu$,$\pkt$ be given.
-  %%   Let $r_1,\ldots,r_n = \sigma(t)$.
-  %%   There are two cases, either the packet hits a row, or hits no row. Explore each case separately.
-  %%   \begin{enumerate}
-  %%   \item[\textsc{Hit}]
-  %%     \textitP{}
-
-  %%   \item[\textsc{Miss}] In this case, there is no $r_i$ such that
-  %%     $\satisfy\sigma\pkt\mu r_i.\matches = t.\keys$, which means that the LHS
-  %%     of the toplevel choice operator should evaluate to $\{\}$ \todo{compute
-  %%       this}, and the RHS evaluates to $\{\mu,\pkt\}$\todo{compute this}, which
-  %%     is precisely the result we get when executing
-  %%     $\denote{t.apply()}^\sigma~\mu~\pkt$.
-
-  %%   \end{enumerate}
-
-  %% \end{enumerate}
-\end{proof}
-
-\subsection{Ghost State}
-
-Following p4v, we want construct a ghostly header $\gamma$ from a program $p$
-and use it to instrument the program to capture all dataplane behaviors.
-
+\begin{figure}[H]
   \[
-\begin{array}{l>{\triangleq}cl}
-  \multicolumn{3}{l}{\ghostly :: \Cmd \to \Cmd} \\
-  \multicolumn{3}{l}{\ghostly :: \Prog \to \GCL} \\
-  \ghostly(h.f := e)
-  && h.f := e \\
-  \ghostly(\assert b)
-  && \assert b\\
-  \ghostly(\assume b)
-  && \assume b \\
-  \ghostly(c_1;c_2)
-  && \ghostly(c_1);\ghostly(c_2) \\
-  \ghostly(c_1\choice c_2)
-  && \ghostly(c_1) \choice \ghostly(c_1)\\
-  \ghostly(t.\apply())
-  &&\\
-  \multicolumn{3}{c}{
-    \begin{array}[t]{>{\displaystyle\qquad}l}
-       \gamma.t_{\mathit{reach}} := [1]_1; \\
-       \gamma.t_{\mathit{read}(0)} := t.\keys[0]; \\
-       \vdots \\
-       \gamma.t_{\mathit{read}(|t.\keys|-1)} := t.\keys[|t.\keys|-1]; \\[1em]
-       \left(\gamma.t_{\mathit{hit}} := [0]_1 \right)
-       \choice \left(\bigchoice_{j = 1}^{|t.\action|-1}{
-         \left(\begin{array}{l}
-           \gamma.t_{\mathit{hit}} := [1]_1 ;\\
-           \gamma.t_{\mathit{action}} := [1]_1;\\
-           t.\action[j](\overrightarrow{\gamma.t_{\mathit{data}}})
-         \end{array}\right)
-       }\right)
-     \end{array}}
+  \begin{array}{r >{\triangleq}c l}
+    \WP(h.f := e, b)
+    && b[e/h.f] \\
+    %% \WP(\havoc \rho, b)
+    %% && \forall \rho.\,b \\
+    \WP(\assert b_a, b)
+    && b_a \wedge b \\
+    \WP(\assume b_a, b)
+    && b_a \Rightarrow b \\
+    \WP(t.\apply,b) && \textrm{\textcolor{red}{undefined}} \\
+    \WP(c_1;c_2, b)
+    && \WP(c_1,\WP(c_2,b)) \\
+    \WP(c_1 \choice c_2, b)
+    && \WP(c_1,b) \wedge \WP(c_2,b)
   \end{array}
   \]
-
-And now what we want to prove is
-
-\begin{theorem}{Instrumentation sufficiency}
-  For every well-formed program $c \in \Cmd$, every state $\sigma : \Value \to
-  \Row^*$, there is some $b_\sigma \in \BExpr$ such that for every model $\mu
-  \in \Model$ and every $\pkt \in \Pkt$,
-  \[\denote{c}^\sigma\,\mu\,\pkt =_{\not\gamma} \denote{\ghostly(c);\assume b}^\sigma\,\mu\,\pkt\]
-\end{theorem}
-
-\begin{proof}
-  Proceed by induction on $c$:
-  \begin{enumerate}[align=left]
-  \item[$(h.f:=e)$] Let $b_\sigma = \TRUE$; immediate.
-  \item[$(\assume b)$] Let $b_\sigma = \TRUE$; immediate.
-  \item[$(\assert b)$] Let $b_\sigma = \TRUE$; immediate.
-  \item[$(c_1;c_2)$] By IHs. $b_\sigma = b_1 \wedge b_2$ because $b_1$ and $b_2$ refer to disjoint variables.
-  \item[$(c_1 \choice c_2)$] \textit{sim}.
-  \item[$(t.\apply())$]
-
-    Let $b_\sigma$ be defined as follows:
-    \[
-    b_\sigma = \bigvee_{r \in \sigma(t.\id)}
-    \begin{array}{>{\displaystyle}l}
-      \gamma
-    \end{array}
-    \]
-
-    We can compute as follows:
-    \[\begin{array}{>{\displaystyle}ll}
-    \denote{\ghostly(c)}^\sigma\,\mu\,\pkt\\
-    =\denote{
-        \begin{array}{>{\displaystyle}l}
-          \gamma.t_{\mathit{reach}} := [1]_1; \\
-          \gamma.t_{\mathit{read}(0)} := t.\keys[0]; \\
-          \vdots \\
-          \gamma.t_{\mathit{read}(|t.\keys|-1)} := t.\keys[|t.\keys|-1]; \\[1em]
-          \left(\gamma.t_{\mathit{hit}} := [0]_1 \right)
-          \choice \left(\bigchoice_{j = 1}^{|t.\action|-1}
-            \left(\begin{array}{l}
-              \gamma.t_{\mathit{hit}} := [1]_1 ;\\
-              \gamma.t_{\mathit{action}} := j;\\
-              t.\action[j](\overrightarrow{\gamma.t_{\mathit{data}}})
-            \end{array}\right)
-          \right)
-    \end{array}}^\sigma\,\mu\,\pkt \\[2em]
-    =\denote{
-        \begin{array}{>{\displaystyle}l}
-          \left(\gamma.t_{\mathit{hit}} := [0]_1 \right)
-          \choice \left(\bigchoice_{j = 1}^{|t.\action|-1}
-            \left(\begin{array}{l}
-              \gamma.t_{\mathit{hit}} := [1]_1 ;\\
-              \gamma.t_{\mathit{action}} := j;\\
-              t.\action[j](\overrightarrow{\gamma.t_{\mathit{data}}})
-            \end{array}\right)
-          \right)
-    \end{array}}^\sigma\!\mu\,\pkt' \\
-    \qquad\qquad \text{where}~\pkt' = \pkt\left\{\begin{array}{l}
-    \gamma.t_{\mathit{reach}} \mapsto [1]_1,\\
-    \overrightarrow{\gamma.t_{\textit{key}} \mapsto \edenote{t.\keys}\,\pkt\,\mu}
-    \end{array}\right\} \\
-    =\denote{
-      \left(\gamma.t_{\mathit{hit}} := [0]_1 \right)}^\sigma\!\mu\,\pkt'
-    \cup
-    \denote{\left(\bigchoice_{j = 1}^{|t.\action|-1}
-            \left(\begin{array}{l}
-              \gamma.t_{\mathit{hit}} := [1]_1 ;\\
-              \gamma.t_{\mathit{action}} := j;\\
-              t.\action[j](\overrightarrow{\gamma.t_{\mathit{data}}})
-            \end{array}\right)
-            \right)
-    }^\sigma\!\mu\,\pkt' \\
-    = \left\{(\mu,\pkt'\left\{\gamma.t_{\mathit{hit}} \mapsto [0]_1\right\})\right\}
-    \cup
-        \denote{\left(\bigchoice_{j = 1}^{|t.\action|-1}
-          \left(\begin{array}{l}
-            \gamma.t_{\mathit{hit}} := [1]_1 ;\\
-            \gamma.t_{\mathit{action}} := j;\\
-            t.\action[j](\overrightarrow{\gamma.t_{\mathit{data}}})
-          \end{array}\right)
-          \right)
-        }^\sigma\!\mu\,\pkt' \\
-  = \left\{(\mu,\pkt'\left\{\gamma.t_{\mathit{hit}} \mapsto [0]_1\right\})\right\}
-  \cup
-  \left(\bigcup_{j = 1}^{|t.\action|-1}
-  \denote{t.\action[j](\overrightarrow{\gamma.t_{\mathit{data}}})}^\sigma\,\mu\,\pkt'\left\{
-  \begin{array}{l}
-    \gamma.t_{\textit{hit}} \mapsto [1]_1,\\
-    \gamma.t_{\textit{action}} \mapsto j
-  \end{array}
-  \right\}
-  \right) \\
-
-    \end{array}
-\]
-
-  \end{enumerate}
-\end{proof}
-
-\subsection{Tables as nondeterministic choice}
-
-Ive said before that the nondeterministic choice is an overapproximation of
-the control plane behavior. It's a slightly inaccurate claim in that it
-correctly captures all possible control plane behaviors, but it doesn't have a
-handle on constrain control plane behavior.
-
-A simplified version of P4v's encoding is shown in Figure~\ref{fig:p4vlike}. It
-omits the key-reads commands and hit/miss predicates.
-
-\begin{figure}
-  \[\begin{array}{l >{\triangleq}cl}
-  \multicolumn{3}{l}{\nondet : \Cmd \to \Cmd} \\
-  \multicolumn{3}{l}{\nondet : \Prog \to \Instr} \\
-  \nondet(h.f := e) && h.f := e \\
-  \nondet(\assume b) && \assume b \\
-  \nondet(\assert b) && \assert b \\
-  \nondet(c_1;c_2) && \nondet(c_1;c_2) \\
-  \nondet(c_1 \choice c_2) && \nondet(c_1) \choice \nondet(c_2) \\
-  %% \nondet(\choose \rho t c) && \textcolor{red}{\text{undefined}} \\
-  \nondet(t.\apply()) && \\
-  \multicolumn{3}{>{\qquad}r}{
-    \displaystyle
-    \bigchoice_{i = 0}^{|t.\action|-1}t.\action[i]()
-  }
-  \end{array}\]
-  \caption{A \texttt{p4v}-like instrumentation. For simplicity of reasoning
-    about variables, assume that actions take no action data. }
-  \label{fig:p4vlike}
+  \caption{Dijkstra's weakest precondition function}
+  \label{fig:wp}
 \end{figure}
 
-The theorem below demonstrates the over-approximation. Compare with the Instrumentation Equivalence Theorem~\ref{thm:instr-equiv}.
+Now, we need to relate states $\sigma$ and the computed $\WP(c\instr,\TRUE)$.
+Note that since there are no havocs introduced in $c\instr$, all symbolic rows
+$\rho$, and header values $h.f$ are unbound.
 
-\begin{theorem}[\texttt{p4v} over-approximates]
-  \label{thm:p4v-over}
-  For every $c \in \Prog$, well-formed state $\sigma : \Value \to \Row^* $,
-  model $\mu \in \Model$, and packet $\pkt \in \Pkt$,
-  $\denote{c}^\sigma\,\mu\,\pkt \subseteq \denote{\nondet(c)}^\sigma\,\mu\,\pkt$
-\end{theorem}
+This means that we must check the validity of \[\state\sigma_c \Rightarrow
+\WP(c\instr,\TRUE)\]
 
-\begin{proof}
-\todo[inline]{By induction on c}.
-\end{proof}
-
-We can witness non-equality by the following counter-example. Consider a program
-$t.\apply()$ where $t = \{\keys=[h.f];\actions=[\lambda().\SKIP, \lambda(). h.f :=
-  99]\}$. Consider an input packet with $\pkt(h.f) = 47$, and an empty input
-model and state that selects a single action.
-
-Then $\denote{t.\apply()}^\emptyset\,\emptyset\,\pkt = \{(\mu,\pkt)\}$.
-
-Whereas
-\[\begin{array}{l}
-\denote{\nondet(t.\apply())}^\emptyset\,\emptyset\,\pkt \\
-= \denote{
-    \displaystyle
-    \bigchoice_{i = 0}^{|t.\action|-1} t.\action[i]()
-}^\emptyset\,\emptyset\,\pkt \\
-= \denote{\SKIP}^\emptyset\,\emptyset\,\pkt \cup \denote{h.f := 99}^\emptyset\,\emptyset\,\pkt\\
-= \{(\emptyset,\pkt),(\emptyset,\pkt\{h.f\mapsto 99\})\}
-\end{array}\]
-
-Finally, $\{(\emptyset,\pkt)\} \subset
-\{(\emptyset,\pkt),(\emptyset,\pkt\{h.f\mapsto 99\})\}$, exemplifies a strict
-subset behavior described by Theorem~\ref{thm:p4v-over}.
-
+\clearpage
 \section{The Research Problem}
@@
   \textbf{Conservative Problem.} Given a command $c \in \Cmd$, compute the weakest test
-  $b$ such that if a state $\sigma : \Value \to \Row^*$ the following
-  judgement is defined and satisfied \[\satisfy\sigma\emptyset\emptyset b,\] then \[\forall \pkt \in \Pkt.~
-  \denote{c}^\sigma \pkt \neq \error \]
+  $b$ such that if, for a state $\sigma : \Value \to \Row^*$, the following
+  judgement is valid \[\state\sigma_c \Rightarrow b,\] then \[\forall \pkt \in \Pkt.~
+  \denote{c}^\sigma \emptyset\,\pkt \neq \error \]
 \end{quote}
@@
   \textbf{Flexible Problem.} Given a command $c \in \Cmd$, compute a test
-  $b_{\mathit{cp}}$, and a test $b_{\mathit{pkt}}$ such that
-  for every packet $\pkt \in \Pkt$ such that
-  \[\satisfy \emptyset \pkt \emptyset b_{\mathit{pkt}},\]
-  and for every state $\sigma : \Value \to \Row^*$ such that
-  \[\satisfy \sigma \pkt \emptyset b_{\mathit{cp}},\]
-  then \[\denote{c}^\sigma ~\emptyset~\pkt \neq \error \]
+  $b_{\mathit{cp}}$, and a test $b_{\mathit{pkt}}$ such that for every packet
+  $\pkt \in \Pkt$ and state $\sigma : \Value \to \Row^*$, if the following
+  formula is valid
+  \[b_{\mathit{pkt}} \wedge \state\sigma_c \Rightarrow b_{\mathit{cp}},\]
+  then \[\denote{c}^\sigma\emptyset\,\pkt \neq \error \]
 \end{quote}
@@
 
-We can observe that the \textbf{Conservative Problem} is simply solvable. We
-sketch the argument here.
+We can observe that the \textbf{Conservative Problem} is simply solvable by
+reduction to Quantifier Elimination. In fact Quantifier elimination is one of
+the key driving forces of the Dillig/Dillig/Millstein abduction work (\textsc{Hola},
+\textsc{Explain}). We sketch the argument here.
 
@@
 \begin{proof}[Proof Sketch.]
-  Given a command $c \in \Cmd$. Compute the instrumented command $c[?]$. We know
-  it is equivalent to $c$ by Theorem~\ref{thm:instr-equiv}. Then $\forall
-  \overrightarrow{h.f}.\WP(c[?],\TRUE)[\overrightarrow{h.f}]$,\todo{how to write this?} is such a
-  formula $b$ by Dijkstra '75. To make $b$ a controller formula, then we simply
-  need to remove the dataplane variables via a QE procedure.
+  Given a command $c \in \Cmd$. Compute the instrumented command $c\instr$. Let
+  $\overrightarrow{h.f}$ describe all the header fields occurring in $c\instr$. Write
+  $b[\overrightarrow{h.f}] = \WP(c\instr,\TRUE)$. Observe that $(\forall
+  \overrightarrow{h.f}.\;b[\overrightarrow{h.f}]) \Rightarrow
+  b[\overrightarrow{h.f}]$ is valid. Then the desired control plane formula is $\qe(\forall \overrightarrow{h.f}.\;b[\overrightarrow{h.f}])$.
 \end{proof}
 
-This doesn't work in general. Standard QE procedures are exponential at best, and
-enumerating all eth/ipv4 src/dst addresses already takes an intractable amount
-of time.
+This \emph{probably} doesn't work in general. Standard QE procedures are
+exponential at best, and enumerating all eth/ipv4 src/dst addresses already
+takes an intractable amount of time.
+
+The Dilligs' line of work uses Presburger arithmetic QE to implement the
+quantifier elimination scheme. We could directly adopt this, however networking
+programs make heavy use of bitwise operators like XOR, so we need to take a
+different strategy.
+StackOverflow\footnote{https://stackoverflow.com/questions/21293278/mathematical-arithmetic-representation-of-xor}
+says that you can implement $a \mathbin{\texttt{XOR}} b$ as $a + b - ab(1 + a +
+b - ab)$, but that still takes us out of PA, and casting between the naturals
+and fixed-width bitvectors seems dangerous.
 
@@
 
-We may be able to generate a template that looks like
+We may be able to generate a boolean template constructed from the following
+comparisons
 
-\[\begin{array}{>{\displaystyle}c}
-  {[\exists \rho \in t.~ \rho.\keys[0]~\square_{\{=,\neq\}}~?_{v_1}\wedge \rho.\data[0]~\square_{\{=,\neq\}}~?_{v_2} ]?} \\
-  \wedge\\
-  {[\forall \rho \in t.~\rho.\keys[0]~\square_{\{=,\neq\}}~?_{v_3} \Rightarrow \rho.\data[0]~\square_{\{=,\neq\}}~?_{v_4}]?}
+\[
+\begin{array}{l}
+  \rho.\keys[i]~\square_{\{=,\neq\}}~?_{v_3} \Rightarrow \rho.\data[i]~\square_{\{=,\neq\}}~?_{v_4} \\
+
+  \rho.\keys[i]~\square_{\{=,\neq\}}\rho.\data[i]
 \end{array}\]

### 2021-05-18 inference.tex
Create topology

@@
 \usepackage[margin=1in]{geometry}
-\usepackage{halloweenmath}
 \usepackage{amsmath, amssymb, amsthm, array, stmaryrd}
@@
 \usepackage{enumitem}
+\usepackage{hyperref}
+\usepackage{soul}
 
@@
 \newcommand{\SymbRow}{\mathsf{SymbRow}}
+\newcommand{\State}{\mathsf{State}}
 
@@
 \newtheorem{conjecture}{Conjecture}
+\newtheorem{definition}{Definition}
 
@@
     \multicolumn{3}{l}{\Value} \\
-    v & ::= & [n]_w  & \textit{Bitvector Literal} & (a,g,p,\iota) \\
+    v & ::= & [n]_w  & \textit{Bitvector Literal} \\
     \multicolumn{3}{l}{\SymbRow} \\
@@
       \data = \List[x]
-    \end{array}\right\} & \textit{Symbolic Row} & (\iota)\\
+    \end{array}\right\} & \textit{Symbolic Row} \\
     \multicolumn{3}{l}{\BVExpr} \\
-    e & ::= & v & \textit{Value} & (a,g,p,\iota)\\
-      & \mid & h.f & \textit{Header Access} & (a,g,p,\iota)\\
-      & \mid & \rho.\matches[i] & \textit{Symbolic Match} & (\iota)\\
-      & \mid & \rho.\action & \textit{Symbolic Action} & (\iota) \\
-      & \mid & \rho.\data[i] & \textit{Symbolic Action Data} & (\iota) \\
-      & \mid & x & \textit{Action Data Variable} & (a^*) \\
-      & \mid & e \binop e & \textit{Arbitrary Binary Operation} & (a,g,p,\iota) \\
-    & \mid & \unop e & \textit{Arbitrary Unary Operation} & (a,g,p,\iota) \\
+    e & ::= & v & \textit{Value} \\
+      & \mid & h.f & \textit{Header Access}\\
+      & \mid & \rho.\matches[i] & \textit{Symbolic Match}\\
+      & \mid & \rho.\action & \textit{Symbolic Action}\\
+      & \mid & \rho.\data[i] & \textit{Symbolic Action Data}\\
+      & \mid & x & \textit{Action Data Variable}\\
+      & \mid & e \binop e & \textit{Arbitrary Binary Operation}\\
+    & \mid & \unop e & \textit{Arbitrary Unary Operation} \\
     \multicolumn{3}{l}{\Table} \\
@@
     t.\actions = \List[\lambda \vec x.\,a]\\
-    \end{array}\right\} & \textit{Table} & (p)\\
+    \end{array}\right\} & \textit{Table}\\
     \multicolumn{3}{l}{\BExpr} \\
-    b & ::= & \FALSE  & \textit{Absurdity}& (a,g,p,\iota)\\
-      & \mid & b \Rightarrow b & \textit{Implication} & (a,g,p,\iota)\\
-    & \mid & e = e & \textit{Bitvector Equality} & (a,g,p,\iota)\\
-    & \mid & \rho \in t & \textit{Table membership} & (\iota)\\
-    & \mid & \forall \rho.~b & \textit{Row Universality} & (\iota)\\
+    b & ::= & \FALSE  & \textit{Absurdity}\\
+      & \mid & b \Rightarrow b & \textit{Implication}\\
+    & \mid & e = e & \textit{Bitvector Equality}\\
+    & \mid & \rho \in t & \textit{Table membership}\\
+    %% & \mid & \forall \rho.~b & \textit{Row Universality}\\
     \multicolumn{4}{l}{c \in \Cmd, a \in \Action, g \in \GCL, p \in \Prog, \iota \in \Instr } \\
-    c & ::=  & h.f := e & \textit{Assignment} & (g,p,\iota) \\
+    c & ::=  & h.f := e & \textit{Assignment}\\
       %% & \mid & \havoc \rho & \textit{Row Havoc} & (\iota) \\
-      & \mid & \assert b & \textit{Assertion} & (g,p,\iota)\\
-      & \mid & \assume b & \textit{Assumption} & (g,p,\iota) \\
-      & \mid & t.\apply() & \textit{Table Application} & (p)\\
-      & \mid & c;c & \textit{Sequential Composition} & (a,g,p,\iota) \\
-      & \mid & c \choice c & \textit{Nondeterministic Choice} & (g,p,\iota)\\
+      & \mid & \assert b & \textit{Assertion}\\
+      & \mid & \assume b & \textit{Assumption} \\
+      & \mid & t.\apply() & \textit{Table Application} \\
+      & \mid & c;c & \textit{Sequential Composition} \\
+      & \mid & c \choice c & \textit{Nondeterministic Choice} \\
       %% & \mid & \{c\} & \textit{Scope} & (g,p,\iota)\\
@@
 \]
-\caption{Grammar of Commands $c \in \Cmd$. Legal action indicated by $a \in
-  \Action$, legal instrumented command variants indicated by $\iota \in \Instr$,
-  dataplane programs indicated by $p \in \Prog$, and table-free commands
-  indicated by $g \in \GCL$.}
+\caption{Grammar of Commands $c \in \Cmd$.}
 \label{fig:grammar}
@@
 
-\begin{figure}
-  \[\begin{array}{l>{\iff}cl}
-  \satisfy \sigma \pkt \mu \FALSE && \textit{pigs fly} \\
-  \satisfy \sigma \pkt \mu b_1 \Rightarrow b_2 &&
-  \notsatisfy \sigma \pkt \mu  b_1~\textit{or}~
-  \satisfy \sigma \pkt \mu b_2\\
-  \satisfy \sigma \pkt \mu e_1 = e_2 &&
-  \edenote{e_1}~\pkt~\mu = \edenote{e_2}~\pkt~\mu, \qquad
-  \text{(note strict semantics)} \\
-  \satisfy\sigma\pkt\mu \rho \in t. && \mu(\rho) \in \sigma(t.\id) \\
-  \satisfy \sigma \pkt \mu \forall \rho. ~b &&
-  \forall r.
-  \satisfy \sigma \pkt {\mu\{\rho \mapsto r\}} b
-  \end{array}
-  \]
-  \todo[inline]{Clarify $\error$ states}
-  \caption{Satisfaction for $\BExpr$.}
-  \label{fig:bool-satis}
-\end{figure}
-
 \paragraph{Command Evaluation}
@@
   \[\begin{array}{l >{\triangleq}cl}
-  \multicolumn{3}{l}{\denote{c}^\sigma : \Model \to \Pkt \to \mathcal{P}(\Model \times \Pkt)_\error }\\
+  \multicolumn{3}{l}{\denote{c}^\sigma : \Pkt \to \mathcal{P}(\Pkt)_\error }\\
     \denote{h.f := e}^\sigma\,\mu\,\pkt && \{(\mu, \pkt\{h \mapsto v\} \mid v \in \edenote{e}\,\mu\,\pkt\})\} \\
@@
 
-This means that we must check the validity of \[\state\sigma_c \Rightarrow
-\WP(c\instr,\TRUE)\]
+\begin{conjecture}
+  For every $c \in \Cmd$, and $\sigma \in \State$, the following formula is valid
+  \[\state\sigma_c \Rightarrow  \WP(c\instr,\TRUE) \iff \WP(c[\sigma], \TRUE)\]
+\end{conjecture}
 
@@
 Now we can formally define the research problems that we want to solve. We have
-some ideal goals, that are probably not solvable in general.
+an ideal, that is probably not solvable in general. First we begin by defining a control plane formula.
+
+\begin{definition}[Control Plane Formulae ]
+  A control plane formula $\varphi$ is a boolean expression in $\BExpr$ that
+  contains no header accesses $h.f$.
+\end{definition}
 
 \begin{quote}
-  \textbf{Conservative Problem.} Given a command $c \in \Cmd$, compute the weakest test
-  $b$ such that if, for a state $\sigma : \Value \to \Row^*$, the following
-  judgement is valid \[\state\sigma_c \Rightarrow b,\] then \[\forall \pkt \in \Pkt.~
-  \denote{c}^\sigma \emptyset\,\pkt \neq \error \]
+  \textbf{Conservative Problem.} Given a command $c \in \Cmd$, compute (the
+  weakest) control plane formula $\varphi$ such that if, for a state $\sigma :
+  \Value \to \Row^*$, the following judgement is valid \[\state\sigma_c
+  \Rightarrow \varphi,\] then \[\forall \pkt \in \Pkt.~ \denote{c}^\sigma
+  \emptyset\,\pkt \neq \error \]
 \end{quote}
 
-However, this $b$ may be $\FALSE$ for many realistic programs $c$, so we may
+However, this $\phi$ may be $\FALSE$ for many realistic programs $c$, so we may
 prefer to solve the following problem:
@@
 \begin{quote}
-  \textbf{Flexible Problem.} Given a command $c \in \Cmd$, compute a test
-  $b_{\mathit{cp}}$, and a test $b_{\mathit{pkt}}$ such that for every packet
-  $\pkt \in \Pkt$ and state $\sigma : \Value \to \Row^*$, if the following
-  formula is valid
-  \[b_{\mathit{pkt}} \wedge \state\sigma_c \Rightarrow b_{\mathit{cp}},\]
+  \textbf{Flexible Problem.} Given a command $c \in \Cmd$, compute (the weakest)
+  control plane formula $\varphi$, and a test $b$ such that for every
+  packet $\pkt \in \Pkt$ and state $\sigma : \Value \to \Row^*$, if the
+  following formula is valid
+  \[b \wedge \state\sigma_c \Rightarrow \varphi,\]
   then \[\denote{c}^\sigma\emptyset\,\pkt \neq \error \]
@@
 
-\section{Reduction to QE}
+\section{Quantifier Elimination?}
 
 We can observe that the \textbf{Conservative Problem} is simply solvable by
-reduction to Quantifier Elimination. In fact Quantifier elimination is one of
-the key driving forces of the Dillig/Dillig/Millstein abduction work (\textsc{Hola},
+reduction to Quantifier Elimination. In fact. Quantifier elimination is one of
+the key driving forces of the Dilligs' abduction work (\textsc{Hola},
 \textsc{Explain}). We sketch the argument here.
@@
 
-This \emph{probably} doesn't work in general. Standard QE procedures are
+\paragraph{Discussion} This \emph{probably} doesn't work in general. Standard QE procedures are
 exponential at best, and enumerating all eth/ipv4 src/dst addresses already
-takes an intractable amount of time.
+takes an intractable amount of time. It's also not clear (to me) that this
+solution is weakest.
 
@@
 different strategy.
-StackOverflow\footnote{https://stackoverflow.com/questions/21293278/mathematical-arithmetic-representation-of-xor}
-says that you can implement $a \mathbin{\texttt{XOR}} b$ as $a + b - ab(1 + a +
-b - ab)$, but that still takes us out of PA, and casting between the naturals
-and fixed-width bitvectors seems dangerous.
 
-\section{Can Templates save us?}
+\section{Data-Driven Precondition Inference?}
+
+It seems that a solution to our problem can also be found in work by
+Padhi/Sharma/Millstein entitled \textit{Data-Driven Precondition Inference with
+  Learned Features}\todo{This won't make a ton of sense unless you've
+  read this paper}. Their algorithm uses a subroutine,
+$\textsc{VPreGen}(c,Q,G)$, which takes code $c$, a boolean predicate $Q$,
+and a set of passing test inputs $G$ and produces a predicate $P$ such that
+$P(t) = \TRUE$ for all $t \in G$, and $\{P\}\,C\,\{Q\}$.
+
+This gets us most of the way there. Naively, let's consider the space of inputs
+to be $\State \times \Pkt$. Then let $b_P$ be a boolean expression such that
+$b_p = \textsc{VPreGen}(c\instr,\TRUE,\emptyset)$. Then we know $b_P \Rightarrow
+\WP(c\instr,\TRUE)$. However, $b_p$ is not \emph{a priori} a control-plane
+formula.
+
+There's a trivial fix: modify the feature learning algorithm to only generate
+features that are control-plane formulae. The procedure $\textsc{FeatureLearn}$
+generates features from a space of operations described by the procedure
+$\textsc{GetOperations}()$. If we simply restrict \textsc{GetOperations} so that
+it only returns features that are control-plane formulae, then the generated
+expression $b_p$ is guaranteed to be a control plane formula. It may also be
+prudent to restrict the input space $G$ to be a subset of $\State$, rather than
+of $\State \times \Pkt$, but it doesn't seem strictly necessary.
+
+\paragraph{Discussion} This doesn't quite achieve our goal, because the
+synthesized solution is not necessarily the weakest. Further, its not clear to
+what extent this procedure relies on a non-empty set of examples $G$. Nor is it
+immediately clear where we would get such inputs. Currently each iteration of
+the \textsc{VPreGen} algorithm augments the set of examples with a failing
+example -- if it turns out that good inputs are necessary, it could be augmented
+with a similar solver call to generate a positive example (e.g.
+$\textsc{Sat}(\WP(c\instr,\TRUE))$).
+
+\paragraph{Relation to \texttt{bf4}} The \texttt{bf4} algorithm is similar to
+\textsc{VPreGen}, using formulae rather than examples. Rather than learn features, \texttt{bf4} uses a static set of
+control-plane features (which they call ``atoms''). Then, rather than learn a
+candidate boolean expression by abstracting these atoms and using a set of
+concrete positive and negative examples, \texttt{bf4} uses Z3's unsat-core to
+refine its candidate solution. Finally, \texttt{bf4} is interested in
+bug-finding rather than verification so it computes necessary conditions rather
+than sufficient conditions. The following table summarizes the differences
+
+\begin{center}
+\begin{tabular}{lll}
+  \toprule
+  & $\textsc{VPreGen}(c,Q,G)$ & \texttt{bf4's} $\textsc{INFER}(G,B,\mathcal P)$ \\
+  \midrule \\
+  Solution ($\varphi$)
+  & $\{\varphi\} C \{Q\}$, and $\forall t \in G.\,\varphi(t)$
+  & $G \models \phi$, minimizing(?) $|\{\eta\mid B
+  \wedge \varphi\}|$ \\[1.5em]
+  Features
+  & dynamically learns features
+  & features (aka \emph{atoms}) are heuristically defined ($\mathcal P$) \\[1.5em]
+  Candidates &
+  \begin{tabular}{>{\!\!\!}l}
+    abstracts features; \\
+    learns boolean expression; \\
+    concretizes features
+  \end{tabular}
+  & iteratively negates Z3's unsat-core\\
+  \bottomrule
+\end{tabular}
+\end{center}
+
+\section{Can we use templates?}
 
 One option is to use a kind of constrained template-based synthesis, where we
-generate a catalogue of temaplates based on syntactic features of the target table.
+generate a catalogue of templates based on syntactic features of the target table.
 
@@
 \begin{enumerate}[align=left]
-\item[\textbf{Keys}.] Should generalize by changing the LHS comparisons to $
+\item[\textit{Keys}.] Should generalize by changing the LHS comparisons to $
   \bigwedge_i[\rho.\keys[i]~\square_{\{=,\neq\}}~?_{v_i}]?$
-\item[\textbf{Action Data}.] Should generalize by changing the RHS comparisons to $\bigwedge_i[\rho.\data[i]~\square_{\{=,\neq\}}~?_{v_i}]?$
+\item[\textit{Action Data}.] Should generalize by changing the RHS comparisons to $\bigwedge_i[\rho.\data[i]~\square_{\{=,\neq\}}~?_{v_i}]?$
 \item[\textbf{Actions}.] We need to generalize the template across multiple actions. This should be as simple as adding $\rho.\action = i$ on the RHS of each implication above.
-\item[\textbf{Sequence}.] I'm not sure how to generalize sequentially composed tables.
+\item[\textit{Sequence}.] I'm not sure how to generalize sequentially composed tables.
   Perhaps a larger catalogue of templates conditioned on structural information:
@@
 
-\item[\textbf{Nondeterminism} ($c_1 \choice c_2$).] To solve the \textbf{Conservative
+\item[\textit{Nondeterminism} ($c_1 \choice c_2$).] To solve the \textbf{Conservative
   Problem} we simply and the templates together, but to solve the
@@
 \end{document}
+m

### 2021-05-18 inference.tex
clean up, remove vestiges of quantifiers

@@
 \newcommand{\qe}{\textsc{Qe}}
+\newcommand{\dom}{\textrm{dom}}
 
@@
     e & ::= & v & \textit{Value} \\
-      & \mid & h.f & \textit{Header Access}\\
-      & \mid & \rho.\matches[i] & \textit{Symbolic Match}\\
-      & \mid & \rho.\action & \textit{Symbolic Action}\\
-      & \mid & \rho.\data[i] & \textit{Symbolic Action Data}\\
-      & \mid & x & \textit{Action Data Variable}\\
+      & \mid & x & \textit{Variable}\\
       & \mid & e \binop e & \textit{Arbitrary Binary Operation}\\
@@
     & \mid & e = e & \textit{Bitvector Equality}\\
-    & \mid & \rho \in t & \textit{Table membership}\\
+    %% & \mid & \rho \in t & \textit{Table membership}\\
     %% & \mid & \forall \rho.~b & \textit{Row Universality}\\
@@
 Our dataplane programs will reason about structured representations of packets,
-that is, they will be transformations on sets of packet headers. In programs,
-packet headers $h \in \Hdr$ are essentially structs with fields $f \in
-\Field$. Write $h.f$ for a header $h$ and a field $f$ is the value $v$ of $f$
-in header $h$.
+that is, they will be transformations on sets of packet headers. In P4 programs,
+packet headers $h \in \Hdr$ are essentially structs with fields $f \in \Field$.
+Write $h.f$ for a header $h$ and a field $f$ is the value $v$ of $f$ in header
+$h$. Our simple command language just treats ``h.f'' as a variable.
 
@@
 \end{enumerate}
-For now we assume that the default action is the no-operation command (see
-$\SKIP$ below). We use $|l|$ to indicate the number of elements in a list or vector $l$
-such as $t.\keys$ or $t.\actions$.
+For now we assume that the default action is always specified by the controller.
+That is, programs never `miss' in tables. See the paragraph on state below. We use $|l|$ to indicate the number of
+elements in a list or vector $l$ such as $t.\keys$ or $t.\actions$.
 
 \paragraph{Symbolic Table Rows}
-A symbolic table row $\rho \in \SymbRow$ represents a symbolic row. Like tables,
+A symbolic table row $\rho \in \SymbRow$ represents a symbolic row. These rows are represented in our programs as simple variables $x$. Like tables,
 rows are also records with the following fields \todo{Well formedness constraint?}
@@
 \[\begin{array}{l}
-  \TRUE \triangleq \TRUE \Rightarrow \TRUE \\
+  \TRUE \triangleq \FALSE \Rightarrow \FALSE \\
   \neg b \triangleq b \Rightarrow \FALSE \\
@@
 
-We define $\SKIP$ as the following syntactic sugar $\assume \TRUE$.
-
-We express a collection of invariants on commands as subsets of the set $\Cmd$.
-It is convenient to describe these as subgrammars. First, action-commands $a \in
-\Action$ contain only non-symbolic assignment and sequence. If an action
-command has variables, indicated by the starred variant $a*$, it is called
-\emph{open}, if it has none it is called \emph{closed}.
+We define $\SKIP$ as syntactic sugar for $\assume \TRUE$.
 
@@
 \paragraph{Packets}
-A packet $\pkt$ is a map from header fields to values. Use the syntax $\pkt[h.f
-  \mapsto v]$ to represent the packet that is equivalent to $\pkt$ on all header
-fields except $h.f$, which maps to $v$.
+A packet $\pkt$ is a map from header fields to values. Use the syntax $\pkt[x
+  \mapsto v]$ to represent the packet that is equivalent to $\pkt$ on all variables
+fields except $x$, which maps to $v$.
 
@@
 following conditions hold:
-\begin{enumerate}
-\item[\textsc{Total}] For every $t \in \tables(c)$ $\pkt \in \pkt$ there exists a row $r \in \sigma(t.\id)$ such that $r.matches = \pkt(t.keys)$.
-\item[\textsc{WellFormedRows}] For every $t\in \tables(c)$, every row in $\sigma(t.\id)$ is well-formed.
-\item[\textsc{NoShadow}] For every $t \in \tables(c)$, with $r_1,\ldots,r_n =
+\begin{enumerate}[align=left]
+\item[(\textsc{Total})] For every $t \in \tables(c)$ $\pkt \in \pkt$ there exists a row $r \in \sigma(t.\id)$ such that $r.matches = \pkt(t.keys)$.
+\item[(\textsc{WellFormedRows})] For every $t\in \tables(c)$, every row in $\sigma(t.\id)$ is well-formed.
+\item[(\textsc{NoShadow})] For every $t \in \tables(c)$, with $r_1,\ldots,r_n =
   \sigma(t.\id)$, for every $1 \leq i \leq n$, there does not exist $1 \leq j <
@@
 \end{enumerate}
-\paragraph{Row Models}
-We also need models $\mu \in \Model$ to indicate which symbolic rows corresponds
-to which concrete rows, i.e. $\Model = \SymbRow \rightharpoonup \Row$. We write
-$\mu\{\rho \mapsto r\}$ to extend $\mu$ by mapping $\rho$ to $r$, and write $\mu
-\setminus \rho$, to remove $\rho$ from the map.
+
+Importantly, we assume that all state is \emph{total}, that is, no packet locally misses
+in the table. This is formally defined in Definition~\ref{def:total-state}.
+
+\begin{definition}[Total State]
+  \label{def:total-state}
+  A state $\sigma$ is total if for every $t \in \dom(\sigma)$, the following
+  formula is valid:
+  \[\bigwedge_{r \in \sigma(t.\id)} t.\keys = r.\matches \]
+\end{definition}
+
+Henceforth we assume that all states are total.
 
@@
   \[\begin{array}{l>{\triangleq}cl}
-  \multicolumn{3}{l}{\edenote{e} : \Model \to \Pkt \rightharpoonup \Value }\\
-    \edenote{v}~\mu~\pkt &&
+  \multicolumn{3}{l}{\edenote{e} : \Pkt \to \Value }\\
+    \edenote{v}\pkt &&
     v \\
-    \edenote{h.f}~\mu~\pkt &&
+    \edenote{h.f}\pkt &&
     pkt(h.f)\\
-    \edenote{\rho.\matches[i]}~\mu~\pkt &&
-    \mu(\rho).\matches[i] \\
-    \edenote{\rho.\action}~\mu~\pkt &&
-    \mu(\rho).\action \\
-    \edenote{\rho.\data[i]} &&
-    \mu(\rho).\data[i]\\
-    \edenote{e_1 \binop e_2}~\mu~\pkt&&
-    \edenote{e_1}~\mu~\pkt \binop
-    \edenote{e_2}~\mu~\pkt \\
-    \edenote{\unop e}~\mu~\pkt &&
-    \unop \left(\edenote{e}~\mu~\pkt\right)
+    %% \edenote{\rho.\matches[i]}\pkt &&
+    %% \mu(\rho).\matches[i] \\
+    %% \edenote{\rho.\action}\pkt &&
+    %% \mu(\rho).\action \\
+    %% \edenote{\rho.\data[i]} &&
+    %% \mu(\rho).\data[i]\\
+    \edenote{e_1 \binop e_2}\pkt&&
+    \edenote{e_1}\pkt \binop
+    \edenote{e_2}\pkt \\
+    \edenote{\unop e}\pkt &&
+    \unop \left(\edenote{e}\pkt\right)
   \end{array}
@@
 \paragraph{Boolean Satisfaction}
-The satisfaction relation $\satisfy \sigma \pkt \mu b$ is defined in
-Figure~\ref{fig:bool-satis} in terms of the three kinds of variable contexts,
-packets, models, and states. We also use the strict semantics here, most
-importantly in the equality rule. Sometimes two expressions are considered equal
-if they are both undefined; we do not take this approach, considering the
-equality undefined if either argument is undefined.
+The satisfaction relation $\pkt \models b$ is defined in
+Figure~\ref{fig:bool-satis}. It is completely standard.
+
+\begin{figure}
+  \[
+  \begin{array}{l >{\iff}c l}
+    \pkt \models \FALSE
+    && \mathit{never}\\
+    \pkt \models b_1 \Rightarrow b_2
+    && \pkt \not\models b_1~\mathit{or}~\pkt \models b_2\\
+    \pkt \models e_1 = e_2
+    && \edenote{e_1}\pkt = \edenote{e_2}\pkt\\
+
+  \end{array}
+  \]
+  \caption{The satisfaction relation}
+  \label{fig:bool-satis}
+\end{figure}
 
@@
 \begin{figure}[htp]
-  \[\begin{array}{l >{\triangleq}cl}
+  \[\begin{array}{r >{\triangleq}cl}
   \multicolumn{3}{l}{\denote{c}^\sigma : \Pkt \to \mathcal{P}(\Pkt)_\error }\\
-    \denote{h.f := e}^\sigma\,\mu\,\pkt && \{(\mu, \pkt\{h \mapsto v\} \mid v \in \edenote{e}\,\mu\,\pkt\})\} \\
-    \denote{\assume b}^\sigma\,\mu\,\pkt &&
+    \denote{h.f := e}^\sigma\pkt && \{\pkt\{h \mapsto \edenote{e}\pkt\}\} \\
+    \denote{\assume b}^\sigma\pkt &&
     \begin{cases}
-      \{(\mu,\pkt)\} & \satisfy \sigma \pkt \mu b \\
+      \{\pkt\} & \pkt \models b \\
       \{\} & \textit{otherwise} \\
     \end{cases}\\
-    \denote{\assert b}^\sigma\,\mu\,\pkt &&
+    \denote{\assert b}^\sigma\pkt &&
     \begin{cases}
-      \{(\mu,\pkt)\} & \satisfy \sigma \pkt \mu b \\
+      \{\pkt\} & \pkt \models b \\
       \error & \textit{otherwise}
@@
     %% \{(\mu\{\rho \mapsto r\}, \pkt) \mid r \in \Row\} \\
-    \denote{c_1;c_2}^\sigma\,\mu\,\pkt&&
-    \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2}^\sigma\,\mu'\,\pkt' \\
-    \denote{c_1 \choice c_2}^\sigma\,\mu\,\pkt &&
-    \denote{c_1}^\sigma\,\mu\,\pkt \cup \denote{c_2}^\sigma\,\mu\,\pkt \\
+    \denote{c_1;c_2}^\sigma\pkt&&
+    \bigcup_{\pkt' \in \denote{c_1}^\sigma\pkt}\denote{c_2}^\sigma\pkt' \\
+    \denote{c_1 \choice c_2}^\sigma\pkt &&
+    \denote{c_1}^\sigma\pkt \cup \denote{c_2}^\sigma\pkt \\
     %% \denote{\{c\}}^\sigma\,\mu\,\pkt &&
     %% \{(\mu,\pkt') \mid (\mu',\pkt') \in \denote{c}^\sigma\,\mu\,\pkt \}\\
-    \denote{t.\apply()}^\sigma\,\mu\,\pkt &&
+    \denote{t.\apply()}^\sigma\,\pkt &&
     \begin{cases}
-      \denote{c[\vec d/\vec x]}^\sigma\,\mu\,\pkt,
+      \denote{c[\vec d/\vec x]}^\sigma\pkt,
       & \text{where}~\lambda \vec x.~c = r.\action, \vec d = r.\data \\
-      & \text{where}~r = \argmin_i\{\satisfy \sigma \pkt \mu b_i\} \\
+      & \text{where}~r = \argmin_i\{\pkt \models b_i\} \\
       & \text{where}~b_i = \sigma(t.\id)[i].\matches = t.\keys \\
-      \{(\mu,\pkt)\}, & \mathit{otherwise}
+      \{\pkt\}, & \mathit{otherwise}
     \end{cases}
@@
 \begin{theorem}[Table Substitution]
-  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, model
-  $\mu \in \Model$, and packet $\pkt \in \Pkt$, then for any state $\tau :
-  \Value \to \Row^*$,
-  \[\denote{c}^\sigma~\mu~\pkt = \denote{c[\sigma]}^\tau~\mu~\pkt\]
+  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, and
+  packet $\pkt \in \Pkt$, then for any state $\tau : \Value \to \Row^*$,
+  \[\denote{c}^\sigma\pkt = \denote{c[\sigma]}^\tau\pkt\]
 \end{theorem}
@@
   \label{thm:instr-equiv}
-  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, model
-  $\mu \in \Model$, and packet $\pkt \in \Pkt$, \[\denote{c}^\sigma~\mu~\pkt = \denote{\assume \state\sigma_c; c\instr}^\sigma~\mu~\pkt.\]
+  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, and
+  packet $\pkt \in \Pkt$, \[\denote{c}^\sigma\pkt = \denote{\assume
+    \state\sigma_c; c\instr}^\sigma\pkt.\]
 \end{theorem}
@@
 \begin{conjecture}
-  For every $c \in \Cmd$, and $\sigma \in \State$, the following formula is valid
-  \[\state\sigma_c \Rightarrow  \WP(c\instr,\TRUE) \iff \WP(c[\sigma], \TRUE)\]
+  For every $c \in \Cmd$, and $\sigma \in \State$, the following are equivalent:
+  \begin{itemize}
+  \item \(\state\sigma_c \Rightarrow  \WP(c\instr,\TRUE)\)
+  \item \(\WP(c[\sigma], \TRUE)\)
+  \end{itemize}
 \end{conjecture}
@@
 \begin{quote}
-  \textbf{Conservative Problem.} Given a command $c \in \Cmd$, compute (the
+  \textbf{Inference Problem.} Given a command $c \in \Cmd$, compute (the
   weakest) control plane formula $\varphi$ such that if, for a state $\sigma :
-  \Value \to \Row^*$, the following judgement is valid \[\state\sigma_c
-  \Rightarrow \varphi,\] then \[\forall \pkt \in \Pkt.~ \denote{c}^\sigma
-  \emptyset\,\pkt \neq \error \]
+  \Value \to \Row^*$, the following formula is valid \[\state\sigma_c
+  \Rightarrow \varphi,\] then \[\forall \pkt \in \Pkt.~ \denote{c}^\sigma\pkt \neq \error \]
 \end{quote}
 
-However, this $\phi$ may be $\FALSE$ for many realistic programs $c$, so we may
+However, this $\varphi$ may be $\FALSE$ for many realistic programs $c$, so we may
 prefer to solve the following problem:
@@
 \begin{quote}
-  \textbf{Flexible Problem.} Given a command $c \in \Cmd$, compute (the weakest)
+  \textbf{Repair Problem.} Given a command $c \in \Cmd$, compute (the weakest)
   control plane formula $\varphi$, and a test $b$ such that for every
@@
   following formula is valid
-  \[b \wedge \state\sigma_c \Rightarrow \varphi,\]
-  then \[\denote{c}^\sigma\emptyset\,\pkt \neq \error \]
+  \[\state\sigma_c \Rightarrow \varphi,\]
+  then \[\denote{\assume b; c}^\sigma\pkt \neq \error \]
 \end{quote}
@@
 
-We can observe that the \textbf{Conservative Problem} is simply solvable by
+We can observe that \textbf{Inference} is simply solvable by
 reduction to Quantifier Elimination. In fact. Quantifier elimination is one of
@@
 \begin{conjecture}[Quantifer Elimination]
-  The \textbf{\textrm{Conservative Problem}} is reducable to Quantifer Elimination.
+  The \textbf{\textrm{Inference Problem}} is reducable to Quantifer Elimination.
 \end{conjecture}
@@
 
-\item[\textit{Nondeterminism} ($c_1 \choice c_2$).] To solve the \textbf{Conservative
-  Problem} we simply and the templates together, but to solve the
-  \textbf{Flexible Problem} we may be able to constrain the problem (by
+\item[\textit{Nondeterminism} ($c_1 \choice c_2$).] To solve the
+  \textbf{Inference Problem} we simply `and' the templates together, but to solve
+  the \textbf{Repair Problem} we may be able to constrain the problem (by
   constraining the packets) to either $c_1$ or $c_2$. This would require
@@
 \end{document}
-m

### 2021-05-24 inference.tex
sketch isntrumentation equivalence proof

@@
 \newcommand{\State}{\mathsf{State}}
+\newcommand{\Kind}{\mathsf{Kind}}
 
@@
 \newcommand{\reads}{\mathit{reads}}
+\newcommand{\hitAction}{\mathsf{hitAction}}
+\newcommand{\miss}{\mathsf{miss}}
 
@@
 \DeclareMathOperator*{\bigchoice}{\scalerel*{\choiceop}{\sum}}
-\newcommand{\havoc}[1]{\mathop{\textsf{havoc}}#1}
+\newcommand{\havoc}[1]{\mathop{\mathsf{havoc}}#1}
 \newcommand{\SKIP}{\mathsf{skip}}
+
+% matchkinds
+\newcommand{\exact}{\ensuremath{\mathsf{exact}}}
+\newcommand{\ternary}{\ensuremath{\mathsf{ternary}}}
+\newcommand{\lpm}{\ensuremath{\mathsf{lpm}}}
 
@@
 \newcommand{\dom}{\textrm{dom}}
+\newcommand{\fvs}{\textrm{fvs}}
+\newcommand{\efvs}{\textrm{efvs}}
+\newcommand{\bfvs}{\textrm{bfvs}}
+
+%% operators
+\newcommand{\bvand}{\mathop{\texttt{\&}}}
+\newcommand{\match}{\mathrel{\texttt{{=}\raisebox{-2.5px}{\textasciitilde}}}}
+\newcommand{\matchrow}{\mathsf{matchrow}}
 
@@
     & \mid & \unop e & \textit{Arbitrary Unary Operation} \\
+    \multicolumn{3}{l}{\Kind} \\
+    t & ::= & \exact  & \textit{Exact Match}\\
+    & \mid & \ternary & \textit{Ternary Match} \\
+    & \mid & \lpm & \textit{Longest Prefix Match} \\
     \multicolumn{3}{l}{\Table} \\
@@
     t.\id = v; \\
-    t.\keys = \List[h.f]; \\
+    t.\keys = \List[h.f : \kappa]; \\
     t.\actions = \List[\lambda \vec x.\,a]\\
@@
     \multicolumn{4}{l}{c \in \Cmd, a \in \Action, g \in \GCL, p \in \Prog, \iota \in \Instr } \\
-    c & ::=  & h.f := e & \textit{Assignment}\\
-      %% & \mid & \havoc \rho & \textit{Row Havoc} & (\iota) \\
+    c & ::=  & h.f := e & \textit{Assignment} \\
+      & \mid & \havoc \rho & \textit{Row Havoc} \\
       & \mid & \assert b & \textit{Assertion}\\
@@
     one occurence of each table identifier.
-  \item[($t.\keys$)] A list of header field accesses indicating the match keys.
+  \item[($t.\keys$)] A list of header field accesses indicating the match keys and their match kind.
   \item[($t.\actions$)] A list of actions that can be executed. An action is a lambda
@@
 elements in a list or vector $l$ such as $t.\keys$ or $t.\actions$.
+
+\paragraph{Match Kind}
+There are three match kinds, \exact, \ternary, and \lpm. Exact matches (\exact)
+indicate that the bits in the key must be the same as the match bits. Ternary
+matches (\ternary) specify a match and a mask and the key must be equivalent up
+to masking. Longest prefix matches (\lpm) specify a match and a length
+$l$, the first $l$ bits of the key must be equivalent to the match.
 
@@
   formula is valid:
-  \[\bigwedge_{r \in \sigma(t.\id)} t.\keys = r.\matches \]
+  \[\bigvee_{r \in \sigma(t.\id)} t.\keys = r.\matches \]
 \end{definition}
@@
 Henceforth we assume that all states are total.
+
+\todo[inline]{Can be Removed(?)}
 
@@
 notable differences are the inclusion of assertions, the nondeterminism
-operators, and a cleaner representation of table application.
+operators, support for inexact matches, and a cleaner representation of table
+application. The denotation function $\denote{-}$ is parameterized on a
+$\sigma$, it is used in the definition of $t.\apply()$. If no $\sigma$ is
+provided, the behavior of $t.\apply()$ is undefined.
+
+In evaluating tables, we must condition on and evaluate matches. To do this, we
+define an operator $t.\matchrow : \Row \to \BExpr$ which checks that the lists of
+matches and keys are equivalent up to the corresponding match type.
+
+First, define an auxiliary binary operator $(\match) : (\BVExpr \cup \BVExpr^2 \cup \BVExpr/\BVExpr) \to (\BVExpr \times \Kind) \to \BExpr$ that is $\TRUE$ iff the arguments are equivalent up
+to the match kind $\kappa$. Formally we define this as follows:
+
+\[
+\begin{array}{r<{\!\!\!} l >{\triangleq}cl}
+  e \match e' :& \exact && e = e' \\
+  (e_v,e_m) \match e' :& \ternary && e_v \bvand e_m = e' \bvand e_m \\
+  e/l \match  e' :& \lpm && e \bvand 1^l0^{|e| - l} = e' \bvand 1^l0^{|e| - l} \\
+\end{array}
+\]
+
+Now we can define $t.\matchrow(r)$ as follows
+\[t.\matchrow(r) \triangleq
+   \bigwedge_{i = 0}^n r.\matches[i] \match t.\keys[i]
+\]
 
@@
   \multicolumn{3}{l}{\denote{c}^\sigma : \Pkt \to \mathcal{P}(\Pkt)_\error }\\
-    \denote{h.f := e}^\sigma\pkt && \{\pkt\{h \mapsto \edenote{e}\pkt\}\} \\
+  \denote{x := e}^\sigma\pkt && \{\pkt\{x \mapsto \edenote{e}\pkt\}\} \\
+  \denote{\havoc x}^\sigma\pkt && \{ \pkt\{x \mapsto v\} \mid v \in \Value \} \\
     \denote{\assume b}^\sigma\pkt &&
@@
     \denote{c_1;c_2}^\sigma\pkt&&
-    \bigcup_{\pkt' \in \denote{c_1}^\sigma\pkt}\denote{c_2}^\sigma\pkt' \\
+    %% \bigcup_{\pkt' \in \denote{c_1}^\sigma\pkt}\denote{c_2}^\sigma\pkt' \\
+    {\denote{c_2}^\sigma}^\dagger\left(\denote{c_1}^\sigma\pkt\right)\\
     \denote{c_1 \choice c_2}^\sigma\pkt &&
@@
       \denote{c[\vec d/\vec x]}^\sigma\pkt,
+      & \text{exists}~r = \argmin_i\{\pkt \models b_i\} \\
       & \text{where}~\lambda \vec x.~c = r.\action, \vec d = r.\data \\
-      & \text{where}~r = \argmin_i\{\pkt \models b_i\} \\
-      & \text{where}~b_i = \sigma(t.\id)[i].\matches = t.\keys \\
-      \{\pkt\}, & \mathit{otherwise}
+      & \text{where}~b_i = t.\matchrow(\sigma(t.\id)[i]) \\
+      \{\pkt\}, & \textit{if no such } r \textit{ exists}
     \end{cases}
   \end{array}\]
-  \caption{Semantics of Commands}
+  \caption{Semantics of Commands. The dagger operator $(\cdot^\dagger) : (\Pkt \to
+    \mathcal{P}(\Pkt)_\error) \to \mathcal{P}(\Pkt)_\error \to
+    \mathcal{P}(\Pkt)_\error$ lifts a function that accepts packets to one that
+    accepts elements of to the pointed powerset.}
   \label{fig:com-sem}
@@
   \multicolumn{3}{l}{(\cdot)[\sigma] : \Cmd \to \Cmd} \\
-  (h.f := e)[\sigma] && h.f := e   \\
+  (x := e)[\sigma] && h.f := e \\
+  (\havoc x)[\sigma] && \havoc x \\
   (\assume b)[\sigma] && \assume b \\
@@
     (\\
-    \quad \assume {r_1.\matches = t.\keys};\\
+    \quad \assume {t.\matchrow(r_1)};\\
     \quad t.\action[r_1.\action](r_1.\data)\\
@@
     \phantom)\,\choice~(\\
-    \quad \assume {r_n.\matches = t.\keys};\\
+    \quad \assume {t.\matchrow(r_n)};\\
+    \quad \assume {\bigwedge_{i=1}^{n-1}{\neg t.\matchrow(r_i)}};\\
     \quad t.\action[r_n.\action](r_n.\data)\\
     )\choice (\\
-    \quad \displaystyle \assume {\bigwedge_{i=1}^n r_n.\matches \neq t.\keys};\\
+    \quad \assume {\bigwedge_{i=1}^n \neg t.\matchrow(r_i)};\\
     \quad \SKIP \\
@@
 \begin{theorem}[Table Substitution]
+  \label{thm:subst}
   For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, and
-  packet $\pkt \in \Pkt$, then for any state $\tau : \Value \to \Row^*$,
-  \[\denote{c}^\sigma\pkt = \denote{c[\sigma]}^\tau\pkt\]
+  packet $\pkt \in \Pkt$,
+  \[\denote{c}^\sigma\pkt = \denote{c[\sigma]}\,\pkt\]
 \end{theorem}
 \begin{proof}
-\todo[inline]{By induction on the structure of $c$.}
+   Proceed by induction on the structure of $c$.
+  \begin{enumerate}[align=left]
+    \item[$(c = (x := e))$] Immediate, since $(x:=e)[\sigma] = (x:=e)$.
+
+    \item[$(c = \havoc x)$] Immediate, since $(\havoc x)[\sigma] = \havoc x$.
+
+    \item[$(c = \assume b)$] Immediate, since $(\assume b)[\sigma] = (\assume b)$.
+
+    \item[$(c = \assert b)$] Immediate, since $(\assert b)[\sigma] = (\assert b)$.
+
+    \item[$(c = c_1;c_2)$] Let $\pkt,\sigma$ be arbitrary. Compute as follows
+      \[\begin{array}{ >{\displaystyle}l l}
+        \denote{(c_1;c_2)[\sigma]}\,\pkt \\
+        = \denote{c_1[\sigma];c_2[\sigma]}\,\pkt  & \text{definition of } (\cdot)[\sigma]\\
+        = \bigcup_{\pkt' \in \denote{c_1[\sigma]}\,\pkt} \denote{c_2[\sigma]}\pkt' & \text{definition of } \denote{\cdot}\\
+        = \bigcup_{\pkt' \in \denote{c_1[\sigma]}\,\pkt} \denote{c_2}^\sigma\pkt' & \text{IH}(c_2) \\
+        = \bigcup_{\pkt' \in \denote{c_1}^\sigma\pkt} \denote{c_2}^\sigma\pkt' & \text{IH}(c_1) \\
+        = \denote{c_1;c_2}^\sigma\pkt & \text{definition of } \denote{\cdot}^\sigma
+      \end{array}\]
+
+    \item[$(c = c_1 \choice c_2)$] Let $\pkt, \sigma$ be arbitrary. Compute as follows:
+      \[ \begin{array}{>{\displaystyle}ll}
+          \denote{c_1\choice c_2[\sigma]}\,\pkt\\
+          = \denote{c_1[\sigma] \choice c_2[\sigma]}\,\pkt & \text{definition of } (\cdot)[\sigma]\\
+          = \denote{c_1[\sigma]}\,\pkt \cup \denote{c_2[\sigma]}\,\pkt & \text{definition of } \denote{\cdot}\\
+          = \denote{c_1}^\sigma\pkt \cup \denote{c_2[\sigma]}\,\pkt & \text{IH}(c_1)\\
+          = \denote{c_1}^\sigma\pkt \cup \denote{c_2}^\sigma\,\pkt & \text{IH}(c_2)\\
+          = \denote{c_1\choice c_2}^\sigma\pkt & \text{definition of } \denote{\cdot}^\sigma
+          \end{array}
+      \]
+
+    \item[$(c = t.\apply())$] Let $\pkt,\sigma$ be arbitrary. Let $[r_1, \ldots,
+      r_n] = \sigma(t)$.
+
+      Compute as follows:
+      \[
+      \begin{array}{ll}
+        \denote{(t.\apply())[\sigma]}\,\pkt\\
+        = \denote{
+          \begin{array}{l}
+            \bigchoice_{i=1}^{n}
+            \begin{array}{l}
+              \assume t.\matchrow(r_i);\\
+              \assume \bigwedge_{j=1}^{i-1}\neg t.\matchrow(r_j)\\
+              t.\action[r_i.\action](r_i.data)
+            \end{array} \\
+            \choice \assume \bigwedge_{i=1}^{n} \neg  t.\matchrow(r_j); \SKIP
+        \end{array}}\pkt \\
+        = \begin{array}{l}
+          \bigcup_{i=1}^{|\sigma(t)|}
+          \denote{
+            \begin{array}{l}
+              \assume t.\matchrow(r_i);\\
+              \assume \bigwedge_{j=1}^{i-1}\neg t.\matchrow(r_j)\\
+              t.\action[\sigma(t)[i]](r_i.data)
+            \end{array}}\,\pkt\\
+            \cup \denote{\assume \bigwedge_{i=1}^{n} \neg  t.\matchrow(r_j); \SKIP}\,\pkt
+        \end{array} \\
+        \end{array}
+      \]
+      Now consider the two cases.
+
+      \begin{enumerate}
+      \item[\textit{Case 1}.] Assume that there is some $i \in \{1,\ldots,
+        n\}$ such that $\pkt \models t.\matchrow(r_i)$. Let $i$ be the
+        smallest such $i$. Then by definition $\denote{t.\apply()}^\sigma =
+        \denote{t.\action[r_i.\action](r_i.\data)}^\sigma\,\pkt$. Since it is
+        the smallest, it must be the that $\pkt \not\models t.\matchrow(r_j)$
+        for $1 \leq j < i$.This is equivalent to saying that $\pkt
+        \models\bigwedge_{j=1}^{i-1}\neg t.\matchrow(r_j)$. Consequently,
+        $\denote{t.\action[r_i.\action](r_i.\data)}\,\pkt \subseteq
+        \denote{t.\apply()[\sigma]}$. In fact this is an equality.
+
+        Because $i$ is the smallest index satisfying $\pkt \models
+        t.\matchrow(r_i)$, for all rows $r_j$ with $j < i$, $\denote{\assume
+          t.\matchrow(r_j)} = \emptyset$, and so the following holds: \[
+          \denote{
+          \begin{array}{l}
+            \assume t.\matchrow(r_j);\\
+            \assume \bigwedge_{j=1}^{i-1}\neg t.\matchrow(r_j)\\
+            t.\action[\sigma(t)[i]](r_i.data)
+        \end{array}}\,\pkt = \emptyset
+          \]
+
+        Now, for every index $j > i$, it may be that $\denote{\assume
+          t.\matchrow(r_j)} \neq \emptyset$, but because $\pkt \models
+        t.\matchrow(r_i)$, and $i < j$, we know $\denote{\bigwedge_{j=1}^{i-1}
+          \neg t.\matchrow(r_j)} = \emptyset$, so conclude that \[
+          \denote{
+          \begin{array}{l}
+            \assume t.\matchrow(r_j);\\
+            \assume \bigwedge_{j=1}^{i-1}\neg t.\matchrow(r_j)\\
+            t.\action[\sigma(t)[i]](r_i.data)
+        \end{array}}\,\pkt = \emptyset
+          \]
+
+        Similarly, $\denote{\assume \bigwedge_{i=1}^{n} \neg t.\matchrow(r_j)} =
+        \emptyset$, so conclude that $\denote{\assume \bigwedge_{i=1}^{n} \neg
+          t.\matchrow(r_j);\SKIP} = \emptyset$
+
+        Conclude that $\denote{t.\apply()[\sigma]} =
+        \denote{t.\action[r_i.\action](r_i.\data)}^\sigma\,\pkt$. The result
+        follows by applying the IH to $t.\action[r_i.\action](r_i.\data)$.
+
+        \item[\textit{Case 2.}] Assume that for all $i \in \{1, \ldots, n\}$,
+          $\pkt \not\models t.\matchrow(r_i)$. In this case,
+          $\denote{t.\apply()}^\sigma\,\pkt = \{\pkt\}$. Show that $\denote{t.\apply()[\sigma]}\,\pkt = \{\}\pkt\}$.
+
+          First observe that every $i \in \{1, \ldots, n\}$, we have
+          $\denote{\assume t.\matchrow(r_i)}\,\pkt = \emptyset$, so conclude
+          that
+          \[ \forall i \in \{1,\ldots,n\}.\denote{
+          \begin{array}{l}
+            \assume \bigwedge_{j=1}^{i-1}\neg t.\matchrow(r_j)\\
+            t.\action[\sigma(t)[i]](r_i.\data)
+        \end{array}}\,\pkt = \emptyset
+          \]
+
+          Finally $\denote{\assume \bigwedge_{i=1}^n \neg
+            t.\matchrow(r_i)}\,\pkt = \{\pkt\}$, and $\denote{\SKIP}\,\pkt
+          =\{\pkt\}$, so conclude that $\denote{\assume \bigwedge_{i=1}^n \neg
+            t.\matchrow(r_i);\SKIP}\,\pkt = \{\pkt\}$.
+          Conclude that $\denote{t.\apply()[\sigma]}\pkt = \{pkt\}$.
+      \end{enumerate}
+  \end{enumerate}
 \end{proof}
@@
     \begin{array}[t]{l} \displaystyle
-    \assume \rho_t.\matches = t.\keys; \\
+      \assume t.\matchrow(\rho_t); \\
+      \assume \gamma_t.\keys = t.\keys; \\
     \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
@@
       \assume \rho_t.\action = i;\\
+      \assume \gamma_t.\hitAction = \rho_t.\id;\\
       t.\action[i](\rho_t.\data)
@@
   & \displaystyle\bigwedge_{ t \in \tables(c) }
-  \bigvee_{r_i \in \sigma(t.\id)} \rho = r_i
+  \left(
+  \begin{array}{>{\displaystyle}l}
+  \gamma_t.\miss = 1 \Leftrightarrow \gamma_t.\hitAction > |\sigma(t.\id)| \\
+  \wedge \bigvee_{r_i \in \sigma(t.\id)} \left(\rho = r_i
+  \wedge \left(\gamma_t.\hitAction \geq i \Leftrightarrow \bigwedge_{j=0}^{i-1} t.\matchrow(r_j)[\gamma_t.\keys/t.\keys]\right)\right)\\
+  \end{array}\right)
   \end{array}
@@
 
+\begin{definition}[Projections]
+  For a set of variables $S \subseteq \Var$, and a packet $\pkt$, write
+  $\pkt|_S$ to be projection of $\pkt$ onto $S$. \todo{I guess this means
+    packets are partial functions?} Formally,
+  \[\pkt|_S \triangleq \{x \mapsto \pkt(x) \mid x \in S\}.\]
+  Two packets $\pkt$,
+  $\pkt'$ are equivalent up to $S \subseteq \Var$, written $\pkt =_S \pkt'$, iff
+  $\pkt|_S = \pkt'|_s$. A set of packets $P$ is a subset of $P'$ up to $S$,
+  written $P \subseteq_S P'$ iff for every packet $\pkt \in P$, there exists
+  $\pkt' \in P'$ such that $\pkt =_S \pkt'$. Two sets of packets $P$ and $P'$
+  are equivalent up to $S$, written $P =_S P'$ iff $P \subseteq_S P'$ and $P \subseteq_S P'$.
+\end{definition}
+
+\begin{definition}[Variables]
+  The free variables of a command $c$, written $\fvs(c) \subseteq \Var$, is the
+  set of variables that occur at any point in $c$ (i.e. as the lhs of an
+  assignment, in a boolean expression, a bitvector expression, as a key in a
+  table.) The only bound variables are the action data in tables. Formally:
+  \[\begin{array}{l >{\triangleq}cl}
+  \efvs :: \BVExpr \to \mathcal{P}(\Var) \\
+  \efvs(v) && \{\} \\
+  \efvs(x) && \{x\}\\
+  \efvs(e_1 \binop e_2) && \efvs(e_1) \cup \efvs(e_2)\\
+  \efvs(\unop e) && \efvs(unop) \\[1em]
+  \bfvs :: \BExpr \to \mathcal{P}(\Var) \\
+  \bfvs(\FALSE) && \{\}\\
+  \bfvs(b_1 \Rightarrow b_2) && \bfvs(b_1) \cup \bfvs(b_2)\\
+  \bfvs(e_1 = e_2) && \efvs(e_1) = \efvs(e_2) \\[1em]
+  \fvs :: \Cmd \to \mathcal{P}(\Var) \\
+  \fvs(x := e) && \{x\} \cup \fvs(e)\\
+  \fvs(\havoc x) && \{x\}\\
+  \fvs(\assume b) && \fvs(b)\\
+  \fvs(\assert b) && \fvs(b) \\
+  \fvs(c_1;c_2) && \fvs(c_1) \cup \fvs(c_1) \\
+  \fvs(c_1 \choice c_2) && \fvs(c_1) \cup \fvs(c_2) \\
+  \fvs(t.\apply()) &&
+  t.\keys \cup \bigcup_{\lambda \vec x. c \in t.\actions} (\fvs(c) \setminus \{x_1,\ldots, x_n\})
+  \end{array}\]
+\end{definition}
+
 \begin{samepage}
@@
   \label{thm:instr-equiv}
-  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, and
-  packet $\pkt \in \Pkt$, \[\denote{c}^\sigma\pkt = \denote{\assume
-    \state\sigma_c; c\instr}^\sigma\pkt.\]
+  Given a set of variables $V \subseteq \Var$ that contains no ghost variables
+  or symbolic variables. For every $c \in \Cmd$ such that $\fvs(c) \subseteq V$, well-formed state
+  $\sigma : \Value \to \Row^*$, and packet $\pkt \in \Pkt$,
+  %% \[\denote{c}^\sigma\pkt =_{\fvs(c)} \denote{\assume\state\sigma_c;c\instr}\,\pkt.\]
+
+  \[\denote{c}^\sigma\,\pkt =_V \denote{\havoc \vec{\rho_t}; \havoc \vec{\gamma_t}; \assume{\state\sigma}_t; c\instr}\,\pkt  \]
 \end{theorem}
-\begin{proof}
-\todo[inline]{do the proof}
-\end{proof}
 \end{samepage}
 
-\subsection{Predicate Semantics}
+\begin{proof}
+  Proceed by induction on $c$.
+  \begin{enumerate}[align=left]
+  \item[$(c = x:=e)$] Immediate.
+  \item[$(c = \havoc x)$] Immediate.
+  \item[$(c = \assume b)$] Immediate.
+  \item[$(c = \assert b)$] Immediate.
+  \item[$(c = c_1;c_2)$] Let $\vec{\rho_t}$ be the symbolic rows for tables $t
+    \in \tables(c_1;c_2)$. We can write $\vec\rho_t$ as $\vec{\rho_t} = \vec{\rho_{t_1}} \vec{\rho_{t_2}}$
+    where $\vec{\rho_{t_i}}$ is the set of symbolic rows in $c_i$. Let
+    $\vec{\gamma_t}$ be the ghost variables rows for tables $t \in
+    \tables(c_1;c_2)$. We can write $\vec{\gamma_t}$ as $\vec{\gamma_t} = \vec{\gamma_{t_1}} \vec{\gamma_{t_2}}$
+    where $\vec{\gamma_{t_i}}$ is the set of ghost variables in $c_i$. Compute as
+    follows:
+    \[\begin{array}{ll}
+    \denote{\havoc{\vec{\rho_t}}; \havoc{\vec\gamma_t}; \assume{\state\sigma_{c_1;c_2}};(c_1;c_2)\instr}\,\pkt
+    & \\
+    = \denote{\havoc{\vec{\rho_t}}; \havoc{\vec\gamma_t}; \assume{\state\sigma_{c_1;c_2}}; c_1\instr;c_2\instr}\,\pkt
+    & \text{definition of }(\cdot)\instr \\
+    = \denote{\havoc{\vec{\rho_t}}; \havoc{\vec\gamma_t}; \assume{\state\sigma_{c_1}};\assume{\state\sigma_{c_2}};c_1\instr;c_2\instr}\,\pkt
+    & \text{LEMMAs INVOLVING SIGMA-DISTRIB \& ASSUME-DISTRIB} \\
+    = \denote{\havoc{\vec{\rho_{t_1}}}\havoc{\vec{\rho_{t_2}}}; \havoc{\vec\gamma_{t_1}}; \havoc{\vec\gamma_{t_2}}; \assume{\state\sigma_{c_1}};\assume{\state\sigma_{c_2}};c_1\instr;c_2\instr}\,\pkt
+    & \text{by definition} \\
+    = \denote{\havoc{\vec{\rho_{t_1}}};\havoc{\vec{\rho_{t_2}}}; \havoc{\vec\gamma_{t_1}}; \havoc{\vec\gamma_{t_2}}; \assume{\state\sigma_{c_1}};c_1\instr;\assume{\state\sigma_{c_2}};c_2\instr}\,\pkt
+    & \text{BY FREE-VARIABLE-ASSUMPTION COMMUTATIVITY LEMMA} \\
+    = \denote{\havoc{\vec{\rho_{t_1}}}; \havoc{\vec\gamma_{t_1}};\havoc{\vec{\rho_{t_2}}}; \havoc{\vec\gamma_{t_2}}; \assume{\state\sigma_{c_1}};c_1\instr;\assume{\state\sigma_{c_2}};c_2\instr}\,\pkt
+    & \text{By Commutativity of }\havoc\\
+    = \denote{(\havoc{\vec{\rho_{t_1}}}; \havoc{\vec\gamma_{t_1}}; \assume{\state\sigma_{c_1}};c_1\instr);
+      (\havoc{\vec{\rho_{t_2}}}; \havoc{\vec\gamma_{t_2}};\assume{\state\sigma_{c_2}};c_2\instr)}\,\pkt
+    & \text{BY FREE-VARIABLE HAVOC COMMUTE LEMMA}\\
+    = \denote{\havoc{\vec{\rho_{t_2}}}; \havoc{\vec\gamma_{t_2}};\assume{\state\sigma_{c_2}};c_2\instr}^\dagger
+    \left(\denote{\havoc{\vec{\rho_{t_1}}}; \havoc{\vec\gamma_{t_1}}; \assume{\state\sigma_{c_1}};c_1\instr}\,\pkt\right)
+    & \text{BY DEFINITION} \\
+    =_V {\denote{c_2}^\sigma}^\dagger \left(\denote{\havoc{\vec{\rho_{t_1}}}; \havoc{\vec\gamma_{t_1}}; \assume{\state\sigma_{c_1}};c_1\instr}\,\pkt\right)
+    & \text{by IH}(c_2) \text{ since } \fvs(c_2) \subseteq \fvs(c_1;c_2) \subseteq V \\
+    =_V {\denote{c_2}^\sigma}^\dagger \left({\denote{c_1}^\sigma}^\dagger\pkt\right)
+    & \text{by IH}(c_1) \text{ since } \fvs(c_1) \subseteq \fvs(c_1;c_2) \subseteq V \\
+    = \denote{c_1;c_2}^\sigma\pkt
+    & \text{by definition of }\denote{-}^\sigma\\
+    \end{array}\]
+
+    THE RESULT FOLLOWS BY EQUALITY IMPLIES PROJECTED EQUALITY LEMMA
+
+  \item[$(c = c_1\choice c_2)$] Let $\vec{\rho_t}$ be the symbolic rows for tables $t
+    \in \tables(c_1;c_2)$. We can write $\vec\rho_t$ as $\vec{\rho_t} = \vec{\rho_{t_1}} \vec{\rho_{t_2}}$
+    where $\vec{\rho_{t_i}}$ is the set of symbolic rows in $c_i$. Let
+    $\vec{\gamma_t}$ be the ghost variables rows for tables $t \in
+    \tables(c_1;c_2)$. We can write $\vec{\gamma_t}$ as $\vec{\gamma_t} = \vec{\gamma_{t_1}} \vec{\gamma_{t_2}}$
+    where $\vec{\gamma_{t_i}}$ is the set of ghost variables in $c_i$. Compute as
+    follows:
+    \[\begin{array}{ll}
+    \denote{\havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}};\assume{\state\sigma_{c_1\choice c_2}}; (c_1 \choice c_2)\instr}\,\pkt\\
+    & \\
+    = \denote{\havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}};\assume{\state\sigma_{c_1\choice c_2}}; (c_1\instr \choice c_2\instr)}\,\pkt\\
+    & \text{by definition of }(\cdot)\instr\\
+    = \denote{\havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}};\assume{\state\sigma_{c_1} \wedge \state\sigma_{c_2}}; (c_1\instr \choice c_2\instr)}\,\pkt
+    & \text{BY LEMMA RELATING [] and STATE}\\
+    = \denote{\havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; (c_1\instr  \choice   c_2\instr)}\,\pkt
+    & \text{BY WEDGE-ASSUME LEMMA}\\
+    = \denote{\havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; c_1\instr
+      \choice \havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
+    & \text{BY PRODUCTS(;) AND SUMS([]) LEMMA} \\
+    = \denote{\havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_1}}; c_1\instr
+      \choice \havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
+    & \text{BY UNUSED-VARIABLES -> TRIVIAL ASSUMPTIONS LEMMA} \\
+    = \denote{\havoc{\vec{\rho_{t_1}}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_1}}; c_1\instr
+      \choice \havoc{\vec{\rho_{t_2}}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
+    & \text{BY UNUSEDVARIABLE HAVOC ELIMINATION LEMMA } \\
+    = \denote{\havoc{\vec{\rho_{t_1}}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_1}}; c_1\instr}\,\pkt
+    \cup \denote{\havoc{\vec{\rho_{t_2}}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
+    & \text{by definition of }\denote{\cdot} \\
+    =_V \denote{c_1}^\sigma\pkt \cup \denote{\havoc{\vec{\rho_{t_2}}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
+    & \text{By IH}(c_1) \text{ since } \fvs(c_1) \subseteq \fvs(c_1 \choice c_2) \subseteq V \\
+    =_V \denote{c_1}^\sigma\pkt \cup \denote{c_2}^\sigma\pkt
+    & \text{By IH}(c_2) \text{ since } \fvs(c_2) \subseteq \fvs(c_1 \choice c_2) \subseteq V \\
+    = \denote{c_1 \choice c_2}^\sigma\pkt
+    & \text{by definition of }\denote{\cdot}^\sigma
+    \end{array}
+    \]
+
+    THE RESULT FOLLOWS BY EQUALITY IMPLIES PROJECTED EQUALITY LEMMA
+
+  \item[$(c = t.\apply())$] Let $\pkt$ and $\sigma$ be arbitrary. Let
+    $[r_1,\ldots, r_n] = \sigma(t)$. There are two cases, corresponding to hit
+    and miss.
+    \begin{enumerate}
+    \item[\textit{Case 1}.] Assume that there is some $1 \leq i \leq n$ such
+      that $\pkt \models t.\matchrow(r_i)$. In fact, assume that $i$ is the
+      smallest such $i$. Then, by definition, $\denote{t.\apply()}^\sigma\,\pkt
+      = \denote{t.\action[r_i.\action](r_i.\data)}^\sigma\,\pkt$. Show that the
+      RHS is $V$-equal to this set. Let $P = \{\pkt\{\rho_t.* \mapsto
+      v\}\{\gamma_t.* \mapsto v' \mid v,v' \in \Value\} \}$ \todo{clean this up}.
+      Compute as follows:
+
+      \[\begin{array}{lcl}
+      \denote{\havoc \rho_t; \havoc \gamma_t; \assume {\state\sigma_{t.\apply()}}; t.\apply\instr}\,\pkt
+      & \\
+      = \denote{\assume {\state\sigma_{t.\apply()}}; t.\apply()\instr}^\dagger P
+      & \text{by definition of }\denote{\cdot}^\sigma \\
+      = \denote{t.\apply()\instr}^\dagger \{\pkt \mid \pkt \in P, \pkt \models \state\sigma_{t.\apply()} \}
+      & \text{by definition of }\denote{\cdot}^\sigma \\
+      =\denote{t.\apply()\instr}^\dagger
+      \left\{\pkt\,\,
+      \begin{array}{|l}
+        \pkt(\gamma_t.\miss) = 1 \Leftrightarrow \pkt(\gamma_t.\hitAction) > |\sigma(t)| \\
+        \pkt \models \displaystyle\bigvee_{r_i \in \sigma(t)}
+        \begin{array}{l}
+          \rho = r_i\\
+          \wedge (\gamma_t.\hitAction \geq i\\
+          \phantom\wedge \Rightarrow\bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)[\gamma_t.\keys / t.\keys]))
+        \end{array}
+      \end{array}
+      \right\}
+      & \text{by definition of } \models \text{ and }\state\sigma_{t.\apply()} \\
+      = \denote{  \begin{array}{l} \displaystyle
+          \assume t.\matchrow(\rho_t); \\
+          \assume \gamma_t.\keys = t.\keys; \\
+          \displaystyle \bigchoice_{a=0}^{|t.\action| -1}
+          \left(\begin{array}{l}
+            \assume \rho_t.\action = a;\\
+            \assume \gamma_t.\hitAction = \rho_t.\id;\\
+            t.\action[a](\rho_t.\data)
+          \end{array}\right)
+      \end{array}}
+      \left\{\pkt \in P\,\,
+      \begin{array}{|l}
+        \pkt(\gamma_t.\miss) = 1 \Leftrightarrow \pkt(\gamma_t.\hitAction) > |\sigma(t)|, \\
+        \pkt \models \displaystyle\bigvee_{r_i \in \sigma(t)}
+        \begin{array}{l}
+          \rho = r_i\\
+          \wedge (\gamma_t.\hitAction \geq i\\
+          \phantom\wedge \Rightarrow\bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)[\gamma_t.\keys / t.\keys]))
+        \end{array}
+      \end{array}
+      \right\}
+      & \text{by definition of } (\cdot)\instr \\
+      = \denote{  \begin{array}{l} \displaystyle
+          \displaystyle \bigchoice_{a=0}^{|t.\action| -1}
+          \left(\begin{array}{l}
+            \assume \rho_t.\action = a;\\
+            \assume \gamma_t.\hitAction = \rho_t.\id;\\
+            t.\action[a](\rho_t.\data)
+          \end{array}\right)
+      \end{array}}
+      \left\{\pkt \in P\,\,
+      \begin{array}{|l}
+        \pkt(\gamma_t.\miss) = 1 \Leftrightarrow \pkt(\gamma_t.\hitAction) > |\sigma(t)|, \\
+        \pkt \models \displaystyle\bigvee_{r_i \in \sigma(t)}
+        \begin{array}{l}
+          \rho = r_i\\
+          \wedge (\gamma_t.\hitAction \geq i\\
+          \phantom\wedge \Rightarrow\bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)[\gamma_t.\keys / t.\keys]))
+        \end{array} \\
+        \pkt \models t.\matchrow(\rho_t), \\
+        \pkt(\gamma_t.\keys) = \pkt(t.\keys) \\
+      \end{array}
+      \right\}
+      & \text{by definition of } \denote{\cdot} \& \models \\
+      = \displaystyle \bigcup_{a=0}^{|t.\action| -1}
+      \denote{t.\action[a](\rho_t.\data)}
+      \left\{\pkt \in P\,\,
+      \begin{array}{|l}
+        \pkt(\gamma_t.\miss) = 1 \Leftrightarrow \pkt(\gamma_t.\hitAction) > |\sigma(t)|, \\
+        \pkt \models \displaystyle\bigvee_{r_i \in \sigma(t)}
+        \begin{array}{l}
+          \rho = r_i\\
+          \wedge (\gamma_t.\hitAction \geq i\\
+          \phantom\wedge \Rightarrow\bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)[\gamma_t.\keys / t.\keys]))
+        \end{array} \\
+        \pkt \models t.\matchrow(\rho_t), \\
+        \pkt(\gamma_t.\keys) = \pkt(t.\keys), \\
+        \pkt(\rho_t.\action) = a,\\
+        \pkt(\gamma_t.\hitAction) = \pkt(\rho_t.\id),\\
+      \end{array}
+      \right\}
+      & \text{by definition of } \denote{\cdot} \& \models \\
+
+      = \displaystyle \bigcup_{a=0}^{|t.\action| -1}
+      \denote{t.\action[a](\rho_t.\data)}
+      \left\{\pkt \in P\,\,
+      \begin{array}{|l}
+        \pkt(\gamma_t.\miss) = 1 \Leftrightarrow \pkt(\gamma_t.\hitAction) > |\sigma(t)|, \\
+        \pkt \models \displaystyle\bigvee_{r_i \in \sigma(t)}
+        \begin{array}{l}
+          \rho = r_i\\
+          \wedge (\gamma_t.\hitAction \geq i\Rightarrow\bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)))
+        \end{array} \\
+        \pkt \models t.\matchrow(\rho_t), \\
+        \pkt(\gamma_t.\keys) = \pkt(t.\keys), \\
+        \pkt(\rho_t.\action) = a,\\
+        \pkt(\gamma_t.\hitAction) = \pkt(\rho_t.\id),\\
+      \end{array}
+      \right\}
+      & \text{since } \pkt(\gamma_t.\keys) = \pkt(t.\keys), \\
+      =\displaystyle \bigcup_{a=0}^{|t.\action| -1}
+      \denote{t.\action[a](\rho_t.\data)}
+      \left\{\pkt \in P\,\,
+      \begin{array}{|l}
+        \pkt(\gamma_t.\miss) = 1 \Leftrightarrow \pkt(\gamma_t.\hitAction) > |\sigma(t)|, \\
+        \pkt \models \displaystyle\bigvee_{r_i \in \sigma(t)}
+        \begin{array}{l}
+          \rho = r_i\\
+          \wedge t.\matchrow(r_i) \\
+          \wedge r_i.\action = a \\
+          \wedge \gamma_t.\hitAction = i \\
+          \wedge (\gamma_t.\hitAction \geq i\Rightarrow\bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)))
+        \end{array} \\
+        \pkt(\gamma_t.\keys) = \pkt(t.\keys), \\
+      \end{array}
+      \right\}
+      & \text{by taut} \\
+      =\displaystyle \bigcup_{a=0}^{|t.\action| -1}
+      \denote{t.\action[a](\rho_t.\data)}
+      \left\{\pkt \in P\,\,
+      \begin{array}{|l}
+        \pkt(\gamma_t.\miss) = 1 \Leftrightarrow \pkt(\gamma_t.\hitAction) > |\sigma(t)|, \\
+        \pkt \models \displaystyle\bigvee_{r_i \in \sigma(t)}
+        \begin{array}{l}
+          \rho = r_i\\
+          \wedge t.\matchrow(r_i) \\
+          \wedge r_i.\action = a \\
+          \wedge \gamma_t.\hitAction = i \\
+          \wedge \bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j))
+        \end{array} \\
+        \pkt(\gamma_t.\keys) = \pkt(t.\keys), \\
+      \end{array}
+      \right\}
+      & \text{by modus ponens} \\
+      =\displaystyle \bigcup_{a=0}^{|t.\action| -1}
+      \denote{t.\action[a](\rho_t.\data)}
+      \left\{\pkt \in P\,\,
+      \begin{array}{|l}
+        \pkt(\gamma_t.\miss) = 1 \Leftrightarrow \pkt(\gamma_t.\hitAction) > |\sigma(t)|, \\
+        \pkt(\rho) = r_i, \\
+        \pkt \models t.\matchrow(r_i), \\
+        r_i.\action = a, \\
+        \gamma_t.\hitAction = i, \\
+        \pkt \models \bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)), \\
+        \pkt(\gamma_t.\keys) = \pkt(t.\keys), \\
+      \end{array}
+      \right\}
+      & \text{by assumption about } i, \text{ and definition of } \models
+      \end{array}\]
+      \[\begin{array}{ll}
+      =\denote{t.\action[r_i.\action](r_i.\data)}
+      \left\{\pkt \in P\,\,
+      \begin{array}{|l}
+        \pkt(\gamma_t.\miss) = 1 \Leftrightarrow \pkt(\gamma_t.\hitAction) > |\sigma(t)|, \\
+        \pkt(\rho) = r_i, \\
+        \pkt \models t.\matchrow(r_i), \\
+        r_i.\action = a, \\
+        \gamma_t.\hitAction = i, \\
+        \pkt \models \bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)), \\
+        \pkt(\gamma_t.\keys) = \pkt(t.\keys), \\
+      \end{array}
+      \right\}
+      & \text{by definition of} \denote{\cdot}\\
+      =_V \denote{t.\action[r_i.\action](r_i.\data)} \pkt
+      & \text{SINCE ACTIONS ARE INSTRUMENTATION FREE AND } V \text{CONTAINS NO INSTRUMENATION VARIABLES} \\
+      =_V \denote{t.\action[r_i.\action](r_i.\data)}^\sigma\pkt
+      & \text{by IH}(t.\action[r_i.\action](r_i.\data)) \\
+      \end{array}\]
+
+    \item[\text{Case 2}.] \todo[inline]{Add defaults.}
+
+    \end{enumerate}
+
+  \end{enumerate}
+\end{proof}
+
+\subsection{Generation of Verification Conditions}
 
@@
     && b[e/h.f] \\
-    %% \WP(\havoc \rho, b)
-    %% && \forall \rho.\,b \\
+    \WP(\havoc x, b)
+    && \forall x.\,b \\
     \WP(\assert b_a, b)
@@
 
-\begin{conjecture}
-  For every $c \in \Cmd$, and $\sigma \in \State$, the following are equivalent:
-  \begin{itemize}
-  \item \(\state\sigma_c \Rightarrow  \WP(c\instr,\TRUE)\)
-  \item \(\WP(c[\sigma], \TRUE)\)
-  \end{itemize}
-\end{conjecture}
+\begin{theorem}[VC Correctness]
+  Given a set of variables $V$, for every $c \in \Cmd$ such that $\fvs(c) = V$,
+  and $\sigma \in \State$, the following formula is valid:
+  \[ \WP(c[\sigma], \TRUE) \iff (\forall \vec{\rho_t}. \forall \vec{\gamma_t}. \state\sigma_c \Rightarrow  \WP(c\instr,\TRUE)\]
+\end{theorem}
+
+\begin{proof}
+  Let $c$ and $\sigma$ be arbitary.
+  Theorem~\ref{thm:instr-equiv} gives the equivalence
+  \[\forall \pkt. \denote{c}^\sigma\,\pkt =_{\fvs(c)} \denote{\havoc \vec{\rho_t}; \havoc{\vec{\gamma_t}}; \assume{\state\sigma_c}; c\instr}\,\pkt\]
+
+  By Theorem~\ref{thm:subst}, we get
+
+  \[\forall \pkt.\denote{c[\sigma]}\,\pkt =_{\fvs(c)}  \denote{\havoc \vec{\rho_t}; \havoc{\vec{\gamma_t}}; \assume{\state\sigma_c}; c\instr}\,\pkt\]
+
+  By Dijkstra '75, we can see
+
+  \[\WP(c[\sigma], \TRUE) \iff  \WP(\havoc \vec{\rho_t}; \havoc{\vec{\gamma_t}}; \assume{\state\sigma_c}; c\instr, \TRUE))\]
+
+  By definition of $\WP$, this is equivalent to
+
+  \[\WP(c[\sigma],\TRUE) \iff \left(\forall \vec{\rho_t}.\forall \vec{\gamma_t}.\state\sigma_c \Rightarrow \WP(c\instr, \TRUE)\right)\]
+
+\end{proof}
 
@@
 
-\section{Quantifier Elimination?}
+\section{Abduction}
 
@@
 
-\paragraph{Relation to \texttt{bf4}} The \texttt{bf4} algorithm is similar to
+\paragraph{Relation to \texttt{bf4}}  \texttt{bf4} algorithm is similar to
 \textsc{VPreGen}, using formulae rather than examples. Rather than learn features, \texttt{bf4} uses a static set of

### 2021-05-25 inference.tex
finish sketching proof

@@
 \paragraph{Commands}
-Data plane programs $c \in \mathsf{Cmd}$ can be assignments ($h.f := e$),
+Data plane programs $c \in \mathsf{Cmd}$ can be assignments ($h.f := e$), havoc statements ($\havoc x$),
 assumptions ($\assume b$), assertions ($\assert b$), table application
@@
 
-We define $\SKIP$ as syntactic sugar for $\assume \TRUE$.
+We define $\SKIP$ as syntactic sugar for $\assume \TRUE$. We also lift havoc to
+sets of variables $\havoc \{x_1,\ldots,x_n\} = \SKIP; \havoc{x_1}; \cdots; \havoc{x_n}$.
+
+It is sensible to use sets here (rather than lists), because havocs commute:
+
+\begin{lemma}[Havocs Commute]
+  For any two variables $\{x_1,x_2\} \subseteq \Var$,
+  \[\forall \sigma.\forall \pkt. \denote{\havoc{x_1};\havoc{x_2}}^\sigma\pkt = \denote{\havoc{x_2};\havoc{x_2}}^\sigma\pkt \]
+\end{lemma}
+
+\begin{proof}
+  Let $x_1, x_2,\sigma,\pkt$ be given.
+  Compute as follows:
+  \[
+  \begin{array}{ll}
+    \denote{\havoc{x_1};\havoc{x_2}}^\sigma\pkt\\
+    = {\denote{\havoc{x_2}}^\sigma}^\dagger\left(\denote{\havoc{x_1}}\right)^\sigma\pkt
+    & \\
+    = {\denote{\havoc{x_2}}^\sigma}^\dagger\{\pkt\{x_1 \mapsto v_1\} \mid v_1 \in \Value\}
+    & \\
+    = \{\pkt\{x_1 \mapsto v_1\}\{x_2 \mapsto v_2\} \mid v_1,v_2 \in \Value\}
+    & \\
+    = \{\pkt\{x_2 \mapsto v_2\}\{x_1 \mapsto v_1\} \mid v_1,v_2 \in \Value\}
+    & \\
+    = {\denote{\havoc{x_1}}^\sigma}^\dagger\{\pkt\{x_2 \mapsto v_2\} \mid v_2 \in \Value\}
+    & \\
+    = {\denote{\havoc{x_1}}^\sigma}^\dagger\left(\denote{\havoc{x_2}}^\sigma\pkt\right)\}
+    & \\
+    = \denote{\havoc{x_2};\havoc{x_2}}^\sigma\pkt
+  \end{array}
+  \]
+\end{proof}
 
@@
     \left(\begin{array}{l}
+      \assume \gamma_t.\miss = 0;\\
       \assume \rho_t.\action = i;\\
@@
     \end{array}\right)\\
+    \choice (\assume \gamma_t.\miss = 1; \SKIP)  \\
   \end{array}}
@@
   \gamma_t.\miss = 1 \Leftrightarrow \gamma_t.\hitAction > |\sigma(t.\id)| \\
-  \wedge \bigvee_{r_i \in \sigma(t.\id)} \left(\rho = r_i
-  \wedge \left(\gamma_t.\hitAction \geq i \Leftrightarrow \bigwedge_{j=0}^{i-1} t.\matchrow(r_j)[\gamma_t.\keys/t.\keys]\right)\right)\\
+  \wedge \bigvee_{r_i \in \sigma(t.\id)} \rho = r_i  \\
+  \wedge \bigwedge_{r_i \in \sigma(t.\id)} \gamma_t.\hitAction = i \Leftrightarrow \rho_t = r_i \\
+  \wedge \bigwedge_{r_i \in \sigma(t.\id)} \left(\gamma_t.\hitAction \geq i \Leftrightarrow \bigwedge_{j=0}^{i-1} \neg t.\matchrow(r_j)[\gamma_t.\keys/t.\keys]\right)\\
   \end{array}\right)
@@
   Given a set of variables $V \subseteq \Var$ that contains no ghost variables
-  or symbolic variables. For every $c \in \Cmd$ such that $\fvs(c) \subseteq V$, well-formed state
-  $\sigma : \Value \to \Row^*$, and packet $\pkt \in \Pkt$,
-  %% \[\denote{c}^\sigma\pkt =_{\fvs(c)} \denote{\assume\state\sigma_c;c\instr}\,\pkt.\]
-
-  \[\denote{c}^\sigma\,\pkt =_V \denote{\havoc \vec{\rho_t}; \havoc \vec{\gamma_t}; \assume{\state\sigma}_t; c\instr}\,\pkt  \]
+  or symbolic variables. For every $c \in \Cmd$ such that $\fvs(c) \subseteq V$,
+  well-formed state $\sigma : \Value \to \Row^*$, and packet $\pkt \in \Pkt$,
+  where $R = \{\rho_t \mid t \in \tables(c)\}$, and $G = \{\gamma_t \mid t \in
+  \tables(c)\}$,
+
+  \[\denote{c}^\sigma\,\pkt =_V \denote{\havoc {R \cup G}; \assume{\state\sigma}_t; c\instr}\,\pkt  \]
 \end{theorem}
@@
   \begin{enumerate}[align=left]
-  \item[$(c = x:=e)$] Immediate.
-  \item[$(c = \havoc x)$] Immediate.
-  \item[$(c = \assume b)$] Immediate.
-  \item[$(c = \assert b)$] Immediate.
-  \item[$(c = c_1;c_2)$] Let $\vec{\rho_t}$ be the symbolic rows for tables $t
-    \in \tables(c_1;c_2)$. We can write $\vec\rho_t$ as $\vec{\rho_t} = \vec{\rho_{t_1}} \vec{\rho_{t_2}}$
-    where $\vec{\rho_{t_i}}$ is the set of symbolic rows in $c_i$. Let
-    $\vec{\gamma_t}$ be the ghost variables rows for tables $t \in
-    \tables(c_1;c_2)$. We can write $\vec{\gamma_t}$ as $\vec{\gamma_t} = \vec{\gamma_{t_1}} \vec{\gamma_{t_2}}$
-    where $\vec{\gamma_{t_i}}$ is the set of ghost variables in $c_i$. Compute as
-    follows:
+  \item[$(c = x:=e)$] Immediate, since $R = \emptyset = G$, $\assume{\state\sigma}_{x:=e} = \TRUE$, and $(x:=e)\instr = (x:=e)$.
+  \item[$(c = \havoc x)$] \textit{sim}.
+  \item[$(c = \assume b)$] \textit{sim}.
+  \item[$(c = \assert b)$] \textit{sim}.
+  \item[$(c = c_1;c_2)$] Let $R_i = \{\rho_t \mid t \in \tables(c_i)\}$ for $i =
+    1,2$. Notice that $R = R_1 \cup R_2$. Similarly, let $G_i = \{\gamma_t \mid
+    t \in \tables(c_i)\}$ for $i = 1,2$. Notice that $G = G_1 \cup G_2$. Compute
+    as follows:
     \[\begin{array}{ll}
-    \denote{\havoc{\vec{\rho_t}}; \havoc{\vec\gamma_t}; \assume{\state\sigma_{c_1;c_2}};(c_1;c_2)\instr}\,\pkt
+    \denote{\havoc {R \cup G}; \assume{\state\sigma_{c_1;c_2}};(c_1;c_2)\instr}\,\pkt
     & \\
-    = \denote{\havoc{\vec{\rho_t}}; \havoc{\vec\gamma_t}; \assume{\state\sigma_{c_1;c_2}}; c_1\instr;c_2\instr}\,\pkt
+    = \denote{\havoc{R\cup G}; \assume{\state\sigma_{c_1;c_2}}; c_1\instr;c_2\instr}\,\pkt
     & \text{definition of }(\cdot)\instr \\
-    = \denote{\havoc{\vec{\rho_t}}; \havoc{\vec\gamma_t}; \assume{\state\sigma_{c_1}};\assume{\state\sigma_{c_2}};c_1\instr;c_2\instr}\,\pkt
+    = \denote{\havoc{R\cup G}; \assume{\state\sigma_{c_1}};\assume{\state\sigma_{c_2}};c_1\instr;c_2\instr}\,\pkt
     & \text{LEMMAs INVOLVING SIGMA-DISTRIB \& ASSUME-DISTRIB} \\
-    = \denote{\havoc{\vec{\rho_{t_1}}}\havoc{\vec{\rho_{t_2}}}; \havoc{\vec\gamma_{t_1}}; \havoc{\vec\gamma_{t_2}}; \assume{\state\sigma_{c_1}};\assume{\state\sigma_{c_2}};c_1\instr;c_2\instr}\,\pkt
+    = \denote{\havoc{R_1};\havoc{G_1}; \havoc{R_2};\havoc{G_2}; \assume{\state\sigma_{c_1}};\assume{\state\sigma_{c_2}};c_1\instr;c_2\instr}\,\pkt
     & \text{by definition} \\
-    = \denote{\havoc{\vec{\rho_{t_1}}};\havoc{\vec{\rho_{t_2}}}; \havoc{\vec\gamma_{t_1}}; \havoc{\vec\gamma_{t_2}}; \assume{\state\sigma_{c_1}};c_1\instr;\assume{\state\sigma_{c_2}};c_2\instr}\,\pkt
+    = \denote{\havoc{R_1};\havoc{G_1}; \havoc{R_2}; \havoc{G_2}; \assume{\state\sigma_{c_1}};c_1\instr;\assume{\state\sigma_{c_2}};c_2\instr}\,\pkt
     & \text{BY FREE-VARIABLE-ASSUMPTION COMMUTATIVITY LEMMA} \\
-    = \denote{\havoc{\vec{\rho_{t_1}}}; \havoc{\vec\gamma_{t_1}};\havoc{\vec{\rho_{t_2}}}; \havoc{\vec\gamma_{t_2}}; \assume{\state\sigma_{c_1}};c_1\instr;\assume{\state\sigma_{c_2}};c_2\instr}\,\pkt
-    & \text{By Commutativity of }\havoc\\
-    = \denote{(\havoc{\vec{\rho_{t_1}}}; \havoc{\vec\gamma_{t_1}}; \assume{\state\sigma_{c_1}};c_1\instr);
-      (\havoc{\vec{\rho_{t_2}}}; \havoc{\vec\gamma_{t_2}};\assume{\state\sigma_{c_2}};c_2\instr)}\,\pkt
+    = \denote{(\havoc{R_1}; \havoc{G_1}; \assume{\state\sigma_{c_1}};c_1\instr);
+      (\havoc{R_2}; \havoc{G_2};\assume{\state\sigma_{c_2}};c_2\instr)}\,\pkt
     & \text{BY FREE-VARIABLE HAVOC COMMUTE LEMMA}\\
-    = \denote{\havoc{\vec{\rho_{t_2}}}; \havoc{\vec\gamma_{t_2}};\assume{\state\sigma_{c_2}};c_2\instr}^\dagger
-    \left(\denote{\havoc{\vec{\rho_{t_1}}}; \havoc{\vec\gamma_{t_1}}; \assume{\state\sigma_{c_1}};c_1\instr}\,\pkt\right)
+    = \denote{\havoc{R_2}; \havoc{G_2};\assume{\state\sigma_{c_2}};c_2\instr}^\dagger
+    \left(\denote{\havoc{R_1}; \havoc{G_1}; \assume{\state\sigma_{c_1}};c_1\instr}\,\pkt\right)
     & \text{BY DEFINITION} \\
@@
 
-  \item[$(c = c_1\choice c_2)$] Let $\vec{\rho_t}$ be the symbolic rows for tables $t
-    \in \tables(c_1;c_2)$. We can write $\vec\rho_t$ as $\vec{\rho_t} = \vec{\rho_{t_1}} \vec{\rho_{t_2}}$
-    where $\vec{\rho_{t_i}}$ is the set of symbolic rows in $c_i$. Let
-    $\vec{\gamma_t}$ be the ghost variables rows for tables $t \in
-    \tables(c_1;c_2)$. We can write $\vec{\gamma_t}$ as $\vec{\gamma_t} = \vec{\gamma_{t_1}} \vec{\gamma_{t_2}}$
-    where $\vec{\gamma_{t_i}}$ is the set of ghost variables in $c_i$. Compute as
-    follows:
+  \item[$(c = c_1\choice c_2)$] Let $R_i = \{\rho_t \mid t \in \tables(c_i)\}$ for $i =
+    1,2$. Notice that $R = R_1 \cup R_2$. Similarly, let $G_i = \{\gamma_t \mid
+    t \in \tables(c_i)\}$ for $i = 1,2$. Notice that $G = G_1 \cup G_2$. Compute
+    as follows:
     \[\begin{array}{ll}
-    \denote{\havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}};\assume{\state\sigma_{c_1\choice c_2}}; (c_1 \choice c_2)\instr}\,\pkt\\
+    \denote{\havoc{R \cup G};\assume{\state\sigma_{c_1\choice c_2}}; (c_1 \choice c_2)\instr}\,\pkt\\
     & \\
-    = \denote{\havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}};\assume{\state\sigma_{c_1\choice c_2}}; (c_1\instr \choice c_2\instr)}\,\pkt\\
+    = \denote{\havoc{R \cup G};\assume{\state\sigma_{c_1\choice c_2}}; (c_1\instr \choice c_2\instr)}\,\pkt\\
     & \text{by definition of }(\cdot)\instr\\
-    = \denote{\havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}};\assume{\state\sigma_{c_1} \wedge \state\sigma_{c_2}}; (c_1\instr \choice c_2\instr)}\,\pkt
+    = \denote{\havoc{R \cup G};\assume{\state\sigma_{c_1} \wedge \state\sigma_{c_2}}; (c_1\instr \choice c_2\instr)}\,\pkt
     & \text{BY LEMMA RELATING [] and STATE}\\
-    = \denote{\havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; (c_1\instr  \choice   c_2\instr)}\,\pkt
+    = \denote{\havoc{R \cup G};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; (c_1\instr  \choice   c_2\instr)}\,\pkt
     & \text{BY WEDGE-ASSUME LEMMA}\\
-    = \denote{\havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; c_1\instr
-      \choice \havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
+    = \denote{\havoc{R \cup G};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; c_1\instr
+      \choice \havoc{R \cup G};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
     & \text{BY PRODUCTS(;) AND SUMS([]) LEMMA} \\
-    = \denote{\havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_1}}; c_1\instr
-      \choice \havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
+    = \denote{\havoc{R \cup G}; \assume{\state\sigma_{c_1}}; c_1\instr
+      \choice \havoc{R \cup G}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
     & \text{BY UNUSED-VARIABLES -> TRIVIAL ASSUMPTIONS LEMMA} \\
-    = \denote{\havoc{\vec{\rho_{t_1}}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_1}}; c_1\instr
-      \choice \havoc{\vec{\rho_{t_2}}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
+    =_V \denote{\havoc{R_1 \cup G_1}; \assume{\state\sigma_{c_1}}; c_1\instr
+      \choice \havoc{R_2 \cup G_2}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
     & \text{BY UNUSEDVARIABLE HAVOC ELIMINATION LEMMA } \\
-    = \denote{\havoc{\vec{\rho_{t_1}}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_1}}; c_1\instr}\,\pkt
-    \cup \denote{\havoc{\vec{\rho_{t_2}}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
+    = \denote{\havoc{R_1 \cup G_1}; \assume{\state\sigma_{c_1}}; c_1\instr}\,\pkt
+    \cup \denote{\havoc{R_2 \cup G_2}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
     & \text{by definition of }\denote{\cdot} \\
@@
   \item[$(c = t.\apply())$] Let $\pkt$ and $\sigma$ be arbitrary. Let
-    $[r_1,\ldots, r_n] = \sigma(t)$. There are two cases, corresponding to hit
-    and miss.
-    \begin{enumerate}
-    \item[\textit{Case 1}.] Assume that there is some $1 \leq i \leq n$ such
-      that $\pkt \models t.\matchrow(r_i)$. In fact, assume that $i$ is the
-      smallest such $i$. Then, by definition, $\denote{t.\apply()}^\sigma\,\pkt
-      = \denote{t.\action[r_i.\action](r_i.\data)}^\sigma\,\pkt$. Show that the
-      RHS is $V$-equal to this set. Let $P = \{\pkt\{\rho_t.* \mapsto
+    $[r_1,\ldots, r_n] = \sigma(t)$.
+
+    Let $P = \{\pkt\{\rho_t.* \mapsto
       v\}\{\gamma_t.* \mapsto v' \mid v,v' \in \Value\} \}$ \todo{clean this up}.
@@
       \[\begin{array}{lcl}
-      \denote{\havoc \rho_t; \havoc \gamma_t; \assume {\state\sigma_{t.\apply()}}; t.\apply\instr}\,\pkt
+      \denote{\havoc R \cup G; \assume {\state\sigma_{t.\apply()}}; t.\apply\instr}\,\pkt
       & \\
@@
       & \text{by definition of }\denote{\cdot}^\sigma \\
-      =\denote{t.\apply()\instr}^\dagger
-      \left\{\pkt\,\,
-      \begin{array}{|l}
-        \pkt(\gamma_t.\miss) = 1 \Leftrightarrow \pkt(\gamma_t.\hitAction) > |\sigma(t)| \\
-        \pkt \models \displaystyle\bigvee_{r_i \in \sigma(t)}
-        \begin{array}{l}
-          \rho = r_i\\
-          \wedge (\gamma_t.\hitAction \geq i\\
-          \phantom\wedge \Rightarrow\bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)[\gamma_t.\keys / t.\keys]))
-        \end{array}
-      \end{array}
-      \right\}
-      & \text{by definition of } \models \text{ and }\state\sigma_{t.\apply()} \\
       = \denote{  \begin{array}{l} \displaystyle
-          \assume t.\matchrow(\rho_t); \\
           \assume \gamma_t.\keys = t.\keys; \\
@@
           \left(\begin{array}{l}
+            \assume t.\matchrow(\rho_t); \\
+            \assume \gamma_t.\miss = 0;\\
             \assume \rho_t.\action = a;\\
@@
             t.\action[a](\rho_t.\data)
-          \end{array}\right)
+          \end{array}\right) \\
+          \choice (\assume \gamma_t.\miss = 1;\SKIP) \\
       \end{array}}
-      \left\{\pkt \in P\,\,
-      \begin{array}{|l}
-        \pkt(\gamma_t.\miss) = 1 \Leftrightarrow \pkt(\gamma_t.\hitAction) > |\sigma(t)|, \\
-        \pkt \models \displaystyle\bigvee_{r_i \in \sigma(t)}
-        \begin{array}{l}
-          \rho = r_i\\
-          \wedge (\gamma_t.\hitAction \geq i\\
-          \phantom\wedge \Rightarrow\bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)[\gamma_t.\keys / t.\keys]))
-        \end{array}
-      \end{array}
-      \right\}
-      & \text{by definition of } (\cdot)\instr \\
-      = \denote{  \begin{array}{l} \displaystyle
-          \displaystyle \bigchoice_{a=0}^{|t.\action| -1}
-          \left(\begin{array}{l}
-            \assume \rho_t.\action = a;\\
-            \assume \gamma_t.\hitAction = \rho_t.\id;\\
-            t.\action[a](\rho_t.\data)
-          \end{array}\right)
-      \end{array}}
-      \left\{\pkt \in P\,\,
-      \begin{array}{|l}
-        \pkt(\gamma_t.\miss) = 1 \Leftrightarrow \pkt(\gamma_t.\hitAction) > |\sigma(t)|, \\
-        \pkt \models \displaystyle\bigvee_{r_i \in \sigma(t)}
-        \begin{array}{l}
-          \rho = r_i\\
-          \wedge (\gamma_t.\hitAction \geq i\\
-          \phantom\wedge \Rightarrow\bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)[\gamma_t.\keys / t.\keys]))
-        \end{array} \\
-        \pkt \models t.\matchrow(\rho_t), \\
-        \pkt(\gamma_t.\keys) = \pkt(t.\keys) \\
-      \end{array}
-      \right\}
-      & \text{by definition of } \denote{\cdot} \& \models \\
-      = \displaystyle \bigcup_{a=0}^{|t.\action| -1}
+      \{\pkt \mid \pkt \in P, \pkt \models \state\sigma_{t.\apply()} \}
+      & \text{by definition of }(\cdot)\instr \\
+      = \displaystyle\bigcup_{a=0}^{|t.\action| - 1}
       \denote{t.\action[a](\rho_t.\data)}
-      \left\{\pkt \in P\,\,
-      \begin{array}{|l}
-        \pkt(\gamma_t.\miss) = 1 \Leftrightarrow \pkt(\gamma_t.\hitAction) > |\sigma(t)|, \\
-        \pkt \models \displaystyle\bigvee_{r_i \in \sigma(t)}
-        \begin{array}{l}
-          \rho = r_i\\
-          \wedge (\gamma_t.\hitAction \geq i\\
-          \phantom\wedge \Rightarrow\bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)[\gamma_t.\keys / t.\keys]))
-        \end{array} \\
-        \pkt \models t.\matchrow(\rho_t), \\
-        \pkt(\gamma_t.\keys) = \pkt(t.\keys), \\
-        \pkt(\rho_t.\action) = a,\\
-        \pkt(\gamma_t.\hitAction) = \pkt(\rho_t.\id),\\
-      \end{array}
-      \right\}
-      & \text{by definition of } \denote{\cdot} \& \models \\
-
-      = \displaystyle \bigcup_{a=0}^{|t.\action| -1}
-      \denote{t.\action[a](\rho_t.\data)}
-      \left\{\pkt \in P\,\,
-      \begin{array}{|l}
-        \pkt(\gamma_t.\miss) = 1 \Leftrightarrow \pkt(\gamma_t.\hitAction) > |\sigma(t)|, \\
-        \pkt \models \displaystyle\bigvee_{r_i \in \sigma(t)}
-        \begin{array}{l}
-          \rho = r_i\\
-          \wedge (\gamma_t.\hitAction \geq i\Rightarrow\bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)))
-        \end{array} \\
-        \pkt \models t.\matchrow(\rho_t), \\
-        \pkt(\gamma_t.\keys) = \pkt(t.\keys), \\
-        \pkt(\rho_t.\action) = a,\\
-        \pkt(\gamma_t.\hitAction) = \pkt(\rho_t.\id),\\
-      \end{array}
-      \right\}
-      & \text{since } \pkt(\gamma_t.\keys) = \pkt(t.\keys), \\
-      =\displaystyle \bigcup_{a=0}^{|t.\action| -1}
-      \denote{t.\action[a](\rho_t.\data)}
-      \left\{\pkt \in P\,\,
-      \begin{array}{|l}
-        \pkt(\gamma_t.\miss) = 1 \Leftrightarrow \pkt(\gamma_t.\hitAction) > |\sigma(t)|, \\
-        \pkt \models \displaystyle\bigvee_{r_i \in \sigma(t)}
-        \begin{array}{l}
-          \rho = r_i\\
-          \wedge t.\matchrow(r_i) \\
-          \wedge r_i.\action = a \\
-          \wedge \gamma_t.\hitAction = i \\
-          \wedge (\gamma_t.\hitAction \geq i\Rightarrow\bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)))
-        \end{array} \\
-        \pkt(\gamma_t.\keys) = \pkt(t.\keys), \\
-      \end{array}
-      \right\}
-      & \text{by taut} \\
-      =\displaystyle \bigcup_{a=0}^{|t.\action| -1}
-      \denote{t.\action[a](\rho_t.\data)}
-      \left\{\pkt \in P\,\,
-      \begin{array}{|l}
-        \pkt(\gamma_t.\miss) = 1 \Leftrightarrow \pkt(\gamma_t.\hitAction) > |\sigma(t)|, \\
-        \pkt \models \displaystyle\bigvee_{r_i \in \sigma(t)}
-        \begin{array}{l}
-          \rho = r_i\\
-          \wedge t.\matchrow(r_i) \\
-          \wedge r_i.\action = a \\
-          \wedge \gamma_t.\hitAction = i \\
-          \wedge \bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j))
-        \end{array} \\
-        \pkt(\gamma_t.\keys) = \pkt(t.\keys), \\
-      \end{array}
-      \right\}
-      & \text{by modus ponens} \\
-      =\displaystyle \bigcup_{a=0}^{|t.\action| -1}
-      \denote{t.\action[a](\rho_t.\data)}
-      \left\{\pkt \in P\,\,
-      \begin{array}{|l}
-        \pkt(\gamma_t.\miss) = 1 \Leftrightarrow \pkt(\gamma_t.\hitAction) > |\sigma(t)|, \\
-        \pkt(\rho) = r_i, \\
-        \pkt \models t.\matchrow(r_i), \\
-        r_i.\action = a, \\
-        \gamma_t.\hitAction = i, \\
-        \pkt \models \bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)), \\
-        \pkt(\gamma_t.\keys) = \pkt(t.\keys), \\
-      \end{array}
-      \right\}
-      & \text{by assumption about } i, \text{ and definition of } \models
+      \left\{\pkt \in P\,\, \begin{array}{|ll}
+        \pkt \models \state\sigma_{t.\apply()}, & \pkt \models t.\matchrow(\rho_t)\\
+        \pkt \models \gamma_t.\keys = t.\keys, & \pkt \models \rho_t.\action = a \\
+        \multicolumn{2}{|l}{\pkt \models \gamma_t.\hitAction = \rho_t.\id}
+      \end{array}\right\}
+      & \text{by definition of }\denote{\cdot}\\
+      \qquad \cup\,\,\{\pkt \in P \mid \pkt \models \state\sigma_{t.\apply()} \wedge \gamma_t.\keys = t.\keys \wedge \gamma_t.\miss = 1 \}
       \end{array}\]
-      \[\begin{array}{ll}
-      =\denote{t.\action[r_i.\action](r_i.\data)}
-      \left\{\pkt \in P\,\,
-      \begin{array}{|l}
-        \pkt(\gamma_t.\miss) = 1 \Leftrightarrow \pkt(\gamma_t.\hitAction) > |\sigma(t)|, \\
-        \pkt(\rho) = r_i, \\
-        \pkt \models t.\matchrow(r_i), \\
-        r_i.\action = a, \\
-        \gamma_t.\hitAction = i, \\
-        \pkt \models \bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)), \\
-        \pkt(\gamma_t.\keys) = \pkt(t.\keys), \\
-      \end{array}
-      \right\}
-      & \text{by definition of} \denote{\cdot}\\
-      =_V \denote{t.\action[r_i.\action](r_i.\data)} \pkt
-      & \text{SINCE ACTIONS ARE INSTRUMENTATION FREE AND } V \text{CONTAINS NO INSTRUMENATION VARIABLES} \\
-      =_V \denote{t.\action[r_i.\action](r_i.\data)}^\sigma\pkt
-      & \text{by IH}(t.\action[r_i.\action](r_i.\data)) \\
-      \end{array}\]
-
-    \item[\text{Case 2}.] \todo[inline]{Add defaults.}
-
+
+       There are two cases, corresponding to hit and miss. Let $M$ be the `miss
+       set' i.e.
+       \[M \triangleq \{\pkt \in P \mid \pkt \models \state\sigma_{t.\apply()} \wedge \gamma_t.\keys = t.\keys \wedge \gamma_t.\miss = 1 \},\]
+
+       and let $H_a$ be the input `hit-set' for action $a$, i.e.
+      \[ H_a \triangleq \left\{\pkt \in P\,\, \begin{array}{|ll}
+        \pkt \models \state\sigma_{t.\apply()}, & \pkt \models t.\matchrow(\rho_t)\\
+        \pkt \models \gamma_t.\keys = t.\keys, & \pkt \models \rho_t.\action = a \\
+        \multicolumn{2}{|l}{\pkt \models \gamma_t.\hitAction = \rho_t.\id}
+      \end{array}\right\}\]
+
+    \begin{enumerate}
+    \item[\textit{Case 1}.] Assume that there is some $1 \leq i \leq n$ such
+      that $\pkt \models t.\matchrow(r_i)$. In fact, assume that $i$ is the
+      smallest such $i$. Then, by definition, $\denote{t.\apply()}^\sigma\,\pkt
+      = \denote{t.\action[r_i.\action](r_i.\data)}^\sigma\,\pkt$. Show that
+      $M \cup \bigcup_{a=0}^{|t.\action|-1}\denote{t.\action[a](\rho_t.\data)} H_a$ is $V$-equal to this set.
+
+      Show that $M = \emptyset$, by showing that the refinement is contradictory.
+
+      Assume $\state\sigma_{t.\apply()}$, $\gamma_t.\keys = t.\keys$ and
+      $\gamma_t.\miss = 1$. First expand the definition of
+      $\state\sigma_{t.\apply()}$, which gives us $\gamma_t.\miss = 1
+      \Leftrightarrow \gamma_t.\hitAction > n$, $\bigvee_{i}^n \rho = r_i$,
+      $\bigwedge_{i=1}^n \gamma_t.\hitAction = i \Leftrightarrow \rho_t = r_i$,
+      and
+
+      \[\bigwedge_{i=1}^n \left(\gamma_t.\hitAction \geq i \Leftrightarrow \bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)[\gamma_t.\keys/t.\keys]\right)\]
+
+      Since we know $\gamma_t.\keys = t.\keys$, we can equivalently write
+      \[\bigwedge_{i=1}^n \left(\gamma_t.\hitAction \geq i \Leftrightarrow \bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)\right)\]
+
+      Since $\gamma_t.\miss = 1$ we know that $\gamma_t.\hitAction > n$, which
+      means that for all $1 \leq j \leq n$, $\neg t.\matchrow(r_j)$, which
+      contradicts the assumption that $t.\matchrow(r_j)$. Thus the refinement is
+      contradictory and $M = \emptyset$. \hfill \checkmark
+
+      Now, show that $\forall \pkt \in H_a. \rho_t = r_i$, by showing that the
+      refinement implies that $\rho_t = r_i \wedge \bigwedge_{j \neq i}\rho_t \neq r_j$. We assume the following facts:
+      \[ \state\sigma_{t.\apply()} \wedge \gamma_t = t.\keys \wedge \gamma.\hitAction = \rho_t.\id \wedge t.\matchrow(\rho_t) \wedge
+      \rho_t.\action = a\]
+
+      By unfolding the definition of $\state\sigma_{t.\apply()}$, we also know that
+      $\gamma.\miss = 1 \Leftrightarrow \gamma_t.\hitAction > n$, $\bigvee_{i=1}^n \rho = r_i$, $\bigwedge_{i=1}^n \gamma_t.\hitAction = i \Rightarrow \rho_t = r_i$, and
+      \[\bigwedge_{i=1}^n \left( \gamma_t.\hitAction \geq i \Leftrightarrow \bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)[\gamma_t.\keys/t.\keys] \right)\]
+
+      As before, since $\gamma_t.\keys = t.\keys$ we can write this as
+      \[\bigwedge_{i=1}^n \left( \gamma_t.\hitAction \geq i \Leftrightarrow \bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)\right)\]
+
+      By assumption, we know that $t.\matchrow(r_i)$, and for $1 \leq j < i$,
+      $\neg t.\matchrow(r_j)$. Combining this with the above fact gives
+      $\gamma_t.\hitAction = i$. Then since $\bigwedge_{i=1}^n
+      \gamma_t.\hitAction = i \Leftrightarrow \rho_t = r_i$,
+      conclude $\rho_t = r_i \wedge \bigwedge_{i\neq j} \rho_t\neq r_i$. \hfill \checkmark
+
+      Now, we can write
+      \[\bigcup_{a=0}^{|t.\action| - 1}\denote{t.\action[a](\rho_t.\data)}^\dagger H_a
+       = \denote{t.\action[r_i.\action](r_i.\data)}^\dagger H_a
+       \]
+
+       Which is $V$-equivalent to $\denote{t.\action[r_i.\action](r_i.\data)}\,\pkt$. \todo{PROVE THIS LEMMA}
+       The result follows by the IH on $t.\action[r_i.\action](r_i.\data)$.
+
+     \item[\text{Case 2}.]Assume that there is no $1 \leq i \leq n$ such that
+       $\pkt \models t.\matchrow(r_i)$. Then, by definition,
+       $\denote{t.\apply()}^\sigma\,\pkt = \{\pkt\}$. Show that
+       $M \cup\bigcup_{a=0}^{|t.\action|-1}\denote{t.\action[a](\rho_t.\data)} H_a$ is $V$-equal to this $\{\pkt\}$.
+
+        Let $\pkt' \in P$ such that $\pkt' \models \state\sigma_{t.\apply}$,
+        $\pkt' \models \gamma_t.\keys$, and $\gamma_t.\miss =1$. Expanding the
+        definition of $\state\sigma_{t.\apply()}$ further we get that $\pkt'
+        \models \gamma_t.\miss = 1 \Leftrightarrow \gamma_t.\hitAction > n$, forall $0
+        \leq i \leq n$, $\pkt' \models \gamma_t.\hitAction = i \Leftrightarrow
+        \rho_t = r_i$, and $\pkt' \models \gamma_t.\hitAction \geq i \Leftrightarrow
+        \bigwedge_{j=1}^{i-1} \neg.\matchrow(r_j)$ (since $\pkt' \models
+        \gamma_t.\keys = t.\keys$).
+
+        Our assumption gives us that $\bigwedge_{j=1}^n \neg t.\matchrow(r_j)$,
+        which means that $\gamma_t.\hitAction > n$, and that $\gamma_t.\miss =
+        1$. Conclude that $H_a = \emptyset$ and $M \neq \emptyset$. Since $P =_V
+        \{\pkt\}$, and $M$ is a nonempty refinement of $P$\todo{PROVE THIS
+          LEMMA?}, conclude that $M =_V \{\pkt\}$. The result follows.
     \end{enumerate}
@@
 Now, we would like to be able to translate the result from
-Theorem~\ref{thm:instr-equiv} into a
-symbolic representation. To do this, we'll use the weakest precondition, defined
-in Figure~\ref{fig:wp}
+Theorem~\ref{thm:instr-equiv} into a symbolic representation. To do this, we'll
+use the weakest precondition, defined in Figure~\ref{fig:wp}
 
@@
 \begin{theorem}[VC Correctness]
+  \label{thm:vc-correct}
   Given a set of variables $V$, for every $c \in \Cmd$ such that $\fvs(c) = V$,
   and $\sigma \in \State$, the following formula is valid:
-  \[ \WP(c[\sigma], \TRUE) \iff (\forall \vec{\rho_t}. \forall \vec{\gamma_t}. \state\sigma_c \Rightarrow  \WP(c\instr,\TRUE)\]
+  \[ \forall V. \WP(c[\sigma], \TRUE) \iff \forall V. (\forall \vec{\rho_t}. \forall \vec{\gamma_t}. \state\sigma_c \Rightarrow  \WP(c\instr,\TRUE)\]
 \end{theorem}
@@
 
-  By Dijkstra '75, we can see
-
-  \[\WP(c[\sigma], \TRUE) \iff  \WP(\havoc \vec{\rho_t}; \havoc{\vec{\gamma_t}}; \assume{\state\sigma_c}; c\instr, \TRUE))\]
+  By Dijkstra, we can see
+
+  \[\forall V. \WP(c[\sigma], \TRUE) \iff \forall V. \WP(\havoc \vec{\rho_t}; \havoc{\vec{\gamma_t}}; \assume{\state\sigma_c}; c\instr, \TRUE))\]
 
@@
 
-  \[\WP(c[\sigma],\TRUE) \iff \left(\forall \vec{\rho_t}.\forall \vec{\gamma_t}.\state\sigma_c \Rightarrow \WP(c\instr, \TRUE)\right)\]
+  \[ \forall V. \WP(c[\sigma],\TRUE) \iff \forall V. \left(\forall \vec{\rho_t}.\forall \vec{\gamma_t}.\state\sigma_c \Rightarrow \WP(c\instr, \TRUE)\right)\]
 
@@
 
-We can observe that \textbf{Inference} is simply solvable by
-reduction to Quantifier Elimination. In fact. Quantifier elimination is one of
-the key driving forces of the Dilligs' abduction work (\textsc{Hola},
-\textsc{Explain}). We sketch the argument here.
-
-\begin{conjecture}[Quantifer Elimination]
-  The \textbf{\textrm{Inference Problem}} is reducable to Quantifer Elimination.
-\end{conjecture}
-
-\begin{proof}[Proof Sketch.]
-  Given a command $c \in \Cmd$. Compute the instrumented command $c\instr$. Let
-  $\overrightarrow{h.f}$ describe all the header fields occurring in $c\instr$. Write
-  $b[\overrightarrow{h.f}] = \WP(c\instr,\TRUE)$. Observe that $(\forall
-  \overrightarrow{h.f}.\;b[\overrightarrow{h.f}]) \Rightarrow
-  b[\overrightarrow{h.f}]$ is valid. Then the desired control plane formula is $\qe(\forall \overrightarrow{h.f}.\;b[\overrightarrow{h.f}])$.
+We can observe that \textbf{Inference} is simply solvable by reduction to
+Quantifier Elimination (QE). In fact, QE is one of the key driving forces of the
+Dilligs' abduction work (\textsc{Hola}, \textsc{Explain}). We sketch the
+argument here.
+
+In the following we assume a terminating function $\qe : \Var \times \BExpr \to
+\BExpr$ such that for $V \in \Var$ and $b \in \BExpr$, $\bfvs(\qe(V,b)) \cap V =
+\emptyset$ and $\qe(V,b) \Leftrightarrow \forall V. b$.
+
+\begin{theorem}[Quantifer Elimination]
+  For every instrumentation-free command $c \in \Cmd$, the formula $\qe(\fvs(c), \WP(c\instr, \TRUE))$ is a solution to the \textbf{Inference Problem}.
+\end{theorem}
+
+\begin{proof}
+  Let $c \in \Cmd$ be instrumentation-free. Let $R$ be the set of symbolic row
+  variables in $c\instr$, and let $G$ be the set of ghost variables in $c\instr$,
+  Let $\varphi = \qe(\fvs(c),\WP(c\instr, \TRUE))$. Since $\bfvs(\varphi) \cap
+  \fvs(c) = \emptyset$, it is trivial to show \todo{do it} that $\varphi$ is a
+  control plane formula.
+
+  Now, let $\sigma$ be such that $\state\sigma_c \Rightarrow \varphi$ is valid,
+  i.e. $\forall R. \forall G. \state\sigma_c \Rightarrow \varphi$ is valid. By
+  assumption about $\qe$, this is equivalent to $\forall R.\forall G.
+  \state\sigma_c \Rightarrow \WP(c\instr, \TRUE)$. By
+  Theorem~\ref{thm:vc-correct}, $\WP(c[\sigma],\TRUE)$ is valid. By Dijkstra,
+  $\forall \pkt \in \Pkt. \denote{c[\sigma]} \neq \error$. Conclude $\forall
+  \pkt \in \Pkt. \denote{c}^\sigma \neq \error$, by Theorem~\ref{thm:subst}.
+
 \end{proof}
-
-\paragraph{Discussion} This \emph{probably} doesn't work in general. Standard QE procedures are
-exponential at best, and enumerating all eth/ipv4 src/dst addresses already
-takes an intractable amount of time. It's also not clear (to me) that this
-solution is weakest.
-
-The Dilligs' line of work uses Presburger arithmetic QE to implement the
-quantifier elimination scheme. We could directly adopt this, however networking
-programs make heavy use of bitwise operators like XOR, so we need to take a
-different strategy.
 

### 2021-05-25 inference.tex
finish sketching proof

@@
   $\pkt' \in P'$ such that $\pkt =_S \pkt'$. Two sets of packets $P$ and $P'$
-  are equivalent up to $S$, written $P =_S P'$ iff $P \subseteq_S P'$ and $P \subseteq_S P'$.
+  are $S$-equivalent, written $P =_S P'$ iff $P \subseteq_S P'$ and $P \subseteq_S P'$.
 \end{definition}
+
+\begin{lemma}[Projection]
+  For $c \in \Cmd$, $\fvs(c) \subseteq V \subseteq \Var$ and $P \subseteq \Pkt$ such that $\forall
+  \pkt_1,\pkt_2 \in P, \pkt_1 =_V \pkt_2$, $\denote{c} P =_V \denote{c} P|_V$.
+
+\end{lemma}
 

### 2021-05-25 inference.tex
close proof holes

@@
       \{\pkt\} & \pkt \models b \\
-      \{\} & \textit{otherwise} \\
+      \{\} & \textit{otherwis
+e} \\
     \end{cases}\\
@@
 \end{figure}
+
+\subsection{Lemmas}
+
+\begin{lemma}[Assume-Currying]
+  \label{lem:assume-currying}
+  For $b_1,b_2 \in \BExpr$, $\sigma : \Value \to \Row^*$,
+  \[\forall \pkt. \denote{\assume {b_1\wedge b_2}}^\sigma \pkt = \denote{\assume {b_1};\assume {b_2}}^\sigma \pkt\]
+\end{lemma}
+\begin{proof}
+  Let $b_1,b_2,\sigma, \pkt$ be arbitrary. Compute using the definitions of $\denote\cdot$ and $\models$.
+  \[\begin{array}{ll}
+  \denote{\assume {b_1 \wedge b_2}}^\sigma \pkt \\
+  = \{\pkt \mid \pkt \models b_1 \wedge b_2\} \\
+  = \{\pkt \mid \pkt \models b_1, \pkt \models b_2\} \\
+  = {\denote{\assume{b_1}}^\sigma}^\dagger\{\pkt \mid \pkt \models b_1\} \\
+  = \denote{\assume{b_1}; \assume{b_2}}^\sigma \pkt
+  \end{array}\]
+\end{proof}
+
+\begin{lemma}[Products and Sums]
+  \label{lem:prod-sum}
+  For $c_1,c_2, c_3 \in \Cmd$, $\sigma : \Value \to \Row^*$, $\pkt \in \Pkt$.
+  $\denote{c_1;(c_2 \choice c_3)}^\sigma\pkt = \denote{c_1;c_2 \choice
+    c_1;c_3}^\sigma\pkt$.
+\end{lemma}
+\begin{proof}
+  Let $c_1,c_2,c_3$, $\sigma$, and $\pkt$ be given as above. Compute:
+  \[
+  \begin{array}{ll}
+  \denote{c_1;(c_2 \choice c_3)}^\sigma \pkt \\
+  = {\denote{c_2 \choice c_3}^\sigma}^\dagger\left(\denote{c_1}^\sigma\pkt\right) \\
+  = {\denote{c_2}^\sigma}^\dagger\left(\denote{c_1}^\sigma \pkt\right)
+  \cup {\denote{c_3}^\sigma}^\dagger\left(\denote{c_1}^\sigma \pkt\right) \\
+  = {\denote{c_1;c_2}^\sigma}\pkt\cup {\denote{c_1;c_3}^\sigma}\pkt
+  \end{array}
+  \]
+\end{proof}
+
+\begin{lemma}[Disjoint Commutativity]
+  \label{lem:disj-comm}
+  For $c_1,c_2 \in \Cmd$, and $\sigma : \Value \to \Row^*$ if $\fvs(c_1) \cap
+  \fvs(c_2) = \emptyset$, $ \forall \pkt \in \Pkt. \denote{c_1;c_2}^\sigma\pkt =
+  \denote{c_2;c_1}^\sigma \pkt$.
+\end{lemma}
+\begin{proof}
+  %% Let $c_1,c_2$, $\sigma$, and $\pkt$ be given as above. Assume $\fvs(c_1) \cap
+  %% \fvs(c_2) = \emptyset$. Compute:
+  %% \[
+  %% \begin{array}{ll}
+  %%   \denote{c_1;c_2}^\sigma \pkt =
+
+  %% \end{array}
+  %% \]
+  \todo[inline]{prove}
+\end{proof}
 
@@
   \[\begin{array}{r cl}
-  \multicolumn{3}{l}{\state\sigma_c : (\Value \to \Row^*) \to \BExpr} \\
-  \state\sigma_c &\triangleq
-  & \displaystyle\bigwedge_{ t \in \tables(c) }
-  \left(
-  \begin{array}{>{\displaystyle}l}
+  \multicolumn{3}{l}{\state\sigma^t \in \BExpr} \\
+  \state\sigma^t & \triangleq &
+  \left(\begin{array}{>{\displaystyle}l}
   \gamma_t.\miss = 1 \Leftrightarrow \gamma_t.\hitAction > |\sigma(t.\id)| \\
@@
   \wedge \bigwedge_{r_i \in \sigma(t.\id)} \left(\gamma_t.\hitAction \geq i \Leftrightarrow \bigwedge_{j=0}^{i-1} \neg t.\matchrow(r_j)[\gamma_t.\keys/t.\keys]\right)\\
-  \end{array}\right)
+  \end{array}\right)\\\\
+
+  \multicolumn{3}{l}{\state\sigma_c \in \BExpr} \\
+  \state\sigma_c &\triangleq
+  & \displaystyle\bigwedge_{ t \in \tables(c) } \state\sigma^t
   \end{array}
@@
 \end{figure}
+
+\begin{lemma}[Symbolic State distributes over Sequence]
+  \label{lem:sigma-distrib-seq}
+  For $\sigma : \Value \to \Row^*$, $c_1,c_2 \in \Cmd$,
+  $\state\sigma_{c_1;c_2} \Leftrightarrow \state{\sigma}_{c_1}\wedge\state{\sigma}_{c_2}$
+\end{lemma}
+
+\begin{proof}
+  Let $\sigma : \Value \to \Row^*$ and $c_1,c_2\in \Cmd$ be given. Notice that
+  $\tables(c_1;c_2) = \tables(c_1) \cup \tables(c_2)$. Compute:
+  \[
+  \begin{array}{l >{\displaystyle}ll}
+    \state\sigma_{c_1;c_2}
+    & = \bigwedge_{ t \in \tables(c_1;c_2) } \state\sigma^t \\
+    & = \bigwedge_{ t \in \tables(c_1)\cup \tables(c_2) } \state\sigma^t \\
+    & = \bigwedge_{t \in \tables(c_1)} \state\sigma^t \wedge \bigwedge_{t \in \tables(c_2)} \state\sigma^t \\
+    & = \state\sigma_{c_1} \wedge \state\sigma_{c_2}
+  \end{array}
+  \]
+\end{proof}
+
+\begin{lemma}[Symbolic State distributes over Choice]
+  \label{lem:sigma-distrib-choice}
+  For $\sigma : \Value \to \Row^*$, $c_1,c_2 \in \Cmd$,
+  $\state\sigma_{c_1 \choice c_2} \Leftrightarrow \state{\sigma}_{c_1}\wedge\state{\sigma}_{c_2}$
+\end{lemma}
+
+\begin{proof}
+  Let $\sigma : \Value \to \Row^*$ and $c_1,c_2\in \Cmd$ be given. Notice that
+  $\tables(c_1 \choice c_2) = \tables(c_1) \cup \tables(c_2)$. Compute:
+  \[
+  \begin{array}{l >{\displaystyle}ll}
+    \state\sigma_{c_1 \choice c_2}
+    & = \bigwedge_{ t \in \tables(c_1 \choice c_2) } \state\sigma^t \\
+    & = \bigwedge_{ t \in \tables(c_1)\cup \tables(c_2) } \state\sigma^t \\
+    & = \bigwedge_{t \in \tables(c_1)} \state\sigma^t \wedge \bigwedge_{t \in \tables(c_2)} \state\sigma^t \\
+    & = \state\sigma_{c_1} \wedge \state\sigma_{c_2}
+  \end{array}
+  \]
+\end{proof}
 
@@
   \[\pkt|_S \triangleq \{x \mapsto \pkt(x) \mid x \in S\}.\]
-  Two packets $\pkt$,
-  $\pkt'$ are equivalent up to $S \subseteq \Var$, written $\pkt =_S \pkt'$, iff
-  $\pkt|_S = \pkt'|_s$. A set of packets $P$ is a subset of $P'$ up to $S$,
-  written $P \subseteq_S P'$ iff for every packet $\pkt \in P$, there exists
-  $\pkt' \in P'$ such that $\pkt =_S \pkt'$. Two sets of packets $P$ and $P'$
-  are $S$-equivalent, written $P =_S P'$ iff $P \subseteq_S P'$ and $P \subseteq_S P'$.
+  Two packets $\pkt$, $\pkt'$ are equivalent up to $S \subseteq \Var$, written
+  $\pkt =_S \pkt'$, iff $\pkt|_S = \pkt'|_s$. A set of packets $P$ is a subset
+  of $P'$ up to $S$, written $P \subseteq_S P'$ iff for every packet $\pkt \in
+  P$, there exists $\pkt' \in P'$ such that $\pkt =_S \pkt'$. Two sets of
+  packets $P$ and $P'$ are $S$-equivalent, written $P =_S P'$ iff $P \subseteq_S
+  P'$ and $P \subseteq_S P'$. A set $P \subseteq \Pkt$ is a $V$-equivalence
+  class if $\forall \pkt_1, \pkt_2 \in P$, $\pkt_1 =_V \pkt_2$.
 \end{definition}
 
-\begin{lemma}[Projection]
-  For $c \in \Cmd$, $\fvs(c) \subseteq V \subseteq \Var$ and $P \subseteq \Pkt$ such that $\forall
-  \pkt_1,\pkt_2 \in P, \pkt_1 =_V \pkt_2$, $\denote{c} P =_V \denote{c} P|_V$.
-
+\begin{lemma}[Projection Union-Distributivity]
+  \label{lem:proj-union-set}
+  For all $P,Q \subseteq \Pkt$, $V \subseteq \Var$,
+  $(P\cup Q)|_V = P|_V \cup Q|_V$
 \end{lemma}
+
+\begin{proof}
+  Let $P,Q \subseteq \Pkt$ and $V \subseteq \Var$ be given. Compute:
+
+  \[\begin{array}{ll}
+  (P\cup Q)|_V\\
+  = \{\pkt|_V \mid \pkt \in P\cup Q\}\\
+  = \{\pkt|_V \mid \pkt \in P\}\cup \{\pkt|_V \mid \pkt \in Q\}\\
+  = P|_V \cup Q|_V
+  \end{array}\]
+\end{proof}
+
+\begin{lemma}[Projection Idempotence]
+  \label{lem:proj-idem}
+  For all $\pkt \in \Pkt$, $V \subseteq \Var$,
+  $(\pkt|_V)|_V = \pkt|_V$
+\end{lemma}
+
+\begin{proof}
+  Let $\pkt \in \Pkt$ and $V \subseteq \Var$ be given.
+  \[\begin{array}{ll}
+    (\pkt|_V)|_V \\
+    = \{x \mapsto (\pkt|_V)(x) \mid x \in V\} & \text{by definition}\\
+    = \{x \mapsto (\{y \mapsto \pkt(y) \mid y \in V\})(x) \mid x \in V\} & \text{by definition}\\
+    = \{x \mapsto \pkt(x) \mid x \in V\} & \beta\text{-reduction}\\
+    = \pkt|_V & \text{by definition}
+  \end{array}\]
+\end{proof}
+
+\begin{lemma}[Projection Idempotence for Sets]
+  \label{lem:proj-idem-set}
+  For all $P \subseteq \Pkt$, $V \subseteq \Var$,
+  $(P|_V)|_V = P|_V$
+\end{lemma}
+
+\begin{proof}
+  Let $P \subseteq \Pkt$ and $V \subseteq \Var$ be given. Compute:
+  \[\begin{array}{ll}
+  (P|_V)|_V \\
+  = \{\pkt|_V \mid \pkt \in P|_V\}
+  & \text{by definition of } (\cdot)|_V : \mathcal{P}(\Pkt) \to \mathcal{P}(\Pkt) \\
+  = \{\pkt|_V \mid \pkt \in \{\pkt'|_V \mid \pkt'\in P\}\}
+  & \text{by definition of } (\cdot)|_V : \mathcal{P}(\Pkt) \to \mathcal{P}(\Pkt) \\
+  = \{(\pkt|_V)|_V \mid \pkt \in P \}
+  & \text{by set theory} \\
+  = \{\pkt|_V \mid \pkt \in P\} & \text{by Lemma~\ref{lem:proj-idem}} \\
+  = P|_V & \text{by definition of } (\cdot)|_V : \mathcal{P}(\Pkt) \to \mathcal{P}(\Pkt)
+  \end{array}\]
+\end{proof}
+
+\begin{lemma}{Subset Implies Projected Subset}
+  \label{lem:subset-proj}
+  For $P,Q \subseteq \Pkt$, and $ V \subseteq \Var$,
+  If $P \subseteq Q$ then $P \subseteq_V Q$.
+\end{lemma}
+
+\begin{proof}
+Let $P,Q \subseteq \Pkt$ and $V \subseteq \Var$ be given. Assume $P \subseteq Q$,
+to show that $P \subseteq_V Q$.
+
+Let $\pkt \in P$. Since $P \subseteq Q$, $\pkt \in Q$. Notice that $\pkt =_V
+\pkt$, so $P \subseteq_V Q$.
+\end{proof}
+
+\begin{lemma}{Equality Implies Projected Equality}
+  \label{lem:equal-proj}
+  For $P,Q \subseteq \Pkt$, and $ V \subseteq \Var$,
+  If $P = Q$ then $P =_V Q$.
+\end{lemma}
+
+\begin{proof}
+  By Lemma~\ref{lem:subset-proj}
+\end{proof}
+
+\begin{lemma}[Expression Projection]
+  \label{lem:proj-expr}
+  For every set of variables $V \subseteq \Var$, packet $\pkt \in \Pkt$ and
+  expression $e \in \BVExpr$, such that $\efvs(e) \subseteq V$, $\edenote{e}\pkt =
+  \edenote{e}(\pkt|_V)$.
+\end{lemma}
+
+\begin{proof}
+  Let $V \subseteq \Var$ be given.
+  Proceed by induction on $e$, leaving $\pkt$ arbitrary.
+  \begin{enumerate}[align=left]
+  \item[($e = v$)] Let $\pkt$ be given. $\edenote{v}\pkt = v = \edenote{v}{\pkt|_V}$
+  \item[($e = x$)] Let $\pkt$ be given. Note that $\efvs(x) = \{x\} \subseteq V$. Compute:
+    \[\begin{array}{ll}
+    \edenote{x}(\pkt|_V) \\
+    = (\pkt|_V)(x)
+    & \text{by definition of }\edenote\cdot\\
+    = \{y \mapsto \pkt(y) \mid y \in V\}(x)
+    & \text{by definition of }(\cdot)|_V : \Pkt \to \Pkt\\
+    = \pkt(x) & \text{since } x \in V
+    \end{array}
+    \]
+  \item[($e = e_1 \binop e_2$)] Let $\pkt$ be given. Compute:
+    \[\begin{array}{ll}
+    \edenote{e_1 \binop e_2}\pkt \\
+    = \edenote{e_1}\pkt \binop \edenote{e_2}\pkt & \text{definition of }\edenote\cdot \\
+    = \edenote{e_1}(\pkt|_V) \binop \edenote{e_2}(\pkt|_V) & \text{by IHS since }\efvs(e_i) \subseteq \efvs(e_1 \binop e_2) \subseteq V, i = 1,2 \\
+    = \edenote{e_1 \binop e_2}(\pkt|_V) & \text{definition of }\edenote\cdot
+    \end{array}\]
+  \item[($e = \unop e$)] Let $\pkt$ be given. Compute:
+    \[\begin{array}{ll}
+    \edenote{\unop e}\pkt \\
+    = \unop \edenote{e}\pkt & \text{definition of }\edenote\cdot \\
+    = \unop \edenote{e}(\pkt|_V) & \text{by IH, since } \efvs(e) = \efvs(\unop e) \subseteq V \\
+    = \edenote{\unop e}(\pkt|_V) & \text{definition of }\edenote\cdot
+    \end{array}\]
+  \end{enumerate}
+\end{proof}
+
+\begin{lemma}[Satisfaction Projection]
+  \label{lem:proj-satis}
+  For every set of variables $V \subseteq \Var$, packet $\pkt \in \Pkt$ and
+  boolean expression $b \in \BExpr$, such that $\bfvs(b) \subseteq V$,
+  $\pkt \models b$ iff $\pkt|_V \models b$.
+\end{lemma}
+
+\begin{proof}
+  Let $V \subseteq \Var$ be given.
+  Proceed by induction on $b$, leaving $\pkt$ arbitrary.
+  \begin{enumerate}[align=left]
+  \item[($b = \FALSE$)] Immediate.
+  \item[($b = b_1 \Rightarrow b_2$)]
+    Let $\pkt$ be given. Compute:
+    \[\begin{array}{ll}
+    \pkt \models b_1 \Rightarrow b_2 \\
+    \iff \pkt \not\models b_1 \text{or} \pkt\models b_2 & \text{by definition of } \models   \\
+    \iff \pkt|_V \not\models b_1 \text{or} \pkt|_V \models b_2  & \text{by IHs, since} \bfvs(b_i) \subseteq \bfvs(b_1 \Rightarrow b_2) \subseteq V, i = 1,2  \\
+    \iff \pkt|_V \models b_1 \Rightarrow b_2
+    & \text{by definition of } \models
+    \end{array}\]
+
+  \item[($b = e_1 = e_2$)] Let $\pkt$ be given. Compute:
+    \[\begin{array}{ll}
+    \pkt \models e_1 = e_2 \\
+    \iff \edenote{e_1}\pkt = \edenote{e_2}\pkt
+    & \text{by definition of }\models \\
+    \iff \edenote{e_1}(\pkt|_V) = \edenote{e_2}(\pkt|_V)
+    & \text{by Lemma~\ref{lem:proj-expr}, since } \efvs(e_i) \subseteq \bfvs(e_1 = e_2) \subseteq V, i = 1,2 \\
+    \iff \pkt|_V \models e_1 = e_2
+    & \text{by definition of }\models
+    \end{array}\]
+
+  \end{enumerate}
+
+\end{proof}
+
+\begin{lemma}[Projection-Denotation Commutativity]
+  \label{lem:proj-den-comm}
+  For a closed action $c \in \Cmd$, $\fvs(c) \subseteq V \subseteq \Var$ and
+  $\pkt \in \Pkt$, $\denote{c} \pkt =_V \denote{c} \pkt|_V$.
+\end{lemma}
+
+\begin{proof}
+  Proceed by induction on $c$, leaving $\pkt$ general.
+  \begin{enumerate}[align=left]
+  \item[$(c = x:=e)$] Let $\pkt$ be given. Compute:
+    \[\begin{array}{ll}
+    (\denote{x:=e} (\pkt|_V))|_V\\
+    = ((\pkt|_V)\{x:=\edenote{e}(\pkt|_V)\})|_V
+    & \text{definition} \\
+    = ((\pkt|_V)\{x := \edenote{e}\pkt\})|_V
+    & \text{By Lemma~\ref{lem:proj-expr} since } \fvs(e) \subseteq \fvs(x := e) \subseteq V \\
+    = (\pkt\{x := \edenote{e}\pkt\}|_V)|_V
+    & \text{since } x \in \fvs(x:=e) \subseteq V \\
+    = ((\denote{x:=e}\pkt)|_V)|_V
+    & \text{by definition of }\denote{\cdot}\\
+    = (\denote{x:=e}\pkt)|_V
+    & \text{by Lemma~\ref{lem:proj-idem}}
+    \end{array}\]
+    The result follows by definition of $=_V$.
+
+  \item[$(c = \havoc x)$] Contradictory; $\havoc x$ is not an action.
+  \item[$(c = \assume b)$]
+    Let $\pkt$ be given. Compute:
+    \[\begin{array}{ll}
+    (\denote{\assume b} \pkt)|_V \\
+    = (\{\pkt \mid \pkt \models b\})|_V
+    & \text{definition of}\denote\cdot\\
+    = ((\{\pkt \mid \pkt \models b\})|_V)|_V
+    & \text{By Lemma~\ref{lem:proj-idem-set}} \\
+    = (\{\pkt|_V \mid \pkt \models b\})|_V
+    & \text{definition of }(\cdot)|_V : \mathcal P (\Pkt) \to \mathcal P (\Pkt) \\
+    = (\{\pkt|_V \mid \pkt|_V \models b\})|_V
+    & \text{By Lemma~\ref{lem:proj-satis} since } \fvs(b) \subseteq \fvs(\assume b) \subseteq V \\
+    = (\denote{\assume b}(\pkt|_V))|_V
+    & \text{by definition of}\denote\cdot
+    \end{array}
+    \]
+    The result follows by definition of $=_V$.
+
+  \item[$(c = \assert b)$] \textit{sim.}
+  \item[$(c = c_1;c_2)$] Let $\pkt$ be given. Compute:
+    \[\begin{array}{>{\displaystyle}ll}
+    (\denote{c_1;c_2}\pkt)|_V \\
+    = (\denote{c_2}^\dagger(\denote{c_1}\pkt))|_V
+    & \text{definition of }\denote\cdot \\
+    = \left(\bigcup_{\pkt' \in \denote{c_1} \pkt} \denote{c_2}\pkt'\right)|_V
+    & \text{definition of }(\cdot)^\dagger \\
+    = \bigcup_{\pkt' \in \denote{c_1} \pkt} (\denote{c_2}\pkt')|_V
+    & \text{by Lemma~\ref{lem:proj-union-set}} \\
+    = \bigcup_{\pkt' \in \denote{c_1}\pkt} (\denote{c_2}(\pkt'|_V))|_V
+    & \text{IH}(c_2)  \\
+    =\bigcup_{\pkt' \in (\denote{c_1}\pkt)|_V} (\denote{c_2} \pkt')|_V
+    & \text{by set theory} \\
+    = \bigcup_{\pkt' \in (\denote{c_1}(\pkt|_V))|_V} (\denote{c_2} \pkt')|_V
+    & \text{by IH}(c_1) \\
+    = \bigcup_{\pkt' \in \denote{c_1}(\pkt|_V)} (\denote{c_2} (\pkt'|_V))|_V
+    & \text{by set theory} \\
+    = \bigcup_{\pkt' \in \denote{c_1}(\pkt|_V)} ((\denote{c_2} \pkt')|_V)|_V
+    & \text{by IH}(c_2) \\
+    = \bigcup_{\pkt' \in \denote{c_1}(\pkt|_V)} (\denote{c_2} \pkt')|_V
+    & \text{by Lemma~\ref{lem:proj-idem-set}} \\
+    = \left(\bigcup_{\pkt' \in \denote{c_1}(\pkt|_V)} \denote{c_2}\pkt'\right)|_V
+    & \text{by Lemma~\ref{lem:proj-union-set}} \\
+    = \denote{c_2}^\dagger(\denote{c_1}(\pkt|_V))|_V
+    & \text{By definition of }(\cdot)^\dagger\\
+    = (\denote{c_1;c_2}(\pkt|_V))|_V
+    & \text{By definition of }\denote\cdot
+    \end{array}\]
+    The result follows by definition of $=_V$.
+
+  \item[$(c = c_1 \choice c_2)$] Let $\pkt$ be given.
+    \[\begin{array}{ll}
+    (\denote{c_1\choice c_2}\pkt)|_V \\
+    = (\denote{c_1}\pkt \cup \denote{c_2}\pkt)|_V
+    & \text{By definition of }\denote\cdot \\
+    = (\denote{c_1}\pkt)|_V \cup (\denote{c_2}\pkt)|_V
+    & \text{By Lemma~\ref{lem:proj-union-set}} \\
+    = (\denote{c_1}(\pkt|_V))|_V \cup (\denote{c_2}(\pkt|_V))|_V
+    & \text{By IHs} \\
+    = ((\denote{c_1}(\pkt|_V) \cup \denote{c_2}(\pkt|_V))|_V)|_V
+    & \text{By Lemma~\ref{lem:proj-union-set}} \\
+    = (\denote{c_1}(\pkt|_V) \cup \denote{c_2}(\pkt|_V))|_V
+    & \text{By Lemma~\ref{lem:proj-idem-set}}\\
+    = (\denote{c_1 \choice c_2}(\pkt|_V))|_V
+    & \text{By definition of }\denote\cdot \\
+    \end{array}\]
+
+    The result follows by definition of $=_V$.
+
+  \item[$(c = t.\apply())$] Contradictory; $t.\apply()$ is not an action.
+  \end{enumerate}
+
+\end{proof}
 
@@
     = \denote{\havoc{R\cup G}; \assume{\state\sigma_{c_1}};\assume{\state\sigma_{c_2}};c_1\instr;c_2\instr}\,\pkt
-    & \text{LEMMAs INVOLVING SIGMA-DISTRIB \& ASSUME-DISTRIB} \\
+    & \text{By Lemmas~\ref{lem:sigma-distrib-seq} \&~\ref{lem:assume-currying}} \\
     = \denote{\havoc{R_1};\havoc{G_1}; \havoc{R_2};\havoc{G_2}; \assume{\state\sigma_{c_1}};\assume{\state\sigma_{c_2}};c_1\instr;c_2\instr}\,\pkt
@@
     = \denote{\havoc{R_1};\havoc{G_1}; \havoc{R_2}; \havoc{G_2}; \assume{\state\sigma_{c_1}};c_1\instr;\assume{\state\sigma_{c_2}};c_2\instr}\,\pkt
-    & \text{BY FREE-VARIABLE-ASSUMPTION COMMUTATIVITY LEMMA} \\
+    & \text{By Lemma~\ref{lem:disj-comm}} \\
     = \denote{(\havoc{R_1}; \havoc{G_1}; \assume{\state\sigma_{c_1}};c_1\instr);
       (\havoc{R_2}; \havoc{G_2};\assume{\state\sigma_{c_2}};c_2\instr)}\,\pkt
-    & \text{BY FREE-VARIABLE HAVOC COMMUTE LEMMA}\\
+    & \text{By Lemma~\ref{lem:disj-comm}}\\
     = \denote{\havoc{R_2}; \havoc{G_2};\assume{\state\sigma_{c_2}};c_2\instr}^\dagger
     \left(\denote{\havoc{R_1}; \havoc{G_1}; \assume{\state\sigma_{c_1}};c_1\instr}\,\pkt\right)
-    & \text{BY DEFINITION} \\
+    & \text{by definition} \\
     =_V {\denote{c_2}^\sigma}^\dagger \left(\denote{\havoc{\vec{\rho_{t_1}}}; \havoc{\vec\gamma_{t_1}}; \assume{\state\sigma_{c_1}};c_1\instr}\,\pkt\right)
@@
 
-    THE RESULT FOLLOWS BY EQUALITY IMPLIES PROJECTED EQUALITY LEMMA
+    The result follows by Lemma~\ref{lem:equal-proj}.
 
@@
     = \denote{\havoc{R \cup G};\assume{\state\sigma_{c_1} \wedge \state\sigma_{c_2}}; (c_1\instr \choice c_2\instr)}\,\pkt
-    & \text{BY LEMMA RELATING [] and STATE}\\
+    & \text{By Lemma~\ref{lem:sigma-distrib-choice}}\\
     = \denote{\havoc{R \cup G};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; (c_1\instr  \choice   c_2\instr)}\,\pkt
-    & \text{BY WEDGE-ASSUME LEMMA}\\
+    & \text{By Lemma~\ref{lem:assume-currying}}\\
     = \denote{\havoc{R \cup G};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; c_1\instr
       \choice \havoc{R \cup G};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
-    & \text{BY PRODUCTS(;) AND SUMS([]) LEMMA} \\
+    & \text{By Lemma~\ref{lem:prod-sum}} \\
     = \denote{\havoc{R \cup G}; \assume{\state\sigma_{c_1}}; c_1\instr
@@
     & \text{by definition of }\denote{\cdot} \\
-    =_V \denote{c_1}^\sigma\pkt \cup \denote{\havoc{\vec{\rho_{t_2}}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
+    =_V \denote{c_1}^\sigma\pkt \cup \denote{\havoc{R_2 \cup G_2}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
     & \text{By IH}(c_1) \text{ since } \fvs(c_1) \subseteq \fvs(c_1 \choice c_2) \subseteq V \\
@@
       \[\bigcup_{a=0}^{|t.\action| - 1}\denote{t.\action[a](\rho_t.\data)}^\dagger H_a
-       = \denote{t.\action[r_i.\action](r_i.\data)}^\dagger H_a
-       \]
-
-       Which is $V$-equivalent to $\denote{t.\action[r_i.\action](r_i.\data)}\,\pkt$. \todo{PROVE THIS LEMMA}
-       The result follows by the IH on $t.\action[r_i.\action](r_i.\data)$.
+       = \denote{t.\action[r_i.\action](r_i.\data)}^\dagger H_{r_i.\action}
+       \]
+
+       which is $V$-equivalent to
+       $\denote{t.\action[r_i.\action](r_i.\data)}\,\pkt$ by
+       Lemma~\ref{lem:proj-den-comm}. The result follows by the IH on
+       $t.\action[r_i.\action](r_i.\data)$ and Lemma~\ref{lem:equal-proj}.
 
@@
         which means that $\gamma_t.\hitAction > n$, and that $\gamma_t.\miss =
-        1$. Conclude that $H_a = \emptyset$ and $M \neq \emptyset$. Since $P =_V
-        \{\pkt\}$, and $M$ is a nonempty refinement of $P$\todo{PROVE THIS
-          LEMMA?}, conclude that $M =_V \{\pkt\}$. The result follows.
+        1$. Conclude that $H_a = \emptyset$ and $M \neq \emptyset$. Since
+        $\emptyset \neq M \subseteq P =_V \{\pkt\}$, $M =_V \{\pkt\}$ by
+        Lemma~\ref{lem:subset-proj}. The result follows by
+        Lemma~\ref{lem:equal-proj}.
     \end{enumerate}

### 2021-05-26 inference.tex
more lemmas

@@
 % commands
-\newcommand{\assert}{\mathop{\mathsf{assert}}}
-\newcommand{\assume}{\mathop{\mathsf{assume}}}
+\newcommand{\assert}{\mathop{\mathsf{ast}}}
+\newcommand{\assume}{\mathop{\mathsf{asm}}}
 \newcommand{\apply}{\mathsf{apply}}
@@
 \DeclareMathOperator*{\bigchoice}{\scalerel*{\choiceop}{\sum}}
-\newcommand{\havoc}[1]{\mathop{\mathsf{havoc}}#1}
+\newcommand{\havoc}[1]{\mathop{\mathsf{hvc}}#1}
 \newcommand{\SKIP}{\mathsf{skip}}
@@
 
-\begin{definition}[Total State]
-  \label{def:total-state}
-  A state $\sigma$ is total if for every $t \in \dom(\sigma)$, the following
-  formula is valid:
-  \[\bigvee_{r \in \sigma(t.\id)} t.\keys = r.\matches \]
-\end{definition}
-
-Henceforth we assume that all states are total.
-
-\todo[inline]{Can be Removed(?)}
+%% \begin{definition}[Total State]
+%%   \label{def:total-state}
+%%   A state $\sigma$ is total if for every $t \in \dom(\sigma)$, the following
+%%   formula is valid:
+%%   \[\bigvee_{r \in \sigma(t.\id)} t.\keys = r.\matches \]
+%% \end{definition}
+
+%% Henceforth we assume that all states are total.
+
+%% \todo[inline]{Can be Removed(?)}
 
@@
 
+\subsection{Definitions}
+
+\begin{definition}[Variables]
+  The free variables of a command $c$, written $\fvs(c) \subseteq \Var$, is the
+  set of variables that occur at any point in $c$ (i.e. as the lhs of an
+  assignment, in a boolean expression, a bitvector expression, as a key in a
+  table.) The only bound variables are the action data in tables. Formally:
+  \[\begin{array}{l >{\triangleq}cl}
+  \efvs :: \BVExpr \to \mathcal{P}(\Var) \\
+  \efvs(v) && \{\} \\
+  \efvs(x) && \{x\}\\
+  \efvs(e_1 \binop e_2) && \efvs(e_1) \cup \efvs(e_2)\\
+  \efvs(\unop e) && \efvs(unop) \\[1em]
+  \bfvs :: \BExpr \to \mathcal{P}(\Var) \\
+  \bfvs(\FALSE) && \{\}\\
+  \bfvs(b_1 \Rightarrow b_2) && \bfvs(b_1) \cup \bfvs(b_2)\\
+  \bfvs(e_1 = e_2) && \efvs(e_1) = \efvs(e_2) \\[1em]
+  \fvs :: \Cmd \to \mathcal{P}(\Var) \\
+  \fvs(x := e) && \{x\} \cup \fvs(e)\\
+  \fvs(\havoc x) && \{x\}\\
+  \fvs(\assume b) && \fvs(b)\\
+  \fvs(\assert b) && \fvs(b) \\
+  \fvs(c_1;c_2) && \fvs(c_1) \cup \fvs(c_1) \\
+  \fvs(c_1 \choice c_2) && \fvs(c_1) \cup \fvs(c_2) \\
+  \fvs(t.\apply()) &&
+  t.\keys \cup \bigcup_{\lambda \vec x. c \in t.\actions} (\fvs(c) \setminus \{x_1,\ldots, x_n\})
+  \end{array}\]
+\end{definition}
+
+\begin{definition}[Projections]
+  For a set of variables $S \subseteq \Var$, and a packet $\pkt$, write
+  $\pkt|_S$ to be projection of $\pkt$ onto $S$. \todo{I guess this means
+    packets are partial functions?} Formally,
+  \[\pkt|_S \triangleq \{x \mapsto \pkt(x) \mid x \in S\}.\]
+  Two packets $\pkt$, $\pkt'$ are equivalent up to $S \subseteq \Var$, written
+  $\pkt =_S \pkt'$, iff $\pkt|_S = \pkt'|_s$. A set of packets $P$ is a subset
+  of $P'$ up to $S$, written $P \subseteq_S P'$ iff for every packet $\pkt \in
+  P$, there exists $\pkt' \in P'$ such that $\pkt =_S \pkt'$. Two sets of
+  packets $P$ and $P'$ are $S$-equivalent, written $P =_S P'$ iff $P \subseteq_S
+  P'$ and $P \subseteq_S P'$. A set $P \subseteq \Pkt$ is a $V$-equivalence
+  class if $\forall \pkt_1, \pkt_2 \in P$, $\pkt_1 =_V \pkt_2$.
+\end{definition}
+
+\begin{lemma}[Projection Union-Distributivity]
+  \label{lem:proj-union-set}
+  For all $P,Q \subseteq \Pkt$, $V \subseteq \Var$,
+  $(P\cup Q)|_V = P|_V \cup Q|_V$
+\end{lemma}
+
+\begin{proof}
+  Let $P,Q \subseteq \Pkt$ and $V \subseteq \Var$ be given. Compute:
+
+  \[\begin{array}{ll}
+  (P\cup Q)|_V\\
+  = \{\pkt|_V \mid \pkt \in P\cup Q\}\\
+  = \{\pkt|_V \mid \pkt \in P\}\cup \{\pkt|_V \mid \pkt \in Q\}\\
+  = P|_V \cup Q|_V
+  \end{array}\]
+\end{proof}
+
+\begin{lemma}[Projection Idempotence]
+  \label{lem:proj-idem}
+  For all $\pkt \in \Pkt$, $V \subseteq \Var$,
+  $(\pkt|_V)|_V = \pkt|_V$
+\end{lemma}
+
+\begin{proof}
+  Let $\pkt \in \Pkt$ and $V \subseteq \Var$ be given.
+  \[\begin{array}{ll}
+    (\pkt|_V)|_V \\
+    = \{x \mapsto (\pkt|_V)(x) \mid x \in V\} & \text{by definition}\\
+    = \{x \mapsto (\{y \mapsto \pkt(y) \mid y \in V\})(x) \mid x \in V\} & \text{by definition}\\
+    = \{x \mapsto \pkt(x) \mid x \in V\} & \beta\text{-reduction}\\
+    = \pkt|_V & \text{by definition}
+  \end{array}\]
+\end{proof}
+
+\begin{lemma}[Projection Idempotence for Sets]
+  \label{lem:proj-idem-set}
+  For all $P \subseteq \Pkt$, $V \subseteq \Var$,
+  $(P|_V)|_V = P|_V$
+\end{lemma}
+
+\begin{proof}
+  Let $P \subseteq \Pkt$ and $V \subseteq \Var$ be given. Compute:
+  \[\begin{array}{ll}
+  (P|_V)|_V \\
+  = \{\pkt|_V \mid \pkt \in P|_V\}
+  & \text{by definition of } (\cdot)|_V : \mathcal{P}(\Pkt) \to \mathcal{P}(\Pkt) \\
+  = \{\pkt|_V \mid \pkt \in \{\pkt'|_V \mid \pkt'\in P\}\}
+  & \text{by definition of } (\cdot)|_V : \mathcal{P}(\Pkt) \to \mathcal{P}(\Pkt) \\
+  = \{(\pkt|_V)|_V \mid \pkt \in P \}
+  & \text{by set theory} \\
+  = \{\pkt|_V \mid \pkt \in P\} & \text{by Lemma~\ref{lem:proj-idem}} \\
+  = P|_V & \text{by definition of } (\cdot)|_V : \mathcal{P}(\Pkt) \to \mathcal{P}(\Pkt)
+  \end{array}\]
+\end{proof}
+
+\begin{lemma}[Subset Implies Projected Subset]
+  \label{lem:subset-proj}
+  For $P,Q \subseteq \Pkt$, and $ V \subseteq \Var$,
+  If $P \subseteq Q$ then $P \subseteq_V Q$.
+\end{lemma}
+
+\begin{proof}
+Let $P,Q \subseteq \Pkt$ and $V \subseteq \Var$ be given. Assume $P \subseteq Q$,
+to show that $P \subseteq_V Q$.
+
+Let $\pkt \in P$. Since $P \subseteq Q$, $\pkt \in Q$. Notice that $\pkt =_V
+\pkt$, so $P \subseteq_V Q$.
+\end{proof}
+
+\begin{lemma}[Equality Implies Projected Equality]
+  \label{lem:equal-proj}
+  For $P,Q \subseteq \Pkt$, and $ V \subseteq \Var$,
+  If $P = Q$ then $P =_V Q$.
+\end{lemma}
+
+\begin{proof}
+  By Lemma~\ref{lem:subset-proj}
+\end{proof}
+
+\begin{lemma}[Expression Projection]
+  \label{lem:proj-expr}
+  For every set of variables $V \subseteq \Var$, packet $\pkt \in \Pkt$ and
+  expression $e \in \BVExpr$, such that $\efvs(e) \subseteq V$, $\edenote{e}\pkt =
+  \edenote{e}(\pkt|_V)$.
+\end{lemma}
+
+\begin{proof}
+  Let $V \subseteq \Var$ be given.
+  Proceed by induction on $e$, leaving $\pkt$ arbitrary.
+  \begin{enumerate}[align=left]
+  \item[($e = v$)] Let $\pkt$ be given. $\edenote{v}\pkt = v = \edenote{v}{\pkt|_V}$
+  \item[($e = x$)] Let $\pkt$ be given. Note that $\efvs(x) = \{x\} \subseteq V$. Compute:
+    \[\begin{array}{ll}
+    \edenote{x}(\pkt|_V) \\
+    = (\pkt|_V)(x)
+    & \text{by definition of }\edenote\cdot\\
+    = \{y \mapsto \pkt(y) \mid y \in V\}(x)
+    & \text{by definition of }(\cdot)|_V : \Pkt \to \Pkt\\
+    = \pkt(x) & \text{since } x \in V
+    \end{array}
+    \]
+  \item[($e = e_1 \binop e_2$)] Let $\pkt$ be given. Compute:
+    \[\begin{array}{ll}
+    \edenote{e_1 \binop e_2}\pkt \\
+    = \edenote{e_1}\pkt \binop \edenote{e_2}\pkt & \text{definition of }\edenote\cdot \\
+    = \edenote{e_1}(\pkt|_V) \binop \edenote{e_2}(\pkt|_V) & \text{by IHS since }\efvs(e_i) \subseteq \efvs(e_1 \binop e_2) \subseteq V, i = 1,2 \\
+    = \edenote{e_1 \binop e_2}(\pkt|_V) & \text{definition of }\edenote\cdot
+    \end{array}\]
+  \item[($e = \unop e$)] Let $\pkt$ be given. Compute:
+    \[\begin{array}{ll}
+    \edenote{\unop e}\pkt \\
+    = \unop \edenote{e}\pkt & \text{definition of }\edenote\cdot \\
+    = \unop \edenote{e}(\pkt|_V) & \text{by IH, since } \efvs(e) = \efvs(\unop e) \subseteq V \\
+    = \edenote{\unop e}(\pkt|_V) & \text{definition of }\edenote\cdot
+    \end{array}\]
+  \end{enumerate}
+\end{proof}
+
+\begin{lemma}[Satisfaction Projection]
+  \label{lem:proj-satis}
+  For every set of variables $V \subseteq \Var$, packet $\pkt \in \Pkt$ and
+  boolean expression $b \in \BExpr$, such that $\bfvs(b) \subseteq V$,
+  $\pkt \models b$ iff $\pkt|_V \models b$.
+\end{lemma}
+
+\begin{proof}
+  Let $V \subseteq \Var$ be given.
+  Proceed by induction on $b$, leaving $\pkt$ arbitrary.
+  \begin{enumerate}[align=left]
+  \item[($b = \FALSE$)] Immediate.
+  \item[($b = b_1 \Rightarrow b_2$)]
+    Let $\pkt$ be given. Compute:
+    \[\begin{array}{ll}
+    \pkt \models b_1 \Rightarrow b_2 \\
+    \iff \pkt \not\models b_1 \text{or} \pkt\models b_2 & \text{by definition of } \models   \\
+    \iff \pkt|_V \not\models b_1 \text{or} \pkt|_V \models b_2  & \text{by IHs, since} \bfvs(b_i) \subseteq \bfvs(b_1 \Rightarrow b_2) \subseteq V, i = 1,2  \\
+    \iff \pkt|_V \models b_1 \Rightarrow b_2
+    & \text{by definition of } \models
+    \end{array}\]
+
+  \item[($b = e_1 = e_2$)] Let $\pkt$ be given. Compute:
+    \[\begin{array}{ll}
+    \pkt \models e_1 = e_2 \\
+    \iff \edenote{e_1}\pkt = \edenote{e_2}\pkt
+    & \text{by definition of }\models \\
+    \iff \edenote{e_1}(\pkt|_V) = \edenote{e_2}(\pkt|_V)
+    & \text{by Lemma~\ref{lem:proj-expr}, since } \efvs(e_i) \subseteq \bfvs(e_1 = e_2) \subseteq V, i = 1,2 \\
+    \iff \pkt|_V \models e_1 = e_2
+    & \text{by definition of }\models
+    \end{array}\]
+
+  \end{enumerate}
+
+\end{proof}
+
+\begin{lemma}[Projection-Denotation Commutativity]
+  \label{lem:proj-den-comm}
+  For a closed action $c \in \Cmd$, $\fvs(c) \subseteq V \subseteq \Var$ and
+  $\pkt \in \Pkt$, $\denote{c} \pkt =_V \denote{c} \pkt|_V$.
+\end{lemma}
+
+\begin{proof}
+  Proceed by induction on $c$, leaving $\pkt$ general.
+  \begin{enumerate}[align=left]
+  \item[$(c = x:=e)$] Let $\pkt$ be given. Compute:
+    \[\begin{array}{ll}
+    (\denote{x:=e} (\pkt|_V))|_V\\
+    = ((\pkt|_V)\{x:=\edenote{e}(\pkt|_V)\})|_V
+    & \text{definition} \\
+    = ((\pkt|_V)\{x := \edenote{e}\pkt\})|_V
+    & \text{By Lemma~\ref{lem:proj-expr} since } \fvs(e) \subseteq \fvs(x := e) \subseteq V \\
+    = (\pkt\{x := \edenote{e}\pkt\}|_V)|_V
+    & \text{since } x \in \fvs(x:=e) \subseteq V \\
+    = ((\denote{x:=e}\pkt)|_V)|_V
+    & \text{by definition of }\denote{\cdot}\\
+    = (\denote{x:=e}\pkt)|_V
+    & \text{by Lemma~\ref{lem:proj-idem}}
+    \end{array}\]
+    The result follows by definition of $=_V$.
+
+  \item[$(c = \havoc x)$] Contradictory; $\havoc x$ is not an action.
+  \item[$(c = \assume b)$]
+    Let $\pkt$ be given. Compute:
+    \[\begin{array}{ll}
+    (\denote{\assume b} \pkt)|_V \\
+    = (\{\pkt \mid \pkt \models b\})|_V
+    & \text{definition of}\denote\cdot\\
+    = ((\{\pkt \mid \pkt \models b\})|_V)|_V
+    & \text{By Lemma~\ref{lem:proj-idem-set}} \\
+    = (\{\pkt|_V \mid \pkt \models b\})|_V
+    & \text{definition of }(\cdot)|_V : \mathcal P (\Pkt) \to \mathcal P (\Pkt) \\
+    = (\{\pkt|_V \mid \pkt|_V \models b\})|_V
+    & \text{By Lemma~\ref{lem:proj-satis} since } \fvs(b) \subseteq \fvs(\assume b) \subseteq V \\
+    = (\denote{\assume b}(\pkt|_V))|_V
+    & \text{by definition of}\denote\cdot
+    \end{array}
+    \]
+    The result follows by definition of $=_V$.
+
+  \item[$(c = \assert b)$] \textit{sim.}
+  \item[$(c = c_1;c_2)$] Let $\pkt$ be given. Compute:
+    \[\begin{array}{>{\displaystyle}ll}
+    (\denote{c_1;c_2}\pkt)|_V \\
+    = (\denote{c_2}^\dagger(\denote{c_1}\pkt))|_V
+    & \text{definition of }\denote\cdot \\
+    = \left(\bigcup_{\pkt' \in \denote{c_1} \pkt} \denote{c_2}\pkt'\right)|_V
+    & \text{definition of }(\cdot)^\dagger \\
+    = \bigcup_{\pkt' \in \denote{c_1} \pkt} (\denote{c_2}\pkt')|_V
+    & \text{by Lemma~\ref{lem:proj-union-set}} \\
+    = \bigcup_{\pkt' \in \denote{c_1}\pkt} (\denote{c_2}(\pkt'|_V))|_V
+    & \text{IH}(c_2)  \\
+    =\bigcup_{\pkt' \in (\denote{c_1}\pkt)|_V} (\denote{c_2} \pkt')|_V
+    & \text{by set theory} \\
+    = \bigcup_{\pkt' \in (\denote{c_1}(\pkt|_V))|_V} (\denote{c_2} \pkt')|_V
+    & \text{by IH}(c_1) \\
+    = \bigcup_{\pkt' \in \denote{c_1}(\pkt|_V)} (\denote{c_2} (\pkt'|_V))|_V
+    & \text{by set theory} \\
+    = \bigcup_{\pkt' \in \denote{c_1}(\pkt|_V)} ((\denote{c_2} \pkt')|_V)|_V
+    & \text{by IH}(c_2) \\
+    = \bigcup_{\pkt' \in \denote{c_1}(\pkt|_V)} (\denote{c_2} \pkt')|_V
+    & \text{by Lemma~\ref{lem:proj-idem-set}} \\
+    = \left(\bigcup_{\pkt' \in \denote{c_1}(\pkt|_V)} \denote{c_2}\pkt'\right)|_V
+    & \text{by Lemma~\ref{lem:proj-union-set}} \\
+    = \denote{c_2}^\dagger(\denote{c_1}(\pkt|_V))|_V
+    & \text{By definition of }(\cdot)^\dagger\\
+    = (\denote{c_1;c_2}(\pkt|_V))|_V
+    & \text{By definition of }\denote\cdot
+    \end{array}\]
+    The result follows by definition of $=_V$.
+
+  \item[$(c = c_1 \choice c_2)$] Let $\pkt$ be given.
+    \[\begin{array}{ll}
+    (\denote{c_1\choice c_2}\pkt)|_V \\
+    = (\denote{c_1}\pkt \cup \denote{c_2}\pkt)|_V
+    & \text{By definition of }\denote\cdot \\
+    = (\denote{c_1}\pkt)|_V \cup (\denote{c_2}\pkt)|_V
+    & \text{By Lemma~\ref{lem:proj-union-set}} \\
+    = (\denote{c_1}(\pkt|_V))|_V \cup (\denote{c_2}(\pkt|_V))|_V
+    & \text{By IHs} \\
+    = ((\denote{c_1}(\pkt|_V) \cup \denote{c_2}(\pkt|_V))|_V)|_V
+    & \text{By Lemma~\ref{lem:proj-union-set}} \\
+    = (\denote{c_1}(\pkt|_V) \cup \denote{c_2}(\pkt|_V))|_V
+    & \text{By Lemma~\ref{lem:proj-idem-set}}\\
+    = (\denote{c_1 \choice c_2}(\pkt|_V))|_V
+    & \text{By definition of }\denote\cdot \\
+    \end{array}\]
+
+    The result follows by definition of $=_V$.
+
+  \item[$(c = t.\apply())$] Contradictory; $t.\apply()$ is not an action.
+  \end{enumerate}
+
+\end{proof}
+
 \subsection{Lemmas}
@@
 
+\begin{lemma}[Pseudo-Parametricity]
+  \label{lem:parametricity}
+  For $c \in \Cmd$, $\sigma : \State$, $y \in \Var$, $\pkt \in \Pkt$, if $y
+  \not\in \fvs(c)$, then $\forall \pkt' \in \denote{c}^\sigma \pkt$, $\pkt'(y) =
+  \pkt(y)$.
+\end{lemma}
+
+\begin{proof}
+  Let $\sigma : \State $, $y \in \Var$, and $\pkt \in \Pkt$ be given.
+  Proceed by induction on $c$.
+  \begin{enumerate}[align=left]
+  \item[($c = x := e$)] Assume $y \not \in \fvs(x:=e)$, that is, $y \neq x$.
+    Compute $\denote{x:=e}^\sigma \pkt = \{\pkt\{ x \mapsto \edenote e
+    \pkt\}\}$. Observe that $\pkt\{ x \mapsto \edenote e \pkt\}(y) = \pkt(y)$
+    since $y \neq x$. The result follows.
+
+  \item[($c = \havoc x$)] Assume $y \not \in \fvs(\havoc x)$, that is $y \neq
+    x$. Compute $\denote{\havoc x}^\sigma\pkt = \{\pkt\{x \mapsto v\} | v \in
+    \Value\}$. let $v \in \Value$ be arbitrary. Observe that since $x \neq y$,
+    $\pkt\{x \mapsto v\}(y) = \pkt(y)$. The result follows.
+
+  \item[($c = \assume b$)] Assume $y \not \in \fvs(\assume b)$, that is $y
+    \not\in \bfvs(b)$. Compute $\denote{\assume b}^\sigma \pkt = \{\pkt \mid
+    \pkt \models b\}$. If $\pkt \not\models b$ the result holds vacuously.
+    Otherwise the result follows from $\pkt(y) = \pkt(y)$.
+
+  \item[$(c = \assert b)$] \textit{sim}.
+
+  \item[$(c = c_1;c_2)$] Assume $y \not \in \fvs(c_1;c_2)$. Consequently $y
+    \not\in \fvs(c_1)$, and $y \not \in \fvs(c_2)$. By IH$(c_1)$, $\forall
+    \pkt_1 \in \denote{c_1}^\sigma(\pkt)$, $\pkt_1(y) = pkt(y)$. Pick an
+    arbitrary $\pkt_1 \in \denote{c_1}^\sigma(\pkt)$. By IH$(c_2)$, $\forall
+    \pkt_2 \in \denote{c_1}^\sigma(\pkt_1)$, $\pkt_2(y) = \pkt_1(y)$. By
+    transitivity, we have $\pkt_2(y) = \pkt(y)$. Conclude that $\forall \pkt_2
+    \in {\denote{c_2}^\sigma}^\dagger (\denote{c_1}^\sigma \pkt)$, $\pkt_2(y) =
+    \pkt_1(y)$.
+
+  \item[$(c = c_1 \choice c_2)$] Assume $y \not \in \fvs(c_1 \choice c_2)$.
+    Consequently, $y \not \in \fvs(c_1)$ and $y \not \in \fvs(c_2)$.
+    By IH$(c_i)$, $\forall \pkt' \in \denote(c_i)^\sigma\pkt$, for $i = 1,2$.
+    By definition of $\denote{\cdot}$, the result follows.
+
+  \item[$(t.\apply())$] Assume $y \not\in \fvs(t.\apply())$.
+    There are two cases. Either $\pkt$ hits in the table or it misses.
+    \begin{enumerate}[align=left]
+    \item[\textit{Case 1}.] Assume that there is a row $r$ such that $\pkt
+      \models t.\matchrow(r)$. Then $\denote{t.\apply()}^\sigma \pkt =
+      \denote{t.\action[r.\action](r.\data)} \pkt$. The result follows by the
+      IH.
+
+    \item[\textit{Case 2}.] Conversely, there is no such row, in which case
+      $\denote{t.\apply()}^\sigma\pkt = \{\pkt\}$, in which case the result is
+      trivial.
+    \end{enumerate}
+  \end{enumerate}
+\end{proof}
+
+\begin{lemma}[Unused-Variable Expression Noninterference]
+  \label{lem:exp-param-noninter}
+  For $e \in \BVExpr$, $\pkt \in \Pkt$,
+  $\{v_1,v_2\} \subseteq \Value$, $y \in \Var \setminus \efvs(e) $,
+  $\edenote{e}\pkt\{y \mapsto v_1\} = \edenote{e}pkt\{y \mapsto v_2\}$.
+\end{lemma}
+\begin{proof}
+  Let $\pkt$, $v_1$, $v_2$, and $y$ be given as above. Proceed by induction on
+  $e$.
+  \begin{enumerate}[align=left]
+    \item[$(e = v)$] Compute: $\edenote{v}\pkt\{y
+      \mapsto v_1\} = v = \edenote{v}pkt\{y \mapsto v_2\}$.
+    \item[$(e = x)$] Since $y \not\in \efvs(x)$, conclude $y \neq x$. Compute:
+      $\edenote{x}\pkt\{y \mapsto v_i\} = \pkt\{y \mapsto v_i\}(x) = \pkt(x) $
+      for $i = 1,2$. The result follows.
+
+    \item[$(e = e_1 \binop e_2)$]
+      Note that $\efvs(e_i) \subseteq \efvs(e_1 \binop e_2)$ for $i = 1,2$. Compute
+      \[\begin{array}{ll}
+      \edenote{e_1 \binop e_2}\pkt\{y \mapsto v_1\} \\
+      = \edenote{e_1}\pkt{y \mapsto v_1} \binop \edenote{e_2}\pkt\{y \mapsto v_1\} \\
+      = \edenote{e_1}\pkt{y \mapsto v_2} \binop \edenote{e_2}\pkt\{y \mapsto v_2\} & \text{By IH}(e_i), i = 1,2\\
+      = \denote{e_1 \binop e_2}\pkt\{y \mapsto v_2\}
+      \end{array}
+      \]
+
+    \item[$(e = \unop e)$]
+      Note that $\efvs(e) \subseteq \efvs(\unop e)$. Compute:
+      \[\begin{array}{ll}
+      \edenote{\unop e}\pkt\{y \mapsto v_1\} \\
+      = \unop \edenote{e}\pkt\{y \mapsto v_1\} \\
+      = \unop \denote{e}\pkt\{y \mapsto v_2\} & \textit{by IH}(e)\\
+      = \denote{\unop e}\pkt\{y \mapsto v_2\}
+      \end{array}\]
+  \end{enumerate}
+\end{proof}
+
+\begin{lemma}[Unused-Variable Boolean Noninterference]
+  \label{lem:bool-noninter}
+  For $b \in \BExpr$, $\pkt \in \Pkt$,
+  $\{v_1,v_2\} \subseteq \Value$, $y \in \Var \setminus \bfvs(b) $,
+  $\pkt\{y \mapsto v_1\} \models b \iff pkt\{y \mapsto v_2\} \models b$.
+\end{lemma}
+
+\begin{proof}
+  Let $\pkt$, $v_1$, $v_2$, and $y$ be given as above.
+  Proceed by induction on $b$:
+  \begin{enumerate}[align=left]
+  \item[($b = \FALSE$)] Trivial.
+  \item[($b = b_1 \Rightarrow b_2$)] Since $y \not \in\bfvs(b_1 \Rightarrow b_2) =
+    \bfvs(b_1) \cup \bfvs(b_2)$, $y \not \in \bfvs(b_i)$, for $i = 1,2$. Compute
+    as follows:
+    \[\begin{array}{ll}
+    \pkt\{y \mapsto v_1\} \models b_1 \Rightarrow b_2 \\
+    \iff \pkt\{y \mapsto v_1\} \not\models b_1 \text{ or } \pkt\{y \mapsto v_1\} \models b_2 \\
+    \iff \pkt\{y \mapsto v_2\} \not\models b_1 \text{ or } \pkt\{y \mapsto v_2\} \models b_2
+    & \text{by IHs since } y \not \in \bfvs(b_i), i = 1,2 \\
+    \iff \pkt\{y \mapsto v_2\} \models b_1 \Rightarrow b_2
+    \end{array}\]
+
+  \item[$(b = e_1 = e_2)$] Since $y \not \in\bfvs(e_1 = e_2) = \efvs(e_1) \cup
+    \efvs(e_2)$, $y \not \in \bfvs(e_i)$ for $i =1,2$. Compute as follows:
+    \[\begin{array}{ll}
+    \pkt\{y \mapsto v_1\} \models e_1 = e_2
+    \iff \edenote{e_1}\pkt\{y \mapsto v_1\} = \edenote{e_2}\pkt\{y \mapsto v_1\} \\
+    \iff\edenote{e_1}\pkt\{y \mapsto v_2\} = \edenote{e_2}\pkt\{y \mapsto v_2\} & \text{by Lemma~\ref{lem:exp-param-noninter}} \\
+    \iff \pkt\{y\mapsto v_2\} \models e_1 = e_2
+    \end{array}
+    \]
+  \end{enumerate}
+\end{proof}
+
+\begin{lemma}[Unused Variable Noninterference]
+  \label{lem:param-noninter}
+  For $c \in \Cmd$, $\sigma : \State$, $\pkt \in \Pkt$,
+  $\{v_1,v_2\} \subseteq \Value$, $y \in \Var \setminus \fvs(c)$,
+  $\denote{c}^\sigma \pkt\{y \mapsto v\} =_{\Var \setminus \{y\}}\denote{c}^\sigma\pkt\{y \mapsto v\}$.
+\end{lemma}
+
+\begin{proof}
+  Let $\sigma$, $v_1$, $v_2$, and $y$ be given.
+  Proceed by induction on $c$, leaving $\pkt$ and $x$ general.
+  \begin{enumerate}[align=left]
+  \item[$(c = z := e)$]
+    Let $\pkt$ be given. Since $y \not\in \fvs(z := e)$ we know that $y \neq z$.
+    Compute:
+    \[\begin{array}{ll}
+    \pkt' \in \denote{z:=e}^\sigma (\pkt\{y \mapsto v\}) \\
+    \iff \pkt' \in \{\pkt\{y \mapsto v\}\{z \mapsto \edenote{e}(\pkt\{y\mapsto v\})\}\}\\
+    \iff \pkt' = \pkt\{y \mapsto v\}\{z \mapsto \edenote{e}(\pkt\{y\mapsto v\})\}
+    \end{array}
+    \]
+    Similarly, compute:
+    \[\begin{array}{ll}
+    \pkt'' \in \denote{z:=e}^\sigma (\pkt\{y \mapsto v'\})\\
+    \iff \pkt'' \in \{\pkt\{y \mapsto v'\}\{z \mapsto \edenote{e}(\pkt\{y\mapsto v'\})\}\}\\
+    \iff \pkt'' = \pkt\{y \mapsto v'\}\{z \mapsto \edenote{e}(\pkt\{y\mapsto v'\})\}
+    \end{array}
+    \]
+
+    We need to show that
+    \[\forall x \in \Var \setminus \{y\}.
+    \pkt\{y \mapsto v'\}\{z \mapsto \edenote{e}(\pkt\{y\mapsto v'\})(x)
+    = \pkt\{y \mapsto v\}\{z \mapsto \edenote{e}(\pkt\{y\mapsto v\})\}(x)
+    \]
+
+    Equivalently we can show
+    \[\forall x \in \Var \setminus \{y\}.
+    \pkt\{z \mapsto \edenote{e}(\pkt\{y\mapsto v'\})(x)
+    = \pkt\{z \mapsto \edenote{e}(\pkt\{y\mapsto v\})\}(x)
+    \]
+
+    Let $x \neq y$ be given. The result is immediate if $x \neq z$. Otherwise we
+    must show that $\edenote{e}(\pkt\{y \mapsto v\}) = \edenote{e}(\pkt\{y
+    \mapsto v'\})$. Since $y \not \in \efvs(e)$, this follows by
+    Lemma~\ref{lem:exp-param-noninter}.
+
+  \item[$(c = \havoc z)$] Let $\pkt$ by given. Since $y \not \in \fvs(\havoc z)$, we know $y \neq z$.
+    Compute:
+    \[\begin{array}{ll}
+    \denote{\havoc z}^\sigma\pkt\{y \mapsto v_1\} \\
+    = \{\pkt\{y \mapsto v_1\}\{z \mapsto v_z\} \mid v_z \in \Value\}\\
+    =_{\Var \setminus \{y\}} \{\pkt\{y \mapsto v_2\}\{z \mapsto v_z\} \mid v_z \in \Value\} \\
+    = \denote{\havoc z}^\sigma\pkt\{y \mapsto v_2\}
+    \end{array}
+    \]
+
+    The result follows by Lemma~\ref{lem:equal-proj}.
+
+  \item[$(c = \assume b)$]
+    Let $\pkt$ by given. Since $\bfvs(b) \subseteq \fvs(\assume b)$, $y \not\in \bfvs(b)$.
+    Compute \[
+    \begin{array}{ll}
+      \denote{\assume b}^\sigma \pkt\{y \mapsto v_1\} \\
+      = \{\pkt \mid \pkt\{y \mapsto v_1\} \models b\}\\
+      =_{\Var \setminus \{y\}} \{\pkt \mid \pkt\{y \mapsto v_2\} \models b\} & \text{by Lemma~\ref{lem:bool-noninter}, since }y \not\in \bfvs(b) \\
+      = \denote{\assume b}^\sigma \pkt\{y \mapsto v_2\}
+    \end{array}
+    \]
+
+    The result follows by Lemma~\ref{lem:equal-proj}.
+
+    \item[$(c = \assert b)$] \textit{sim}.
+    \item[$(c = c_1;c_2)$] Observe that $\fvs(c_i) \subseteq \fvs(c_1;c_2)$ for
+      $i = 1,2$, which enables both IHS. Observe that:
+      \[\begin{array}{ll}
+      \denote{c_1;c_2}^\sigma\pkt\{y \mapsto v_1\}
+      = {\denote{c_2}^\sigma}^\dagger\left(\denote{c_1}^\sigma\pkt\{y\mapsto v_1\}\right) \\
+      \end{array}
+      \]
+      If $c_1$ denotes $\error$, the result is immediate, so assume otherwise.
+
+      Let $\{i,j\} \subseteq \{1,2\}$. Let $\pkt_i \in
+      \denote{c_1}^\sigma\pkt\{y \mapsto v_i\}$ be arbitary. By
+      Lemma~\ref{lem:parametricity}, we know that $\pkt_i(y) = v_i$. Further
+      IH$(c_1)$ tells us that there is some $\pkt_j \in
+      \denote{c_1}^\sigma\pkt\{y\mapsto v_j\}$ such that $\pkt_1 =_{\Var
+        \setminus \{y\}} \pkt_j$. Again, IH$(c_1)$ tells us that $\pkt_j(y) =
+      v$. Consequently, we can write $\pkt_i$ and $\pkt_j$ in terms of a another
+      packet $\pkt'$, i.e $\pkt_k = \pkt'\{y \mapsto v_k\}$ for $k = i,j$. Then
+      by IH$(c_2)$, $\denote{c_2}^\sigma\pkt'\{y \mapsto v_i\} =_{\Var \setminus \{y\}}
+      \denote{c_2}^\sigma \pkt'\{y \mapsto v_j\}$. Conclude that
+      $\denote{c_1}^\sigma\pkt\{y \mapsto v_i\} \subseteq_{\Var \setminus \{y\} }
+      \denote{c_1}^\sigma\pkt\{y \mapsto v_j\}$.
+
+      We just proved that the projected subset is bidrectional, i.e. that
+      $\denote{c_1}^\sigma\pkt\{y \mapsto v_1\} =_{\Var \setminus\{y\}}
+      \denote{c_1}^\sigma\pkt\{y \mapsto v_2\}$
+
+    \item[$(c = c_1 \choice c_2)$] Observe that $\fvs(c_i) \subseteq \fvs(c_1 \choice c_2)$ for
+      $i = 1,2$, which enables both IHs. Compute as follows
+      \[\begin{array}{ll}
+      \denote{c_1 \choice c_2}^\sigma\pkt\{y \mapsto v_1\} \\
+      = \denote{c_1}^\sigma\pkt\{y \mapsto v_1\} \cup \denote{c_2}^\sigma\pkt\{y \mapsto v_1\} \\
+      =_{\Var \setminus \{y\}} \denote{c_1}^\sigma\pkt\{y \mapsto v_2\} \cup \denote{c_2}^\sigma\pkt\{y \mapsto v_2\}
+      & \text{by IHs} \\
+      = \denote{c_1 \choice c_2}
+      \end{array}
+      \]
+
+      The result follows by Lemma~\ref{lem:equal-proj}.
+
+    \item[$(c = t.\apply())$]
+      There are two cases, corresponding to hitting and missing.
+      \begin{enumerate}[align=left]
+      \item[\textit{Case 1}.] Assume there is some $r \in \sigma(t)$ such that
+        $\pkt\{y \mapsto v_1\} \models t.\matchrow(r)$, and that $r$ is the
+        minimally-indexed such row. Then by Lemma~\ref{lem:bool-noninter},
+        $\pkt\{y \mapsto v_2\} \models t.\matchrow(r)$, and is the minimally indexed such row. Then, since
+        $\fvs(t.\action[r.\action](r.\data)) \subseteq \fvs(t.\apply())$, the
+        result follows by IH$(t.\action[r.\action](r.\data))$.
+
+        \item[\textit{Case 2}.] Assume there is no $r \in \sigma(t)$ such that
+          $\pkt\{y \mapsto v_1\} \models t.\matchrow(r)$. Then by
+          Lemma~\ref{lem:bool-noninter}, there is now $r \in \sigma(t)$ such
+          that $\pkt\{y \mapsto v_2\} \models t.\matchrow(r)$. Then by definitions:
+          \[\denote{t.\apply()}^\sigma \pkt\{y \mapsto v_1\} = \{\pkt\{y \mapsto v_1\}\} = \{\pkt\{y \mapsto v_2\}\} = \denote{t.\apply()}^\sigma \pkt\{y \mapsto v_2\}\]
+          The result follows by Lemma~\ref{lem:equal-proj}.
+      \end{enumerate}
+
+  \end{enumerate}
+\end{proof}
+
 \begin{lemma}[Disjoint Commutativity]
   \label{lem:disj-comm}
-  For $c_1,c_2 \in \Cmd$, and $\sigma : \Value \to \Row^*$ if $\fvs(c_1) \cap
+  For $c_1,c_2 \in \Cmd$, and $\sigma \in \State$ if $\fvs(c_1) \cap
   \fvs(c_2) = \emptyset$, $ \forall \pkt \in \Pkt. \denote{c_1;c_2}^\sigma\pkt =
@@
 \begin{proof}
-  %% Let $c_1,c_2$, $\sigma$, and $\pkt$ be given as above. Assume $\fvs(c_1) \cap
-  %% \fvs(c_2) = \emptyset$. Compute:
-  %% \[
-  %% \begin{array}{ll}
-  %%   \denote{c_1;c_2}^\sigma \pkt =
-
-  %% \end{array}
-  %% \]
-  \todo[inline]{prove}
+  \todo[inline]{todo}
 \end{proof}
@@
 
-\begin{definition}[Projections]
-  For a set of variables $S \subseteq \Var$, and a packet $\pkt$, write
-  $\pkt|_S$ to be projection of $\pkt$ onto $S$. \todo{I guess this means
-    packets are partial functions?} Formally,
-  \[\pkt|_S \triangleq \{x \mapsto \pkt(x) \mid x \in S\}.\]
-  Two packets $\pkt$, $\pkt'$ are equivalent up to $S \subseteq \Var$, written
-  $\pkt =_S \pkt'$, iff $\pkt|_S = \pkt'|_s$. A set of packets $P$ is a subset
-  of $P'$ up to $S$, written $P \subseteq_S P'$ iff for every packet $\pkt \in
-  P$, there exists $\pkt' \in P'$ such that $\pkt =_S \pkt'$. Two sets of
-  packets $P$ and $P'$ are $S$-equivalent, written $P =_S P'$ iff $P \subseteq_S
-  P'$ and $P \subseteq_S P'$. A set $P \subseteq \Pkt$ is a $V$-equivalence
-  class if $\forall \pkt_1, \pkt_2 \in P$, $\pkt_1 =_V \pkt_2$.
-\end{definition}
-
-\begin{lemma}[Projection Union-Distributivity]
-  \label{lem:proj-union-set}
-  For all $P,Q \subseteq \Pkt$, $V \subseteq \Var$,
-  $(P\cup Q)|_V = P|_V \cup Q|_V$
-\end{lemma}
-
-\begin{proof}
-  Let $P,Q \subseteq \Pkt$ and $V \subseteq \Var$ be given. Compute:
-
-  \[\begin{array}{ll}
-  (P\cup Q)|_V\\
-  = \{\pkt|_V \mid \pkt \in P\cup Q\}\\
-  = \{\pkt|_V \mid \pkt \in P\}\cup \{\pkt|_V \mid \pkt \in Q\}\\
-  = P|_V \cup Q|_V
-  \end{array}\]
-\end{proof}
-
-\begin{lemma}[Projection Idempotence]
-  \label{lem:proj-idem}
-  For all $\pkt \in \Pkt$, $V \subseteq \Var$,
-  $(\pkt|_V)|_V = \pkt|_V$
-\end{lemma}
-
-\begin{proof}
-  Let $\pkt \in \Pkt$ and $V \subseteq \Var$ be given.
-  \[\begin{array}{ll}
-    (\pkt|_V)|_V \\
-    = \{x \mapsto (\pkt|_V)(x) \mid x \in V\} & \text{by definition}\\
-    = \{x \mapsto (\{y \mapsto \pkt(y) \mid y \in V\})(x) \mid x \in V\} & \text{by definition}\\
-    = \{x \mapsto \pkt(x) \mid x \in V\} & \beta\text{-reduction}\\
-    = \pkt|_V & \text{by definition}
-  \end{array}\]
-\end{proof}
-
-\begin{lemma}[Projection Idempotence for Sets]
-  \label{lem:proj-idem-set}
-  For all $P \subseteq \Pkt$, $V \subseteq \Var$,
-  $(P|_V)|_V = P|_V$
-\end{lemma}
-
-\begin{proof}
-  Let $P \subseteq \Pkt$ and $V \subseteq \Var$ be given. Compute:
-  \[\begin{array}{ll}
-  (P|_V)|_V \\
-  = \{\pkt|_V \mid \pkt \in P|_V\}
-  & \text{by definition of } (\cdot)|_V : \mathcal{P}(\Pkt) \to \mathcal{P}(\Pkt) \\
-  = \{\pkt|_V \mid \pkt \in \{\pkt'|_V \mid \pkt'\in P\}\}
-  & \text{by definition of } (\cdot)|_V : \mathcal{P}(\Pkt) \to \mathcal{P}(\Pkt) \\
-  = \{(\pkt|_V)|_V \mid \pkt \in P \}
-  & \text{by set theory} \\
-  = \{\pkt|_V \mid \pkt \in P\} & \text{by Lemma~\ref{lem:proj-idem}} \\
-  = P|_V & \text{by definition of } (\cdot)|_V : \mathcal{P}(\Pkt) \to \mathcal{P}(\Pkt)
-  \end{array}\]
-\end{proof}
-
-\begin{lemma}{Subset Implies Projected Subset}
-  \label{lem:subset-proj}
-  For $P,Q \subseteq \Pkt$, and $ V \subseteq \Var$,
-  If $P \subseteq Q$ then $P \subseteq_V Q$.
-\end{lemma}
-
-\begin{proof}
-Let $P,Q \subseteq \Pkt$ and $V \subseteq \Var$ be given. Assume $P \subseteq Q$,
-to show that $P \subseteq_V Q$.
-
-Let $\pkt \in P$. Since $P \subseteq Q$, $\pkt \in Q$. Notice that $\pkt =_V
-\pkt$, so $P \subseteq_V Q$.
-\end{proof}
-
-\begin{lemma}{Equality Implies Projected Equality}
-  \label{lem:equal-proj}
-  For $P,Q \subseteq \Pkt$, and $ V \subseteq \Var$,
-  If $P = Q$ then $P =_V Q$.
-\end{lemma}
-
-\begin{proof}
-  By Lemma~\ref{lem:subset-proj}
-\end{proof}
-
-\begin{lemma}[Expression Projection]
-  \label{lem:proj-expr}
-  For every set of variables $V \subseteq \Var$, packet $\pkt \in \Pkt$ and
-  expression $e \in \BVExpr$, such that $\efvs(e) \subseteq V$, $\edenote{e}\pkt =
-  \edenote{e}(\pkt|_V)$.
-\end{lemma}
-
-\begin{proof}
-  Let $V \subseteq \Var$ be given.
-  Proceed by induction on $e$, leaving $\pkt$ arbitrary.
-  \begin{enumerate}[align=left]
-  \item[($e = v$)] Let $\pkt$ be given. $\edenote{v}\pkt = v = \edenote{v}{\pkt|_V}$
-  \item[($e = x$)] Let $\pkt$ be given. Note that $\efvs(x) = \{x\} \subseteq V$. Compute:
-    \[\begin{array}{ll}
-    \edenote{x}(\pkt|_V) \\
-    = (\pkt|_V)(x)
-    & \text{by definition of }\edenote\cdot\\
-    = \{y \mapsto \pkt(y) \mid y \in V\}(x)
-    & \text{by definition of }(\cdot)|_V : \Pkt \to \Pkt\\
-    = \pkt(x) & \text{since } x \in V
-    \end{array}
-    \]
-  \item[($e = e_1 \binop e_2$)] Let $\pkt$ be given. Compute:
-    \[\begin{array}{ll}
-    \edenote{e_1 \binop e_2}\pkt \\
-    = \edenote{e_1}\pkt \binop \edenote{e_2}\pkt & \text{definition of }\edenote\cdot \\
-    = \edenote{e_1}(\pkt|_V) \binop \edenote{e_2}(\pkt|_V) & \text{by IHS since }\efvs(e_i) \subseteq \efvs(e_1 \binop e_2) \subseteq V, i = 1,2 \\
-    = \edenote{e_1 \binop e_2}(\pkt|_V) & \text{definition of }\edenote\cdot
-    \end{array}\]
-  \item[($e = \unop e$)] Let $\pkt$ be given. Compute:
-    \[\begin{array}{ll}
-    \edenote{\unop e}\pkt \\
-    = \unop \edenote{e}\pkt & \text{definition of }\edenote\cdot \\
-    = \unop \edenote{e}(\pkt|_V) & \text{by IH, since } \efvs(e) = \efvs(\unop e) \subseteq V \\
-    = \edenote{\unop e}(\pkt|_V) & \text{definition of }\edenote\cdot
-    \end{array}\]
-  \end{enumerate}
-\end{proof}
-
-\begin{lemma}[Satisfaction Projection]
-  \label{lem:proj-satis}
-  For every set of variables $V \subseteq \Var$, packet $\pkt \in \Pkt$ and
-  boolean expression $b \in \BExpr$, such that $\bfvs(b) \subseteq V$,
-  $\pkt \models b$ iff $\pkt|_V \models b$.
-\end{lemma}
-
-\begin{proof}
-  Let $V \subseteq \Var$ be given.
-  Proceed by induction on $b$, leaving $\pkt$ arbitrary.
-  \begin{enumerate}[align=left]
-  \item[($b = \FALSE$)] Immediate.
-  \item[($b = b_1 \Rightarrow b_2$)]
-    Let $\pkt$ be given. Compute:
-    \[\begin{array}{ll}
-    \pkt \models b_1 \Rightarrow b_2 \\
-    \iff \pkt \not\models b_1 \text{or} \pkt\models b_2 & \text{by definition of } \models   \\
-    \iff \pkt|_V \not\models b_1 \text{or} \pkt|_V \models b_2  & \text{by IHs, since} \bfvs(b_i) \subseteq \bfvs(b_1 \Rightarrow b_2) \subseteq V, i = 1,2  \\
-    \iff \pkt|_V \models b_1 \Rightarrow b_2
-    & \text{by definition of } \models
-    \end{array}\]
-
-  \item[($b = e_1 = e_2$)] Let $\pkt$ be given. Compute:
-    \[\begin{array}{ll}
-    \pkt \models e_1 = e_2 \\
-    \iff \edenote{e_1}\pkt = \edenote{e_2}\pkt
-    & \text{by definition of }\models \\
-    \iff \edenote{e_1}(\pkt|_V) = \edenote{e_2}(\pkt|_V)
-    & \text{by Lemma~\ref{lem:proj-expr}, since } \efvs(e_i) \subseteq \bfvs(e_1 = e_2) \subseteq V, i = 1,2 \\
-    \iff \pkt|_V \models e_1 = e_2
-    & \text{by definition of }\models
-    \end{array}\]
-
-  \end{enumerate}
-
-\end{proof}
-
-\begin{lemma}[Projection-Denotation Commutativity]
-  \label{lem:proj-den-comm}
-  For a closed action $c \in \Cmd$, $\fvs(c) \subseteq V \subseteq \Var$ and
-  $\pkt \in \Pkt$, $\denote{c} \pkt =_V \denote{c} \pkt|_V$.
-\end{lemma}
-
-\begin{proof}
-  Proceed by induction on $c$, leaving $\pkt$ general.
-  \begin{enumerate}[align=left]
-  \item[$(c = x:=e)$] Let $\pkt$ be given. Compute:
-    \[\begin{array}{ll}
-    (\denote{x:=e} (\pkt|_V))|_V\\
-    = ((\pkt|_V)\{x:=\edenote{e}(\pkt|_V)\})|_V
-    & \text{definition} \\
-    = ((\pkt|_V)\{x := \edenote{e}\pkt\})|_V
-    & \text{By Lemma~\ref{lem:proj-expr} since } \fvs(e) \subseteq \fvs(x := e) \subseteq V \\
-    = (\pkt\{x := \edenote{e}\pkt\}|_V)|_V
-    & \text{since } x \in \fvs(x:=e) \subseteq V \\
-    = ((\denote{x:=e}\pkt)|_V)|_V
-    & \text{by definition of }\denote{\cdot}\\
-    = (\denote{x:=e}\pkt)|_V
-    & \text{by Lemma~\ref{lem:proj-idem}}
-    \end{array}\]
-    The result follows by definition of $=_V$.
-
-  \item[$(c = \havoc x)$] Contradictory; $\havoc x$ is not an action.
-  \item[$(c = \assume b)$]
-    Let $\pkt$ be given. Compute:
-    \[\begin{array}{ll}
-    (\denote{\assume b} \pkt)|_V \\
-    = (\{\pkt \mid \pkt \models b\})|_V
-    & \text{definition of}\denote\cdot\\
-    = ((\{\pkt \mid \pkt \models b\})|_V)|_V
-    & \text{By Lemma~\ref{lem:proj-idem-set}} \\
-    = (\{\pkt|_V \mid \pkt \models b\})|_V
-    & \text{definition of }(\cdot)|_V : \mathcal P (\Pkt) \to \mathcal P (\Pkt) \\
-    = (\{\pkt|_V \mid \pkt|_V \models b\})|_V
-    & \text{By Lemma~\ref{lem:proj-satis} since } \fvs(b) \subseteq \fvs(\assume b) \subseteq V \\
-    = (\denote{\assume b}(\pkt|_V))|_V
-    & \text{by definition of}\denote\cdot
-    \end{array}
-    \]
-    The result follows by definition of $=_V$.
-
-  \item[$(c = \assert b)$] \textit{sim.}
-  \item[$(c = c_1;c_2)$] Let $\pkt$ be given. Compute:
-    \[\begin{array}{>{\displaystyle}ll}
-    (\denote{c_1;c_2}\pkt)|_V \\
-    = (\denote{c_2}^\dagger(\denote{c_1}\pkt))|_V
-    & \text{definition of }\denote\cdot \\
-    = \left(\bigcup_{\pkt' \in \denote{c_1} \pkt} \denote{c_2}\pkt'\right)|_V
-    & \text{definition of }(\cdot)^\dagger \\
-    = \bigcup_{\pkt' \in \denote{c_1} \pkt} (\denote{c_2}\pkt')|_V
-    & \text{by Lemma~\ref{lem:proj-union-set}} \\
-    = \bigcup_{\pkt' \in \denote{c_1}\pkt} (\denote{c_2}(\pkt'|_V))|_V
-    & \text{IH}(c_2)  \\
-    =\bigcup_{\pkt' \in (\denote{c_1}\pkt)|_V} (\denote{c_2} \pkt')|_V
-    & \text{by set theory} \\
-    = \bigcup_{\pkt' \in (\denote{c_1}(\pkt|_V))|_V} (\denote{c_2} \pkt')|_V
-    & \text{by IH}(c_1) \\
-    = \bigcup_{\pkt' \in \denote{c_1}(\pkt|_V)} (\denote{c_2} (\pkt'|_V))|_V
-    & \text{by set theory} \\
-    = \bigcup_{\pkt' \in \denote{c_1}(\pkt|_V)} ((\denote{c_2} \pkt')|_V)|_V
-    & \text{by IH}(c_2) \\
-    = \bigcup_{\pkt' \in \denote{c_1}(\pkt|_V)} (\denote{c_2} \pkt')|_V
-    & \text{by Lemma~\ref{lem:proj-idem-set}} \\
-    = \left(\bigcup_{\pkt' \in \denote{c_1}(\pkt|_V)} \denote{c_2}\pkt'\right)|_V
-    & \text{by Lemma~\ref{lem:proj-union-set}} \\
-    = \denote{c_2}^\dagger(\denote{c_1}(\pkt|_V))|_V
-    & \text{By definition of }(\cdot)^\dagger\\
-    = (\denote{c_1;c_2}(\pkt|_V))|_V
-    & \text{By definition of }\denote\cdot
-    \end{array}\]
-    The result follows by definition of $=_V$.
-
-  \item[$(c = c_1 \choice c_2)$] Let $\pkt$ be given.
-    \[\begin{array}{ll}
-    (\denote{c_1\choice c_2}\pkt)|_V \\
-    = (\denote{c_1}\pkt \cup \denote{c_2}\pkt)|_V
-    & \text{By definition of }\denote\cdot \\
-    = (\denote{c_1}\pkt)|_V \cup (\denote{c_2}\pkt)|_V
-    & \text{By Lemma~\ref{lem:proj-union-set}} \\
-    = (\denote{c_1}(\pkt|_V))|_V \cup (\denote{c_2}(\pkt|_V))|_V
-    & \text{By IHs} \\
-    = ((\denote{c_1}(\pkt|_V) \cup \denote{c_2}(\pkt|_V))|_V)|_V
-    & \text{By Lemma~\ref{lem:proj-union-set}} \\
-    = (\denote{c_1}(\pkt|_V) \cup \denote{c_2}(\pkt|_V))|_V
-    & \text{By Lemma~\ref{lem:proj-idem-set}}\\
-    = (\denote{c_1 \choice c_2}(\pkt|_V))|_V
-    & \text{By definition of }\denote\cdot \\
-    \end{array}\]
-
-    The result follows by definition of $=_V$.
-
-  \item[$(c = t.\apply())$] Contradictory; $t.\apply()$ is not an action.
-  \end{enumerate}
-
-\end{proof}
-
-\begin{definition}[Variables]
-  The free variables of a command $c$, written $\fvs(c) \subseteq \Var$, is the
-  set of variables that occur at any point in $c$ (i.e. as the lhs of an
-  assignment, in a boolean expression, a bitvector expression, as a key in a
-  table.) The only bound variables are the action data in tables. Formally:
-  \[\begin{array}{l >{\triangleq}cl}
-  \efvs :: \BVExpr \to \mathcal{P}(\Var) \\
-  \efvs(v) && \{\} \\
-  \efvs(x) && \{x\}\\
-  \efvs(e_1 \binop e_2) && \efvs(e_1) \cup \efvs(e_2)\\
-  \efvs(\unop e) && \efvs(unop) \\[1em]
-  \bfvs :: \BExpr \to \mathcal{P}(\Var) \\
-  \bfvs(\FALSE) && \{\}\\
-  \bfvs(b_1 \Rightarrow b_2) && \bfvs(b_1) \cup \bfvs(b_2)\\
-  \bfvs(e_1 = e_2) && \efvs(e_1) = \efvs(e_2) \\[1em]
-  \fvs :: \Cmd \to \mathcal{P}(\Var) \\
-  \fvs(x := e) && \{x\} \cup \fvs(e)\\
-  \fvs(\havoc x) && \{x\}\\
-  \fvs(\assume b) && \fvs(b)\\
-  \fvs(\assert b) && \fvs(b) \\
-  \fvs(c_1;c_2) && \fvs(c_1) \cup \fvs(c_1) \\
-  \fvs(c_1 \choice c_2) && \fvs(c_1) \cup \fvs(c_2) \\
-  \fvs(t.\apply()) &&
-  t.\keys \cup \bigcup_{\lambda \vec x. c \in t.\actions} (\fvs(c) \setminus \{x_1,\ldots, x_n\})
-  \end{array}\]
-\end{definition}
-
 \begin{samepage}
@@
 
+%% \begin{lemma}[Universal Quantifier Weakness]
+%%   For $b \in \BExpr$ and $V \subseteq \bfvs(b)$, $\forall V. b$ is the weakest
+%%   formula on variables $\bfvs(b) \setminus V$ that implies $b$.
+%% \end{lemma}
+%% \begin{proof}
+%%   Elementary.
+%% \end{proof}
+
 \begin{theorem}[Quantifer Elimination]
@@
 
-\end{proof}
-
+  \todo[inline]{weakest?}
+\end{proof}
+
+\clearpage
 \section{Data-Driven Precondition Inference?}

### 2021-05-27 inference.tex
prove disjoint commutativity

@@
 
+\begin{lemma}[Set Pseudo-Parametricity]
+  \label{lem:parametricity-set}
+  For $c \in \Cmd$, $\sigma : \State$, $y \in \Var$, $\pkt \in \Pkt$, if $Y
+  \subseteq \Var \setminus \fvs(c)$, then $\forall \pkt' \in \denote{c}^\sigma
+  \pkt$, $\pkt'(y) = \pkt(y)$.
+\end{lemma}
+
+\begin{proof}
+By Lemma~\ref{lem:parametricity}.
+\end{proof}
+
 \begin{lemma}[Unused-Variable Expression Noninterference]
@@
     \[\begin{array}{ll}
-    \pkt\{y \mapsto v_1\} \models e_1 = e_2
+    \pkt\{y \mapsto v_1\} \models e_1 = e_2 \\
     \iff \edenote{e_1}\pkt\{y \mapsto v_1\} = \edenote{e_2}\pkt\{y \mapsto v_1\} \\
@@
 
+\begin{lemma}[Unused Variable Set Noninterference]
+  \label{lem:param-noninter-set}
+  For $c \in \Cmd$, $\sigma : \State$, $\pkt \in \Pkt$,
+  $\{v_1,v_2\} \subseteq \Value$, $Y \subseteq \Var \setminus \fvs(c)$,
+  $\denote{c}^\sigma \pkt\{y \mapsto v\} =_{\Var \setminus Y}\denote{c}^\sigma\pkt\{y \mapsto v\}$.
+\end{lemma}
+
+\begin{proof}
+  By Lemma~\ref{lem:param-noninter}.
+\end{proof}
+
 \begin{lemma}[Disjoint Commutativity]
@@
 \begin{proof}
-  \todo[inline]{todo}
+  Let $c_1, c_2$, $\sigma$, and $\pkt$ be given as above.
+  Let $\{i,j\} = \{1,2\}$. Observe:
+  \[\begin{array}{ll}
+  \denote{c_i;c_j}^\sigma \pkt
+  = {\denote{c_j}\sigma}^\dagger\left(\denote{c_i}^\sigma \pkt\right)
+  \end{array}\]
+
+  Let $V = \Var \setminus \fvs(c_1;c_2)$. Let $P_1 = \denote{c_1}^\sigma \pkt$,
+  and $P_2 = \denote{c_2}^\sigma \pkt$. By Lemma~\ref{lem:parametricity-set}, the following hold
+  \[P_1 =_{\fvs(c_2) \cup V} \{\pkt\}\] \[P_2 =_{\fvs(c_1) \cup V} \{\pkt\}\]
+  So, Let $P_{12} = {\denote{c_2}^\sigma}^\dagger P_1$, and $P_{21} =
+  {\denote{c_1}^\sigma}^\dagger P_2$. Again, by
+  Lemma~\ref{lem:parametricity-set}, the following equivalences hold
+  \[P_{12} =_{\fvs(c_1) \cup V} P_1\] \[P_{21} =_{\fvs(c_2) \cup V} P_2\]
+  Now by Lemma~\ref{lem:proj-set-subset}, we know $P_{12} =_V \{\pkt\} = P_{21}$.
+
+  Now by Lemma~\ref{lem:param-noninter-set} we know
+  \[P_{21} =_{\fvs(c_1)} P_1\] \[P_{12} = \fvs(c_2) P_2\]
+
+  Then Lemma~\ref{lem:proj-set-subset} gives
+  \[P_{12} =_{\fvs(c_1)} {P_1}\] \[P_{21} =_{\fvs(c_2)} {P_2}\]
+
+  Finally Lemma~\ref{lem:union} proves our goal:
+  \[P_{12} =_{\fvs(c_1) \cup \fvs(c_2) \cup V} P_{21}\]
+
 \end{proof}

### 2021-05-28 inference.tex
close holes in proofs

--- 2021-05-26 inference.tex
@@
 \newcommand{\SymbRow}{\mathsf{SymbRow}}
+\newcommand{\Ghost}{\mathsf{Ghost}}
 \newcommand{\State}{\mathsf{State}}
@@
 % Theorems
+\theoremstyle{plain}
 \newtheorem{theorem}{Theorem}
@@
 \newtheorem{conjecture}{Conjecture}
+\newtheorem{corollary}{Corollary}
+
+\theoremstyle{definition}
 \newtheorem{definition}{Definition}
+
+\theoremstyle{remark}
+\newtheorem{remark}{Remark}
 
@@
 
+\paragraph{Variables}
+Dataplane programs use variables $x \in \Var$ to represent packet state.
+
+\paragraph{Values}
+Values $v \in \Value$ are bitvector literals, written $[n]_w$, where $n$
+is a natural value and $w$ is the width. We say that $[n]_w = [m]_l$ is
+undefined if $w \neq l$, and otherwise is true iff $n \equiv m \mod 2^w$.
+
 \paragraph{Packets}
 Our dataplane programs will reason about structured representations of packets,
-that is, they will be transformations on sets of packet headers. In P4 programs,
-packet headers $h \in \Hdr$ are essentially structs with fields $f \in \Field$.
-Write $h.f$ for a header $h$ and a field $f$ is the value $v$ of $f$ in header
-$h$. Our simple command language just treats ``h.f'' as a variable.
+that is, they will be transformations on sets of packet headers. We define
+Packets to be functions from variables to values, that is $\Pkt = \Value^\Var$.
+Use the syntax $\pkt\{x \mapsto v\}$ to represent the packet that is equivalent
+to $\pkt$ on all variables fields except $x$, which maps to $v$. Use $\pkt(x)$
+to look up the value of $x$ in packet $\pkt$.
 
@@
 \begin{enumerate}[align=left]
-  \item[($t.\id$)] A table's unique identifier. A well-formed program can only have
-    one occurence of each table identifier.
+  \item[($t.\id$)] A table's unique identifier. A well-formed program can only
+    have one occurence of each table identifier.
   \item[($t.\keys$)] A list of header field accesses indicating the match keys and their match kind.
@@
 \paragraph{Symbolic Table Rows}
-A symbolic table row $\rho \in \SymbRow$ represents a symbolic row. These rows are represented in our programs as simple variables $x$. Like tables,
-rows are also records with the following fields \todo{Well formedness constraint?}
+A symbolic table row $\rho_t \in \SymbRow \subseteq \Var$ represents a symbolic row for table
+$t$. These rows are represented in our programs as simple variables. Like
+tables, rows are also ``records'' with the following fields \todo{well formedness constraint?}
 \begin{enumerate}[align=left]
-  \item[($\rho.\matches$)] A list variable of length representing the possible match keys.
-  \item[($\rho.\action$)] A variable representing the action choice
-  \item[($\rho.\data$)] A list variable representing the possible action data
+  \item[($\rho_t.\matches$)] A list variable of length representing the possible match keys.
+  \item[($\rho_t.\action$)] A variable representing the action choice
+  \item[($\rho_t.\data$)] A list variable representing the possible action data
+  \item[($\rho_t.\id$)] A single variable representing the row's index.
 \end{enumerate}
@@
 
-\paragraph{Values}
-Values $v \in \BVExpr$ are bitvector literals, written $[n]_w$, where $n$
-is a natural value and $w$ is the width. We say that $[n]_w = [m]_l$ is
-undefined if $w \neq l$, and otherwise is true iff $n \equiv m \mod 2^w$.
+For a set of tables $T \subseteq \Table$, we absue set-builder notation to write the shorthand
+$\{\rho_t.*| t \in T \}$ to indicate the set $\{\rho_t.f | t \in T, f \in
+\{\action, \id, \matches[0], \ldots, \matches[n_t], \data[0], ldots, \data[m_t]
+\}\}$ where $n_t = |t.\keys|$ and $m_t = \max\{|\vec x| \mid\lambda \vec x. c
+\in t.\actions\}$
+
+\paragraph{Ghost Variables}
+A ghost varaible $\gamma_t \in \Ghost \subseteq \Var$ represents some ghost
+state for table $t$. These ghost variables are represented as simple variables
+in our program. They are ``records'' with the following fields
+\begin{enumerate}[align=left]
+\item[($\gamma_t.\miss$)] Equal to $[1]_1$ if the table packet misses in the table, and $[0]_1$ otherwise.
+\item[$(\gamma_t.\keys)$] A list variable that captures the values of the keys at table-application time.
+\end{enumerate}
+
+For a set of tables $T \subseteq \Table$, we absue set-builder notation to write
+the shorthand $\{\gamma_t.*| t \in T \}$ to indicate the set $\{\gamma_t.f | t \in
+T, f \in \{\miss, \keys[0], \ldots, \keys[n_t]\}\}$ where $n_t = |t.\keys|$.
 
@@
 \paragraph{Commands}
-Data plane programs $c \in \mathsf{Cmd}$ can be assignments ($h.f := e$), havoc statements ($\havoc x$),
-assumptions ($\assume b$), assertions ($\assert b$), table application
-($t.\apply()$), sequential composition ($c_1;c_2$ for two commands $c_1$ and $c_2$), and
-nondeterministic choice ($c_1 \choice c_2$ for two commands $c_1$ and $c_2$).
+Data plane programs $c \in \mathsf{Cmd}$ can be assignments ($h.f := e$), havoc
+statements ($\havoc x$), assumptions ($\assume b$), assertions ($\assert b$),
+table application ($t.\apply()$), sequential composition ($c_1;c_2$ for two
+commands $c_1$ and $c_2$), and nondeterministic choice ($c_1 \choice c_2$ for
+two commands $c_1$ and $c_2$). Note that table application, havoc, and
+nondeterministic choice are is prohibited in action bodies.
+
+\begin{definition}
+  \label{def:well-formed-command}
+   A command $c \in \Cmd$ is \emph{well-formed} if tables only occur once in the
+   program and each has a unique $t.\id$.
+\end{definition}
 
@@
 
-\paragraph{Packets}
-A packet $\pkt$ is a map from header fields to values. Use the syntax $\pkt[x
-  \mapsto v]$ to represent the packet that is equivalent to $\pkt$ on all variables
-fields except $x$, which maps to $v$.
-
 \paragraph{State}
 On-switch state is represented by a mapping from tables identifiers $t.\id$ to
-lists of rows $[r_1,\ldots,r_n] \in \Row(t)^*$.
+lists of rows $[r_1,\ldots,r_n] \in \Row(t)^*$. I.e. the set of states $\State =
+\Value \to \Row^*$
 
 A similar to a symbolic row $\rho$, a concrete row $r$ comprises a list of match
-values $r.\matches$, an action identifier $r.\action$, and a list of action data
-$r.\data$. A concrete row $r$ is in the set $\Row(t)$ if it is well formed for a
-table $t$ iff the following conditions hold:
+values $r.\matches$, an action identifier $r.\action$, a list of action data
+$r.\data$, and an index $r.\id$. A concrete row $r$ is in the set $\Row(t)$ if
+it is well-formed for a table $t$ iff the following conditions hold:
 \begin{enumerate}
@@
 \item $|r.\data| = |\vec x|$ where $\lambda \vec x. c = t.\actions[r.\action]$
-\end{enumerate}\
-
-A state map $\sigma : \Value \to \Row^*$ is well-formed for a command $c$ iff the
-following conditions hold:
+\end{enumerate}
+
+A state map $\sigma : \Value \to \Row^*$ is well-formed for a command $c$ iff
+the following conditions hold:
 \begin{enumerate}[align=left]
-\item[(\textsc{Total})] For every $t \in \tables(c)$ $\pkt \in \pkt$ there exists a row $r \in \sigma(t.\id)$ such that $r.matches = \pkt(t.keys)$.
 \item[(\textsc{WellFormedRows})] For every $t\in \tables(c)$, every row in $\sigma(t.\id)$ is well-formed.
-\item[(\textsc{NoShadow})] For every $t \in \tables(c)$, with $r_1,\ldots,r_n =
-  \sigma(t.\id)$, for every $1 \leq i \leq n$, there does not exist $1 \leq j <
-  i$ such that $r_i.\matches = r_j.\matches$.
+\item[(\textsc{SelfRef})] For every $t \in tables(c)$, such that $[r_1,\ldots,
+  r_n] = \sigma(t.\id)$, $r_i.\id = i$ for all $1 \leq i \leq n$.
 \end{enumerate}
 
-Importantly, we assume that all state is \emph{total}, that is, no packet locally misses
-in the table. This is formally defined in Definition~\ref{def:total-state}.
-
-%% \begin{definition}[Total State]
-%%   \label{def:total-state}
-%%   A state $\sigma$ is total if for every $t \in \dom(\sigma)$, the following
-%%   formula is valid:
-%%   \[\bigvee_{r \in \sigma(t.\id)} t.\keys = r.\matches \]
-%% \end{definition}
-
-%% Henceforth we assume that all states are total.
-
-%% \todo[inline]{Can be Removed(?)}
+Because of the well-formedness assumption about commands, overload $\sigma(t)
+\triangleq \sigma(t.\id)$.
 
@@
 
-\begin{definition}[Variables]
+\begin{definition}[Free Variables]
   The free variables of a command $c$, written $\fvs(c) \subseteq \Var$, is the
@@
 
-\begin{lemma}[Subset Implies Projected Subset]
+\begin{lemma}[Subset Packet Projection]
+  \label{lem:subset-pkt-proj}
+ Let $\pkt_1,\pkt_2 \in \Pkt$ and $U \subseteq V \subseteq \Var$,
+  If $\pkt_1 =_V \pkt_2$, then $\pkt_1 =_U \pkt_2$.
+\end{lemma}
+\begin{proof}
+  Let $\pkt_1$,$\pkt_2$, $U$, $V$ be given as above. Assume $\pkt_1 =_V \pkt_2$.
+  Let $x \in U$, show that $\pkt_1(x) = \pkt_2(x)$. Since $U \subseteq V$, $x
+  \in V$, so by assumption $\pkt_1(x) = \pkt_2(x)$.
+\end{proof}
+
+\begin{lemma}[Subset Projection Subset]
+  \label{lem:subset-proj-subset}
+  For $P,Q \subseteq \Pkt$ and $U \subseteq V \subseteq \Var$,
+  If $P \subseteq_V Q$, then $P \subseteq_U Q$.
+\end{lemma}
+
+\begin{proof}
+  Let $P$,$Q$,$U$,$V$ be given as above. Assume $P \subseteq_V Q$. Show $P
+  \subseteq_U Q$. Let $\pkt \in P$. Since $P \subseteq_V Q$, we know that there
+  is some $\pkt' \in Q$ such that $\pkt =_V \pkt'$. By Lemma~\ref{lem:subset-pkt-proj},
+  $\pkt =_U \pkt'$. The result follows.
+\end{proof}
+
+\begin{corollary}[Subset Implies Projected Subset]
   \label{lem:subset-proj}
@@
   If $P \subseteq Q$ then $P \subseteq_V Q$.
-\end{lemma}
-
-\begin{proof}
-Let $P,Q \subseteq \Pkt$ and $V \subseteq \Var$ be given. Assume $P \subseteq Q$,
-to show that $P \subseteq_V Q$.
-
-Let $\pkt \in P$. Since $P \subseteq Q$, $\pkt \in Q$. Notice that $\pkt =_V
-\pkt$, so $P \subseteq_V Q$.
-\end{proof}
-
-\begin{lemma}[Equality Implies Projected Equality]
+\end{corollary}
+
+\begin{lemma}[Subset Projection Equality]
+  \label{lem:subset-proj-equal}
+  For $P,Q \subseteq \Pkt$ and $U \subseteq V \subseteq \Var$,
+  If $P =_V Q$, then $P =_U Q$.
+\end{lemma}
+
+\begin{proof}
+  Let $P$,$Q$,$U$,$V$ be given as above. Assume $P =_V Q$. Show $P =_U Q$. Let
+  $\pkt \in P$. Since $P =_V Q$, we know that $P \subseteq_V Q$ and $Q
+  \subseteq_V P$. By Lemma~\ref{lem:subset-proj-subset},$P \subseteq_U Q$ and $Q
+  \subseteq_U P$. Conclude $P =_U Q$.
+\end{proof}
+
+\begin{corollary}[Equality Implies Projected Equality]
   \label{lem:equal-proj}
-  For $P,Q \subseteq \Pkt$, and $ V \subseteq \Var$,
-  If $P = Q$ then $P =_V Q$.
-\end{lemma}
-
-\begin{proof}
-  By Lemma~\ref{lem:subset-proj}
+  For $P,Q \subseteq \Pkt$, and $ V \subseteq \Var$, If $P = Q$ then $P =_V Q$.
+\end{corollary}
+
+\begin{lemma}[Union Packet Projection]
+  \label{lem:union-pkt-proj}
+  For $\pkt_1,\pkt_2 \in \Pkt$, and $U, V \subseteq \Pkt$, if $\pkt_1 =_U
+  \pkt_2$ and $\pkt_1 =_V \pkt_2$, then $\pkt_1 =_{U \cup V} \pkt_2$.
+\end{lemma}
+
+\begin{proof}
+  Let $\pkt_1$, $\pkt_2$, $U$, $V$ be given as above. Assume $\pkt_1 =_U \pkt_2$
+  and $\pkt_1 =_V \pkt_2$. Let $x \in U \cup V$. There are two cases. If $x \in
+  U$, the assumption $\pkt_1 =_U \pkt_2 $ gives $\pkt_1(x) = \pkt_2(x)$.
+  Conversely if $x \in V$, the assumption $\pkt_1 =_V \pkt_2$ gives $\pkt_1(x) =
+  \pkt_2(x)$. Conclude $\pkt_1 =_{U \cup V} \pkt_2$.
 \end{proof}
@@
 \end{lemma}
-
 \begin{proof}
@@
 
+\begin{lemma}[Input-Congruence]
+  \label{lem:input-congruence}
+  For $c,c',c_0 \in \Cmd$,
+  if for all $\sigma \in \State$ and $\pkt \in \Pkt$, $\denote{c}^\sigma \pkt = \denote{c'}^\sigma \pkt$,
+  then for all $\sigma \in \State$ and $\pkt \in \Pkt$, $\denote{c_0;c}^\sigma \pkt = \denote{c_0;c'}^\sigma\pkt$.
+\end{lemma}
+\begin{proof}
+  Let $c$, $c'$, $c_0$, $\sigma$ and $\pkt$ be given as above.
+  Let $\pkt_0 \in \denote{c_0}^\sigma \pkt$. By assumption $\denote{c}^\sigma \pkt_0 = \denote{c'}^\sigma \pkt_0$.
+  Conclude ${\denote{c}^\sigma}^\dagger\left(\denote{c_0}^\sigma\pkt\right) = {\denote{c'}^\sigma}^\dagger\left(\denote{c_0}^\sigma\pkt\right)$. The result follows
+
+\end{proof}
+
+\begin{lemma}[Congruence]
+  \label{lem:congruence}
+  For $c,c',c_a,c_b \in \Cmd$,
+  if for all $\sigma \in \State$ and $\pkt \in \Pkt$, $\denote{c}^\sigma \pkt = \denote{c'}^\sigma \pkt$,
+  then for all $\sigma \in \State$ and $\pkt \in \Pkt$, $\denote{c_a;c;c_b}^\sigma \pkt = \denote{c_a;c';c_b}^\sigma\pkt$.
+\end{lemma}
+
+\begin{proof}
+  Let $\sigma \in \State$ and $\pkt \in \Pkt$ be arbitrary. Compute:
+  \[\begin{array}{ll}
+  \denote{c_a;c;c_b}^\sigma \pkt \\
+  = {\denote{c_b}^\sigma}^\dagger\left({\denote{c}^\sigma}^\dagger\left(\denote{c_a}^\sigma \pkt\right)\right)\\
+  = {\denote{c_b}^\sigma}^\dagger\left({\denote{c'}^\sigma}^\dagger\left(\denote{c_a}^\sigma \pkt\right)\right)
+  & \text{By Lemma~\ref{lem:input-congruence}}\\
+  =\denote{c_a;c;c_b}^\sigma \pkt \\
+  \end{array}\]
+\end{proof}
+
+\begin{remark}
+  The proofs that follow rely so heavily on Lemma~\ref{lem:congruence}, that we omit recording it's use for simplicity.
+\end{remark}
+
 \begin{lemma}[Assume-Currying]
@@
   \label{lem:parametricity}
-  For $c \in \Cmd$, $\sigma : \State$, $y \in \Var$, $\pkt \in \Pkt$, if $y
+  For $c \in \Cmd$, $\sigma \in \State$, $y \in \Var$, $\pkt \in \Pkt$, if $y
   \not\in \fvs(c)$, then $\forall \pkt' \in \denote{c}^\sigma \pkt$, $\pkt'(y) =
@@
 \begin{proof}
-  \todo[inline]{todo}
+  Let $c_1, c_2$, $\sigma$, and $\pkt$ be given as above. Let $V = \Var
+  \setminus (\fvs(c_1)\cup \fvs(c_2))$Let $P_i = \denote{c_i}^\sigma \pkt$ for
+  $i = 1,2$, and $P_{ij} = { \denote{c_j}^\sigma}^\dagger P_i$ for $\{i,j\} =
+  \{1,2\}$. Note that $P_{ij} = \denote{c_i;c_j}^\sigma \pkt$.
+
+  It suffices to show, without loss of generality, for $\pkt' \in P_{12}$,
+  there exists $\pkt_1 \in P_1$ and $\pkt_2 \in P_2$ such that
+  \[\pkt'(x) = \begin{cases}
+    \pkt_1(x),& \text{if } x \in \fvs(c_1)\\
+    \pkt_2(x),& \text{if } x \in \fvs(c_2)\\
+    \pkt(x), & \text{if } x \in V
+  \end{cases} \]
+
+  Let $\pkt' \in P_{12}$ be arbitrary. By definition, there exists $\pkt_1 \in
+  P_1$ such that $\pkt' \in \denote{c_2}^\sigma \pkt_1$. By
+  Lemma~\ref{lem:parametricity}, $\pkt'=_{\fvs(c_1)} \pkt_1$, so our first goal
+  is done. Further, by Lemma~\ref{lem:parametricity}, $\pkt_1 =_{\fvs(c_2)}
+  \pkt(x)$. So by Lemma~\ref{lem:param-noninter}, there is $\pkt_2 \in P_2$
+  such that $\pkt'(x) =_{\fvs(c_2)} \pkt_2(x)$. Then by
+  Lemma~\ref{lem:parametricity}, we know that $\pkt'(x) =_V \pkt(x)$. The result
+  follows.
 \end{proof}
@@
   \]
-  \caption{Symbolic Table instrumentation. For simplicity, we
+  \caption{Symbolic Table Instrumentation. For simplicity, we
     assume each action in a table has the same number (and types) of parameters}
@@
 
-  \multicolumn{3}{l}{\state\sigma_c \in \BExpr} \\
-  \state\sigma_c &\triangleq
-  & \displaystyle\bigwedge_{ t \in \tables(c) } \state\sigma^t
-  \end{array}
-  \]
-  \caption{Symbolic State Encoding}
+  \multicolumn{3}{l}{\state\sigma_T \in \BExpr} \\
+  \state\sigma_T &\triangleq
+  & \displaystyle\bigwedge_{ t \in T } \state\sigma^t
+  \end{array}
+  \]
+  \caption{Symbolic State Encoding. Where $T \subseteq \Table$. We write
+    $\state\sigma_c$ where $c \in \Cmd$ as shorthand for $\state\sigma_{\tables(c)}$.}
   \label{fig:table-instrument}
 \end{figure}
+
+\begin{lemma}[Symbolic State Distributes]
+  \label{lem:sigma-distrib}
+  For $\sigma : \Value \to \Row^*$, $T,T_1,T_2 \subseteq \Table$ such that $T = T_1 \subseteq T_2$,
+  $\state\sigma_{T} \Leftrightarrow \state{\sigma}_{T_1}\wedge\state{\sigma}_{T_2}$
+\end{lemma}
+
+\begin{proof}
+  Let $\sigma$, $T$, $T_1$, $T_2$ be given as above. Assume $T = T_1 \subseteq T_2$.
+  Compute:
+  \[\state\sigma_T\Leftrightarrow
+  \left(\bigwedge_{t \in T} \state\sigma^t\right) \Leftrightarrow
+  \left(\bigwedge_{t \in T_1 \cup T_2} \state\sigma^t\right) \Leftrightarrow
+  \left(\bigwedge_{t \in T_1} \state\sigma^t \wedge \bigwedge_{t \in T_2}\state\sigma^t \right) \Leftrightarrow
+  \state\sigma_{T_1} \wedge \state\sigma_{T_2}
+  \]
+
+\end{proof}
 
@@
 \begin{proof}
-  Let $\sigma : \Value \to \Row^*$ and $c_1,c_2\in \Cmd$ be given. Notice that
-  $\tables(c_1;c_2) = \tables(c_1) \cup \tables(c_2)$. Compute:
-  \[
-  \begin{array}{l >{\displaystyle}ll}
-    \state\sigma_{c_1;c_2}
-    & = \bigwedge_{ t \in \tables(c_1;c_2) } \state\sigma^t \\
-    & = \bigwedge_{ t \in \tables(c_1)\cup \tables(c_2) } \state\sigma^t \\
-    & = \bigwedge_{t \in \tables(c_1)} \state\sigma^t \wedge \bigwedge_{t \in \tables(c_2)} \state\sigma^t \\
-    & = \state\sigma_{c_1} \wedge \state\sigma_{c_2}
-  \end{array}
-  \]
-\end{proof}
-
-\begin{lemma}[Symbolic State distributes over Choice]
+  By Lemma~\ref{lem:sigma-distrib} since $\tables(c_1;c_2) = \tables(c_1) \cup \tables(c_2)$.
+\end{proof}
+
+\begin{lemma}[Symbolic State Distributes Over Choice]
   \label{lem:sigma-distrib-choice}
@@
 \begin{proof}
-  Let $\sigma : \Value \to \Row^*$ and $c_1,c_2\in \Cmd$ be given. Notice that
-  $\tables(c_1 \choice c_2) = \tables(c_1) \cup \tables(c_2)$. Compute:
-  \[
-  \begin{array}{l >{\displaystyle}ll}
-    \state\sigma_{c_1 \choice c_2}
-    & = \bigwedge_{ t \in \tables(c_1 \choice c_2) } \state\sigma^t \\
-    & = \bigwedge_{ t \in \tables(c_1)\cup \tables(c_2) } \state\sigma^t \\
-    & = \bigwedge_{t \in \tables(c_1)} \state\sigma^t \wedge \bigwedge_{t \in \tables(c_2)} \state\sigma^t \\
-    & = \state\sigma_{c_1} \wedge \state\sigma_{c_2}
-  \end{array}
-  \]
+  By Lemma~\ref{lem:sigma-distrib} since $\tables(c_1;c_2) = \tables(c_1) \cup \tables(c_2)$.
 \end{proof}
@@
   Given a set of variables $V \subseteq \Var$ that contains no ghost variables
-  or symbolic variables. For every $c \in \Cmd$ such that $\fvs(c) \subseteq V$,
+  or symbolic variables. For every well-formed $c \in \Cmd$ such that $\fvs(c) \subseteq V$,
   well-formed state $\sigma : \Value \to \Row^*$, and packet $\pkt \in \Pkt$,
-  where $R = \{\rho_t \mid t \in \tables(c)\}$, and $G = \{\gamma_t \mid t \in
-  \tables(c)\}$,
-
-  \[\denote{c}^\sigma\,\pkt =_V \denote{\havoc {R \cup G}; \assume{\state\sigma}_t; c\instr}\,\pkt  \]
+  where $T \supseteq \tables(c)$, $R \supseteq \{\rho_t \mid t \in T\}$, and $G \supseteq \{\gamma_t \mid t \in
+  T\}$,
+
+  \[\denote{c}^\sigma\,\pkt =_V \denote{\havoc {R \cup G}; \assume{\state\sigma}_T; c\instr}\,\pkt  \]
 \end{theorem}
@@
   \item[$(c = \assert b)$] \textit{sim}.
-  \item[$(c = c_1;c_2)$] Let $R_i = \{\rho_t \mid t \in \tables(c_i)\}$ for $i =
-    1,2$. Notice that $R = R_1 \cup R_2$. Similarly, let $G_i = \{\gamma_t \mid
-    t \in \tables(c_i)\}$ for $i = 1,2$. Notice that $G = G_1 \cup G_2$. Compute
+  \item[$(c = c_1;c_2)$] Note that well-formedness\todo{define well formedness as single-use tables} implies $\tables(c_1) \cap
+    \tables(c_2) = \emptyset$. Let $T_2 = \tables(c_2)$ and $T_1 = T \setminus
+    T_2$. Let $R_2 = \{\rho_t.* \mid t \in T_2\}$ and $R_1 = R \setminus R_2$.
+    Let $G_2 = \{\rho_t.* \mid t \in T_2\}$ and $G_1 = G \setminus G_2$. Compute
     as follows:
     \[\begin{array}{ll}
-    \denote{\havoc {R \cup G}; \assume{\state\sigma_{c_1;c_2}};(c_1;c_2)\instr}\,\pkt
+    \denote{\havoc {R \cup G}; \assume{\state\sigma_{T}};(c_1;c_2)\instr}\,\pkt
     & \\
-    = \denote{\havoc{R\cup G}; \assume{\state\sigma_{c_1;c_2}}; c_1\instr;c_2\instr}\,\pkt
+    = \denote{\havoc{R\cup G}; \assume{\state\sigma_{T}}; c_1\instr;c_2\instr}\,\pkt
     & \text{definition of }(\cdot)\instr \\
-    = \denote{\havoc{R\cup G}; \assume{\state\sigma_{c_1}};\assume{\state\sigma_{c_2}};c_1\instr;c_2\instr}\,\pkt
+    = \denote{\havoc{R\cup G}; \assume{\state\sigma_{T_1}};\assume{\state\sigma_{T_2}};c_1\instr;c_2\instr}\,\pkt
     & \text{By Lemmas~\ref{lem:sigma-distrib-seq} \&~\ref{lem:assume-currying}} \\
-    = \denote{\havoc{R_1};\havoc{G_1}; \havoc{R_2};\havoc{G_2}; \assume{\state\sigma_{c_1}};\assume{\state\sigma_{c_2}};c_1\instr;c_2\instr}\,\pkt
+    = \denote{\havoc{R_1 \cup G_1}; \havoc{R_2 \cup G_2}; \assume{\state\sigma_{T_1}};\assume{\state\sigma_{T_2}};c_1\instr;c_2\instr}\,\pkt
+    & \text{By definition } \\
+    = \denote{\havoc{R_1\cup G_1}; \havoc{R_2 \cup G_2}; \assume{\state\sigma_{T_1}};c_1\instr;\assume{\state\sigma_{T_2}};c_2\instr}\,\pkt
+    & \text{By Lemma~\ref{lem:disj-comm} and } \tables(c_1) \cap \tables(c_2) = \emptyset \\
+    = \denote{(\havoc{R_1 \cup G_1}; \assume{\state\sigma_{T_1}};c_1\instr);
+      (\havoc{R_2 \cup G_2};\assume{\state\sigma_{T_2}};c_2\instr)}\,\pkt
+    & \text{By Lemma~\ref{lem:disj-comm} and } \tables(c_1) \cap \tables(c_2) = \emptyset\\
+    = \denote{\havoc{R_2 \cup G_2};\assume{\state\sigma_{T_2}};c_2\instr}^\dagger
+    \left(\denote{\havoc{R_1 \cup G_1}; \assume{\state\sigma_{T_1}};c_1\instr}\,\pkt\right)
     & \text{by definition} \\
-    = \denote{\havoc{R_1};\havoc{G_1}; \havoc{R_2}; \havoc{G_2}; \assume{\state\sigma_{c_1}};c_1\instr;\assume{\state\sigma_{c_2}};c_2\instr}\,\pkt
-    & \text{By Lemma~\ref{lem:disj-comm}} \\
-    = \denote{(\havoc{R_1}; \havoc{G_1}; \assume{\state\sigma_{c_1}};c_1\instr);
-      (\havoc{R_2}; \havoc{G_2};\assume{\state\sigma_{c_2}};c_2\instr)}\,\pkt
-    & \text{By Lemma~\ref{lem:disj-comm}}\\
-    = \denote{\havoc{R_2}; \havoc{G_2};\assume{\state\sigma_{c_2}};c_2\instr}^\dagger
-    \left(\denote{\havoc{R_1}; \havoc{G_1}; \assume{\state\sigma_{c_1}};c_1\instr}\,\pkt\right)
-    & \text{by definition} \\
-    =_V {\denote{c_2}^\sigma}^\dagger \left(\denote{\havoc{\vec{\rho_{t_1}}}; \havoc{\vec\gamma_{t_1}}; \assume{\state\sigma_{c_1}};c_1\instr}\,\pkt\right)
+    =_V {\denote{c_2}^\sigma}^\dagger \left(\denote{\havoc{R_1 \cup G_1}; \assume{\state\sigma_{T_1}};c_1\instr}\,\pkt\right)
     & \text{by IH}(c_2) \text{ since } \fvs(c_2) \subseteq \fvs(c_1;c_2) \subseteq V \\
@@
 
-  \item[$(c = c_1\choice c_2)$] Let $R_i = \{\rho_t \mid t \in \tables(c_i)\}$ for $i =
-    1,2$. Notice that $R = R_1 \cup R_2$. Similarly, let $G_i = \{\gamma_t \mid
-    t \in \tables(c_i)\}$ for $i = 1,2$. Notice that $G = G_1 \cup G_2$. Compute
-    as follows:
-    \[\begin{array}{ll}
-    \denote{\havoc{R \cup G};\assume{\state\sigma_{c_1\choice c_2}}; (c_1 \choice c_2)\instr}\,\pkt\\
+  \item[$(c = c_1\choice c_2)$]
+    \[\begin{array}{ll}
+    \denote{\havoc{R \cup G};\assume{\state\sigma_{T}}; (c_1 \choice c_2)\instr}\,\pkt
     & \\
-    = \denote{\havoc{R \cup G};\assume{\state\sigma_{c_1\choice c_2}}; (c_1\instr \choice c_2\instr)}\,\pkt\\
+    = \denote{\havoc{R \cup G};\assume{\state\sigma_{T}}; (c_1\instr \choice c_2\instr)}\,\pkt
     & \text{by definition of }(\cdot)\instr\\
-    = \denote{\havoc{R \cup G};\assume{\state\sigma_{c_1} \wedge \state\sigma_{c_2}}; (c_1\instr \choice c_2\instr)}\,\pkt
-    & \text{By Lemma~\ref{lem:sigma-distrib-choice}}\\
-    = \denote{\havoc{R \cup G};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; (c_1\instr  \choice   c_2\instr)}\,\pkt
-    & \text{By Lemma~\ref{lem:assume-currying}}\\
-    = \denote{\havoc{R \cup G};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; c_1\instr
-      \choice \havoc{R \cup G};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
+    %% = \denote{\havoc{R \cup G};\assume{\state\sigma_{T_1} \wedge \state\sigma_{T_2}}; (c_1\instr \choice c_2\instr)}\,\pkt
+    %% & \text{By Lemma~\ref{lem:sigma-distrib-choice}}\\
+    %% = \denote{\havoc{R \cup G};\assume{\state\sigma_{T_1}}; \assume{\state\sigma_{T_2}}; (c_1\instr  \choice   c_2\instr)}\,\pkt
+    %% & \text{By Lemma~\ref{lem:assume-currying}}\\
+    = \denote{\havoc{R \cup G};\assume{\state\sigma_{T}}; c_1\instr
+      \choice \havoc{R \cup G};\assume{\state\sigma_{T}}; c_2\instr}\,\pkt
     & \text{By Lemma~\ref{lem:prod-sum}} \\
-    = \denote{\havoc{R \cup G}; \assume{\state\sigma_{c_1}}; c_1\instr
-      \choice \havoc{R \cup G}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
-    & \text{BY UNUSED-VARIABLES -> TRIVIAL ASSUMPTIONS LEMMA} \\
-    =_V \denote{\havoc{R_1 \cup G_1}; \assume{\state\sigma_{c_1}}; c_1\instr
-      \choice \havoc{R_2 \cup G_2}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
-    & \text{BY UNUSEDVARIABLE HAVOC ELIMINATION LEMMA } \\
-    = \denote{\havoc{R_1 \cup G_1}; \assume{\state\sigma_{c_1}}; c_1\instr}\,\pkt
-    \cup \denote{\havoc{R_2 \cup G_2}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
+    %% = \denote{\havoc{R \cup G}; \assume{\state\sigma_{T_1}}; c_1\instr
+    %%   \choice \havoc{R \cup G}; \assume{\state\sigma_{T_2}}; c_2\instr}\,\pkt
+    %% & \text{BY UNUSED-VARIABLES -> TRIVIAL ASSUMPTIONS LEMMA} \\
+    %% =_V \denote{\havoc{R_1 \cup G_1}; \assume{\state\sigma_{c_1}}; c_1\instr
+    %%   \choice \havoc{R_2 \cup G_2}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
+    %% & \text{BY UNUSEDVARIABLE HAVOC ELIMINATION LEMMA } \\
+    = \denote{\havoc{R \cup G}; \assume{\state\sigma_{T}}; c_1\instr}\,\pkt
+    \cup \denote{\havoc{R \cup G}; \assume{\state\sigma_{T}}; c_2\instr}\,\pkt
     & \text{by definition of }\denote{\cdot} \\
-    =_V \denote{c_1}^\sigma\pkt \cup \denote{\havoc{R_2 \cup G_2}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
-    & \text{By IH}(c_1) \text{ since } \fvs(c_1) \subseteq \fvs(c_1 \choice c_2) \subseteq V \\
+    =_V \denote{c_1}^\sigma\pkt \cup \denote{\havoc{R \cup G}; \assume{\state\sigma_{T}}; c_2\instr}\,\pkt
+    & \text{By IH}(c_1) \text{ since } \fvs(c_1) \subseteq \fvs(c_1 \choice c_2) \subseteq V, \tables(c_1) \subseteq T \\
     =_V \denote{c_1}^\sigma\pkt \cup \denote{c_2}^\sigma\pkt
-    & \text{By IH}(c_2) \text{ since } \fvs(c_2) \subseteq \fvs(c_1 \choice c_2) \subseteq V \\
+    & \text{By IH}(c_2) \text{ since } \fvs(c_2) \subseteq \fvs(c_1 \choice c_2) \subseteq V, \tables(c_2) \subseteq T \\
     = \denote{c_1 \choice c_2}^\sigma\pkt
@@
     Let $P = \{\pkt\{\rho_t.* \mapsto
-      v\}\{\gamma_t.* \mapsto v' \mid v,v' \in \Value\} \}$ \todo{clean this up}.
+      v\}\{\gamma_t.* \mapsto v' \mid v,v' \in \Value\} \}$.
       Compute as follows:
@@
   \label{thm:vc-correct}
-  Given a set of variables $V$, for every $c \in \Cmd$ such that $\fvs(c) = V$,
+  Given a set of variables $V$, for every well-formed $c \in \Cmd$ such that $\fvs(c) = V$,
   and $\sigma \in \State$, the following formula is valid:
-  \[ \forall V. \WP(c[\sigma], \TRUE) \iff \forall V. (\forall \vec{\rho_t}. \forall \vec{\gamma_t}. \state\sigma_c \Rightarrow  \WP(c\instr,\TRUE)\]
+  \[ \forall V. \WP(c[\sigma], \TRUE) \iff \forall V. (\forall R. \forall G. \state\sigma_c \Rightarrow  \WP(c\instr,\TRUE)\]
+  where $R = \{\rho_t.* \mid t \in \tables(c)\}$ and $G = \{\gamma_t.*\mid t \in \tables (c)\}$.
 \end{theorem}
@@
 \begin{proof}
-  Let $c$ and $\sigma$ be arbitary.
+  Let $V$, $c$, and $\sigma$ be given as above.
   Theorem~\ref{thm:instr-equiv} gives the equivalence
-  \[\forall \pkt. \denote{c}^\sigma\,\pkt =_{\fvs(c)} \denote{\havoc \vec{\rho_t}; \havoc{\vec{\gamma_t}}; \assume{\state\sigma_c}; c\instr}\,\pkt\]
+  \[\forall \pkt. \denote{c}^\sigma\,\pkt =_{\fvs(c)} \denote{\havoc{R \cup G}; \assume{\state\sigma_c}; c\instr}\,\pkt\]
 
@@
 
-  \[\forall \pkt.\denote{c[\sigma]}\,\pkt =_{\fvs(c)}  \denote{\havoc \vec{\rho_t}; \havoc{\vec{\gamma_t}}; \assume{\state\sigma_c}; c\instr}\,\pkt\]
+  \[\forall \pkt.\denote{c[\sigma]}\,\pkt =_{\fvs(c)}  \denote{\havoc{R \cup G}; \assume{\state\sigma_c}; c\instr}\,\pkt\]
 
@@
 
-  \[\forall V. \WP(c[\sigma], \TRUE) \iff \forall V. \WP(\havoc \vec{\rho_t}; \havoc{\vec{\gamma_t}}; \assume{\state\sigma_c}; c\instr, \TRUE))\]
+  \[\forall V. \WP(c[\sigma], \TRUE) \iff \forall V. \WP(\havoc{R \cup G}; \assume{\state\sigma_c}; c\instr, \TRUE))\]
 
@@
 
-  \[ \forall V. \WP(c[\sigma],\TRUE) \iff \forall V. \left(\forall \vec{\rho_t}.\forall \vec{\gamma_t}.\state\sigma_c \Rightarrow \WP(c\instr, \TRUE)\right)\]
+  \[ \forall V. \WP(c[\sigma],\TRUE) \iff \forall V. \left(\forall R.\forall G.\state\sigma_c \Rightarrow \WP(c\instr, \TRUE)\right)\]
 
@@
 
-%% \begin{lemma}[Universal Quantifier Weakness]
-%%   For $b \in \BExpr$ and $V \subseteq \bfvs(b)$, $\forall V. b$ is the weakest
-%%   formula on variables $\bfvs(b) \setminus V$ that implies $b$.
-%% \end{lemma}
-%% \begin{proof}
-%%   Elementary.
-%% \end{proof}
+\begin{lemma}[Universal Quantifier Weakness]
+  \label{lem:quantified-weakest}
+  For $b \in \BExpr$ and $V \subseteq \Var$, $\forall V. b$ is the weakest
+  formula on variables $\Var \setminus V$ that implies $b$.
+\end{lemma}
+\begin{proof}
+  Let $b$ and $V$ be given. Assume there is some $b' \Rightarrow b$ such that
+  $\bfvs(b') \subseteq \Var \setminus V$. Show that $b' \Rightarrow \forall
+  V.b$.
+
+  \[\begin{array}{rll}
+   1. & b' \Rightarrow b & \textit{assumption} \\
+   2. & b' & \textit{Assumption} \\
+   3. & V & \textit{arbitrary} \\
+   4. & b & \textit{modus ponens (1,2)} \\
+   5. & \forall V.b & \forall\textit{-intro (3-4)}\\
+   6. & b' \Rightarrow \forall V. b & \Rightarrow\textit{-intro (2-5)}
+  \end{array}\]
+
+\end{proof}
 
 \begin{theorem}[Quantifer Elimination]
-  For every instrumentation-free command $c \in \Cmd$, the formula $\qe(\fvs(c), \WP(c\instr, \TRUE))$ is a solution to the \textbf{Inference Problem}.
+  For $D \subseteq \Var$ a set of dataplane variables and $C \subseteq \Var$ a
+  set of control plane variables such that $D \cap C = \emptyset$, for every well-formed
+  command $c \in \Cmd$ such that $\fvs(c) \subseteq D$, the formula $\qe(D,
+  \WP(c\instr, \TRUE))$ is a solution to the \textbf{Inference Problem}.
 \end{theorem}
@@
 \begin{proof}
-  Let $c \in \Cmd$ be instrumentation-free. Let $R$ be the set of symbolic row
-  variables in $c\instr$, and let $G$ be the set of ghost variables in $c\instr$,
-  Let $\varphi = \qe(\fvs(c),\WP(c\instr, \TRUE))$. Since $\bfvs(\varphi) \cap
-  \fvs(c) = \emptyset$, it is trivial to show \todo{do it} that $\varphi$ is a
-  control plane formula.
+  Let $D$, $C$, and $c$ be given as above. Let $R \subseteq C$ be the set of
+  symbolic row variables in $c\instr$, and let $G \subseteq C$ be the set of
+  ghost variables in $c\instr$. Notice that $\fvs(c\instr) = \fvs(c) \cup G \cup
+  R$. Let $\varphi = \qe(D,\WP(c\instr, \TRUE))$. Since
+  $\bfvs(\WP(c\instr,\TRUE)) \subseteq \fvs(c\instr)$, and $\bfvs(\varphi) \cap
+  D = \emptyset$, we know that $\varphi$ is a control plane formula.
 
@@
   assumption about $\qe$, this is equivalent to $\forall R.\forall G.
-  \state\sigma_c \Rightarrow \WP(c\instr, \TRUE)$. By
-  Theorem~\ref{thm:vc-correct}, $\WP(c[\sigma],\TRUE)$ is valid. By Dijkstra,
+  \state\sigma_c \Rightarrow \forall D. \WP(c\instr, \TRUE)$. Since
+  $\bfvs(\state\sigma_c) \cap D = \emptyset$, we can equivalently write $\forall
+  D. \forall R. \forall G. \state\sigma_c \Rightarrow \WP(c\instr,\TRUE)$. By
+  Theorem~\ref{thm:vc-correct}, $\forall D. \WP(c[\sigma],\TRUE)$ is valid. By Dijkstra,
   $\forall \pkt \in \Pkt. \denote{c[\sigma]} \neq \error$. Conclude $\forall
@@
 
-  \todo[inline]{weakest?}
+  Now, by Theorem~\ref{thm:vc-correct} it suffices to show that $\varphi$ is the
+  weakest control plane formula stronger than $\WP(c[?], \TRUE)$. By
+  Lemma~\ref{lem:quantified-weakest}, $\forall D. \WP(c\instr, \TRUE)$ is the
+  weakest condition on variables $C$ that is still stronger than
+  $\WP(c\instr,\TRUE)$. Since $\varphi \Leftrightarrow \forall D. \WP(c\instr,
+  \TRUE)$, $\varphi$ is the weakest condition stronger than $\WP(c\instr,\TRUE)$.
+
 \end{proof}

### 2021-05-28 inference.tex
fix projections

@@
 \paragraph{Expressions}
-The semantics of expressions is shown in Figure~\label{fig:sem-expr}. Note that
+The semantics of expressions is shown in Figure~\ref{fig:sem-expr}. Note that
 here we implicitly use the strict semantics for definedness, that is, operators
@@
 \begin{definition}[Projections]
-  For a set of variables $S \subseteq \Var$, and a packet $\pkt$, write
-  $\pkt|_S$ to be projection of $\pkt$ onto $S$. \todo{I guess this means
-    packets are partial functions?} Formally,
-  \[\pkt|_S \triangleq \{x \mapsto \pkt(x) \mid x \in S\}.\]
-  Two packets $\pkt$, $\pkt'$ are equivalent up to $S \subseteq \Var$, written
-  $\pkt =_S \pkt'$, iff $\pkt|_S = \pkt'|_s$. A set of packets $P$ is a subset
-  of $P'$ up to $S$, written $P \subseteq_S P'$ iff for every packet $\pkt \in
-  P$, there exists $\pkt' \in P'$ such that $\pkt =_S \pkt'$. Two sets of
-  packets $P$ and $P'$ are $S$-equivalent, written $P =_S P'$ iff $P \subseteq_S
-  P'$ and $P \subseteq_S P'$. A set $P \subseteq \Pkt$ is a $V$-equivalence
-  class if $\forall \pkt_1, \pkt_2 \in P$, $\pkt_1 =_V \pkt_2$.
+ Two packets $\pkt$, $\pkt'$ are equivalent up to $V \subseteq \Var$, written
+ $\pkt =_V \pkt'$, iff $\forall x \in V \pkt(x) = \pkt'(x)$. A set of packets
+ $P$ is a subset of $P'$ up to $S$, written $P \subseteq_S P'$ iff for every
+ packet $\pkt \in P$, there exists $\pkt' \in P'$ such that $\pkt =_S \pkt'$.
+ Two sets of packets $P$ and $P'$ are $S$-equivalent, written $P =_S P'$ iff $P
+ \subseteq_S P'$ and $P \subseteq_S P'$. A set $P \subseteq \Pkt$ is a
+ $V$-equivalence class if $\forall \pkt_1, \pkt_2 \in P$, $\pkt_1 =_V \pkt_2$.
 \end{definition}
 
-\begin{lemma}[Projection Union-Distributivity]
-  \label{lem:proj-union-set}
-  For all $P,Q \subseteq \Pkt$, $V \subseteq \Var$,
-  $(P\cup Q)|_V = P|_V \cup Q|_V$
-\end{lemma}
-
-\begin{proof}
-  Let $P,Q \subseteq \Pkt$ and $V \subseteq \Var$ be given. Compute:
-
-  \[\begin{array}{ll}
-  (P\cup Q)|_V\\
-  = \{\pkt|_V \mid \pkt \in P\cup Q\}\\
-  = \{\pkt|_V \mid \pkt \in P\}\cup \{\pkt|_V \mid \pkt \in Q\}\\
-  = P|_V \cup Q|_V
-  \end{array}\]
-\end{proof}
-
-\begin{lemma}[Projection Idempotence]
-  \label{lem:proj-idem}
-  For all $\pkt \in \Pkt$, $V \subseteq \Var$,
-  $(\pkt|_V)|_V = \pkt|_V$
-\end{lemma}
-
-\begin{proof}
-  Let $\pkt \in \Pkt$ and $V \subseteq \Var$ be given.
-  \[\begin{array}{ll}
-    (\pkt|_V)|_V \\
-    = \{x \mapsto (\pkt|_V)(x) \mid x \in V\} & \text{by definition}\\
-    = \{x \mapsto (\{y \mapsto \pkt(y) \mid y \in V\})(x) \mid x \in V\} & \text{by definition}\\
-    = \{x \mapsto \pkt(x) \mid x \in V\} & \beta\text{-reduction}\\
-    = \pkt|_V & \text{by definition}
-  \end{array}\]
-\end{proof}
-
-\begin{lemma}[Projection Idempotence for Sets]
-  \label{lem:proj-idem-set}
-  For all $P \subseteq \Pkt$, $V \subseteq \Var$,
-  $(P|_V)|_V = P|_V$
-\end{lemma}
-
-\begin{proof}
-  Let $P \subseteq \Pkt$ and $V \subseteq \Var$ be given. Compute:
-  \[\begin{array}{ll}
-  (P|_V)|_V \\
-  = \{\pkt|_V \mid \pkt \in P|_V\}
-  & \text{by definition of } (\cdot)|_V : \mathcal{P}(\Pkt) \to \mathcal{P}(\Pkt) \\
-  = \{\pkt|_V \mid \pkt \in \{\pkt'|_V \mid \pkt'\in P\}\}
-  & \text{by definition of } (\cdot)|_V : \mathcal{P}(\Pkt) \to \mathcal{P}(\Pkt) \\
-  = \{(\pkt|_V)|_V \mid \pkt \in P \}
-  & \text{by set theory} \\
-  = \{\pkt|_V \mid \pkt \in P\} & \text{by Lemma~\ref{lem:proj-idem}} \\
-  = P|_V & \text{by definition of } (\cdot)|_V : \mathcal{P}(\Pkt) \to \mathcal{P}(\Pkt)
-  \end{array}\]
-\end{proof}
+\begin{remark}
+  Standard equality and subset at all of these types is equivalent to projected
+  equality/subset on the universal set of variables, i.e. $(=) = (=_\Var)$ and
+  $(\subseteq) = (\subseteq_\Var)$.
+\end{remark}
 
@@
 \begin{corollary}[Subset Implies Projected Subset]
-  \label{lem:subset-proj}
+  \label{cor:subset-proj}
   For $P,Q \subseteq \Pkt$, and $ V \subseteq \Var$,
@@
 
-\begin{lemma}[Subset Projection Equality]
-  \label{lem:subset-proj-equal}
+\begin{lemma}[Subset Projection Equals]
+  \label{lem:subset-proj-equals}
   For $P,Q \subseteq \Pkt$ and $U \subseteq V \subseteq \Var$,
@@
 \begin{proof}
-  Let $P$,$Q$,$U$,$V$ be given as above. Assume $P =_V Q$. Show $P =_U Q$. Let
-  $\pkt \in P$. Since $P =_V Q$, we know that $P \subseteq_V Q$ and $Q
-  \subseteq_V P$. By Lemma~\ref{lem:subset-proj-subset},$P \subseteq_U Q$ and $Q
-  \subseteq_U P$. Conclude $P =_U Q$.
+  Let $P$,$Q$,$U$,$V$ be given as above. Assume $P =_V Q$. By definition we have
+  $P \subseteq_V Q$ and $Q \subseteq_V P$. By
+  Lemma~\ref{lem:subset-proj-subset}, these become $P \subseteq_U Q$ and $Q\subseteq_U P$
+  Conclude $P =_U Q$.
 \end{proof}
@@
 \begin{corollary}[Equality Implies Projected Equality]
-  \label{lem:equal-proj}
+  \label{cor:equal-proj}
   For $P,Q \subseteq \Pkt$, and $ V \subseteq \Var$, If $P = Q$ then $P =_V Q$.
@@
 
-\begin{lemma}[Union Packet Projection]
-  \label{lem:union-pkt-proj}
-  For $\pkt_1,\pkt_2 \in \Pkt$, and $U, V \subseteq \Pkt$, if $\pkt_1 =_U
-  \pkt_2$ and $\pkt_1 =_V \pkt_2$, then $\pkt_1 =_{U \cup V} \pkt_2$.
-\end{lemma}
-
-\begin{proof}
-  Let $\pkt_1$, $\pkt_2$, $U$, $V$ be given as above. Assume $\pkt_1 =_U \pkt_2$
-  and $\pkt_1 =_V \pkt_2$. Let $x \in U \cup V$. There are two cases. If $x \in
-  U$, the assumption $\pkt_1 =_U \pkt_2 $ gives $\pkt_1(x) = \pkt_2(x)$.
-  Conversely if $x \in V$, the assumption $\pkt_1 =_V \pkt_2$ gives $\pkt_1(x) =
-  \pkt_2(x)$. Conclude $\pkt_1 =_{U \cup V} \pkt_2$.
-\end{proof}
-
-\begin{lemma}[Expression Projection]
-  \label{lem:proj-expr}
-  For every set of variables $V \subseteq \Var$, packet $\pkt \in \Pkt$ and
-  expression $e \in \BVExpr$, such that $\efvs(e) \subseteq V$, $\edenote{e}\pkt =
-  \edenote{e}(\pkt|_V)$.
-\end{lemma}
-\begin{proof}
-  Let $V \subseteq \Var$ be given.
-  Proceed by induction on $e$, leaving $\pkt$ arbitrary.
-  \begin{enumerate}[align=left]
-  \item[($e = v$)] Let $\pkt$ be given. $\edenote{v}\pkt = v = \edenote{v}{\pkt|_V}$
-  \item[($e = x$)] Let $\pkt$ be given. Note that $\efvs(x) = \{x\} \subseteq V$. Compute:
-    \[\begin{array}{ll}
-    \edenote{x}(\pkt|_V) \\
-    = (\pkt|_V)(x)
-    & \text{by definition of }\edenote\cdot\\
-    = \{y \mapsto \pkt(y) \mid y \in V\}(x)
-    & \text{by definition of }(\cdot)|_V : \Pkt \to \Pkt\\
-    = \pkt(x) & \text{since } x \in V
-    \end{array}
-    \]
-  \item[($e = e_1 \binop e_2$)] Let $\pkt$ be given. Compute:
-    \[\begin{array}{ll}
-    \edenote{e_1 \binop e_2}\pkt \\
-    = \edenote{e_1}\pkt \binop \edenote{e_2}\pkt & \text{definition of }\edenote\cdot \\
-    = \edenote{e_1}(\pkt|_V) \binop \edenote{e_2}(\pkt|_V) & \text{by IHS since }\efvs(e_i) \subseteq \efvs(e_1 \binop e_2) \subseteq V, i = 1,2 \\
-    = \edenote{e_1 \binop e_2}(\pkt|_V) & \text{definition of }\edenote\cdot
-    \end{array}\]
-  \item[($e = \unop e$)] Let $\pkt$ be given. Compute:
-    \[\begin{array}{ll}
-    \edenote{\unop e}\pkt \\
-    = \unop \edenote{e}\pkt & \text{definition of }\edenote\cdot \\
-    = \unop \edenote{e}(\pkt|_V) & \text{by IH, since } \efvs(e) = \efvs(\unop e) \subseteq V \\
-    = \edenote{\unop e}(\pkt|_V) & \text{definition of }\edenote\cdot
-    \end{array}\]
-  \end{enumerate}
-\end{proof}
-
-\begin{lemma}[Satisfaction Projection]
-  \label{lem:proj-satis}
-  For every set of variables $V \subseteq \Var$, packet $\pkt \in \Pkt$ and
-  boolean expression $b \in \BExpr$, such that $\bfvs(b) \subseteq V$,
-  $\pkt \models b$ iff $\pkt|_V \models b$.
-\end{lemma}
-
-\begin{proof}
-  Let $V \subseteq \Var$ be given.
-  Proceed by induction on $b$, leaving $\pkt$ arbitrary.
-  \begin{enumerate}[align=left]
-  \item[($b = \FALSE$)] Immediate.
-  \item[($b = b_1 \Rightarrow b_2$)]
-    Let $\pkt$ be given. Compute:
-    \[\begin{array}{ll}
-    \pkt \models b_1 \Rightarrow b_2 \\
-    \iff \pkt \not\models b_1 \text{or} \pkt\models b_2 & \text{by definition of } \models   \\
-    \iff \pkt|_V \not\models b_1 \text{or} \pkt|_V \models b_2  & \text{by IHs, since} \bfvs(b_i) \subseteq \bfvs(b_1 \Rightarrow b_2) \subseteq V, i = 1,2  \\
-    \iff \pkt|_V \models b_1 \Rightarrow b_2
-    & \text{by definition of } \models
-    \end{array}\]
-
-  \item[($b = e_1 = e_2$)] Let $\pkt$ be given. Compute:
-    \[\begin{array}{ll}
-    \pkt \models e_1 = e_2 \\
-    \iff \edenote{e_1}\pkt = \edenote{e_2}\pkt
-    & \text{by definition of }\models \\
-    \iff \edenote{e_1}(\pkt|_V) = \edenote{e_2}(\pkt|_V)
-    & \text{by Lemma~\ref{lem:proj-expr}, since } \efvs(e_i) \subseteq \bfvs(e_1 = e_2) \subseteq V, i = 1,2 \\
-    \iff \pkt|_V \models e_1 = e_2
-    & \text{by definition of }\models
-    \end{array}\]
-
-  \end{enumerate}
-
-\end{proof}
-
 \begin{lemma}[Projection-Denotation Commutativity]
   \label{lem:proj-den-comm}
-  For a closed action $c \in \Cmd$, $\fvs(c) \subseteq V \subseteq \Var$ and
-  $\pkt \in \Pkt$, $\denote{c} \pkt =_V \denote{c} \pkt|_V$.
-\end{lemma}
-
-\begin{proof}
-  Proceed by induction on $c$, leaving $\pkt$ general.
-  \begin{enumerate}[align=left]
-  \item[$(c = x:=e)$] Let $\pkt$ be given. Compute:
-    \[\begin{array}{ll}
-    (\denote{x:=e} (\pkt|_V))|_V\\
-    = ((\pkt|_V)\{x:=\edenote{e}(\pkt|_V)\})|_V
-    & \text{definition} \\
-    = ((\pkt|_V)\{x := \edenote{e}\pkt\})|_V
-    & \text{By Lemma~\ref{lem:proj-expr} since } \fvs(e) \subseteq \fvs(x := e) \subseteq V \\
-    = (\pkt\{x := \edenote{e}\pkt\}|_V)|_V
-    & \text{since } x \in \fvs(x:=e) \subseteq V \\
-    = ((\denote{x:=e}\pkt)|_V)|_V
-    & \text{by definition of }\denote{\cdot}\\
-    = (\denote{x:=e}\pkt)|_V
-    & \text{by Lemma~\ref{lem:proj-idem}}
-    \end{array}\]
-    The result follows by definition of $=_V$.
-
-  \item[$(c = \havoc x)$] Contradictory; $\havoc x$ is not an action.
-  \item[$(c = \assume b)$]
-    Let $\pkt$ be given. Compute:
-    \[\begin{array}{ll}
-    (\denote{\assume b} \pkt)|_V \\
-    = (\{\pkt \mid \pkt \models b\})|_V
-    & \text{definition of}\denote\cdot\\
-    = ((\{\pkt \mid \pkt \models b\})|_V)|_V
-    & \text{By Lemma~\ref{lem:proj-idem-set}} \\
-    = (\{\pkt|_V \mid \pkt \models b\})|_V
-    & \text{definition of }(\cdot)|_V : \mathcal P (\Pkt) \to \mathcal P (\Pkt) \\
-    = (\{\pkt|_V \mid \pkt|_V \models b\})|_V
-    & \text{By Lemma~\ref{lem:proj-satis} since } \fvs(b) \subseteq \fvs(\assume b) \subseteq V \\
-    = (\denote{\assume b}(\pkt|_V))|_V
-    & \text{by definition of}\denote\cdot
-    \end{array}
-    \]
-    The result follows by definition of $=_V$.
-
-  \item[$(c = \assert b)$] \textit{sim.}
-  \item[$(c = c_1;c_2)$] Let $\pkt$ be given. Compute:
-    \[\begin{array}{>{\displaystyle}ll}
-    (\denote{c_1;c_2}\pkt)|_V \\
-    = (\denote{c_2}^\dagger(\denote{c_1}\pkt))|_V
-    & \text{definition of }\denote\cdot \\
-    = \left(\bigcup_{\pkt' \in \denote{c_1} \pkt} \denote{c_2}\pkt'\right)|_V
-    & \text{definition of }(\cdot)^\dagger \\
-    = \bigcup_{\pkt' \in \denote{c_1} \pkt} (\denote{c_2}\pkt')|_V
-    & \text{by Lemma~\ref{lem:proj-union-set}} \\
-    = \bigcup_{\pkt' \in \denote{c_1}\pkt} (\denote{c_2}(\pkt'|_V))|_V
-    & \text{IH}(c_2)  \\
-    =\bigcup_{\pkt' \in (\denote{c_1}\pkt)|_V} (\denote{c_2} \pkt')|_V
-    & \text{by set theory} \\
-    = \bigcup_{\pkt' \in (\denote{c_1}(\pkt|_V))|_V} (\denote{c_2} \pkt')|_V
-    & \text{by IH}(c_1) \\
-    = \bigcup_{\pkt' \in \denote{c_1}(\pkt|_V)} (\denote{c_2} (\pkt'|_V))|_V
-    & \text{by set theory} \\
-    = \bigcup_{\pkt' \in \denote{c_1}(\pkt|_V)} ((\denote{c_2} \pkt')|_V)|_V
-    & \text{by IH}(c_2) \\
-    = \bigcup_{\pkt' \in \denote{c_1}(\pkt|_V)} (\denote{c_2} \pkt')|_V
-    & \text{by Lemma~\ref{lem:proj-idem-set}} \\
-    = \left(\bigcup_{\pkt' \in \denote{c_1}(\pkt|_V)} \denote{c_2}\pkt'\right)|_V
-    & \text{by Lemma~\ref{lem:proj-union-set}} \\
-    = \denote{c_2}^\dagger(\denote{c_1}(\pkt|_V))|_V
-    & \text{By definition of }(\cdot)^\dagger\\
-    = (\denote{c_1;c_2}(\pkt|_V))|_V
-    & \text{By definition of }\denote\cdot
-    \end{array}\]
-    The result follows by definition of $=_V$.
-
-  \item[$(c = c_1 \choice c_2)$] Let $\pkt$ be given.
-    \[\begin{array}{ll}
-    (\denote{c_1\choice c_2}\pkt)|_V \\
-    = (\denote{c_1}\pkt \cup \denote{c_2}\pkt)|_V
-    & \text{By definition of }\denote\cdot \\
-    = (\denote{c_1}\pkt)|_V \cup (\denote{c_2}\pkt)|_V
-    & \text{By Lemma~\ref{lem:proj-union-set}} \\
-    = (\denote{c_1}(\pkt|_V))|_V \cup (\denote{c_2}(\pkt|_V))|_V
-    & \text{By IHs} \\
-    = ((\denote{c_1}(\pkt|_V) \cup \denote{c_2}(\pkt|_V))|_V)|_V
-    & \text{By Lemma~\ref{lem:proj-union-set}} \\
-    = (\denote{c_1}(\pkt|_V) \cup \denote{c_2}(\pkt|_V))|_V
-    & \text{By Lemma~\ref{lem:proj-idem-set}}\\
-    = (\denote{c_1 \choice c_2}(\pkt|_V))|_V
-    & \text{By definition of }\denote\cdot \\
-    \end{array}\]
-
-    The result follows by definition of $=_V$.
-
-  \item[$(c = t.\apply())$] Contradictory; $t.\apply()$ is not an action.
-  \end{enumerate}
-
+  For an action $c \in \Cmd$, $\fvs(c) \subseteq V \subseteq \Var$ and $\pkt \in \Pkt$,
+  if $P =_V Q$, then $\denote{c}^\dagger P =_V \denote{c}^\dagger Q$.
+\end{lemma}
+
+\begin{proof}
+  By Lemma~\ref{lem:param-noninter-set}
 \end{proof}
@@
   For $c \in \Cmd$, $\sigma : \State$, $\pkt \in \Pkt$,
-  $\{v_1,v_2\} \subseteq \Value$, $y \in \Var \setminus \fvs(c)$,
+  $v_1,v_2 \in \Value$, $y \in \Var \setminus \fvs(c)$,
   $\denote{c}^\sigma \pkt\{y \mapsto v\} =_{\Var \setminus \{y\}}\denote{c}^\sigma\pkt\{y \mapsto v\}$.
@@
 
-    The result follows by Lemma~\ref{lem:equal-proj}.
+    The result follows by Corollary~\ref{cor:equal-proj}.
 
@@
 
-      The result follows by Lemma~\ref{lem:equal-proj}.
+      The result follows by Corollary~\ref{cor:equal-proj}.
 
@@
           \[\denote{t.\apply()}^\sigma \pkt\{y \mapsto v_1\} = \{\pkt\{y \mapsto v_1\}\} = \{\pkt\{y \mapsto v_2\}\} = \denote{t.\apply()}^\sigma \pkt\{y \mapsto v_2\}\]
-          The result follows by Lemma~\ref{lem:equal-proj}.
+          The result follows by Corollary~\ref{cor:equal-proj}.
       \end{enumerate}
@@
   For $c \in \Cmd$, $\sigma : \State$, $\pkt \in \Pkt$,
-  $\{v_1,v_2\} \subseteq \Value$, $Y \subseteq \Var \setminus \fvs(c)$,
-  $\denote{c}^\sigma \pkt\{y \mapsto v\} =_{\Var \setminus Y}\denote{c}^\sigma\pkt\{y \mapsto v\}$.
+  $V_1,V_2 \subseteq \Value$, $Y \subseteq \Var \setminus \fvs(c)$,
+  $\denote{c}^\sigma \pkt\{Y \mapsto V_1\} =_{\Var \setminus Y}\denote{c}^\sigma\pkt\{Y \mapsto V_2\}$.
 \end{lemma}
@@
     $\state\sigma_c$ where $c \in \Cmd$ as shorthand for $\state\sigma_{\tables(c)}$.}
-  \label{fig:table-instrument}
+  \label{fig:state-encode}
 \end{figure}
@@
   \item[$(c = \assert b)$] \textit{sim}.
-  \item[$(c = c_1;c_2)$] Note that well-formedness\todo{define well formedness as single-use tables} implies $\tables(c_1) \cap
+  \item[$(c = c_1;c_2)$] Note that well-formedness implies $\tables(c_1) \cap
     \tables(c_2) = \emptyset$. Let $T_2 = \tables(c_2)$ and $T_1 = T \setminus
@@
        $\denote{t.\action[r_i.\action](r_i.\data)}\,\pkt$ by
-       Lemma~\ref{lem:proj-den-comm}. The result follows by the IH on
-       $t.\action[r_i.\action](r_i.\data)$ and Lemma~\ref{lem:equal-proj}.
+       Lemma~\ref{lem:proj-den-comm}, since $H_{r_i.\action} =_v \{\pkt\}$. The result follows by the IH on
+       $t.\action[r_i.\action](r_i.\data)$ and Corollary~\ref{cor:equal-proj}.
 
@@
         $\emptyset \neq M \subseteq P =_V \{\pkt\}$, $M =_V \{\pkt\}$ by
-        Lemma~\ref{lem:subset-proj}. The result follows by
-        Lemma~\ref{lem:equal-proj}.
+        Corollary~\ref{cor:subset-proj}. The result follows by
+        Corollary~\ref{cor:equal-proj}.
     \end{enumerate}

### 2021-05-28 inference.tex
polish

@@
   \label{lem:assume-currying}
-  For $b_1,b_2 \in \BExpr$, $\sigma : \Value \to \Row^*$,
+  For $b_1,b_2 \in \BExpr$, $\sigma \in \State$,
   \[\forall \pkt. \denote{\assume {b_1\wedge b_2}}^\sigma \pkt = \denote{\assume {b_1};\assume {b_2}}^\sigma \pkt\]
@@
   \label{lem:prod-sum}
-  For $c_1,c_2, c_3 \in \Cmd$, $\sigma : \Value \to \Row^*$, $\pkt \in \Pkt$.
+  For $c_1,c_2, c_3 \in \Cmd$, $\sigma \in \State$, $\pkt \in \Pkt$.
   $\denote{c_1;(c_2 \choice c_3)}^\sigma\pkt = \denote{c_1;c_2 \choice
@@
 \begin{proof}
-  Let $\sigma : \State $, $y \in \Var$, and $\pkt \in \Pkt$ be given.
+  Let $\sigma \in \State $, $y \in \Var$, and $\pkt \in \Pkt$ be given.
   Proceed by induction on $c$.
@@
   \label{lem:parametricity-set}
-  For $c \in \Cmd$, $\sigma : \State$, $y \in \Var$, $\pkt \in \Pkt$, if $Y
+  For $c \in \Cmd$, $\sigma \in \State$, $y \in \Var$, $\pkt \in \Pkt$, if $Y
   \subseteq \Var \setminus \fvs(c)$, then $\forall \pkt' \in \denote{c}^\sigma
@@
   \label{lem:param-noninter}
-  For $c \in \Cmd$, $\sigma : \State$, $\pkt \in \Pkt$,
+  For $c \in \Cmd$, $\sigma \in \State$, $\pkt \in \Pkt$,
   $v_1,v_2 \in \Value$, $y \in \Var \setminus \fvs(c)$,
@@
   \label{lem:param-noninter-set}
-  For $c \in \Cmd$, $\sigma : \State$, $\pkt \in \Pkt$,
+  For $c \in \Cmd$, $\sigma \in \State$, $\pkt \in \Pkt$,
   $V_1,V_2 \subseteq \Value$, $Y \subseteq \Var \setminus \fvs(c)$,
@@
   \label{thm:subst}
-  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, and
+  For every $c \in \Cmd$, well-formed state $\sigma \in \Value \to \Row^*$, and
   packet $\pkt \in \Pkt$,
@@
 \begin{figure}[htp]
-  %% \begin{minipage}[t]{0.5\textwidth}
-  %% \[\begin{array}{r >{\triangleq}cl}
-  %% \multicolumn{3}{l}{(\cdot)\instr : \Cmd \to \Cmd} \\
-  %% (h.f := e)\instr && h.f := e \\
-  %% (\assume b)\instr && b \\
-  %% (\assert b)\instr && \assert b\\
-  %% (c_1;c_2)\instr && c_1\instr;c_2\instr\\
-  %% (c_1 \choice c_2)\instr && c_1\instr \choice c_2\instr\\
-  %% (\{c\})\instr && \{c\instr\}\\
-  %% (t.\apply())\instr && \\
-  %% \multicolumn{3}{r}{
-  %%   \qquad
-  %%   \begin{array}[t]{l} \displaystyle
-  %%     \{ \\
-  %%     \quad \havoc \rho;\\
-  %%     \quad \assume \rho \in t; \\
-  %%     \quad \assume \rho.\matches = t.\keys; \\
-  %%     \quad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
-  %%     \left(\begin{array}{l}
-  %%       \assume \rho.\action = i;\\
-  %%       t.\action[i](\rho.\data)
-  %%     \end{array}\right)\\
-  %%     \} \choice ( \\
-  %%     \quad \assume \forall \rho. \rho \in t \Rightarrow \rho.\matches \neq t.\keys;\\
-  %%     \quad \SKIP \\
-  %%     )
-  %%   \end{array}}
-  %% \end{array}
-  %% \]
-  %% \end{minipage}\begin{minipage}[t]{0.45\textwidth}
 \[\begin{array}{r >{\triangleq}cl}
@@
   \label{lem:sigma-distrib}
-  For $\sigma : \Value \to \Row^*$, $T,T_1,T_2 \subseteq \Table$ such that $T = T_1 \subseteq T_2$,
+  For $\sigma \in \State$, $T,T_1,T_2 \subseteq \Table$ such that $T = T_1 \subseteq T_2$,
   $\state\sigma_{T} \Leftrightarrow \state{\sigma}_{T_1}\wedge\state{\sigma}_{T_2}$
@@
   \label{lem:sigma-distrib-seq}
-  For $\sigma : \Value \to \Row^*$, $c_1,c_2 \in \Cmd$,
+  For $\sigma \in \State$, $c_1,c_2 \in \Cmd$,
   $\state\sigma_{c_1;c_2} \Leftrightarrow \state{\sigma}_{c_1}\wedge\state{\sigma}_{c_2}$
@@
   \label{lem:sigma-distrib-choice}
-  For $\sigma : \Value \to \Row^*$, $c_1,c_2 \in \Cmd$,
+  For $\sigma \in \State$, $c_1,c_2 \in \Cmd$,
   $\state\sigma_{c_1 \choice c_2} \Leftrightarrow \state{\sigma}_{c_1}\wedge\state{\sigma}_{c_2}$
@@
 \end{proof}
+
+The following theorem relates the instrumented program to a
+formal encoding of the state map $\sigma$, written $\state\sigma_T$ for a set of
+tables $T$. The full definition is in Figure~\ref{fig:state-encode}.
 
@@
   or symbolic variables. For every well-formed $c \in \Cmd$ such that $\fvs(c) \subseteq V$,
-  well-formed state $\sigma : \Value \to \Row^*$, and packet $\pkt \in \Pkt$,
+  well-formed state $\sigma \in \State$, and packet $\pkt \in \Pkt$,
   where $T \supseteq \tables(c)$, $R \supseteq \{\rho_t \mid t \in T\}$, and $G \supseteq \{\gamma_t \mid t \in
@@
     = \denote{\havoc{R_1\cup G_1}; \havoc{R_2 \cup G_2}; \assume{\state\sigma_{T_1}};c_1\instr;\assume{\state\sigma_{T_2}};c_2\instr}\,\pkt
-    & \text{By Lemma~\ref{lem:disj-comm} and } \tables(c_1) \cap \tables(c_2) = \emptyset \\
+    & \text{By Lemma~\ref{lem:disj-comm}}, \\ &\text{since } \tables(c_1) \cap \tables(c_2) = \emptyset \\
     = \denote{(\havoc{R_1 \cup G_1}; \assume{\state\sigma_{T_1}};c_1\instr);
       (\havoc{R_2 \cup G_2};\assume{\state\sigma_{T_2}};c_2\instr)}\,\pkt
-    & \text{By Lemma~\ref{lem:disj-comm} and } \tables(c_1) \cap \tables(c_2) = \emptyset\\
+    & \text{By Lemma~\ref{lem:disj-comm}}, \\ & \text{since } \tables(c_1) \cap \tables(c_2) = \emptyset\\
     = \denote{\havoc{R_2 \cup G_2};\assume{\state\sigma_{T_2}};c_2\instr}^\dagger
@@
     & \text{by definition of }(\cdot)\instr\\
-    %% = \denote{\havoc{R \cup G};\assume{\state\sigma_{T_1} \wedge \state\sigma_{T_2}}; (c_1\instr \choice c_2\instr)}\,\pkt
-    %% & \text{By Lemma~\ref{lem:sigma-distrib-choice}}\\
-    %% = \denote{\havoc{R \cup G};\assume{\state\sigma_{T_1}}; \assume{\state\sigma_{T_2}}; (c_1\instr  \choice   c_2\instr)}\,\pkt
-    %% & \text{By Lemma~\ref{lem:assume-currying}}\\
     = \denote{\havoc{R \cup G};\assume{\state\sigma_{T}}; c_1\instr
@@
     & \text{By Lemma~\ref{lem:prod-sum}} \\
-    %% = \denote{\havoc{R \cup G}; \assume{\state\sigma_{T_1}}; c_1\instr
-    %%   \choice \havoc{R \cup G}; \assume{\state\sigma_{T_2}}; c_2\instr}\,\pkt
-    %% & \text{BY UNUSED-VARIABLES -> TRIVIAL ASSUMPTIONS LEMMA} \\
-    %% =_V \denote{\havoc{R_1 \cup G_1}; \assume{\state\sigma_{c_1}}; c_1\instr
-    %%   \choice \havoc{R_2 \cup G_2}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
-    %% & \text{BY UNUSEDVARIABLE HAVOC ELIMINATION LEMMA } \\
     = \denote{\havoc{R \cup G}; \assume{\state\sigma_{T}}; c_1\instr}\,\pkt
@@
     =_V \denote{c_1}^\sigma\pkt \cup \denote{\havoc{R \cup G}; \assume{\state\sigma_{T}}; c_2\instr}\,\pkt
-    & \text{By IH}(c_1) \text{ since } \fvs(c_1) \subseteq \fvs(c_1 \choice c_2) \subseteq V, \tables(c_1) \subseteq T \\
+    & \text{By IH}(c_1) \text{ since } \fvs(c_1) \subseteq \fvs(c_1 \choice c_2) \subseteq V, \\
+    & \text{and }\tables(c_1) \subseteq T \\
     =_V \denote{c_1}^\sigma\pkt \cup \denote{c_2}^\sigma\pkt
-    & \text{By IH}(c_2) \text{ since } \fvs(c_2) \subseteq \fvs(c_1 \choice c_2) \subseteq V, \tables(c_2) \subseteq T \\
+    & \text{By IH}(c_2) \text{ since } \fvs(c_2) \subseteq \fvs(c_1 \choice c_2) \subseteq V,\\
+    & \text{and }\tables(c_2) \subseteq T \\
     = \denote{c_1 \choice c_2}^\sigma\pkt
@@
   \textbf{Inference Problem.} Given a command $c \in \Cmd$, compute (the
-  weakest) control plane formula $\varphi$ such that if, for a state $\sigma :
-  \Value \to \Row^*$, the following formula is valid \[\state\sigma_c
+  weakest) control plane formula $\varphi$ such that if, for a state $\sigma \in \State$, the following formula is valid \[\state\sigma_c
   \Rightarrow \varphi,\] then \[\forall \pkt \in \Pkt.~ \denote{c}^\sigma\pkt \neq \error \]
@@
   control plane formula $\varphi$, and a test $b$ such that for every
-  packet $\pkt \in \Pkt$ and state $\sigma : \Value \to \Row^*$, if the
+  packet $\pkt \in \Pkt$ and state $\sigma  \in \State$, if the
   following formula is valid

### 2021-07-21 inference.tex
fix typo in instrumentation def
add rewrite rules

@@
 
+%% inference rules
+\newcommand{\andalso}{\quad}
+\newcommand{\pad}[1]{\mbox{}\;#1\;\mbox{}}
+\newcommand{\infrule}[3]{\ensuremath{\frac{\pad{#1}}{\pad{#2}}\;{\textsc{#3}}}}
+
 %% operators
@@
 \newtheorem{corollary}{Corollary}
+\newtheorem{proposition}{Proposition}
 
@@
     \begin{array}[t]{l} \displaystyle
-      \assume t.\matchrow(\rho_t); \\
       \assume \gamma_t.\keys = t.\keys; \\
@@
     \left(\begin{array}{l}
+      \assume t.\matchrow(\rho_t); \\
       \assume \gamma_t.\miss = 0;\\
@@
 
+\subsection{Heuristic Rules}
+
+Here we posit a collection of heuristic elimination rules for quantified
+variables that seem to be effective (at least in some small examples). The goal
+here is to define a collection of rules that will remove quantifiers for a broad
+class of programs.
+
+\begin{figure}[H]
+  \[
+    \begin{array}{>{\displaystyle}c}
+      \infrule{
+      \psi[\forall x. \varphi_{1} \Rightarrow \varphi_{2}] \andalso x \not \in \fvs(\varphi_{1})}{\psi[\varphi_{1} \Rightarrow \forall x. \varphi_{2}]}{ReduceScopeR} \qquad
+
+      \infrule{\psi[\forall x. \varphi_{1} \Rightarrow \varphi_{2}] \andalso x \not \in \fvs(\varphi_{2})}{\psi[(\forall x. \neg \varphi_{1}) \vee \varphi_{2}]}{ReduceScopeL} \\[1em]
+
+      \infrule{\psi[\forall x. \varphi] \andalso x \not \in \fvs(\varphi)}{\psi[\varphi]}{UnusedUVar}
+      \qquad
+      \infrule{\psi[\exists x. \varphi] \andalso x \not \in \fvs(\varphi)}{\psi[\varphi]}{UnusedEVar}  \\[1em]
+
+      \infrule{\psi[\forall x. x = e]}{\FALSE}{UnivEq}
+      \qquad
+      \infrule{\psi[\exists x. x = e]}{\TRUE}{ExtEq}
+    \end{array}
+  \]
+\end{figure}
+
+Now we prove that these rules are sound, that is, applying them to a formulae
+$\psi$ produces a logically equivalent formula $\psi'$.
+
+\begin{proposition}[Rewrite Soundness]
+  The above rules are sound.
+\end{proposition}
+\begin{proof}
+  Consider each rule separately.
+  \begin{enumerate}[align=left]
+    \item[(\textsc{ReduceScopeR})]
+    Let $x$, $\varphi_{1}$ and $\varphi_{2}$ be given such that $x \not \in \fvs(\varphi)$. Let $\mu$ be a (total) model. Show that
+    \[\mu \models \forall x. \varphi_{1} \Rightarrow \varphi_{2} \iff \mu \models \varphi_{1} \Rightarrow \forall x. \varphi_{2}\]
+    Show each direction separately.
+    \begin{enumerate}[align=left]
+      \item[($\Rightarrow$)] Assume $\mu \models \forall x.
+\varphi_{1} \Rightarrow \varphi_{2}$.
+Further, assume $\mu \models \varphi_1$.
+Let $v$ be given to show $\mu[x \mapsto v] \models \varphi_{2}$.
+By the semantics of universal quantification $\mu[x \mapsto v] \models \varphi_{1} \Rightarrow \varphi_{2}$.
+Since $x$ does not occur in $\varphi_{1}$, and $\mu \models \varphi_{1}$ we know $\mu[x \mapsto v] \models \varphi_{1}$.
+By definition of the semantics of implication, we know $\mu[x \mapsto v] \models \varphi_{2}$.
+$\checkmark$
+
+      \item[$(\Leftarrow)$] Assume $\mu \models \varphi_1 \Rightarrow \forall x. \varphi_{2}$.
+      Let $v$ be given to show $\mu[x \mapsto v] \models \varphi_{1} \Rightarrow \varphi_{2}$.
+      Assume $\mu[x \mapsto v] \models \varphi_{1}$ (otherwise we'd be done), to show $\mu[x \mapsto v] \models \varphi_{2}$.
+      This assumption gives us $\mu \models \varphi_{1}$ since $x \not \in \fvs(\varphi_{1})$. Then
+      $\mu \models \forall x. \varphi_{2}$, which means, by definition of the semantics of universal quantification, that $\mu[x \mapsto v] \models \varphi_{2}$. $\checkmark$
+    \end{enumerate}
+
+    \item[(\textsc{ReduceScopeL})]
+    Let $x$, $\varphi_{2}$, and $\varphi_2$ such that $x \not \in \fvs(\varphi_{2})$.
+    Let $\mu$ be a total model, to show that
+    \[\mu \models \forall x. \varphi_{1} \Rightarrow \varphi_{2}
+      \iff
+      \mu \models (\forall x. \neg \varphi_{1}) \vee \varphi_{2}
+    \]
+    Prove each direction separately:
+    \begin{enumerate}[align=left]
+      \item[($\Rightarrow$)]
+      Assume $\mu \models \forall x. \varphi_{1} \Rightarrow \varphi_{2}$.
+      Assume $\mu \models \exists x. \varphi_{1}$ to show $\mu \models \varphi_{2}$.
+      Let $v$ be such that $\mu[x \mapsto v] \models \varphi_{1}$.
+      Then we know that $\mu[x \mapsto v] \models \varphi_{1} \Rightarrow \varphi_{2}$.
+      Further, since $\mu[x \mapsto v] \models \varphi_{1}$, we know $\mu[x \mapsto v] \models \varphi_{2}$. Since $x \not \in \fvs(\varphi_{2})$, conclude that $\mu \models \varphi_{2}$.
+      \hfill $\checkmark$
+
+      \item[($\Leftarrow$)]
+      Assume $\mu \models (\forall x.
+\neg \varphi_{1}) \vee \varphi_{2}$.
+To show that $\mu \models \forall x.
+\varphi_{1} \Rightarrow \varphi_{2}$.
+Let $v$ be arbitrary, and assume that $\mu[x \mapsto v] \models \varphi_{1}$ (otherwise we'd be done), to show that $\mu[x \mapsto v] \models \varphi_{2}$.
+
+      Our assumption gives us two cases. If $\mu \models \varphi_{2}$ then the result follows since $x \not \in \fvs(\varphi_{2})$ means $\mu[x \mapsto v] \models \varphi_{2}$.
+
+      In the second case, $\mu \models (\forall x. \neg \varphi_{1})$, which means that for our $v$,
+      $\mu[x \mapsto v] \not\models \varphi_{1}$. This contradicts the assumption that $\mu[x \mapsto v] \models \varphi_{1}$. And we're done.
+
+    \end{enumerate}
+
+    \item[(\textsc{UnusedUVar})] \todo[inline]{Prove this case}
+    \item[(\textsc{UnusedEVar})] \todo[inline]{Prove this case}
+    \item[(\textsc{UnivEq})] \todo[inline]{prove this case}
+    \item[(\textsc{ExtEq})] \todo[inline]{prove this case}
+
+  \end{enumerate}
+
+\end{proof}
+
 \clearpage

### 2021-07-22 inference.tex
finish rule proofs

@@
 
-      \infrule{\psi[\forall x. x = e]}{\FALSE}{UnivEq}
+      \infrule{\psi[\forall x. x = v]}{\FALSE}{UnivEq}
       \qquad
-      \infrule{\psi[\exists x. x = e]}{\TRUE}{ExtEq}
+      \infrule{\psi[\exists x. x = v]}{\TRUE}{ExtEq}
     \end{array}
@@
 $\psi$ produces a logically equivalent formula $\psi'$.
+
+\begin{lemma}[Model Freedom for Expressions]
+  \label{lem:model-freedom-expressions}
+  For $\mu$, $v$, $x$, $e$, such that $x \not \in \efvs(e)$,
+  $\edenote e \mu = \edenote e \mu[x \mapsto v]$
+\end{lemma}
+\begin{proof}
+  Proceed by induction on $e$.
+  \begin{enumerate}[align=left]
+    \item[($e = v$)] Immediate.
+    \item[($e = y$)] Let $\mu$, $v$ and $x$ be such that $x \not \in \{y\}$.
+    $\edenote y \mu = \mu(y) = \mu[x \mapsto v](y) = \edenote y \mu[x \mapsto v]$.
+    \item[$(e = e_{1} \binop e_{2})$] Let $\mu$, $v$ and $x$ be such that
+    $x \not \in \efvs(e_{1})$ and $x \not \in \efvs(e_{2})$.
+    Then the IHs give $\edenote{e_{i}}\mu = \edenote{e_{i}}\mu[x \mapsto v]$, for $i=1,2$.
+    The result follows by the definition of semantics of binary operations.
+    \item[$(e = \unop e)$] \textit{sim.}
+  \end{enumerate}
+\end{proof}
+
+\begin{lemma}[Model Freedom]
+  \label{lem:model-freedom}
+  For $\mu$, $v$, $x$, $\varphi$ such that $x \not \in \fvs(\varphi)$,
+  $\mu \models \varphi$ iff $\mu[x \mapsto v] \models \varphi$
+\end{lemma}
+\begin{proof}
+  Proceed by induction on $\varphi$.
+  \begin{enumerate}[align=left]
+    \item[$(\FALSE)$] Each direction follows vacuously, there is no such $\mu$, $x$, and $v$.
+    \item[$(\varphi_{1} \Rightarrow \varphi_{2})$] Let $\mu$, $v$, and $x$ be such that $x \not \in \fvs(\varphi_{1} \Rightarrow \varphi_{1})$. Calculate as follows:
+
+    \[\begin{array}{ll}
+        \mu \models \varphi_{1} \Rightarrow \varphi_{2}\\
+    \iff \mu \not \models \varphi_{1} \text{ or } \mu \models \varphi_{2} \\
+    \iff \mu[x \mapsto v] \not\models \varphi_{1} \text{ or } \mu[x \mapsto v] \models \varphi_{2} & \text{By IH}\\
+    \iff\mu[x \mapsto v] \models \varphi_{1} \Rightarrow \varphi_{2}
+      \end{array}\]
+
+    \item[$(\forall y. \varphi)$] Let $\mu$, $v$, and $x$ be such that $x \not \in \fvs(\varphi) \setminus \{y\}$. There are two cases $y = x$ or $y \neq x$.
+    Consider the first case (i.e. where $x = y$):
+    \[ \begin{array}{l}
+         \mu \models \forall x. \varphi\\
+         \iff \forall v'. \mu[x \mapsto v'] \models \varphi \\
+         \iff \forall v'. \mu[x \mapsto v][x \mapsto v'] \models \varphi \\
+         \iff \mu[x \mapsto v] \models \forall x. \varphi \hfill \checkmark
+       \end{array}
+     \]
+
+  Consider the second case (i.e. where $y \neq x$):
+  \[\begin{array}{ll}
+      \mu \models \forall y. \varphi\\
+      \iff \forall v'. \mu[y \mapsto v'] \models \varphi \\
+      \iff \forall v'. \mu[y \mapsto v'][x \mapsto v] \models \varphi & \text{By IHs}\\
+      \iff \forall v'. \mu[x \mapsto v][y \mapsto v'] \models \varphi & \text{Since } y \neq x\\
+      \iff \mu[x \mapsto v] \models \forall y. \varphi \hfill \checkmark
+    \end{array}\]
+
+  \item[$(e_{1} = e_{2})$] Let $\mu$, $v$, and $x$ be such that $x \not \in \efvs(e_{1})\cup\efvs(e_{2})$.
+  Since this also means $x \not \in \efvs(e_{i})$ for $i = 1,2$, Lemma~\ref{lem:model-freedom-expressions} gives $\edenote{e_{i}}\mu = \edenote{e_{i}} \mu[x \mapsto v]$ for $i=1,2$.
+The result is immediate.
+\hfill $\checkmark$
+  \end{enumerate}
+\end{proof}
 
@@
 
-    \item[(\textsc{UnusedUVar})] \todo[inline]{Prove this case}
+    \item[(\textsc{UnusedUVar})] Let $x$ and $\varphi$ be given such that
+    $x \not \in \fvs(\varphi)$. Let $\mu$ be a (total) model. Show that
+    \[\mu \models \forall x. \varphi \iff \mu \models \varphi\]
+    Prove each direction separately:
+    \begin{enumerate}[align=left]
+      \item[($\Rightarrow$)] Assume $\mu \models \forall x. \varphi$ to show
+      $\mu \models \varphi$. Since $\mu$ is total, the assumption gives $\mu[x\mapsto \mu(x)] \models \varphi$. \hfill $\checkmark$
+
+      \item[($\Leftarrow$)] Assume $\mu \models \varphi$, to show $\mu \models \varphi$.
+Let $v$ be arbitrary.
+Since $x \not \in \fvs(\varphi)$, the assumption gives $\mu[x \mapsto v] \models \varphi$.
+\hfill $\checkmark$
+    \end{enumerate}
+
     \item[(\textsc{UnusedEVar})] \todo[inline]{Prove this case}
-    \item[(\textsc{UnivEq})] \todo[inline]{prove this case}
-    \item[(\textsc{ExtEq})] \todo[inline]{prove this case}
+    \item[(\textsc{UnivEq})]
+    Let $x$ and $e$ be given. Let $\mu$ be a (total) model. Show that
+    \[\mu \not\models \forall x. x = v\]
+    Equivalently, show that $\mu \models \exists x. x \neq v$.
+    There are two cases either $v = 0$ or $v \neq 0$. If $v \neq 0$ then $\mu[x \mapsto 0] \models x \neq v$.
+    If $v = 0$ then $\mu[x \mapsto 1] \models x \neq v$. \hfill $\checkmark$
+
+    \item[(\textsc{ExtEq})]
+    Let $x$ and $\varphi$ be given. Let $\mu$ be a total model. Show that
+    \[\mu \models \exists x. x = v\]
+    This is immediate since $\mu[x \mapsto v] \models x = v$.
 

### 2022-01-26 benchmarks/store.py
make time implementation

@@ -1,0 +1,38 @@
+# stores [label][field] = value in possibly existing json file containing experiment results
+
+import argparse
+import subprocess
+import sys
+import os
+import time
+import json
+
+read_dir = './'
+results_file = 'out.json'
+
+
+if len(sys.argv) < 2:
+    print('Usage: python3 sotre.py <label> <field> <value>, e.g. python3 store dir_listing size 32MB')
+
+else:
+    try:
+        with open(results_file, "r") as f:
+            results = json.load(f)
+    except:
+        results = {}
+        
+
+    # label = sys.argv[1]
+    label = sys.argv[1].split("/")[-1:][0]
+    field = sys.argv[2]
+    value = sys.argv[3]
+    if label not in results:
+        results[label] = {}
+    results[label][field] = value
+    
+    with open(results_file, "w") as outfile:
+        json.dump(results, outfile)
+
+# with open(results_file, "r") as f:
+#     results = json.load(f)
+# print(results)

### 2022-01-26 benchmarks/time.py

@@ -1,0 +1,42 @@
+# times commands passed as arguments, saves in a json file
+
+import argparse
+import subprocess
+import sys
+import os
+import time
+import json
+
+read_dir = './'
+results_file = 'out.json'
+
+
+if len(sys.argv) < 2:
+    print('Usage: python3 time.py <label> <command>, e.g. python3 time.py dir_listing ls -l')
+
+else:
+    try:
+        with open(results_file, "r") as f:
+            results = json.load(f)
+    except:
+        results = {}
+        
+
+    # label = sys.argv[1]
+    label = sys.argv[1].split("/")[-1:][0]
+    results[label] = {}
+    start = time.time()
+
+    try:
+        # subprocess.check_call(sys.argv[2:], cwd=read_dir, stderr=subprocess.DEVNULL, stdout=subprocess.DEVNULL)
+        subprocess.check_call(sys.argv[2:], cwd=read_dir)
+
+    except:
+        None
+
+    end = time.time()
+    
+    results[label]['elapsed-ms'] = (end - start) * 1000
+    
+    with open(results_file, "w") as outfile:
+        json.dump(results, outfile)

### 2022-01-26 benchmarks/timebf4.sh

@@ -1,0 +1,124 @@
+#! /usr/bin/bash
+
+test_dir="./examples/bf4_failing"
+includes_dir="./examples/includes"
+success_dir="./examples/bf4_passing"
+nocomp_dir="./examples/bf4_nocompile"
+dup_dir="./examples/bf4_dups"
+log_dir="./examples/logs"
+
+# coloration
+RED='\033[0;31m'
+GRN='\033[0;32m'
+YLW='\033[0;33m'
+BLU='\033[0;34m'
+NC='\033[0m' # No Color
+PASS="${GRN}[PASS]${NC}"
+FAIL="${RED}[FAIL]${NC}"
+BUNK="${YLW}[BUNK]${NC}"
+
+# mkdirs if they dont exist
+if [ ! -d $success_dir ]; then
+    mkdir $success_dir
+fi
+if [ ! -d $nocomp_dir ]; then
+    mkdir $nocomp_dir
+fi
+if [ ! -d $dup_dir ]; then
+    mkdir $dup_dir
+fi
+if [ ! -d $log_dir ]; then
+    mkdir $log_dir
+fi
+
+if [ ! -d $test_dir ]; then
+    echo "error, couldn't find test directory ${test_dir}" >&2
+    exit 1
+fi
+
+
+if [ -z $(which petr4) ]; then
+    echo "error petr4 is not installed" >&2
+    exit 1
+fi
+
+# move everything to the starting directory
+if [ ! -z "$(ls $success_dir)" ]; then
+    mv $success_dir/* $test_dir
+fi
+if [ ! -z "$(ls $nocomp_dir)" ]; then
+    mv $nocomp_dir/* $test_dir
+fi
+
+
+# make capisce
+eval $(opam env)
+make
+
+# if building succeeded run the tests
+if [ $? -eq 0 ]; then
+    # iterate through test programs
+    for f in $test_dir/*.p4; do
+        b=$(basename $f)
+        log=${log_dir}/${b}.log
+        # check if the program works with the petr4 frontend
+        errors=$(petr4 typecheck $f -I $includes_dir > ${log} 2>&1)
+        # if it doesn't move it to the directory of non-compiling programs
+        if [ ! -z "${errors}" ]; then
+            echo -e "$BUNK ${BLU}${b}${NC}"
+            echo -e "\tcat ${log}"
+            mv $f $nocomp_dir
+        else
+            # otherwise run icecap
[…]

### 2022-01-26 benchmarks/store.py
totex.py converts out.json to out.tex, a latex-formatted table with information from m the experiments saved in out.json

@@ -22,7 +22,6 @@ else:
         results = {}
         
 
-    # label = sys.argv[1]
     label = sys.argv[1].split("/")[-1:][0]
     field = sys.argv[2]
     value = sys.argv[3]

### 2022-01-28 benchmarks/timebf4.sh
fixes for latex table generation

@@ -72,15 +72,18 @@ if [ $? -eq 0 ]; then
             # otherwise run icecap
             echo -e -n "[..] ${BLU}${b}${NC}\r"
             python3 ./time.py $f ./icecap infer $f -I $includes_dir --skip-check > ${log} 2>&1
-            python3 ./store.py $f "size" `du -sh ${log}`
+            python3 ./store.py $f "size" `grep -e "(" -o ${log} | wc -l`
+            
             if [ $? -eq 0 ]; then
                 echo -e "$PASS ${BLU}${b}${NC}"
                 echo -e "\tcat ${log}"
                 mv $f $success_dir/
+                python3 ./store.py $f "valid" "x"
             else
                 echo -e "$FAIL ${BLU}${b}${NC}"
                 echo -e "\tcat ${log}"
                 echo -e "\t./icecap infer ${f} -I ${includes_dir} -D"
+                python3 ./store.py $f "valid" "-"
             fi
         fi
     done

### 2022-02-09 benchmarks/plotter.py
post-sigcomm snapshot

@@ -1,0 +1,234 @@
+# -*- coding: utf-8 -*-
+import argparse
+import csv
+import locale
+import time
+import matplotlib as mpl
+# mpl.use('TkAgg')
+import matplotlib.pyplot as plt
+from cycler import cycler
+from collections import defaultdict
+import numpy as np
+import math
+
+locale.setlocale(locale.LC_ALL, 'en_US.UTF8')
+from IPython.display import set_matplotlib_formats
+set_matplotlib_formats('png')
+
+onecolsize = (4, 1.5)   # Tweak based on figure's appearance in the paper
+seaborn_colorblind = cycler('color', ['#0072B2', '#D55E00', '#009E73', '#CC79A7', '#F0E442', '#56B4E9'])
+seaborn_muted = cycler('color', ['#4878CF', '#6ACC65', '#D65F5F', '#B47CC7', '#C4AD66', '#77BEDB'])
+
+def setrcparams():
+  # setup matplotlib rcparams
+  plt.style.use(['seaborn-paper', 'seaborn-colorblind'])
+  # color cyclers
+  seaborn_colorblind = cycler('color', ['#0072B2', '#D55E00', '#009E73', '#CC79A7', '#F0E442', '#56B4E9'])
+  seaborn_muted = cycler('color', ['#4878CF', '#6ACC65', '#D65F5F', '#B47CC7', '#C4AD66', '#77BEDB'])
+
+  plt.rcParams['axes.prop_cycle'] =  seaborn_colorblind + cycler(linestyle=['-', '--', '-.', '--', '-.','-'])
+
+  plt.rcParams['axes.axisbelow'] = True
+  plt.rcParams['axes.edgecolor'] = 'lightgray'
+  plt.rcParams['axes.facecolor'] = 'white'
+  plt.rcParams['axes.spines.left'] = False
+  plt.rcParams['axes.spines.bottom'] = False
+  plt.rcParams["axes.spines.right"] = False
+  plt.rcParams["axes.spines.top"] = False
+  plt.rcParams['axes.grid'] = True
+  plt.rcParams['axes.linewidth'] = 0.1
+
+
+  plt.rcParams['grid.alpha'] = 0.4
+  plt.rcParams['grid.color'] = 'gray'
+  plt.rcParams['grid.linestyle'] = ':'
+  plt.rcParams['grid.linewidth'] = 1.0
+
+  plt.rcParams['hatch.linewidth'] = 1.0
+
+  plt.rcParams['xtick.bottom'] = False
[…]

### 2022-02-09 benchmarks/totex.py

@@ -24,7 +24,7 @@ else:
     except:
         results = {}
         
-fields = ["Source P4 File"]
+fields = ["\\textsc{P4 Program}"]
 first_experiment = list(results)[0]
 fields += list(results[first_experiment])
 
@@ -34,22 +34,30 @@ for experiment in results:
 
 rows = [fields]
 for experiment in experiments:
-    row = ["\\tt{" + experiment.replace("_","\_")[:-3] + "}"] # removes .p4 from files, adds '\' before '_', adds \tt{}
+    row = ["\\texttt{" + experiment.replace("_","\_")[:-3] + "}"] # removes .p4 from files, adds '\' before '_', adds \tt{}
     for value in results[experiment].values():
         if value == "x":
-            row += ['$\\times$']
+            row += ['']
         elif value == "-":
             row += ["$\\bot$"]
         else:
-            row += [f'${float(value):.2f}$']
+            # row += [f'${float(value):.2f}$']
+            row += [float(value)]
     rows.append(row)
 
-rows[0][1] = "Elapsed time (ms)"
-rows[0][2] = "Size (# AST nodes)"
-rows[0][3] = "Satisfiable"
+rows[0][1] = "Time (ms)"
+rows[0][2] = "NT AST Size"
+rows[0][3] = "(= \\texttt{false})"
+
+rows = [rows[0]] + sorted(rows[1:], key=lambda r: r[1])
+rows = [ [ v if isinstance(v, str) else f'${v:.2f}$'
+           for v in row]
+         for row in rows]
+
+
 
 table = texttable.Texttable()
-table.set_cols_align(["l", "l", "l", "l"])
+table.set_cols_align(["l", "l", "l", "c"])
 table.set_cols_valign(["t", "t", "t", "t"])
 table.add_rows(rows)
 

### 2022-08-02 benchmarks/totex.py
update opam file

@@ -16,7 +16,7 @@ output_file = 'out.tex'
 
 if len(sys.argv) != 1:
     print('Usage: python3 totex.py')
-
+    exit
 else:
     try:
         with open(input_file, "r") as f:

### 2023-09-15 popl-rebuttal.md
add rebuttal

# Rebuttal

We wish to thank the reviewers for their time and insightful comments. It seems
that the reviewers largely agreed on the strengths (exciting connection between
databases and data planes) and weaknesses (depth, evaluation and presentation)
of the paper. We discuss our plan for addressing the weaknesses in this
response.

## Overview

Our vision with this paper is to demonstrate the power of database theory in the
domain of data plane programming, by bringing it to bear on the data plane
emulation problem. Specifically we rely on an algorithm from database theory
called the chase, introducing several extensions of the classical algorithm to
solve our problem. Our contributions are as follows:

- An relational encoding of data plane pipeline programs that let us use the chase
to produce: 
    - a semantics for data planes 
    - a semantics for the data plane emulation problem 
- A novel extension called the priority chase that we use to provide: 
    - a priority-aware semantics for data planes with wildcards
    - an algorithm for computing priorities of target programs in the data plane
      emulation problem
    - A symbolic chase algorithm that gives us the first offline solution to the
      data plane emulation problem 
   - The notion of the generative chase and its generative solutions to provide
     a semantics for the data plane emulation problem with action-data.

__Depth__. Reviewer B objected that there is a mismatch between our motivation
in our introduction and the realities of our technical development. Indeed, upon
revisiting our introduction, we agree that our current introduction sets up the
reader to believe that we are going to develop a mathematical correspondence
between the two fields. Instead, our paper is focused on using database theory
to solve an open problem in networking. We devise special variants of key
database theory concepts (dTGDs and EGDs) and use them to define
networking-specific chase algorithms, computing the first offline algorithm for
control plane mapping. We will adjust our overly poetic introduction to set
expectations appropriately.

With this in mind, we believe our work is sufficiently technically deep to merit
publication. We propose 3 new variants of the chase that rely on nonstandard
formalisms for dependencies, and prove them correct, to solve an open question
in networking.

__Presentation__. We would like to thank the authors for their patience in
reviewing this paper, despite its typos and density. Upon revisiting the text,
we agree that the background sections are too dense to effectively onboard
readers unfamiliar with the out-of-domain concepts: the chase and data planes.
It’s also clear that the later sections of the paper are lacking in
intuition-providing prose. Luckily, with four more pages and nearly three weeks
to revise, we can absolutely add the requested detail. We’ve outlined our
revision plan in the Change List section below.

__Evaluation__. Reviewers A and B both pointed out that an implementation and an
evaluation would strengthen the paper. It seems that the key missing piece is
the scalability, which we’ve stated is exponential in the size of the programs.
Indeed, if we were to do an empirical evaluation of this work, we wouldn’t learn
much beyond what we can compute analytically, since we can directly compute the
size of the search space from the inputs. Our contribution here is intended to
be theoretical: we can effectively model problem solutions in networking using
database theory.

All told, we do believe that our formal contribution, comprising a semantic
model of pipelines, and a solution to the control plane mapping problem via the
formal development of the novel symbolic, priority, and generative chases, are
sufficiently interesting for a POPL audience.

## Change List

We propose the following changes to our work to address the questions of Presentation, Depth, and Evaluation summarized above

__Presentation__
- In section 2, we will provide a clear road map, in a figure, for the technical
contributions of the paper to set expectations and help guide the reader through
the technical space.
- For our background sections (especially 3.2 and 3.3), we will provide examples
and take more time to explain the essential concepts. We expect these sections
to at least double in size. 
- We will weave additional signposting, intuition-providing explanations, and
examples into the later technical sections. 
- We will certainly look for places to simplify our notation. For instance, we
can remove the subscripts on $\phi$ and $\psi$ on line 251, and stick to one
notation for collections of variables wherever possible (e.g. unify $\vec x$ and
$\mathbf{x}$). 
- To provide concreteness to our examples, we will limit our use of the
set-theoretic notation for tables and prefer the concrete match-action
representation as on 106-109. 
- We will fix all copy-editing issues (typos and use-before-def errors)

__Depth__.
- We will make our contribution clearer in the introduction. We are not
  proposing an insightful connection between the two fields but rather a
  wrangling of database theory to solve interesting problems in software defined
  networking. For instance, our verbiage about “building a bridge between
  databases and data planes” could indeed be scaled back to avoid misleading
  readers.

__Evaluation__.
- We will modify the case-study figures (e.g. Fig 3) to include the source and
  target instances notated in the match-action style from the introduction.

## Detailed Response

Here we respond inline to the reviewers' comments.

### Reviewer A

> However, the scalability performance of algorithms in this space is known to
> be a challenge, and although the paper presents some nice small clear
> examples, it is not clear if the approach can overcome these scalability
> issues. Using the chase algorithm is a heavy hammer for the examples provided
> in the paper, but its deployment may be justified with suitability large and
> sophisticated pipelines, however this is yet to be determined. A few more
> practical or experimental analyses and results to help underpin this approach
> would certainly merit a higher score for this paper.

While an implementation is out of the question for this work, we are happy to
include the empirical results describing the size of the search space for our
existing case studies.

#### Detailed Feedback

> Page 4, line 187. "tabls" should be "tables".

Thanks! We’ve made this change.

> Page 5, line 204, I tripped up over the use of "Recall" ~ I don't think the
> half program observation was made earlier in this paper? The "recall" is just
> for readers of this paper that have also read Liu et. al. 2018?

Thanks! We’ll remove this use of “Recall”.

> Page 12, line 580. Perhaps this sentence needs rewriting. "From this point on,
> we will elide the bookkeeping associated explicit with priorities when
> possible."

Thanks! This should be “From this point on, we will elide the bookkeeping
associated with explicit priorities when possible”

### Reviewer B

> However, the rest of the paper underdelivers. The theoretical connection
> established between networking and databases seems to be a simple encoding
> that treats everything as relations. Deeper, insightful connections and
> results would make the paper much stronger.

While the paper does start with a simple relational encoding of match-action
tables into relations, this simplicity should not be seen as a downside. In
fact, other relational encodings of data planes either make significant
simplifying assumptions, or only exploit superficial similarities between
match-action tables and database tables. Our work models the core
match-action abstraction, and we are able to show that this approach is valuable
by using it to solve an open problem in networking.

From this point on in the paper, we endeavor to show that thinking of match
action-table pipelines relationally is very powerful. The majority of the paper
is focused on solving the control plane mapping problem, which requires
introducing new variants of the chase (priority, symbolic, generative). Indeed
this is the meat of our contribution.

> Further, my major concern is actually about the presentation which also leads
> to my overall recommendation. If I just peek through the paper, I get the
> rough idea of each section or subsection, but once I zoom in to the technical
> details and formalism, I get lost quite quickly. Overall the text and the
> formulae are interleaved but not really connected. In too many cases,
> terminologies and notations pop out abruptly in the text without prior
> explanation and definition, quite a few mathematical notations are excessive
> unnecessary, and formulae are put in places not very related to the context.
> Examples are needed in many places.

We agree. We can insert signposts and examples where necessary. We will also
work to reduce the notational burden on the reader.

> And the number of typos exceeds the negligible level.

The paper is quite inundated with typos. We apologize deeply for these
oversights. Luckily, these can be fixed in shepherding.

> Some examples are given at the end of the review.

Thanks! We’ve addressed these below.

> Besides the presentation defects, the paper also lacks a solid evaluation
> component. The vast majority of the paper is dedicated to the process of
> encoding toward a formulation that can be handled by the chase algorithm, and
> it is not clear at all what happens beyond that.

As we described above, we are happy to provide the search space size for our
case studies, however we believe that the theoretical contributions suffice for
publication.

> How is the problem solved? Is the problem just sent to an off-the-shelf
> implementation of the chase framework?

Because of our technical changes to the chase, we cannot simply send it off to a
solver.

> Did the paper actually solve the case study?

Yes! The solutions to the case studies are presented in the examples themselves.
We will be sure to highlight this by using the table notation to depict the
solved tables.

> How's the result/performance compared to existing approaches, e.g., Avenir as
> mentioned in the paper?

This is unknown. But we can provide an analytical justification for this paper’s
superiority. Avenir’s runtime is bounded by the number of possible rule
combinations that can be inserted into the target table—its performance relies
heavily on an ad-hoc collection of heuristics. The performance of the chase
algorithms in this paper are bounded by the size of the source instance and the
number of paths in the target program, which is much smaller.

#### Specific Questions

> Besides the encoding presented in the paper, is there any fundamental
> connection between data plane programming and database theory that can be
> leveraged to benefit both areas?

This paper shows one example of how the simple relational encoding can benefit
the networking domain, by solving the open control plane emulation problem. We
believe that there are _many_ database technologies that we can use to benefit
networking, for example data provenance or query optimization.

The other direction is less clear. We expect that efficient compression and
matching algorithms such as those used in networking hardware (e.g. LPM and
TCAM) could allow more efficient data representation in certain application
domains.

#### Detailed Feedback

> line 33: ...we use the [the] chase…

Thanks! This has been fixed.

> line 99: ...where RR is [is] the…

Thanks! This has been fixed.

> line 157: it is not clear what the "cross product" exactly means.

This is a good point. Here we use the term loosely. We intended it to be
evocative, but not technical, since the combine table on lines 160- 167 is not
technically the cross product of the tables.

We construct it thusly. For each match-action row (m0,a0) in the first table,
and each match-action row (m1, a1) in the second table we produce a row that
loosely corresponds to (m0, m1, a0, a1). We then simplify the table following
the idempotence of allow and deny and the fact that any deny will supersede an
allow to simplify the set of actions and rows. These are intuitions from the
reality of network programming that we do not include in our description.

We will remove the use of “cross product” and replace it with more precise and
descriptive prose.

> lines 221-224: the formulae are out of nowhere -- why are they here?

Lines 213 - 220 introduces the concept of semantics of pipelines and then lines
221-224 define them. We will be sure to signpost this by adding the line “We
define the semantics of pipelines in the equations below.” There is also a typo
on line 213 that uses banana brackets instead of Strachey brackets for the
semantics. We will fix this as well.

> line 235: the difference between tuple-generating and equality-generating is
> not clear. More explanation and justification needed to show why these kinds
> of dependencies are permitted.

Briefly, an EGD is a dependency whose head is an equality between variables. A
TGD is a dependency whose head is a logical formula over relations only. Indeed
there is a typo here on line 240. EGDs should be $\forall \mathbf{x}.
\phi(\mathbf{x}) \to x_0 = x_1$, where $x_0, x_1 \in \mathbf{x}$. They are
permitted simply because the chase algorithm is defined explicitly in terms of
these rules. We are happy to provide intuition for why these forms are easy to
define chase algorithms for. Following collective feedback that section 3 is too
dense, we will be sure to pepper in examples and more explanatory prose and
context throughout.

> line 305: $a−j$ => $a_j$

Thanks! This has been fixed.

> line 312: I am confused with A(x,x′) and aj(x)=x′ -- why A becomes aj​? What's
> the difference between A and a?

Following the definitions, an $a_j$ is a syntactic representation of the
actions. $A_j$ is the relational interpretation of the action itself. We use
$a_j(\mathbf{x})=\mathbf{x’}$ as syntactic sugar for $A_j(\mathbf{x},
\mathbf{x’})$ for notational simplicity and familiarity, despite the fact that
it is not syntactically well-defined: $a_j$ is syntax, has no binding construct,
and must be interpreted to be run. In hindsight, this is confusing. We propose
to stick to the $A_j$ notation in the final version.

> lines 340-342: the one-point rules are confusing and I got lost. Explain more.

Thanks! Loosely, the one-point rule says that if we have $x=y => \phi(x)$ we can
rewrite this as $\phi(y)$. It lets us perform the substitutions indicated by the
actions in the body of the dependency on the head. This is why the rules have no
explicit actions, because they’ve been substituted away.

Our purpose here was to simplify the complexity of the formulas. We’ll either
draw this out more carefully or include the literal actions depending on which
is more clear.

> line 418: what is "chase over dependencies"? Not defined anywhere and no idea
> what it is.

Good catch. We don’t define this specifically. In definition 3.2 we define “A
chase tree of $\rho$ over $\Sigma$” where $\Sigma$ is a set of dependencies. It
is common in the chase literature to use the verb “chase” to mean “run the chase
algorithm” or “compute the chase tree”. We combined these without providing such
an explanation. We’ll explain this jargon clearly in section 3.2 and remove
earlier references to it (e.g. on line 172).

> line 497: "chase tree" needs definition and explanation.

“Chase tree” is defined in Definition 3.2 on line 287. We will add more
explanation and a visual example after the definition so it is not so easy to
miss.

> lines 524-527: why the formula is shown here?

These denotational priority semantics for pipelines, as motivated in the
previous paragraph. We will add prose to explicitly signpost the relationship.

>  The ⊴R is not defined.

Thanks for spotting the use-before-def! It is meant to indicate the priority
relationship attached to the relation R. We’ll be sure to fix this during the
revision period.

> line 628: "Recall that Avenir..." You did not mention Avenir anywhere before…

Thanks! We’ve removed “Recall”

> line 662: "Now well" => "Now we'll" ?

Thanks! We’ve made this change.

> line 812: ...were ma[n]y ways...

Thanks! We’ve made this change.

### Reviewer C

> Unfortunately, I found the paper quite hard to follow. It assumes a lot of
> background knowledge on diverse topics such as the chase and network
> data planes, and I don't think that the background section of the paper does
> enough to get a reader on board. I found sections 3.2 and 3.3 quite confusing
> despite already being familiar with the chase.

We apologize for the confusing nature of our exposition. We will include more
detail and explanatory examples in these early sections.

> Many definitions there and elsewhere in the paper are incomplete, confusing,
> seemingly erroneous, or lacking intuition-providing explanations (see detailed
> comments).

We agree that the definitions are confusing and lacking intuition-providing
explanations. As discussed above, we will fix this in the camera ready.

There are also a few that had unfortunate typos which we have already fixed,
however the underlying definitions are themselves correct. We could not find any
that were incomplete. We will make sure to add explanatory prose to clarify the
slightly counter-intuitive nature of the definitions.

#### Detailed Feedback

> 61: "solved the data plane emulation" [problem]

Thanks! We’ve made this change.

> 123: "first-order formulas": this gives the impression that the chase can
> solve any first-order theory, but the chase may not terminate. Also, the chase
> constructs a universal instance; Otherwise, an instance that "has everything
> in it" trivially satisfies any TGDs and EGDs.

Good point. The chase does make certain structural requirements on the set of
first-order formulas. We hope to elide universality in our presentation here as
it is actually not the right notion for us. See the discussion on generative
solutions, though it seems worth mentioning for readers familiar with the chase.

> 172: what does "refines" mean here?

As defined on line 380, a pipeline-instance pair $s[\varsigma]$ refines another
pipeline-instance pair $t[\tau]$ if the semantics of $s[\varsigma]$, when
interpreted as a relation, are a subset of the semantics of $t[\tau]$ when
interpreted as a relation. Informally $s[\varsigma]$ may be defined on fewer
packets, but whenever $s[\varsigma]$ is defined on a packet, then $t[\tau]$ is
defined on that packet _and_ produces an equivalent output packet.

Upon review this early occurrence of "refines" seems out of place and
unnecessary. We'll revise the text to eliminate it.

> 187: "tabls"
> 208: use of bold for vector instead of overline
> 210: "is the case is"

Thanks! We’ve made these changes

> 3.1 could use some examples

We’ll be sure to include some!

> 221: A seems undefined?

The function $\mathcal{A}\llbracket-\rrbracket : \mathsf{Action} -> \mathsf{Pkt} -> \mathsf{Pkt}$ is
defined on lines 222 - 224.

> 237: Are phi and psi arbitrary first order sentences? It seems later they have
> to be in certain forms (disjunctive normal forms for head for example).

Thanks! We’ll add this detail.

> 242: what is x=beta(x)? should it be just beta(x)? It might be simpler to just
> say u=v for u, v in x then use beta.

Precisely. This is a typo. Beta is used in FDGs only.

> 242: what are the differences between FGDs and TGDs? Is it that TGDs don't
> allow equalities?

The head of TGDs can only be conjunctions of relation symbols (i.e. no
equalities). However FDGs are more expressive than just equalities. The heads of
FDGs are quantifier-free sentences over some decidable interpreted theory that
we leave abstract. In our case, it suffices to think of them as sentences in the
theory of quantifier-free fixed-width bit vectors (QFBV).

> 248: Consider citing some of the literature on dTGDs and disjunctive Datalog,
> e.g. https://dl.acm.org/doi/pdf/10.1145/2976736

Thanks! We’ll include this, and others.

> 261: Σs​ and Σt undefined

Thanks! These lines should be moved to section 5.2.

> 268: angle brackets undefined

Thanks! These are taken from [Fagin et al. 2005b] to indicate union over
disjoint schemas. We'll be sure to define them explicitly.

> 269: data exchange deserves at least a brief explanation

Yep! As with the previous two comments, this was a result of some last minute
reorganization. This needs to be moved to section 5.2

> 271: the dTGD form here does not have existential quantifiers. Handling of
> existential quantifiers is usually the most delicate part of a chase
> algorithm.

Good point. The standard treatment of existential quantifiers is insufficient
and indeed incorrect for our setting. We intentionally omit existential
quantifiers until section 9 where we discuss our more restrictive notion of
generative solutions. We’ll make sure to explicitly acknowledge this quirk of
our presentation.

> 280: what about the success case for FGDs?

As with EDGs in the standard chase [Fagin et al. 2005b], there is no success
case. Or rather, since we cannot say that “$d$ can be applied to $\rho$ with
homomorphism $h$”, if $\beta(x)$ succeeds we say “$d$ can_not_ be applied to
$\rho$ with homomorphism $h$”. This simply means that the rule does not apply.
We’ll be sure to draw attention to this subtlety.

> 298: what does x = Var mean?

That’s a typo, should be $\subseteq$, not $\eq$. It’s been fixed.

> 305: R1 or R0​? Seems zero-indexed from line 298

Yes! $R_0$. The formula on line 305 has it correctly as well. It’s been fixed.

> 305: Instead of a−j do you mean aj​?

Thanks! We’ve fixed this.

> 324: What is κ′?

Thanks! That should just be κ. We’ve fixed this.

> 376: what's the intuition of refinement? if a⊑a′, does this mean a' always
> produces more packets than a?

In fact, this line should have been deleted, as a⊑a′ is not used anywhere else.
Actions are total, deterministic functions, so here refinement is actually just
equality.

> 440: How is this different from a normal EGD 
> ∀...,R(k,i0),R(k,i1)=>i0=i1?

The point here is a bit subtle. These two are semantically equivalent, but the
bodies induce different homomorphisms. Homomorphisms are triggered based on the
body of the rule, not the head, so putting the fact that the keys must be equal
in both tables in the body of the rule increases the valid homomorphisms.

We defined it this way once-and-for-all to avoid having to tweak the definition
again in 6.1.

> 6.1: How do priority homomorphisms work for rule bodies with multiple atoms?
> Some examples will be helpful.

It uses a lexicographic order. We’ll include this subtlety, and augment the
example on line 551 to incorporate it.

> 547: is variable i overloaded/shadowed here?

Yep! This is a typo. The subscript should be $j = 1$. We’ve made this change.

> 850: Bu​ should be Bi​

Thanks! We’ve made this change.

> 9.2: Now there are existentials, so termination is no longer obvious. Does
> your algorithm terminate?

Absolutely. As before, we have no cycles of any kind in our dependencies. We’ll
be sure to state this explicitly.

> 1150: I find it easier to understand the problem by thinking of data plane
> emulation as a program synthesis problem, since this gives the intuition of
> why we should use disjunctive dependencies. This connection is made here in
> the related work section but it might be beneficial to mention this early on.

Thanks! We’ll add this analogy to an early section.
